### 2020-09-04 manuscript_Huang.tex
Files initialized.

\documentclass[12pt,aip,cha,reprint,nofootinbib]{revtex4-1}
\usepackage{amsfonts,amssymb,amsmath,times}%
\usepackage{graphicx}
\usepackage{bm}
\usepackage{enumerate}
\usepackage{color}

% \linespread{1}

\begin{document}

\title{Characterizing dynamical transitions by statistical complexity measures based on ordinal pattern transition networks}

\author{Min Huang}
	\affiliation{School of Physics and Electronic Science, East China Normal University, Shanghai 200062, China}

\author{Zhongkui Sun}
	\affiliation{Department of Applied Mathematics, Northwestern Polytechnical University, Xi'an 710072, China}

\author{Jie Zhang}
	\affiliation{Institute of Science and Technology for Brain-Inspired Intellegence, Fudan University, Shanghai 200433,  China}

\author{Reik V. Donner}
        \affiliation{Department of Water, Environment, Construction and Safety, Magdeburg--Stendal University of Applied Sciences, Breitscheidstra{\ss}e 2, 39114 Magdeburg, Germany}
        \affiliation{Potsdam Institute for Climate Impact Research (PIK) -- Member of the Leibniz Society, Telegrafenberg A31, 14473 Potsdam, Germany}

\author{Shuguang Guan}
	\email[corresponding author: ]{sgguan@phy.ecnu.edu.cn}
	\affiliation{School of Physics and Electronic Science, East China Normal University, Shanghai 200062, China}

\author{Yong Zou}
	\email[corresponding author: ]{yzou@phy.ecnu.edu.cn}
	\affiliation{School of Physics and Electronic Science, East China Normal University, Shanghai 200062, China}

\date{\today}

\begin{abstract}
In the recent decade, complex network approaches are emerging as complementary novel ways to perform nonlinear time series analysis, in particular showing many insights that are hidden in the traditional methods. In this work, we focus on one of particular method of characterizing ordinal pattern transition network for time series data. More specifically, we generalize some traditional estimators of statistical complexity measures (SCM), which explicitly disclosing inhomogeneous frequencies of ordinal pattern transitions in the time series networks. To show the usefulness of the estimators, we use SCM to characterize different dynamical transitions, as illustrated by examples from the numerical model of  Logistic map and real time series of fluid experiments and ECG data. Both results of numerical and experimental data demonstrate that the transition frequencies between different ordinal patterns improve the SCM estimations, which are especially important for finite length of time series analysis, hence providing substantial information to the existing complex network approaches for nonlinear time series analysis. 
\end{abstract}

\pacs{05.45.Ac, 05.45.Tp, 89.75.Fb}
\maketitle

\section{Introduction}
In the recent decade, complex network based approaches have emerged as prominent tools for nonlinear time series analysis \cite{Kantz97,Sprott2003}, which have found many interesting applications to experimental data \cite{ZouPR2018}. For instance,

There are several major methods are of particular importance, including recurrence network \cite{MarwanPLA2009,Donner2010a}, visibility graphs \cite{Lacasa2008,Nunez2012}, and transition networks \cite{Nicolis2005} and cycle network \cite{ZhangPRL2006}. Depending on the particular questions at hand, we note that each of these complex network approaches has various variants for different applications, for instance, constructing a single network for a univariate time series or interacting networks (multiplex or multilayer) networks for coupled time series. This line of research is undergoing fast developments for the time being as reviewed in \cite{ZouPR2018}.

The focus of this paper is ordinal pattern transition networks. The advantages and some recent results.

The basic idea of ordinal partition network method originates from identifying ordinal patterns of time series \cite{BandtPRL2002}, which is a well developed concept in nonlinear time series analysis. For instance, permutation entropy and the associated statistical analysis \cite{BandtPRL2002}. In addition, ordinal symbolic representation of time series has found a number of interesting applications in science and engineering, for instance, biomedical recordings \cite{AmigoPTRSA2014}, finance \cite{ZaninChaos2008}, climate sciences \cite{BarreiroChaos2011}. Some recent progress has been comprehensively reviewed in Ref. \cite{AmigoPTRSA2014}. More specifically, given a one-dimensional time series $\{ x(i)\}_{i=1, \cdots, L}$ comprising of $L$ sampled points from a dynamical system, we first reconstruct the corresponding phase space by time delay embedding $\vec{x}(i) = [x(i), x(i+\tau), \cdots, x(i+(D-1)\tau)]$ with dimension $D$ \cite{Takens1981,Kantz97}. Next, we represent each embedded phase point $\vec{x}(i)$ by its corresponding rank order, which is represented by a symbol $\pi_{i}$. When sliding windows from $i=1$ to $N = L - (D - 1)\tau$ in the embedded space, a symbolic representation of the trajectory $\pi_i$ is produced. It is possible to derive information about the dynamics of the underlying system by assessing different probabilities of ordinal patterns which is typical for a time series resulted from a deterministic system. More generally, we obtain a probability distribution $P$ whose elements $p(\pi_{i})$ are the frequencies associated to pattern $\pi_{i}$, $i = 1, \cdots, D!$. Clearly, $P$ provides a significant feasible way to estimate the probability distribution function for a given time series, which plays a crucial role in extracting statistical complexities of the underlying dynamical systems \cite{BandtPRL2002,AmigoPTRSA2014}.

Relatively easy computation of $P$ by ordinal patterns for the underlying system prompts another successful application to compute statistical complexity measure \cite{kowalskiEntropy2011}. Traditionally, statistical complexity measures (SCMs) are defined as the product of the normalized Shannon entropy and disequilibrium \cite{kowalskiEntropy2011}, which capture specific organizational properties of structure and patterns in a process. Following the literature \cite{rossoPRL2007,LopezPLA1995,kowalskiEntropy2011}, in this work we consider Jensen-Shannon SCM which is defined as $\mathcal{C}_{JS}[P] = \mathcal{H}[P]  \cdot  Q_{JS}[P, P_e]$, where $\mathcal{H}[P]$ is the normalized Shannon entropy and $Q_{JS}[P, P_e]$ is the so-called disequilibrium function measuring the distance of the given $P$ to the uniform distribution $P_e$, i.e., by the Jensen-Shannon divergence. Thus, $Q_{JS}[P, P_e]$ is defined as $Q_{JS}[P, P_e] = Q_{0} \cdot JS[P, P_e]$ where $JS[P, P_e]$ is the Jensen-Shannon divergence between $P$ and $P_e$ and $Q_{0}$ is the normalization factor. Furthermore, the complexity-entropy plane $\mathcal{H} \times \mathcal{C}_{JS}$ has been intensively used to distinguish chaotic systems from stochastic ones \cite{rossoPRL2007}. The computation of the complexity-entropy planes has benefit from the ordinal pattern symbolization of time series and has found various applications \cite{rossoPRL2007,kowalskiEntropy2011}. However, in the complexity-entropy analysis, there are some chaotic maps that could be easily confused as random noise because no clear separable regions are available to differentiate dynamics as reported in \cite{BorgesAMC2019}. So far, it remains to be a challenging task to disclose the possible non-monotonic relationship between complexity and entropy \cite{MartinPLA2003}.

Note that, in all above discussions of ordinal pattern analysis, the transition behavior between patterns remains largely unclear. Recently, ordinal partition network representations have been proposed to capture the transition behavior of the ordinal patterns \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}, which sheds novel insight on the standard ordinal symbolic analysis of time series. In particular, each ordinal pattern is considered as a vertex in the graph and a directed weighted edge connecting two patterns in the graph is established depending on the transition frequency. One of advantages of ordinal pattern transition networks is that we can provide pronounced distinction among different time series with short time series \cite{MichaelChaos2015,BorgesAMC2019}.

In this work, we generalize the traditional statistical complexity measures by incorporating pattern transition probabilities. The advantages are the following: (i) All SCMs successfully show distinctions between different dynamics, including time series from numerical models and experimental data; (ii) These generalized SCMs help to capture the consistent relationship between statistical complexity measures and chaoticity; (iii) All SCMs are sensitive to dynamical transitions of period doubling, band merging, inner and outer crisis.

The structure of the paper is organized as follows: In section \ref{sec:OPW}, we show two slightly different ways to obtain the transition matrix of the corresponding ordinal pattern transition networks. Furthermore, based on this matrix, we propose in section \ref{sec:SCM} to compute the statistical complexity measures from three different perspectives including static pattern frequency properties, dynamic pattern transition properties and both. We will use one example to emphasize the important effects of embedding parameters in section \ref{sec:embeddings}. The usefulness of SCMs will be illustrated by four time series of typical dynamical regimes of the Logistic map. In section \ref{sec:plane}, the traditional complexity-entropy planes of each SCM will be discussed while the control parameter is systematically changed. In order to characterize the dynamical transitions, we expand the SCM values versus the control parameter in section \ref{sec:transi}. Finally, we show that SCMs successfully distinguish different dynamical regimes by experimental time series from flow data and human ECG data in \ref{sec:time}. Some discussions will be provided in the last section \ref{sec:con}.

\section{Ordinal pattern transition matrix} \label{sec:OPW}
Most of the current studies compute permutation entropy $\mathcal{S}_O$ focusing on the frequencies of ordinal patterns, which however overlooked the transition behavior between  patterns. To emphasize the transitions between any pair of two patterns, measures of transitional complexity have been further proposed in \cite{zhangSciRep2017,McCullough2017b} to quantify the resulting OPTNs. To this end, we indicate each directed link by its transition frequency $w_{ij} = p({\pi_i \to \pi_j})$, following the time iterations of the series. Calculating the transition frequency for each pattern, we obtain a weighted directed network characterized by a weighted adjacency matrix $W = \{ w_{ij} \}, i, j \in [1, D!]$.

Based on the transition matrix $\mathbf{W}$ of an OPTN, we note that there two slightly different ways to introduce normalization factors in $W$. The first option is to use a global normalization to ensure $\sum_{i,j}^{D!} w_{ij} = 1$ \cite{zhangSciRep2017}, which results in a globally normalized transition entropy as termed below.

The second option of normalization has been proposed by McCullough {\textit{et al}} \cite{McCullough2017b} that they first considered the local out-link transition frequency from pattern $\pi_i$ to $\pi_j$
\begin{equation} \label{eq:localTp}
p_{\pi_i \to \pi_j} = \left \{ \begin{aligned}
				& 0,  \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \text{if} \;\;\; \pi_i = \pi_j \\
				& \frac{w_{ij}}{\sum_{j, j \neq i} w_{ij}}, \;\;\;\;\;\; \text{if} \; \pi_i \neq \pi_j.\\
				\end{aligned}
				\right.
\end{equation}		
Note that the transition frequency of Eq.~\eqref{eq:localTp} is pattern (row-) wise normalized. This case will be referred to as node-wise out-link normalized transition matrix $\mathbf{W}$ in the following. In this way, it easily captures heterogeneous behavior of both static frequencies of different patterns and dynamic transition frequencies between patterns.

In addition, we need to consider self-loops that may affect the numerical estimation of $\mathbf{W}$ which is especially important for time series from continuous systems. More specifically, it has been demonstrated in \cite{zhangSciRep2017} that there are about $99\%$ self-loops in the time series of R\"ossler system integrated by a step size $h = 0.01$, while about $1\%$ non-self-loops are hidden in $\mathbf{W}$. In addition, self-loops are related to the temporal correlation of time series \cite{BorgesAMC2019}. On the other hand, neglecting self-loops in $\mathbf{W}$ emphasizes the transition behavior between different patterns, which follows most of the existing studies of complex networks for computational simplicity and theoretical concerns \cite{CostaADPhy2007}. Therefore, we remove self-loops before computing the weighted matrix $\mathbf{W}$. In other cases of stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results $\mathbf{W}$ \cite{BorgesAMC2019}. We argue that whether to remove or to consider self-loops depends on the particular processes under study. Anyway, we have verified that our results below do not change significantly while considering effects of self-loops.

\section{Statistical complexity measures} \label{sec:SCM}
In this section, we review briefly ordinal pattern frequency based SCM and propose two novel ways to define SCM, which explicitly consider transition frequencies of the resulting ordinal network representations for time series.

$\bullet$ SCM based on permutation entropy of pattern frequencies.

The permutation entropy of a time series is the Shannon entropy of the distribution of patterns $P = \{p_{\pi_1}, p_{\pi_2}, \cdots, p_{\pi_i} \}$, which is computed as
\begin{equation}
\mathcal{S}_{O}[P]= - \sum_{i=1}^{D!} p_{\pi_i} \log p_{\pi_i}. 
\end{equation}
The normalized permutation entropy $\mathcal{H}_O$ is defined as 
\begin{equation} \label{eq:Ho}
\mathcal{H}_{O}[P] = \mathcal{S}_{O}[P] / \mathcal{S}_{O, max}, 
\end{equation}
where $\mathcal{S}_{O, max} = \log D!$ is resulted from the case of uniform distribution of the same number of patterns $P_e = \{1/D!, \cdots, 1/D!\}$.  In the next step, we use some distance definition to measure the given $P$ to the uniform distribution $P_e$, which is the core of statistical complexity measures (SCM).  There are several different distance functions as illustrated in the literature, such as Euclidean norm, Wooters' distance, Kullback-Leibler relative entropy and Jensen-Shannon divergence \cite{kowalskiEntropy2011}. In this work, the difference between the distribution $P$ and $P_e$ is characterized by the Jensen-Shannon divergence $Q_{JS}[P] = Q_0 \cdot JS[P, P_e]$ where $JS[P, P_e]$ is calculated by 
\begin{equation}
JS[P, P_e] = \mathcal{S}_O[\frac{P+P_e}{2}] - \frac{1}{2}\mathcal{S}_O[P] - \frac{1}{2}\mathcal{S}_O[P_e], 
\end{equation}
where $Q_0$ is a normalization factor with its value equal to the inverse of the maximal distance $JS[P, P_e]$, which ensures $Q_{JS}[P] \in [0, 1]$. This maximal distance is obtained when the distribution $P$ has probability one at one pattern and zeros at all other patterns. The general expression of $Q_0$ reads  
\begin{equation} \label{eq:Q0}
Q_0 = -2 \{(\frac{N+1}{N}) \log (N+1) -2 \log (2N) + \log (N)  \}^{-1}
\end{equation}
where $N = D!$ is the number of maximal patterns. For large $D$, we have $Q_0 \approx 1 /  \log 2$.

The function $Q_{JS}[P]$ is different from zero if the system has a kind of privilege over patterns among the accessible ones, reflecting determinism of the system. Finally, the SCM based on permutation entropy of node (pattern) frequencies is defined as 
\begin{equation}
\mathcal{C}_{O}[P] = Q_{JS}[P] \cdot \mathcal{H}_{O}[P].
\end{equation}
This quantity is static in the sense that it quantifies the amount of information stored by ordinal frequencies in the system and its disequilibrium of the observed parts of its accessible states in comparison to uniform distribution \cite{LopezPLA1995}.

$\bullet$ SCM based on transition entropy by globally normalized weighting matrix $W$.

In this case, the sum of all entries of the transition matrix $W$ is 1 and the Shannon entropy of transition frequencies between any pair of ordinal patterns is therefore 
\begin{equation}
\begin{split}
\mathcal{S}_{W}[W] &= - \sum_{i=1}^{D!} \sum_{j = 1; j \neq i}^{D!} w_{ij} \log w_{ij}, \\
& = - \sum_{i=1}^{D!} \sum_{j =1; j \neq i}^{D!} p_{\pi_i \to \pi_j} \log p_{\pi_i \to \pi_j}. 
\end{split}
\end{equation}
The normalized entropy is hence defined 
\begin{equation}
\mathcal{H}_{W}[W] = \mathcal{S}_{W}[W] / \mathcal{S}_{W, max}, 
\end{equation}
where $\mathcal{S}_{W, max} = \log D! (D! - 1)$ and the self transitions are excluded. Furthermore, there are $D! (D! - 1)$ pattern transitions in the uniformly distribution function $P_e$. Thus, in a full analogy with node-based SCM, the Jensen-Shannon distance is therefore computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$, where $JS[W, P_e]$ is computed as 
\begin{equation}
JS [W, P_e] = \mathcal{S}_{W}[\frac{W + P_e}{2}] - \frac{1}{2}\mathcal{S}_{W}[W] - \frac{1}{2}\mathcal{S}_{W}[P_e]. 
\end{equation}
We note that we have the same expression of the normalization factor $Q_0$ while replacing $N = D! (D! - 1)$ in Eq. \eqref{eq:Q0}. This further yields the SCM 
\begin{equation}
\mathcal{C}_{W}[W] = Q_{JS}[W] \cdot \mathcal{H}_{W}[W].
\end{equation}
This quantify is dynamic in the sense that it considers the state transition information stored in the system.

$\bullet$ SCM based on transition entropy by node-wise out-edge normalized
weighting matrix $W$.

In this case, the ordinal transition matrix is node-wised normalized, i.e., row
sum of $W$ is 1. Therefore, we first introduce node-wise out-edge transition entropy which is calculated 
\begin{equation}
\begin{split}
\mathcal{S}_{E}^{\pi_i}[W_{\pi_i}] &= - \sum_{j = 1; j \neq i}^{D!} w_{ij} \log w_{ij},  \\ 
&= - \sum_{j=1, j \neq i}^{D!} p_{\pi_i \to \pi_j} \log p_{\pi_i \to \pi_j}. 
\end{split}
\end{equation}
Furthermore, node-wise normalized transition entropy is 
\begin{equation}
\mathcal{H}_{E}^{\pi_i}[W_{\pi_i}] = \mathcal{S}_{E}^{\pi_i}[W_{\pi_i}] / \mathcal{S}_{E, max}, 
\end{equation}
where $\mathcal{S}_{E, max} = \log (D! - 1)$. Note that this normalization factor is the same for all patterns, namely, $\mathcal{S}_{E, max}^{\pi_i} = \mathcal{S}_{E, max}^{\pi_j}$ and, therefore, the superscript is suppressed in the above definition. In addition, node-wise Jensen-Shannon distance is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$, where the disequilibrium 
\begin{equation}
JS^{\pi_i} [W_{\pi_i}, P_e] = \mathcal{S}_{E}^{\pi_i}[\frac{W_{\pi_i} + P_e}{2}] - \frac{1}{2}\mathcal{S}_{E}^{\pi_i}[W_{\pi_i}] - \frac{1}{2}\mathcal{S}_{E}^{\pi_i}[P_e]. 
\end{equation}
In this case, we note that we have the normalization factor $Q_0$ while replacing $N = D! - 1$ in Eq. \eqref{eq:Q0}. Taking into account node frequencies, the global Jensen-Shannon divergence is computed as 
\begin{equation}
Q_{JS}[W, P_e] = \sum_{i=1}^{D!} p(\pi_i) Q_{JS}^{\pi_i}[W_{\pi_i}], 
\end{equation}
which therefore yields an alternative definition of SCM of the transition network 
\begin{equation}
\mathcal{C}_{E}[W] = Q_{JS}[W, P_e] \cdot \mathcal{H}_{E}[W], 
\end{equation}
where $\mathcal{H}_{E}[W] = \sum_{i} p(\pi_i) \mathcal{H}_{E}^{\pi_i}[W_{\pi_i}]$. This quantity is dynamic in the sense that it quantifies both the static ordinal state frequencies and the dynamic state transition information stored in the system.

\section{Choice of embedding parameters} \label{sec:embeddings}
There are two important algorithmic parameters in constructing ordinal pattern transition networks, namely, embedding parameters of $D$ and $\tau$. In some cases, the choice of $D$ and $\tau$ may strongly change the results since $D$ and $\tau$ should compromise the time scales of the respective time series. In general applications of ordinal pattern analysis (i.e., permutation entropy computation $\mathcal{S}_O$), the choice of optimal $D$ and $\tau$ depends on the field of application, which requires experience of experts as systematically summarized in \cite{Riedl2013}.
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{period4_logisticExample.pdf}
\caption{\small{Example of a period-4 iterative time series ($r= 3.5$ in the Logistic map) to show the effects of embedding parameters $D$ and $\tau$. (a) iterative time series and the corresponding ordinal pattern series by different $D$ and $\tau$, (b) $D = \tau = 3$, (c) $D = \tau = 4$, (d) $D = \tau = 5$, and (e) $D = \tau = 6$. The respective permutation entropy values normalized by $\log D!$ as Eq. \eqref{eq:Ho} are $\mathcal{H}_O = 0.77$, $0$, $0.29$, and $0.11$ as indicated in the insets of each panel. (f-i) show the corresponding ordinal pattern transition networks. The weights of directed edges are suppressed due to two slightly different normalizations are introduced. Note in (g), only self-loops are available since $D = \tau = 4$ result in a unique pattern. Missing patterns are suppressed. }
\label{fig:embed}}
\end{figure*}
Here we use one example to show the important roles of $D$ and $\tau$ for a period-4 time series in Figure \ref{fig:embed}. In this example, we choose $D = \tau $ and the corresponding ordinal pattern series are significantly influenced by these parameter settings, which further result in different permutation entropy values of $\mathcal{H}_O$.  For instance, only one unique ordinal pattern is observed when $D = \tau = 4$ (Fig. \ref{fig:embed}(c, g)) resulting in $\mathcal{H}_O = 0$, which is due to the coverage of one full period by the particular embedding parameters. Other values of $D$ and $\tau$ yield non-zero entropy values. The results of Fig. \ref{fig:embed} seem to be pessimistic since $D$ and $\tau$ will change the positions in the complexity-entropy planes. In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for periodic windows when the control parameter $r$ is varied. In such cases, we should focus more on the transitions between different dynamic regimes. Instead, we show that the results are robust when the control parameters are systematically changed while using the same parameter settings of $D$ and $\tau$.

\section{Results for four typical dynamic regimes} \label{sec:four}
We illustrate the potential of the proposed SCMs to characterize different dynamical regimes and transitions by analysis of the Logistic map $x_{n+1} = r x_n (1 - x_n)$ with the control parameter in the interesting range of $r \in [3.5, 4]$ with a step size of $\Delta r = 0.001$. In the range of the parameter $r$, various dynamical regimes and transitions between them can be found, for instance, period doublings, band merging points, inner and outer crises, intermittency \cite{Kantz97}, which has been served as a pragmatic model for nonlinear time series \cite{MarwanPLA2009}.

First, we show several typical regimes of particular control parameter $r$, i.e., periodic dynamics, band merging, laminar states and outer crisis, we investigate the SCMs in more detail which have been summarized in Tab \ref{tableLog}. We note that several chaos-chaos transitions have been discussed in the following, namely, the band merging corresponds to intermittency, the inner crisis to certain chaos-chaos transition and the outer crisis to fully chaotic dynamics. 
\begin{table}[htb]
    {\begin{tabular}{l  l  l  l  l  l  l  l}
    \hline
    Regime & $r$      & $\mathcal{H}_O$ & $\mathcal{C}_O$ & $\mathcal{H}_W$ & $\mathcal{C}_W$ & $\mathcal{H}_E$   & $\mathcal{C}_E$  \\
    \hline
    Period 3      & 3.83 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \hline
    Band merge    & 3.679 & 0.978 & 0.043 & 0.595 & 577 & 0.211 & 0.205  \\
    \hline
    Laminarity    & 3.791 & 0.986 & 0.029 & 0.633 & 0.609 & 0.280 & 0.269 \\
    \hline
    Outer crisis  & 4.0 & 0.994 & 0.014 & 0.677 & 0.638 & 0.359 & 0.339 \\
    \hline
    \end{tabular}}
   \caption{Statistical complexity measures for four different control parameters of the Logistic map (embedding parameters: $D = \tau = 6$).   \label{tableLog}}    
\end{table}
In Tab. \ref{tableLog}, we have used embeddings $D = \tau = 6$. Therefore, all SCMs values are zeros for period-3 series because each embedded vector covers two full periods (c.f., Fig. \ref{fig:embed}). Other choice of $D$ and $\tau$ often yields non-zero values for period-3 regime as discussed in Fig.\ref{fig:embed}. For other cases of band merging, laminarity and outer crisis, all SCMs values are significantly larger than zeros. The pattern frequency based SCMs of $\mathcal{C}_O$ show smaller values for all regimes, which further lead to the lower right locations in the complexity-entropy planes in Fig. \ref{fig:fBmNoise}(a). However, for transition matrix based SCMs, their entropy $\mathcal{H}_W$, $\mathcal{H}_E$ and complexity $\mathcal{C}_W$, $\mathcal{C}_E$ show more consistent results while considering the chaoticity level of the system. In particular, SCMs are the largest when $r = 4$ in the outer crisis regime, and the SCMs are the smallest when $r = 3.679$ in the band merging case, while the laminarity regime has the intermediate SCM values.

\section{Results on complexity-entropy planes} \label{sec:plane}
In this section, we show the complexity-entropy planes by varying the control parameter $r$ in the illustrative Logistic model.  Note that the Logistic map has been discussed in the framework of SCM \cite{RossoPRE2007,MartinPLA2003}, but only for pattern frequency based complexity. For each control parameter $r$, we construct ordinal pattern transition network from a time series of length $L = 31000$. In order to exclude transient behavior we remove the first $1000$ values from the data series. We have repeated the following calculations for embeddings $D = \tau \in [2, 7]$.

When only one unique ordinal pattern is identified following the time series, we set the SCMs values $0$, which is reasonable since only self-loop is observed. This happens only if $D$ is an integer factor of the period for periodic windows of the Logistic map (e.g., Fig. \ref{fig:embed}(g)).  Another notion is the data requirements for computing entropies and complexity measures. For reliable estimation of SCMs, a condition of the length of time series $L$ larger than $D!$ has been satisfied while $3 \leq D \leq 7$\cite{rossoPRL2007,kowalskiPhyD2007}. In the particular example of Logistic map, one may use longer data lengths (i.e., we have tested it up to $5 \times 10^6$), but this does not improve the estimate very much.

Traditionally, SCM have been applied to distinguish chaos from noise by means of the complexity-entropy plane. In particular, this plane shows the statistical complexity measures ($\mathcal{C}_O$, $\mathcal{C}_W$ and $\mathcal{C}_E$) as a function of the corresponding normalized entropy values ($\mathcal{H}_O$, $\mathcal{H}_W$ and $\mathcal{H}_E$). The above SCMs quantify both randomness and correlation structures of time series, which in consequence result in a non-trivial function of entropy  values. Chaotic systems present maximal complexity while stochastic systems have lower values of complexity, hence presenting in different regions in the complexity-entropy plane\cite{rossoPRL2007}. Furthermore, SCM values lie in a range of a minimum $\mathcal{C}_{min}$ and a maximum $\mathcal{C}_{max}$. A general algorithm for computing the bounds $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$ is provided in \cite{martinPhyA2006}. Figure \ref{fig:fBmNoise} shows the complexity-entropy planes for embedding parameters $D = \tau = 6$. We emphasize that the results in Fig. \ref{fig:fBmNoise} are qualitatively the same when $D$ and $\tau$ are varied in the interval $[3, 7]$, and we choose $D = 4, 5, 6$ as illustrative examples as shown in Fig. \ref{fig:fBmNoise}. Furthermore, the computation of $\mathcal{C}_{max}$ depends on the embedding $D$ since $D$ determines the number of patterns and pattern transitions which are included in the definition of SCMs. However, $\mathcal{C}_{min}$ shows less dependence on $D$. 
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{CompEntropy_LogisticHC.pdf}
\caption{\small{Complexity-entropy planes (Jensen-Shannon divergence) to show statistical complexity measures for example series of the logistic map for varying the control parameter $r$. (a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$. The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy. (a-c) for embedding $D = 4, 5$ and (d-f) for $D = 6$. In (a, d), the bottom-right rectangles of dotted lines highlighted the ambiguous regions for the traditional pattern frequency based SCM analysis. }  \label{fig:fBmNoise}}
\end{figure*}

In this complexity-entropy plane, pattern frequency based SCM values lie around the maximum curve $\mathcal{C}_{max}$. It is clear to see that the displacements of the dots depend on the embedding $D$ as shown in Fig. \ref{fig:fBmNoise}(a, d). In addition, there are many dots in the bottom-right position of the plane, which have been highlighted by rectangles of dotted lines. Indeed, this is a challenging region for SCMs to distinguish chaotic time series from random noise in Fig. \ref{fig:fBmNoise}(a, d), which has been reported in \cite{BorgesAMC2019}, too. It is not an easy task to correctly tell the difference between chaos (i.e., chaotic regimes of large $r$ in the Logistic map) and noise. This is because the SCMs of $\mathcal{H}_O$ and $\mathcal{C}_O$ are static in the sense that only pattern frequencies have been considered in the definitions. The results that  $\mathcal{H}_O$ and $\mathcal{C}_O$ are displaced close to $\mathcal{C}_{max}$ curves have been observed when the embedding $D$ is increased (Fig. \ref{fig:fBmNoise}(a, d)).

For other pattern transition based SCMs of $\mathcal{C}_W$ vs $\mathcal{H}_W$ and $\mathcal{C}_E$ vs $\mathcal{H}_E$, we also find that the dots are distributed close to the respective maximum lines of $\mathcal{C}_{max}$. Increasing the embedding dimension further to $D=6, 7$, the dots are covered by the left uprising branch the parabola curve $\mathcal{C}_{max}$ (Fig. \ref{fig:fBmNoise}(b, d)). The definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes into account the information of both pattern frequencies and pattern transitions, which results in that the dots are closer to $\mathcal{C}_{max}$ in a relative faster manner (Fig. \ref{fig:fBmNoise}(c, f)). The most interesting result is that no points lie in the bottom-right challenging regions of ambiguity. It is the deterministic transition behavior among patterns that makes the distinction from random noise.

%For a better comparison 
%\begin{figure*}
%	\centering 
%	\includegraphics[width=2\columnwidth]{CompEntropy_fBm.pdf}
%\caption{\small{Similar with Fig. \ref{fig:fBmNoise}. Complexity-entropy planes for example series of the fractional Brownian motion for varying Hurst exponent $H$. (a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$. The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy. (a-c) for embedding $D = 4, 5$ and (d-f) for $D = 6$. }  \label{fig:CompEntrFBM}}
%\end{figure*}

Based on the results of Fig. \ref{fig:fBmNoise}, we conclude that pattern transition behavior provides novel insights for defining SCMs, which complement the traditional SCMs.

\section{Characterizing dynamical transitions} \label{sec:transi}
The Logistic map experiences various bifurcation scenarios when the control parameter $r$ is systematically increased. However, these dynamical regimes and regime transitions are hidden in the complexity-entropy planes. In this subsection, we aim to expand the figures by showing complexity measures depending on the control parameter $r$. Our motivation is to verify whether SCMs computed from ordinal pattern transition networks are able to detect the dynamical transitions when the control parameter $r$ is varied with a step size $0.001$. Technically, it is a rather simple idea to check whether the normalized entropy $\mathcal{H}$ and SCMs $\mathcal{C}$ are able to track the changes of dynamics. Furthermore, we compare our results with the spectrum of the Lyapunov exponents.

\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{logisticEntropy.pdf}
\caption{\small{Behavior of different statistical complexity measures for the Logistic map in dependence on the control parameter $r$. (a) bifurcation diagram, (b) the spectrum of the maximal Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{S}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$. (g, h) are entropy $\mathcal{S}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$. Several major periodic windows have been highlighted by grey background. Vertical dotted lines correspond to the cases summarized in Tab. \ref{tableLog}. } \label{fig:bifurcation}}
\end{figure*}

Overall, all SCMs curves follow the changes in the bifurcation diagram, for instance, showing distinctions in periodic windows  (highlighted by grey color in Fig. \ref{fig:bifurcation}). Except $\mathcal{C}_O$, all other SCMs behave in a much similar manner as the Lyapunov exponents. Namely, as the control parameter $r$ is increased, the chaoticity level grows reaching a maximum at $r = 4$, which, therefore, leads to an increase in the statistical complexity. This similar increasing trend is absent in the case of pattern frequency based complexity measure $\mathcal{C}_O$ (Fig. \ref{fig:bifurcation}(c)). Note that it is a known fact that pattern frequency dependent SCMs often do not follow the growth in the chaoticity in the Logistic map \cite{MartinPLA2003}, which can be improved by Wootter's distance function, instead of the Shannon-Jensen divergence that we used in this work. From this point of view, pattern transition frequency based measures of $\mathcal{C}_W$ and $\mathcal{C}_E$ are more reasonable in the sense that they easily track the growth level of chaoticity.

There is another improvement for $\mathcal{H}_E$ and $\mathcal{C}_E$ when the Logistic map presents a period doubling bifurcation at $r = 3.544$ (Fig. \ref{fig:bifurcation}(g, h)). The SCMs of $\mathcal{H}_O$,  $\mathcal{C}_O$, $\mathcal{H}_W$ and $\mathcal{C}_W$ show non-zero values in this period doubling window. Furthermore, there are jumps at the bifurcation point $r = 3.544$, which have been also reported in \cite{BandtPRL2002}. To a large extent, the jumps should not be expected because complexity does not change when $r$ passes this point. We show in Fig. \ref{fig:transient} that the iterative transient points are responsible to affect ordinal pattern definitions, which results in the peculiar jumps of SCMs at $r = 3.544$. For estimations from time series of finite length of this work, $\mathcal{H}_E$ and $\mathcal{C}_E$ are not affected seemly. Certainly, we should not over interpret their capabilities since the numerical inaccuracy would be accumulated in a longer iterative process such that different ordinal patterns are identified. In addition, there are many periodic windows of different periods along the axis of control parameter $r$, which prevents us from using a unique predefined number of iterations as initial transients for time series of different period length. 
\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{period4_exampleTransients.pdf}
\caption{\small{Example to show the transient effects in generating proper bifurcation diagrams of the Logistic map, focusing on the window of period-4 to period-8. The time points get blurred when $r$ is close to $3.544$ when a short transient of $50$ iterates are removed from the analysis (a). In (b) $2000$ initial iterates are removed. }\label{fig:transient}}
\end{figure}
Meanwhile, these jumps have also been observed when the system bifurcates from period-3 to period-6.

\section{Real time series of fluid flows and ECG} \label{sec:time}
In this section, we use two different experimental time series to demonstrate that SCMs successfully characterize the complexity properties of the underlying systems, showing distinct values for different dynamics.

The first example is from a laboratory fluid experiment of Baroclinic instability which is used to study patterns of cyclones and anticyclones in climate systems \cite{Read_jfm_1992,ZouEPJST2008}. Depending on the parameters of rotation rate, temperature difference, viscosity and fluid density, this experimental system exhibits a rich variety of flow regimes. In particular, we focus on: (i) {Stable fluid:} the temperature signal of the stable flow exhibits periodic oscillation as the wave drift patterns arrive at the measured fixed point. (ii) {Quasi-periodic 2- frequency Amplitude vacillation (AV-2):} this case is identified as a 2-frequency quasi-periodic amplitude flux. The wave drift is composed of slow and regular oscillations in the temperature signal, but a fast modulation in the amplitude is also visible. AV-2 is characterized by periodic growth and a drop in wave amplitude with little change in waveform. (iii) {Quasi-periodic 3-frequency amplitude vacillation (AV-3): } We also analyze a quasi-periodic 3-frequency amplitude vacillation (AV-3) time series which is obtained from a three-dimensional direct numerical simulation in an air-filled rotating Baroclinic instability experimental \cite{Read_jfm_1992}. (iv) {Modulated amplitude vacillation (MAV):} this case is identified as a low-dimension stream, chaotically modulated and with waves of varying amplitudes. Amplitude modulation results in a complex temperature signal.

The second example time series comes from the physiological signals of human electrocardiogram (ECG) recordings that are collected from patients in a Coronary Care Unit and the Creighton University Cardiac Center. Ventricular arrhythmia are a very significant but poorly understood dynamical system, showing rich nonlinear properties. In this work, we focus on three different states: normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}. From the data bank of recordings of ECG during various rhythms, we selected 14 time series showing sinus rhythm (SR) prior to onset of arrhythmia, 12 time series of VT and 17 VF. The ECG recordings have been annotated with SR, VT and VF by skilled cardiologists on the basis of waveform morphology. Each recording consists of at least $10,000$ data points (20s) in length and 250 Hz in sampling frequency.

For choosing embedding parameters for both example time series, we follow the suggestions that are proposed in the respective literature \cite{ZouEPJST2008,smallCSF2002}. Although we are facing with the inevitable presence of non-stationary and noise in the experimental data, we have verified that the following results do not change significantly while embedding parameters are varied. In particular, we have treated embedding dimension as a free parameter which is varied in the interval as $D = 2, 3, \dots, 7$, while the embedding delay $\tau$ is the following. In the case of temperature time series from the fluid flows, embedding delay $\tau_{i} = 218s$ for stable solution, $\tau_{ii} = 136.5s$  for AV-2, $\tau_{iii} = 185.625$s for AV-3, and $\tau_{iv} = 118s$ for MAV. In the case of recordings of ECG time series, embedding delay $\tau = 5$. The results have been averaged over all subjects.

Generally speaking, the OPTN based statistical complexity measures of both the flow data and the ECG recordings enhance our understanding of the underlying nonlinear system, which are shown in Fig. \ref{fig:fluid}. More specifically, in the case of fluid data (Fig. \ref{fig:fluid}(a)), complexity measures of $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ show consistent complexity values for stable wave solution (SW), quasi-periodic 2-frequency Amplitude vacillation (AV-2), quasi-periodic 3-frequency Amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV). Namely, the complexity of these four cases has the following order: SW has the smallest complexity since it shows periodic oscillation of trivial recurrences, while MAV has the highest complexity since it is chaotically modulated amplitude vacillation. Quasiperiodic states show inter-mediate complexity values since these two cases have non-trivial recurrences but with low complexity. Interestingly, AV-3 presents higher complexity than AV-2. The relationship of complexity between periodic, quasiperiodic and chaotic solutions still holds for $\mathcal{C}_O$ although the corresponding values become comparable. In particular, it becomes hard for $\mathcal{C}_O$ to tell the difference between AV-2 and AV-3.

Qualitative similar results have been obtained for ECG recording series (Fig. \ref{fig:fluid}(b)). Clearly, different ECG recordings exhibit remarkably distinct complexity values. Specifically, SR recordings show pronounced highest complexity values of $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ since there are rather broad frequency bands between 50 and 100 beats/min in the signals \cite{smallCSF2002}. On the other hand, the complexity of VT has the smallest values comparable to the case of VF. Similarly to the case of flow data, SR shows the smallest value of $\mathcal{C}_O$ while VT and VF have comparable values that are relative higher than SR. 
\begin{figure*}
	\centering 
	\includegraphics[width=2\columnwidth]{fluidExample.pdf}
\caption{\small{Application of SCMs to experimental fluid data (a) and human recordings of ECG (b). The results have been averaged over embedding delays $D = 2, 3, \dots, 7$. In (a), stable wave (SW) $\blacklozenge$, quasiperiodic 2-frequency amplitude vacillation (AV-2) $\bullet$, quasiperiodic 3-frequency amplitude vacillation (AV-3) $\blacktriangle$, and chaotic modulated amplitude vacillation (MAV) $\blacksquare$. In (b), normal sinus rhythm (SR) $\blacktriangle$, ventricular tachycardia (VT) $\blacklozenge$ and ventricular fibrillation (VF) $\bullet$. } \label{fig:fluid}}
\end{figure*}

\section{Discussions} \label{sec:con}
complex network approaches for nonlinear time series analysis \cite{ZouPR2018}.

Some open problems regarding to the ordinal pattern analysis need to be further explored. Length dependence, further numerical performance to show distinctions between different dynamics (various models of maps, long range correlated fractional Brownian motion, fractional Gaussian noise etc)

For periodic windows of the Logistic map, one may use longer data lengths, but does not improve the estimate very much.

Generalization from univariate to bivariate time series analysis will be further discussed. We noticed that this remains to be a largely untouched area for statistical complexity analysis. It would be an interesting task to show the dependence of complexity on interaction.

\section{Acknowledgements}
Parts of this work have been financially supported by the National Natural Science Foundation of China (Grant No. 11872182 and 11835003). Y. Z. further acknowledges the support by Shanghai Municipal Science and Technology Major Project (No.2018SHZDZX01) and ZJ Lab.

\bibliographystyle{unsrt}
\bibliography{ref_Zou}

\end{document}

### 2020-09-29 manuscript_Huang.tex
Updates from Overleaf

@@
         \affiliation{Department of Water, Environment, Construction and Safety, Magdeburg--Stendal University of Applied Sciences, Breitscheidstra{\ss}e 2, 39114 Magdeburg, Germany}
-        \affiliation{Potsdam Institute for Climate Impact Research (PIK) -- Member of the Leibniz Society, Telegrafenberg A31, 14473 Potsdam, Germany}
+        \affiliation{Potsdam Institute for Climate Impact Research (PIK) -- Member of the Leibniz Association, Telegrafenberg A31, 14473 Potsdam, Germany}
 
@@
 \begin{abstract}
-In the recent decade, complex network approaches are emerging as complementary novel ways to perform nonlinear time series analysis, in particular showing many insights that are hidden in the traditional methods.
-In this work, we focus on one of particular method of characterizing ordinal pattern transition network for time series data.
-More specifically, we generalize some traditional estimators of statistical complexity measures (SCM), which explicitly disclosing inhomogeneous frequencies of ordinal pattern transitions in the time series networks.
-To show the usefulness of the estimators, we use SCM to characterize different dynamical transitions, as illustrated by examples from the numerical model of  Logistic map and real time series of fluid experiments and ECG data.
-Both results of numerical and experimental data demonstrate that the transition frequencies between different ordinal patterns improve the SCM estimations, which are especially important for finite length of time series analysis, hence providing substantial information to the existing complex network approaches for nonlinear time series analysis.
+In the recent decade, complex network approaches have been emerging as novel and complementary concepts of nonlinear time series analysis that are able to unveil many features that are hidden to more traditional analysis methods.
+In this work, we focus on one particular approach: the application of ordinal pattern transition network for characterizing time series data.
+More specifically, we generalize some traditional estimators of statistical complexity measures (SCM), which explicitly disclose heterogeneous frequencies of ordinal pattern transitions in the time series networks.
+To demonstrate the usefulness of the estimators, we employ SCM to characterize different dynamical transitions in the paradigmatic model system of the logistic map as well as real-world time series of fluid experiments and ECG data.
+The obtained results for both, artificial and experimental data demonstrate that the consideration of transition frequencies between different ordinal patterns improves the estimation of SCM, which is especially important for the analysis of time series with finite length.
 \end{abstract}
@@
 \maketitle
+
+\textcolor{red}{If this paper is dedicated for submission to Chaos, a lead paragraph needs to be added here.}
 
@@
 
-\section{Acknowledgements}
-Parts of this work have been financially supported by the National Natural Science Foundation of China (Grant No.
+\section*{Acknowledgements}
+Parts of this work have been financially supported by the National Natural Science Foundation of China (grant no.
 11872182 and 11835003).
-Y.
-Z.
-further acknowledges the support by Shanghai Municipal Science and Technology Major Project (No.2018SHZDZX01) and ZJ Lab.
+YZ further acknowledges the support by Shanghai Municipal Science and Technology Major Project (grant no.
+2018SHZDZX01) and ZJ Lab.
 

### 2020-10-28 manuscript_Huang.tex
updated.

@@
 
-\textcolor{red}{If this paper is dedicated for submission to Chaos, a lead paragraph needs to be added here.}
+\begin{quotation}
+In the recent decade, tools of traditional nonlinear time series analysis have been undergoing fast developments benefiting concepts from complex network theories.
+Along this line of research topics, ordinal pattern transition time series network has been providing complement insights to the permutation entropy estimators for time series data.
+Traditionally, permutations based on order patterns are simple and easily implemented tools for estimators of statistical complexity measures (SCMs), which however rely on the static pattern frequencies only.
+Still, much additional knowledge of the order pattern transitions to the SCMs has not been fully considered.
+In this work, we generalize the estimators of SCMs by means of ordinal pattern transition networks which take into account the pattern transition properties explicitly.
+The usefulness of the generalizations will be demonstrated by time series of both model and experimental data.
+\end{quotation}
 
@@
 In the recent decade, complex network based approaches have emerged as prominent tools for nonlinear time series analysis \cite{Kantz97,Sprott2003}, which have found many interesting applications to experimental data \cite{ZouPR2018}.
-For instance,
-
-There are several major methods are of particular importance, including recurrence network \cite{MarwanPLA2009,Donner2010a}, visibility graphs \cite{Lacasa2008,Nunez2012}, and transition networks \cite{Nicolis2005} and cycle network \cite{ZhangPRL2006}.
-Depending on the particular questions at hand, we note that each of these complex network approaches has various variants for different applications, for instance, constructing a single network for a univariate time series or interacting networks (multiplex or multilayer) networks for coupled time series.
-This line of research is undergoing fast developments for the time being as reviewed in \cite{ZouPR2018}.
-
-The focus of this paper is ordinal pattern transition networks. The advantages and some recent results.
+In this line of research, there are several major methods are of particular importance, including recurrence network \cite{MarwanPLA2009,Donner2010a}, visibility graphs \cite{Lacasa2008,Nunez2012}, transition networks \cite{Nicolis2005,MichaelChaos2015} and cycle network \cite{ZhangPRL2006}.
+Depending on the particular questions at hand, we note that each of these complex network approaches has various variants for different applications, for instance, constructing a single network for a univariate time series or interacting networks (multiplex or multilayer) networks for coupled time series, which has been systematically reviewed in \cite{ZouPR2018} recently.
+In this work, we focus on constructing ordinal pattern transition networks for time time series data, which show the advantages of easy implementations and wide applications to various data from different origins, including correlated stochastic processes, EEG and human cardiac series (ECG) \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}.
 
@@
 
-Note that, in all above discussions of ordinal pattern analysis, the transition behavior between patterns remains largely unclear.
+Note that, in all above discussions of ordinal pattern analysis, the transition behavior between patterns remains largely unconsidered.
 Recently, ordinal partition network representations have been proposed to capture the transition behavior of the ordinal patterns \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}, which sheds novel insight on the standard ordinal symbolic analysis of time series.
@@
 \section{Choice of embedding parameters} \label{sec:embeddings}
+{\color{red}We may leave out this section, if we plan a second paper which deals with embedding issue explicitly. }
 There are two important algorithmic parameters in constructing ordinal pattern transition networks, namely, embedding parameters of $D$ and $\tau$.
@@
 \section{Discussions} \label{sec:con}
+In summary, we have proposed to include ordinal pattern transitions into the traditional ordinal pattern frequency analysis, generalizing statistical complexity measures for nonlinear time series analysis.
+In terms of the pattern transitions in the time series networks, permutation entropy relies on the static pattern frequency only.
+Here we suggest two slightly different ways to consider the effects of pattern transitions which are based on (i) a globally normalized weight transition matrix $\mathbf{W}$ and (ii) node-wise normalized transition matrix.
+
 complex network approaches for nonlinear time series analysis \cite{ZouPR2018}.

### 2020-10-30 manuscript_Huang.tex
Continuous Rossler is included.

@@
 \section{Choice of embedding parameters} \label{sec:embeddings}
-{\color{red}We may leave out this section, if we plan a second paper which deals with embedding issue explicitly. }
+{\color{red}We may leave out this section, if we plan a second paper which deals with embedding issue explicitly. Or we have to present the following more precisely. }
 There are two important algorithmic parameters in constructing ordinal pattern transition networks, namely, embedding parameters of $D$ and $\tau$.
@@
 
-%For a better comparison 
-%\begin{figure*}
-%	\centering 
-%	\includegraphics[width=2\columnwidth]{CompEntropy_fBm.pdf}
-%\caption{\small{Similar with Fig.
-\ref{fig:fBmNoise}.
-Complexity-entropy planes for example series of the fractional Brownian motion for varying Hurst exponent $H$.
-(a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
-The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy.
-(a-c) for embedding $D = 4, 5$ and (d-f) for $D = 6$.
-}  \label{fig:CompEntrFBM}}
-%\end{figure*}
-
 Based on the results of Fig. \ref{fig:fBmNoise}, we conclude that pattern transition behavior provides novel insights for defining SCMs, which complement the traditional SCMs.
@@
 Meanwhile, these jumps have also been observed when the system bifurcates from period-3 to period-6.
+
+\section{Transitions in continuous system}
+The previous results have been further verified in the chaotic R\"ossler system \cite{Roessler1976}, namely, $(\dot{x}, \dot{y}, \dot{z}) = (-y-z, x+0.2y, b+z(x-5.7))$, where $b$ is the control parameter.
+There are some challenging problems for constructing OPTNs for time series from continuous systems.
+
+\begin{figure*}
+	\centering 
+	\includegraphics[width=2\columnwidth]{rosslerEntropy.pdf}
+\caption{\small{Similar with the caption of Fig.
+\ref{fig:bifurcation}.
+Behavior of different SCMs for the R\"ossler system varying the control parameter $b$.
+(a) bifurcation diagram, (b) the maximal Lyapunov exponent $\lambda$, (c) pattern frequency based permutation entropy $\mathcal{S}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
+(g, h) are entropy $\mathcal{S}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$.
+Several major periodic windows have been highlighted by grey background.
+Vertical dotted lines correspond to periodic windows.
+} \label{fig:bifurRossler}}
+\end{figure*}
 
@@
 In summary, we have proposed to include ordinal pattern transitions into the traditional ordinal pattern frequency analysis, generalizing statistical complexity measures for nonlinear time series analysis.
-In terms of the pattern transitions in the time series networks, permutation entropy relies on the static pattern frequency only.
-Here we suggest two slightly different ways to consider the effects of pattern transitions which are based on (i) a globally normalized weight transition matrix $\mathbf{W}$ and (ii) node-wise normalized transition matrix.
+The pattern transition properties of the corresponding time series networks provide novel insights for permutation entropy which relies on the static pattern frequency only.
+In addition, we suggest two slightly different ways to consider the effects of pattern transitions which are based on (i) a globally normalized weight transition matrix $\mathbf{W}$ and (ii) node-wise normalized transition matrix.
+Note that the node-wise based SCMs consider both individual pattern frequencies and and the transition frequencies among them.
 

### 2020-10-30 manuscript_Huang.tex
further changes made on Friday.

@@
 The previous results have been further verified in the chaotic R\"ossler system \cite{Roessler1976}, namely, $(\dot{x}, \dot{y}, \dot{z}) = (-y-z, x+0.2y, b+z(x-5.7))$, where $b$ is the control parameter.
-There are some challenging problems for constructing OPTNs for time series from continuous systems.
+There are some challenging problems for constructing OPTNs for time series from continuous systems, for instance, proper choice of sampling frequency and embedding parameters, which must depend on the particular time scales of the system.
+{\color{red} numerical details will be provided soon.
+}
 

### 2020-11-10 manuscript_Huang.tex
conclusion is added.

@@
 Several major periodic windows have been highlighted by grey background.
-Vertical dotted lines correspond to the cases summarized in Tab.
+Vertical dotted lines correspond to the cases as summarized in Tab.
 \ref{tableLog}.
@@
 \section{Transitions in continuous system}
-The previous results have been further verified in the chaotic R\"ossler system \cite{Roessler1976}, namely, $(\dot{x}, \dot{y}, \dot{z}) = (-y-z, x+0.2y, b+z(x-5.7))$, where $b$ is the control parameter.
+The previous results have been further verified in the chaotic R\"ossler system \cite{Roessler1976} while varying the control parameter, namely, $(\dot{x}, \dot{y}, \dot{z}) = (-y-z, x+0.2y, b+z(x-5.7))$, where $b$ is the control parameter.
 There are some challenging problems for constructing OPTNs for time series from continuous systems, for instance, proper choice of sampling frequency and embedding parameters, which must depend on the particular time scales of the system.
-{\color{red} numerical details will be provided soon.
-}
-
+To minimize the discussions on proper choices of further algorithmic parameters, we introduce Poincar\'e sections at $y=0$, $\dot{y}<0$ (for each value of $b$) and construct OPTNs from $N = 10000$ intersection points.
+The results have been shown in Fig.
+\ref{fig:bifurRossler}.
 \begin{figure*}
@@
 Behavior of different SCMs for the R\"ossler system varying the control parameter $b$.
-(a) bifurcation diagram, (b) the maximal Lyapunov exponent $\lambda$, (c) pattern frequency based permutation entropy $\mathcal{S}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
+(a) bifurcation diagram ($x$-component on the Poinar\'e sections), (b) the maximal Lyapunov exponent $\lambda$, (c) pattern frequency based permutation entropy $\mathcal{S}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
 (g, h) are entropy $\mathcal{S}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$.
 Several major periodic windows have been highlighted by grey background.
-Vertical dotted lines correspond to periodic windows.
 } \label{fig:bifurRossler}}
 \end{figure*}
+Rather similar results as Fig. \ref{fig:bifurcation} have been obtained in the chaotic R\"ossler system, identifying different dynamic transitions, where some periodic windows have been highlighted.
 
@@
 
-\section{Discussions} \label{sec:con}
+\section{Conclusions} \label{sec:con}
 In summary, we have proposed to include ordinal pattern transitions into the traditional ordinal pattern frequency analysis, generalizing statistical complexity measures for nonlinear time series analysis.
 The pattern transition properties of the corresponding time series networks provide novel insights for permutation entropy which relies on the static pattern frequency only.
+Therefore, this is another case showing the usefulness of ordinal pattern transition network for time series analysis \cite{ZouPR2018}.
 In addition, we suggest two slightly different ways to consider the effects of pattern transitions which are based on (i) a globally normalized weight transition matrix $\mathbf{W}$ and (ii) node-wise normalized transition matrix.
 Note that the node-wise based SCMs consider both individual pattern frequencies and and the transition frequencies among them.
+We show that these two new statistical complexity measures characterize the complexity properties of deterministic chaos successfully, far away from the lower right bottom of the complexity-entropy plane (Fig.
+\ref{fig:fBmNoise}).
+This helps to obtain deeper understandings of complexity analysis for time series of both deterministic chaos and stochastic systems, for instance, further numerical performance are required to show distinctions between different dynamics including various models of chaotic maps, long range correlated fractional Brownian motion and fractional Gaussian noise etc \cite{BorgesAMC2019,RossoPRE2007}.
+We have demonstrated the usefulness of SCMs by further results of time series for experimental fluid flows and ECG recordings, showing distinctive characteristics for different dynamics.
 
-complex network approaches for nonlinear time series analysis \cite{ZouPR2018}.
-
-Some open problems regarding to the ordinal pattern analysis need to be further explored.
-Length dependence, further numerical performance to show distinctions between different dynamics (various models of maps, long range correlated fractional Brownian motion, fractional Gaussian noise etc)
-
-For periodic windows of the Logistic map, one may use longer data lengths, but does not improve the estimate very much.
-
+Some open problems of traditional nonlinear time series analysis should be further considered, for example, length dependence for reliable estimations of SCMs.
+For iterative processes of maps, one may use longer data lengths, which however does not improve the estimate very much due to the accumulation effects of computer iterations (Fig.
+\ref{fig:transient}).
+For continuous systems, together with embedding parameters of $D$ and $\tau$, the sampling frequency is another important one that needs to be further explored.
 Generalization from univariate to bivariate time series analysis will be further discussed.
 We noticed that this remains to be a largely untouched area for statistical complexity analysis.
-It would be an interesting task to show the dependence of complexity on interaction.
+It will be an interesting task to show the dependence of complexity on interaction.
 

### 2020-11-10 manuscript_Huang.tex
updated.

@@
 \section{Choice of embedding parameters} \label{sec:embeddings}
-{\color{red}We may leave out this section, if we plan a second paper which deals with embedding issue explicitly. Or we have to present the following more precisely. }
 There are two important algorithmic parameters in constructing ordinal pattern transition networks, namely, embedding parameters of $D$ and $\tau$.

### 2020-11-21 manuscript_Huang.tex
updated on 22 Nov.

@@
 \caption{\small{Behavior of different statistical complexity measures for the Logistic map in dependence on the control parameter $r$.
-(a) bifurcation diagram, (b) the spectrum of the maximal Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{S}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
-(g, h) are entropy $\mathcal{S}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$.
+(a) bifurcation diagram, (b) the spectrum of the maximal Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{H}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
+(g, h) are entropy $\mathcal{H}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$.
 Several major periodic windows have been highlighted by grey background.
@@
 Behavior of different SCMs for the R\"ossler system varying the control parameter $b$.
-(a) bifurcation diagram ($x$-component on the Poinar\'e sections), (b) the maximal Lyapunov exponent $\lambda$, (c) pattern frequency based permutation entropy $\mathcal{S}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
-(g, h) are entropy $\mathcal{S}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$.
+(a) bifurcation diagram ($x$-component on the Poinar\'e sections), (b) the maximal Lyapunov exponent $\lambda$, (c) pattern frequency based permutation entropy $\mathcal{H}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
+(g, h) are entropy $\mathcal{H}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$.
 Several major periodic windows have been highlighted by grey background.
@@
 In addition, we suggest two slightly different ways to consider the effects of pattern transitions which are based on (i) a globally normalized weight transition matrix $\mathbf{W}$ and (ii) node-wise normalized transition matrix.
-Note that the node-wise based SCMs consider both individual pattern frequencies and and the transition frequencies among them.
+Note that the node-wise based SCMs consider both individual pattern frequencies and the transition frequencies among them.
 We show that these two new statistical complexity measures characterize the complexity properties of deterministic chaos successfully, far away from the lower right bottom of the complexity-entropy plane (Fig.
 \ref{fig:fBmNoise}).
-This helps to obtain deeper understandings of complexity analysis for time series of both deterministic chaos and stochastic systems, for instance, further numerical performance are required to show distinctions between different dynamics including various models of chaotic maps, long range correlated fractional Brownian motion and fractional Gaussian noise etc \cite{BorgesAMC2019,RossoPRE2007}.
+This helps to obtain deeper understandings of complexity analysis for time series of both deterministic chaos and stochastic systems, for instance, further numerical studies are required to show distinctions of various dynamical systems of chaotic maps, long range correlated fractional Brownian motion and fractional Gaussian noise etc \cite{BorgesAMC2019,RossoPRE2007}.
 We have demonstrated the usefulness of SCMs by further results of time series for experimental fluid flows and ECG recordings, showing distinctive characteristics for different dynamics.

### 2020-11-23 manuscript_Huang.tex
Updates from Overleaf

@@
 \begin{abstract}
-In the recent decade, complex network approaches have been emerging as novel and complementary concepts of nonlinear time series analysis that are able to unveil many features that are hidden to more traditional analysis methods.
-In this work, we focus on one particular approach: the application of ordinal pattern transition network for characterizing time series data.
-More specifically, we generalize some traditional estimators of statistical complexity measures (SCM), which explicitly disclose heterogeneous frequencies of ordinal pattern transitions in the time series networks.
-To demonstrate the usefulness of the estimators, we employ SCM to characterize different dynamical transitions in the paradigmatic model system of the logistic map as well as real-world time series of fluid experiments and ECG data.
-The obtained results for both, artificial and experimental data demonstrate that the consideration of transition frequencies between different ordinal patterns improves the estimation of SCM, which is especially important for the analysis of time series with finite length.
+Complex network approaches have been recently emerging as novel and complementary concepts of nonlinear time series analysis which are able to unveil many features that are hidden to more traditional analysis methods.
+In this work, we focus on one particular approach: the application of ordinal pattern transition networks for characterizing time series data.
+More specifically, we generalize some traditional estimators of statistical complexity measures (SCMs), which explicitly disclose heterogeneous frequencies of ordinal pattern transitions in the time series networks.
+To demonstrate the usefulness of the estimators, we employ SCMs to characterize different dynamical transitions in the logistic map as a paradigmatic model system, as well as real-world time series of fluid experiments and ECG data.
+The obtained results for both, artificial and experimental data demonstrate that the consideration of transition frequencies between different ordinal patterns leads to dynamically meaningful estimates of SCMs, which provide prospective tools for the analysis of time series with finite length.
 \end{abstract}
@@
 \begin{quotation}
-In the recent decade, tools of traditional nonlinear time series analysis have been undergoing fast developments benefiting concepts from complex network theories.
-Along this line of research topics, ordinal pattern transition time series network has been providing complement insights to the permutation entropy estimators for time series data.
-Traditionally, permutations based on order patterns are simple and easily implemented tools for estimators of statistical complexity measures (SCMs), which however rely on the static pattern frequencies only.
-Still, much additional knowledge of the order pattern transitions to the SCMs has not been fully considered.
-In this work, we generalize the estimators of SCMs by means of ordinal pattern transition networks which take into account the pattern transition properties explicitly.
-The usefulness of the generalizations will be demonstrated by time series of both model and experimental data.
+In the recent decade, traditional tools of nonlinear time series analysis have been undergoing fast developments benefiting from concepts from complex network theory.
+Along this line of research topics, ordinal pattern transition networks have been expanding the established concept of permutation entropy and provide complementary insights into the dynamical organization underlying time series data.
+Traditionally, permutations based on order patterns are simple and easy to implement concepts providing estimators of statistical complexity measures (SCMs), which however rely on static pattern frequencies only.
+Yet, much additional information can be exploited by including order pattern transitions into the definitions of SCMs -- an idea that however has not been widely developed and applied so far.
+In this work, we generalize existing estimators of SCMs by means of ordinal pattern transition networks which take into account the pattern transition properties explicitly.
+The usefulness of our generalizations of existing SCMs is demonstrated by using time series of both, model and experimental data.
 \end{quotation}
@@
 \section{Introduction}
-In the recent decade, complex network based approaches have emerged as prominent tools for nonlinear time series analysis \cite{Kantz97,Sprott2003}, which have found many interesting applications to experimental data \cite{ZouPR2018}.
-In this line of research, there are several major methods are of particular importance, including recurrence network \cite{MarwanPLA2009,Donner2010a}, visibility graphs \cite{Lacasa2008,Nunez2012}, transition networks \cite{Nicolis2005,MichaelChaos2015} and cycle network \cite{ZhangPRL2006}.
-Depending on the particular questions at hand, we note that each of these complex network approaches has various variants for different applications, for instance, constructing a single network for a univariate time series or interacting networks (multiplex or multilayer) networks for coupled time series, which has been systematically reviewed in \cite{ZouPR2018} recently.
-In this work, we focus on constructing ordinal pattern transition networks for time time series data, which show the advantages of easy implementations and wide applications to various data from different origins, including correlated stochastic processes, EEG and human cardiac series (ECG) \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}.
+In the recent decade, complex network based approaches have emerged as prominent tools for nonlinear time series analysis \cite{Kantz97,Sprott2003}, which have found many interesting applications to experimental data from various origins \cite{ZouPR2018}.
+In this line of research, there exist several major methods that are of particular importance, including recurrence networks \cite{MarwanPLA2009,Donner2010a}, visibility graphs \cite{Lacasa2008,Nunez2012}, transition networks \cite{Nicolis2005,MichaelChaos2015} and cycle networks \cite{ZhangPRL2006}.
+Depending on the particular questions at hand, each of these complex network approaches exhibit various variants that can be employed for different application purposes.
+For instance, we may either construct a single network representing a univariate time series, or interacting, multiplex or multi-layer networks for coupled time series.
+For a recent systematic review, see Ref.~\cite{ZouPR2018}.
+In this work, we focus on constructing ordinal pattern transition networks (OPTNs) for time series data, which present the advantages of easy implementation and a wide range of existing application to data from different origins, including correlated stochastic processes, neurophysiological (EEG) and human cardiac activity (ECG) series \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019,Subramaniyam2020}.
 
-The basic idea of ordinal partition network method originates from identifying ordinal patterns of time series \cite{BandtPRL2002}, which is a well developed concept in nonlinear time series analysis.
-For instance, permutation entropy and the associated statistical analysis \cite{BandtPRL2002}.
-In addition, ordinal symbolic representation of time series has found a number of interesting applications in science and engineering, for instance, biomedical recordings \cite{AmigoPTRSA2014}, finance \cite{ZaninChaos2008}, climate sciences \cite{BarreiroChaos2011}.
-Some recent progress has been comprehensively reviewed in Ref.
-\cite{AmigoPTRSA2014}.
-More specifically, given a one-dimensional time series $\{ x(i)\}_{i=1, \cdots, L}$ comprising of $L$ sampled points from a dynamical system, we first reconstruct the corresponding phase space by time delay embedding $\vec{x}(i) = [x(i), x(i+\tau), \cdots, x(i+(D-1)\tau)]$ with dimension $D$ \cite{Takens1981,Kantz97}.
-Next, we represent each embedded phase point $\vec{x}(i)$ by its corresponding rank order, which is represented by a symbol $\pi_{i}$.
-When sliding windows from $i=1$ to $N = L - (D - 1)\tau$ in the embedded space, a symbolic representation of the trajectory $\pi_i$ is produced.
-It is possible to derive information about the dynamics of the underlying system by assessing different probabilities of ordinal patterns which is typical for a time series resulted from a deterministic system.
-More generally, we obtain a probability distribution $P$ whose elements $p(\pi_{i})$ are the frequencies associated to pattern $\pi_{i}$, $i = 1, \cdots, D!$.
-Clearly, $P$ provides a significant feasible way to estimate the probability distribution function for a given time series, which plays a crucial role in extracting statistical complexities of the underlying dynamical systems \cite{BandtPRL2002,AmigoPTRSA2014}.
+The basic idea behind the OPTN method originates from identifying ordinal patterns in time series \cite{BandtPRL2002}, which is a well developed concept in nonlinear time series analysis leading to, for instance, the widely employed measure of permutation entropy \cite{BandtPRL2002}.
+In addition, ordinal symbolic representations of time series have found a number of interesting applications in science and engineering, for instance, in biomedical recordings \cite{AmigoPTRSA2014}, finance \cite{ZaninChaos2008}, and even climate \cite{BarreiroChaos2011}.
+Some recent progress has been comprehensively reviewed in Ref.~\cite{AmigoPTRSA2014}.
 
-Relatively easy computation of $P$ by ordinal patterns for the underlying system prompts another successful application to compute statistical complexity measure \cite{kowalskiEntropy2011}.
-Traditionally, statistical complexity measures (SCMs) are defined as the product of the normalized Shannon entropy and disequilibrium \cite{kowalskiEntropy2011}, which capture specific organizational properties of structure and patterns in a process.
-Following the literature \cite{rossoPRL2007,LopezPLA1995,kowalskiEntropy2011}, in this work we consider Jensen-Shannon SCM which is defined as $\mathcal{C}_{JS}[P] = \mathcal{H}[P]  \cdot  Q_{JS}[P, P_e]$, where $\mathcal{H}[P]$ is the normalized Shannon entropy and $Q_{JS}[P, P_e]$ is the so-called disequilibrium function measuring the distance of the given $P$ to the uniform distribution $P_e$, i.e., by the Jensen-Shannon divergence.
-Thus, $Q_{JS}[P, P_e]$ is defined as $Q_{JS}[P, P_e] = Q_{0} \cdot JS[P, P_e]$ where $JS[P, P_e]$ is the Jensen-Shannon divergence between $P$ and $P_e$ and $Q_{0}$ is the normalization factor.
-Furthermore, the complexity-entropy plane $\mathcal{H} \times \mathcal{C}_{JS}$ has been intensively used to distinguish chaotic systems from stochastic ones \cite{rossoPRL2007}.
-The computation of the complexity-entropy planes has benefit from the ordinal pattern symbolization of time series and has found various applications \cite{rossoPRL2007,kowalskiEntropy2011}.
-However, in the complexity-entropy analysis, there are some chaotic maps that could be easily confused as random noise because no clear separable regions are available to differentiate dynamics as reported in \cite{BorgesAMC2019}.
-So far, it remains to be a challenging task to disclose the possible non-monotonic relationship between complexity and entropy \cite{MartinPLA2003}.
+More specifically, given a one-dimensional time series $\{ x_i\}_{i=1, \cdots, L}$ comprising $L$ observations from a dynamical system, we first reconstruct the corresponding phase space by time delay embedding $\vec{x}_i = [x_i, x_{i+\tau}, \ldots, x_{i+(D-1)\tau}]$ with an embedding dimension $D$ \cite{Takens1981,Kantz97}.
+Next, we represent each embedding vector $\vec{x}(i)$ by the corresponding rank order of its components, which is encoded into a symbol $\pi_{i}$.
+Hence, when sliding windows from $i=1$ to $N = L - (D - 1)\tau$ in the embedding space, a symbolic representation of the trajectory $\{\pi_i\}$ is produced.
+It is possible to derive information about the dynamics of the underlying system by assessing the probabilities of occurrence of different ordinal patterns.
+For example, for time series resulting from deterministic dynamical systems, certain patterns may not occur at all.
+More generally, we obtain an empirical probability distribution $P$ whose elements $p_{\pi_{i}}$ are the frequencies associated to the different patterns $\pi_{i}$, $i = 1, \cdots, D!$.
+Clearly, $P$ provides a significant and feasible way to estimate a characteristic probability distribution function for a given time series, which plays a crucial role in extracting statistical complexity information on the underlying dynamical systems \cite{BandtPRL2002,AmigoPTRSA2014}.
 
-Note that, in all above discussions of ordinal pattern analysis, the transition behavior between patterns remains largely unconsidered.
-Recently, ordinal partition network representations have been proposed to capture the transition behavior of the ordinal patterns \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}, which sheds novel insight on the standard ordinal symbolic analysis of time series.
-In particular, each ordinal pattern is considered as a vertex in the graph and a directed weighted edge connecting two patterns in the graph is established depending on the transition frequency.
-One of advantages of ordinal pattern transition networks is that we can provide pronounced distinction among different time series with short time series \cite{MichaelChaos2015,BorgesAMC2019}.
+The conceptually relatively simple computation of $P$ based on ordinal patterns for the underlying system prompts to another successful application of this framework in terms of statistical complexity measures (SCMs) \cite{kowalskiEntropy2011}.
+One common way of defining SCMs is taking the product of a normalized Shannon entropy and an associated measure of disequilibrium \cite{LopezPLA1995,kowalskiEntropy2011}, which captures specific organizational properties of structure and patterns in a process.
+Following the vast amount of previous work on this topic  \cite{rossoPRL2007,LopezPLA1995,kowalskiEntropy2011,Lange2013}, in this work we consider a variant of SCMs based on the concept of Jensen-Shannon divergence, which is defined as $\mathcal{C}_{JS}[P] = \mathcal{H}[P] \cdot  Q_{JS}[P, P_e]$, where $\mathcal{H}[P]$ is the normalized Shannon entropy and $Q_{JS}[P, P_e]$ is the so-called disequilibrium function measuring the distance of the given distribution $P$ to the uniform distribution $P_e$.
+Accordingly, we define $Q_{JS}[P, P_e] = Q_{0} \cdot JS[P, P_e]$ where $JS[P, P_e]$ is the Jensen-Shannon divergence between $P$ and $P_e$ and $Q_{0}$ is a normalization factor.
+Furthermore, the complexity--entropy plane $\mathcal{H} \times \mathcal{C}_{JS}$ has been intensively used to distinguish chaotic systems from stochastic ones \cite{rossoPRL2007}.
+The analysis of complexity--entropy planes has already taken advantage from the ordinal pattern based symbolic encoding of time series and has found various applications \cite{rossoPRL2007,kowalskiEntropy2011}.
+However, in the complexity--entropy analysis, there are some chaotic maps that could be easily confused with random noise because no clearly separable domains of values are available to differentiate those different types of dynamics as reported in \cite{BorgesAMC2019}.
+So far, it remains a challenging task to disclose a possible non-monotonic relationship between complexity and entropy \cite{MartinPLA2003}.
 
-In this work, we generalize the traditional statistical complexity measures by incorporating pattern transition probabilities.
-The advantages are the following: (i) All SCMs successfully show distinctions between different dynamics, including time series from numerical models and experimental data; (ii) These generalized SCMs help to capture the consistent relationship between statistical complexity measures and chaoticity; (iii) All SCMs are sensitive to dynamical transitions of period doubling, band merging, inner and outer crisis.
+It may be noted that in the above discussion on ordinal pattern based time series analysis methods, the transition behavior between patterns has largely remained unconsidered.
+Recently, OPTN representations have been proposed to precisely capture this transition behavior of the ordinal patterns \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}, which opens a broader perspective beyond the standard ordinal symbolic analysis of time series.
+In this framework, each ordinal pattern is considered as a vertex in a graph, and a directed and weighted edge connecting two patterns in the graph is established according to the corresponding transition frequency (i.e., the probability with which a given permutation is followed by a certain other one).
+One of the advantages of the resulting OPTNs is that we can obtain a pronounced distinction between different types of systems based on short time series only \cite{MichaelChaos2015,BorgesAMC2019}.
 
-The structure of the paper is organized as follows: In section \ref{sec:OPW}, we show two slightly different ways to obtain the transition matrix of the corresponding ordinal pattern transition networks.
-Furthermore, based on this matrix, we propose in section \ref{sec:SCM} to compute the statistical complexity measures from three different perspectives including static pattern frequency properties, dynamic pattern transition properties and both.
-We will use one example to emphasize the important effects of embedding parameters in section \ref{sec:embeddings}.
-The usefulness of SCMs will be illustrated by four time series of typical dynamical regimes of the Logistic map.
-In section \ref{sec:plane}, the traditional complexity-entropy planes of each SCM will be discussed while the control parameter is systematically changed.
-In order to characterize the dynamical transitions, we expand the SCM values versus the control parameter in section \ref{sec:transi}.
-Finally, we show that SCMs successfully distinguish different dynamical regimes by experimental time series from flow data and human ECG data in \ref{sec:time}.
-Some discussions will be provided in the last section \ref{sec:con}.
+In this work, we draw upon the recent success of both, complexity--entropy plain analysis and OPTNs, to generalize the traditional SCMs by incorporating pattern transition probabilities.
+The advantages of this approach are as follows: (i) All SCMs successfully show clear distinction between different types of dynamics, which is demonstrated for time series from numerical models and experimental data.
+(ii) The generalized SCMs help to capture a consistent relationship between SCMs and chaoticity.
+(iii) All SCMs are sensitive to dynamical transitions along a bifurcation scenario, including period doubling, band merging, inner and outer crises.
 
-\section{Ordinal pattern transition matrix} \label{sec:OPW}
-Most of the current studies compute permutation entropy $\mathcal{S}_O$ focusing on the frequencies of ordinal patterns, which however overlooked the transition behavior between  patterns.
-To emphasize the transitions between any pair of two patterns, measures of transitional complexity have been further proposed in \cite{zhangSciRep2017,McCullough2017b} to quantify the resulting OPTNs.
-To this end, we indicate each directed link by its transition frequency $w_{ij} = p({\pi_i \to \pi_j})$, following the time iterations of the series.
-Calculating the transition frequency for each pattern, we obtain a weighted directed network characterized by a weighted adjacency matrix $W = \{ w_{ij} \}, i, j \in [1, D!]$.
+The remainder of this paper is organized as follows: In Section~\ref{sec:methods}, we introduce the methodology employed in our work.
+We start by discussing two slightly different ways to obtain the transition matrix between ordinal patterns, which provides the basis for defining the corresponding OPTNs (Section~\ref{sec:OPW}).
+Based on this matrix, we propose in Section~\ref{sec:SCM} to compute SCMs from three different perspectives, including static pattern frequency properties, dynamic pattern transition properties, and a combination of both.
+In Section~\ref{sec:results1}, we then proceed with describing some selected results for the logistic map as as paradigmatic model system exhibiting various types of periodic and chaotic solutions.
+We first discuss in some detail one example of a periodic orbit to emphasize the important effects of embedding parameters in Section~\ref{sec:embeddings}.
+The practical usefulness of SCMs will be further illustrated by studying four exemplary time series from different typical dynamical regimes (Section~\ref{sec:four}).
+In Section~\ref{sec:plane}, the complexity--entropy planes associated with each SCM will be discussed while the control parameter of the logistic map is systematically increased.
+In order to characterize the dynamical transitions in this model, we further study the behavior of the SCM values in dependence on the control parameter in Section~\ref{sec:transi} and show that similar results can also be obtained for time-continuous dynamical systems like the R\"ossler oscillator in Section~\ref{sec:cont}.
+Finally, we show that SCMs successfully distinguish different dynamical regimes in experimental time series from flow data and human ECG measurements in Section~\ref{sec:time}.
+Some concluding remarks and discussions will be provided in Section~\ref{sec:con}.
 
-Based on the transition matrix $\mathbf{W}$ of an OPTN, we note that there two slightly different ways to introduce normalization factors in $W$.
-The first option is to use a global normalization to ensure $\sum_{i,j}^{D!} w_{ij} = 1$ \cite{zhangSciRep2017}, which results in a globally normalized transition entropy as termed below.
+\section{Methods} \label{sec:methods}
 
-The second option of normalization has been proposed by McCullough {\textit{et al}} \cite{McCullough2017b} that they first considered the local out-link transition frequency from pattern $\pi_i$ to $\pi_j$
+\subsection{Ordinal pattern transition matrix} \label{sec:OPW}
+
+Most recent studies employing the concept of permutation entropy $\mathcal{S}_O$ have defined this measure by focusing exclusively on the frequencies of ordinal patterns, which however disregards the transition behavior between subsequent patterns.
+To emphasize the transition frequencies between any pair of patterns, measures of transitional complexity have been further proposed in Refs.~\cite{zhangSciRep2017,McCullough2017b} to quantify the properties of the associated OPTNs.
+To this end, we associate each directed link in an OPTN with its transition frequency $w_{ij} = p({\pi_i \to \pi_j})$.
+Calculating the transition frequencies for each pair of patterns, we obtain a weighted directed network characterized by a weighted adjacency matrix $\mathbf{W} = \{ w_{ij} \}, i, j \in [1, D!]$.
+
+Based on this transition matrix $\mathbf{W}$ of an OPTN, we note that there two slightly different ways to introduce normalization factors.
+The first option is to use a global normalization to ensure $\sum_{i,j}^{D!} w_{ij} = 1$ \cite{zhangSciRep2017}, which results in a globally normalized transition entropy as discussed below.
+The second option has been proposed by McCullough {\textit{et al}}~\cite{McCullough2017b} who first considered the local out-link transition frequency from pattern $\pi_i$ to $\pi_j$
 \begin{equation} \label{eq:localTp}
@@
 \end{equation}
+\noindent
 Note that the transition frequency of Eq.~\eqref{eq:localTp} is pattern (row-) wise normalized.
-This case will be referred to as node-wise out-link normalized transition matrix $\mathbf{W}$ in the following.
-In this way, it easily captures heterogeneous behavior of both static frequencies of different patterns and dynamic transition frequencies between patterns.
+This case will be referred to as the node-wise out-link normalized transition matrix $\mathbf{W}$ in the following.
+In this way, it easily captures the heterogeneous behavior of both (static) occurrence frequencies of different patterns and (dynamic) transition frequencies between patterns.
 
-In addition, we need to consider self-loops that may affect the numerical estimation of $\mathbf{W}$ which is especially important for time series from continuous systems.
-More specifically, it has been demonstrated in \cite{zhangSciRep2017} that there are about $99\%$ self-loops in the time series of R\"ossler system integrated by a step size $h = 0.01$, while about $1\%$ non-self-loops are hidden in $\mathbf{W}$.
-In addition, self-loops are related to the temporal correlation of time series \cite{BorgesAMC2019}.
-On the other hand, neglecting self-loops in $\mathbf{W}$ emphasizes the transition behavior between different patterns, which follows most of the existing studies of complex networks for computational simplicity and theoretical concerns \cite{CostaADPhy2007}.
-Therefore, we remove self-loops before computing the weighted matrix $\mathbf{W}$.
-In other cases of stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results $\mathbf{W}$ \cite{BorgesAMC2019}.
-We argue that whether to remove or to consider self-loops depends on the particular processes under study.
-Anyway, we have verified that our results below do not change significantly while considering effects of self-loops.
+In addition to the proper normalization, we need to consider self-loops that may affect the numerical estimation of $\mathbf{W}$ especially for time series from continuous systems with a given sampling rate.
+For example, it has been demonstrated in Ref.~\cite{zhangSciRep2017} that there are about $99\%$ self-loops in the time series of R\"ossler system when integrated with a step size $h = 0.01$, while the about $1\%$ non-self-loops are hidden in $\mathbf{W}$.
+From this example, one may easily see that self-loops are related to the serial correlation of the time series under study \cite{BorgesAMC2019}.
+On the other hand, neglecting self-loops in $\mathbf{W}$ emphasizes the transition behavior between different patterns, which has therefore been adopted in most existing studies of complex networks for computational simplicity and theoretical concerns \cite{CostaADPhy2007}.
+Accordingly, we also remove self-loops in the present study before computing the weighted matrix $\mathbf{W}$ while acknowledging that in other cases of stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results \cite{BorgesAMC2019}.
+Arguably, whether to remove or to consider self-loops depends on the particular process under study.
+In any of the applications discussed in the rest of this paper, we have ensured that our results do not change qualitatively when considering the effects of self-loops.
 
-\section{Statistical complexity measures} \label{sec:SCM}
-In this section, we review briefly ordinal pattern frequency based SCM and propose two novel ways to define SCM, which explicitly consider transition frequencies of the resulting ordinal network representations for time series.
+\subsection{Statistical complexity measures} \label{sec:SCM}
 
-$\bullet$ SCM based on permutation entropy of pattern frequencies.
+In the following, we will briefly review the existing ordinal pattern frequency based SCM before proposing two novel ways to define SCMs, which explicitly consider transition frequencies of the resulting OPTN representations for time series.
 
-The permutation entropy of a time series is the Shannon entropy of the distribution of patterns $P = \{p_{\pi_1}, p_{\pi_2}, \cdots, p_{\pi_i} \}$, which is computed as
+\subsubsection{SCM based on the permutation entropy of pattern frequencies}
+
+The permutation entropy of a time series is the Shannon entropy of the distribution of patterns $P = \{p_{\pi_1}, p_{\pi_2},\ldots,p_{\pi_i} \}$, which is defined as
 \begin{equation}
@@
 \end{equation}
-The normalized permutation entropy $\mathcal{H}_O$ is defined as 
+The normalized permutation entropy $\mathcal{H}_O$ hence follows as 
 \begin{equation} \label{eq:Ho}
@@
 \end{equation}
-where $\mathcal{S}_{O, max} = \log D!$ is resulted from the case of uniform distribution of the same number of patterns $P_e = \{1/D!, \cdots, 1/D!\}$.
-In the next step, we use some distance definition to measure the given $P$ to the uniform distribution $P_e$, which is the core of statistical complexity measures (SCM).
-There are several different distance functions as illustrated in the literature, such as Euclidean norm, Wooters' distance, Kullback-Leibler relative entropy and Jensen-Shannon divergence \cite{kowalskiEntropy2011}.
-In this work, the difference between the distribution $P$ and $P_e$ is characterized by the Jensen-Shannon divergence $Q_{JS}[P] = Q_0 \cdot JS[P, P_e]$ where $JS[P, P_e]$ is calculated by
+where $\mathcal{S}_{O, max} = \log D!$ corresponds to the case of a uniform distribution of the same number of patterns, i.e., $P_e = \{1/D!, \ldots, 1/D!\}$.
+
+In the next step, we use some appropriate statistical concept to measure the distance of the given empirical distribution $P$ to the uniform distribution $P_e$, which is the essential idea behind the class of SCMs considered in this work.
+Several different distance functions have been considered in the literature, including the Euclidean norm, Wooters' distance, Kullback-Leibler relative entropy and Jensen-Shannon divergence \cite{kowalskiEntropy2011}.
+In this work, the distance between the distributions $P$ and $P_e$ is characterized by the Jensen-Shannon divergence $Q_{JS}[P] = Q_0 \cdot JS[P, P_e]$ with
 \begin{equation}
-JS[P, P_e] = \mathcal{S}_O[\frac{P+P_e}{2}] - \frac{1}{2}\mathcal{S}_O[P] - \frac{1}{2}\mathcal{S}_O[P_e], 
+JS[P, P_e] = \mathcal{S}_O[(P+P_e)/2] - \frac{1}{2}\mathcal{S}_O[P] - \frac{1}{2}\mathcal{S}_O[P_e], 
 \end{equation}
-where $Q_0$ is a normalization factor with its value equal to the inverse of the maximal distance $JS[P, P_e]$, which ensures $Q_{JS}[P] \in [0, 1]$.
-This maximal distance is obtained when the distribution $P$ has probability one at one pattern and zeros at all other patterns.
-The general expression of $Q_0$ reads
+where $Q_0$ is a normalization factor with equals the inverse of the maximal Jensen-Shannon divergence $JS[P, P_e]$, which ensures that $Q_{JS}[P] \in [0, 1]$.
+This maximal distance is obtained when the distribution $P$ has probability one for one pattern and zero for all other patterns.
+The general expression for $Q_0$ reads
 \begin{equation} \label{eq:Q0}
-Q_0 = -2 \{(\frac{N+1}{N}) \log (N+1) -2 \log (2N) + \log (N)  \}^{-1}
+Q_0 = \frac{-2}{\left(\frac{N+1}{N}\right) \log (N+1) -2 \log (2N) + \log (N)},
 \end{equation}
-where $N = D!$ is the number of maximal patterns. For large $D$, we have $Q_0 \approx 1 /  \log 2$.
+where $N = D!$ is the total number of possible patterns. For large $D$, we have $Q_0 \approx 1 /  \log 2$.
 
-The function $Q_{JS}[P]$ is different from zero if the system has a kind of privilege over patterns among the accessible ones, reflecting determinism of the system.
+The function $Q_{JS}[P]$ is different from zero if the system has a kind of preference of certain patterns among the accessible ones, reflecting determinism of the system.
 Finally, the SCM based on permutation entropy of node (pattern) frequencies is defined as
@@
 \end{equation}
-This quantity is static in the sense that it quantifies the amount of information stored by ordinal frequencies in the system and its disequilibrium of the observed parts of its accessible states in comparison to uniform distribution \cite{LopezPLA1995}.
+This quantity is static in the sense that it quantifies the amount of information stored by ordinal frequencies in the system and its disequilibrium of the observed parts of its accessible states in comparison to a uniform distribution \cite{LopezPLA1995}.
 
-$\bullet$ SCM based on transition entropy by globally normalized weighting matrix $W$.
+\subsubsection{SCM based on the transition entropy of the globally normalized weight matrix}
 
-In this case, the sum of all entries of the transition matrix $W$ is 1 and the Shannon entropy of transition frequencies between any pair of ordinal patterns is therefore 
+In this case, the sum of all entries of the transition matrix $\mathbf{W}$ is 1, and the Shannon entropy of transition frequencies between any pair of ordinal patterns is therefore 
 \begin{equation}
@@
 (D!
-- 1)$ and the self transitions are excluded.
+- 1)$ as any self transitions are excluded.
 Furthermore, there are $D!
 (D!
-- 1)$ pattern transitions in the uniformly distribution function $P_e$.
-Thus, in a full analogy with node-based SCM, the Jensen-Shannon distance is therefore computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$, where $JS[W, P_e]$ is computed as
+- 1)$ possible pattern transitions in the uniformly distribution function $P_e$.
+Thus, in a full analogy with the node-based SCM, the Jensen-Shannon distance is computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$ with
 \begin{equation}
-JS [W, P_e] = \mathcal{S}_{W}[\frac{W + P_e}{2}] - \frac{1}{2}\mathcal{S}_{W}[W] - \frac{1}{2}\mathcal{S}_{W}[P_e]. 
+JS [W, P_e] = \mathcal{S}_{W}[(W + P_e)/2] - \frac{1}{2}\mathcal{S}_{W}[W] - \frac{1}{2}\mathcal{S}_{W}[P_e]. 
 \end{equation}
-We note that we have the same expression of the normalization factor $Q_0$ while replacing $N = D! (D! - 1)$ in Eq. \eqref{eq:Q0}. This further yields the SCM 
+We note that we have the same expression for the normalization factor $Q_0$ as above while replacing $N = D! (D! - 1)$ in Eq.~\eqref{eq:Q0}. This further yields the SCM 
 \begin{equation}
@@
 \end{equation}
-This quantify is dynamic in the sense that it considers the state transition information stored in the system.
+This quantity is dynamic in the sense that it considers the state transition information stored in the system.
 
-$\bullet$ SCM based on transition entropy by node-wise out-edge normalized
-weighting matrix $W$.
+\subsubsection{SCM based on the transition entropy of the node-wise out-edge normalized weighting matrix}
 
-In this case, the ordinal transition matrix is node-wised normalized, i.e., row
-sum of $W$ is 1. Therefore, we first introduce node-wise out-edge transition entropy which is calculated 
+We finally consider the case where the ordinal transition matrix is node-wised normalized, i.e., the row sum of $\mathbf{W}$ is 1.
+Accordingly, we again first introduce the node-wise out-edge transition entropy
 \begin{equation}
@@
 \mathcal{S}_{E}^{\pi_i}[W_{\pi_i}] &= - \sum_{j = 1; j \neq i}^{D!} w_{ij} \log w_{ij},  \\ 
-&= - \sum_{j=1, j \neq i}^{D!} p_{\pi_i \to \pi_j} \log p_{\pi_i \to \pi_j}. 
+&= - \sum_{j=1, j \neq i}^{D!} p_{\pi_i \to \pi_j} \log p_{\pi_i \to \pi_j} 
 \end{split}
 \end{equation}
-Furthermore, node-wise normalized transition entropy is 
+and node-wise normalized transition entropy
 \begin{equation}
-\mathcal{H}_{E}^{\pi_i}[W_{\pi_i}] = \mathcal{S}_{E}^{\pi_i}[W_{\pi_i}] / \mathcal{S}_{E, max}, 
+\mathcal{H}_{E}^{\pi_i}[W_{\pi_i}] = \mathcal{S}_{E}^{\pi_i}[W_{\pi_i}] / \mathcal{S}_{E, max} 
 \end{equation}
-where $\mathcal{S}_{E, max} = \log (D!
+with $\mathcal{S}_{E, max} = \log (D!
 - 1)$.
-Note that this normalization factor is the same for all patterns, namely, $\mathcal{S}_{E, max}^{\pi_i} = \mathcal{S}_{E, max}^{\pi_j}$ and, therefore, the superscript is suppressed in the above definition.
-In addition, node-wise Jensen-Shannon distance is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$, where the disequilibrium
+Note that this normalization factor is the same for all patterns, i.e., $\mathcal{S}_{E, max}^{\pi_i} = \mathcal{S}_{E, max}^{\pi_j}$.
+Therefore, the superscript index can be suppressed in the above definition.
+In addition, the node-wise Jensen-Shannon distance is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$ with the disequilibrium measure
+\begin{eqnarray}
+&&JS^{\pi_i} [W_{\pi_i}, P_e] = \nonumber \\
+&&\quad \mathcal{S}_{E}^{\pi_i}[(W_{\pi_i} + P_e)/2] - \frac{1}{2}\mathcal{S}_{E}^{\pi_i}[W_{\pi_i}] - \frac{1}{2}\mathcal{S}_{E}^{\pi_i}[P_e]. 
+\end{eqnarray}
+In this case, we obtain the normalization factor $Q_0$ by replacing $N = D! - 1$ in Eq.~\eqref{eq:Q0}. Taking into account node frequencies, the global Jensen-Shannon divergence is computed as 
 \begin{equation}
-JS^{\pi_i} [W_{\pi_i}, P_e] = \mathcal{S}_{E}^{\pi_i}[\frac{W_{\pi_i} + P_e}{2}] - \frac{1}{2}\mathcal{S}_{E}^{\pi_i}[W_{\pi_i}] - \frac{1}{2}\mathcal{S}_{E}^{\pi_i}[P_e]. 
+Q_{JS}[W, P_e] = \sum_{i=1}^{D!} p_{\pi_i} Q_{JS}^{\pi_i}[W_{\pi_i}], 
 \end{equation}
-In this case, we note that we have the normalization factor $Q_0$ while replacing $N = D!
-- 1$ in Eq.
-\eqref{eq:Q0}.
-Taking into account node frequencies, the global Jensen-Shannon divergence is computed as
-\begin{equation}
-Q_{JS}[W, P_e] = \sum_{i=1}^{D!} p(\pi_i) Q_{JS}^{\pi_i}[W_{\pi_i}], 
-\end{equation}
-which therefore yields an alternative definition of SCM of the transition network 
+which therefore yields an alternative definition of SCM of the OPTN as 
 \begin{equation}
@@
 \end{equation}
-where $\mathcal{H}_{E}[W] = \sum_{i} p(\pi_i) \mathcal{H}_{E}^{\pi_i}[W_{\pi_i}]$.
-This quantity is dynamic in the sense that it quantifies both the static ordinal state frequencies and the dynamic state transition information stored in the system.
+where $\mathcal{H}_{E}[W] = \sum_{i} p_{\pi_i} \mathcal{H}_{E}^{\pi_i}[W_{\pi_i}]$.
+This quantity is dynamic in the sense that it takes both the static ordinal state frequencies and the dynamic state transition information stored in the system into account.
 
-\section{Choice of embedding parameters} \label{sec:embeddings}
-There are two important algorithmic parameters in constructing ordinal pattern transition networks, namely, embedding parameters of $D$ and $\tau$.
-In some cases, the choice of $D$ and $\tau$ may strongly change the results since $D$ and $\tau$ should compromise the time scales of the respective time series.
-In general applications of ordinal pattern analysis (i.e., permutation entropy computation $\mathcal{S}_O$), the choice of optimal $D$ and $\tau$ depends on the field of application, which requires experience of experts as systematically summarized in \cite{Riedl2013}.
+\section{Numerical examples}\label{sec:results1}
+
+\subsection{Choice of embedding parameters} \label{sec:embeddings}
+There are two important algorithmic parameters in constructing OPTNs -- the embedding parameters $D$ and $\tau$.
+The choice of $D$ and $\tau$ may markedly affect the results for the associated SCMs since $D$ and $\tau$ should compromise the essential time scales of the dynamics in the underlying time series.
+In common applications of ordinal pattern analysis (e.g., estimation of the permutation entropy $\mathcal{S}_O$), the choice of optimal (or just pragmatic) values of $D$ and $\tau$ depends on the field of application, which requires experience of experts as systematically summarized in \cite{Riedl2013}.
+
 \begin{figure*}
@@
 	\includegraphics[width=2\columnwidth]{period4_logisticExample.pdf}
-\caption{\small{Example of a period-4 iterative time series ($r= 3.5$ in the Logistic map) to show the effects of embedding parameters $D$ and $\tau$.
-(a) iterative time series and the corresponding ordinal pattern series by different $D$ and $\tau$, (b) $D = \tau = 3$, (c) $D = \tau = 4$, (d) $D = \tau = 5$, and (e) $D = \tau = 6$.
-The respective permutation entropy values normalized by $\log D!$ as Eq.
-\eqref{eq:Ho} are $\mathcal{H}_O = 0.77$, $0$, $0.29$, and $0.11$ as indicated in the insets of each panel.
-(f-i) show the corresponding ordinal pattern transition networks.
-The weights of directed edges are suppressed due to two slightly different normalizations are introduced.
-Note in (g), only self-loops are available since $D = \tau = 4$ result in a unique pattern.
-Missing patterns are suppressed.
-}
+\caption{\small{Example of a period-4 time series demonstrating the effects of different choices of the embedding parameters $D$ and $\tau$.
+(a) Time series comprising 30 iterations of the period-4 orbit.
+The corresponding ordinal pattern series are shown for different combinations of $D$ and $\tau$: (b) $D = \tau = 3$, (c) $D = \tau = 4$, (d) $D = \tau = 5$, and (e) $D = \tau = 6$.
+The respective permutation entropy values normalized by $\log D!$ as in Eq.~\eqref{eq:Ho} are indicated in the insets of each panel.
+Panels (f)-(i) show the corresponding OPTNs.
+The weights of directed edges are not shown (recall that two slightly different normalizations are considered in this work).
+Note that in panel (g), there exist exclusively self-loops since $D = \tau = 4$ results in a unique, constant pattern.
+Missing patterns are suppressed.}
 \label{fig:embed}}
 \end{figure*}
-Here we use one example to show the important roles of $D$ and $\tau$ for a period-4 time series in Figure \ref{fig:embed}.
-In this example, we choose $D = \tau $ and the corresponding ordinal pattern series are significantly influenced by these parameter settings, which further result in different permutation entropy values of $\mathcal{H}_O$.
-For instance, only one unique ordinal pattern is observed when $D = \tau = 4$ (Fig.
-\ref{fig:embed}(c, g)) resulting in $\mathcal{H}_O = 0$, which is due to the coverage of one full period by the particular embedding parameters.
-Other values of $D$ and $\tau$ yield non-zero entropy values.
-The results of Fig.
-\ref{fig:embed} seem to be pessimistic since $D$ and $\tau$ will change the positions in the complexity-entropy planes.
+
+As an illustrative example highlighting the important roles of $D$ and $\tau$ for ordinal pattern based analysis, we consider a period-4 solution of the logistic map 
+\begin{equation}
+x_{i+1}=rx_i(1-x_i) 
+\end{equation}
+\noindent
+with $r= 3.5$.
+For the sake of simplicity, we choose $D = \tau $.
+In this case, Fig.~\ref{fig:embed} demonstrates that the resulting ordinal pattern sequences are significantly affected by the different parameter settings, which further result in different permutation entropy values $\mathcal{H}_O$.
+For instance, only one unique ordinal pattern is observed when $D = \tau = 4$ (Fig.~\ref{fig:embed}(c,g)) resulting in $\mathcal{H}_O = 0$, which is due to the coverage of exactly one full period by the particular embedding parameters.
+Other values of $D$ and $\tau$ however yield non-zero entropy values.
+The results of Fig.~\ref{fig:embed} may raise concern since different choices if $D$ and $\tau$ will change the placement of the system in the complexity--entropy plane.
 In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for periodic windows when the control parameter $r$ is varied.
 In such cases, we should focus more on the transitions between different dynamic regimes.
-Instead, we show that the results are robust when the control parameters are systematically changed while using the same parameter settings of $D$ and $\tau$.
+In the following, we will show that the results are robust when the control parameter is systematically changed while keeping the same parameter settings of $D$ and $\tau$.
 
-\section{Results for four typical dynamic regimes} \label{sec:four}
-We illustrate the potential of the proposed SCMs to characterize different dynamical regimes and transitions by analysis of the Logistic map $x_{n+1} = r x_n (1 - x_n)$ with the control parameter in the interesting range of $r \in [3.5, 4]$ with a step size of $\Delta r = 0.001$.
-In the range of the parameter $r$, various dynamical regimes and transitions between them can be found, for instance, period doublings, band merging points, inner and outer crises, intermittency \cite{Kantz97}, which has been served as a pragmatic model for nonlinear time series \cite{MarwanPLA2009}.
+\subsection{Results for four typical dynamic regimes} \label{sec:four}
 
-First, we show several typical regimes of particular control parameter $r$, i.e., periodic dynamics, band merging, laminar states and outer crisis, we investigate the SCMs in more detail which have been summarized in Tab \ref{tableLog}.
-We note that several chaos-chaos transitions have been discussed in the following, namely, the band merging corresponds to intermittency, the inner crisis to certain chaos-chaos transition and the outer crisis to fully chaotic dynamics.
+We further illustrate the potential of the proposed SCMs to characterize different dynamical regimes and transitions for the logistic map with the control parameter $r$ varying in the range $[3.5, 4]$ with a step size of $\Delta r = 0.001$.
+In this range of $r$, various dynamical regimes and transitions between them can be found, for instance, period doublings, band merging points, inner and outer crises, and intermittency \cite{Kantz97}, which has made this system serving as a paradigmatic model for assessing nonlinear time series analysis methods \cite{MarwanPLA2009}.
+
+As a first step, we consider four representative regimes that can be observed at particular values of the control parameter $r$, corresponding to either periodic dynamics or chaotic dynamics close to band merging, in some laminar state, and at the outer crisis.
+In all examples, we will consistently use the embedding parameters $D = \tau = 6$.
+In general, we emphasize that a higher value of $D$ leads to a larger variety of ordinal patterns and, hence, a more robust statistics.
+On the other hand, for evaluating the statistical properties of either pattern frequencies or transition frequencies, this comes on the cost of requiring increasingly longer time series.
+In the following, we will use realizations of the logistic map comprising \textcolor{red}{$L=xxx$} iterations.
+\textcolor{red}{Yong: please add the used time series length!
+Is it the same as in the next subsection...?}
+
+For all four situations, we investigate the SCMs along with their dual entropy characteristics in detail and summarize the obtained results in Tab.~\ref{tableLog}.
+We note that several types of chaos--chaos transitions will be discussed in the following: the band merging crisis corresponds to intermittency, the inner crisis to certain chaos-chaos transition and the outer crisis to fully chaotic dynamics.
+
 \begin{table}[htb]
@@
     \hline
-    Period 3      & 3.83 & 0 & 0 & 0 & 0 & 0 & 0 \\
+    period-3      & 3.83 & 0 & 0 & 0 & 0 & 0 & 0 \\
     \hline
-    Band merge    & 3.679 & 0.978 & 0.043 & 0.595 & 577 & 0.211 & 0.205  \\
+    band merging    & 3.679 & 0.978 & 0.043 & 0.595 & 0.577 & 0.211 & 0.205  \\
     \hline
-    Laminarity    & 3.791 & 0.986 & 0.029 & 0.633 & 0.609 & 0.280 & 0.269 \\
+    laminar state    & 3.791 & 0.986 & 0.029 & 0.633 & 0.609 & 0.280 & 0.269 \\
     \hline
-    Outer crisis  & 4.0 & 0.994 & 0.014 & 0.677 & 0.638 & 0.359 & 0.339 \\
+    outer crisis  & 4.0 & 0.994 & 0.014 & 0.677 & 0.638 & 0.359 & 0.339 \\
     \hline
     \end{tabular}}
-   \caption{Statistical complexity measures for four different control parameters of the Logistic map (embedding parameters: $D = \tau = 6$).   \label{tableLog}}    
+   \caption{Statistical complexity measures for four different values of the control parameter $r$ of the logistic map (embedding parameters: $D = \tau = 6$).   \label{tableLog}}    
 \end{table}
-In Tab.
-\ref{tableLog}, we have used embeddings $D = \tau = 6$.
-Therefore, all SCMs values are zeros for period-3 series because each embedded vector covers two full periods (c.f., Fig.
-\ref{fig:embed}).
-Other choice of $D$ and $\tau$ often yields non-zero values for period-3 regime as discussed in Fig.\ref{fig:embed}.
-For other cases of band merging, laminarity and outer crisis, all SCMs values are significantly larger than zeros.
-The pattern frequency based SCMs of $\mathcal{C}_O$ show smaller values for all regimes, which further lead to the lower right locations in the complexity-entropy planes in Fig.
-\ref{fig:fBmNoise}(a).
-However, for transition matrix based SCMs, their entropy $\mathcal{H}_W$, $\mathcal{H}_E$ and complexity $\mathcal{C}_W$, $\mathcal{C}_E$ show more consistent results while considering the chaoticity level of the system.
-In particular, SCMs are the largest when $r = 4$ in the outer crisis regime, and the SCMs are the smallest when $r = 3.679$ in the band merging case, while the laminarity regime has the intermediate SCM values.
 
-\section{Results on complexity-entropy planes} \label{sec:plane}
-In this section, we show the complexity-entropy planes by varying the control parameter $r$ in the illustrative Logistic model.
-Note that the Logistic map has been discussed in the framework of SCM \cite{RossoPRE2007,MartinPLA2003}, but only for pattern frequency based complexity.
-For each control parameter $r$, we construct ordinal pattern transition network from a time series of length $L = 31000$.
-In order to exclude transient behavior we remove the first $1000$ values from the data series.
-We have repeated the following calculations for embeddings $D = \tau \in [2, 7]$.
+Due to our choice of $\tau$ and $D$, all SCMs take values of zero for the period-3 series because the temporal distance between each pair of components of the embedding vector covers a multiple of two full periods (cf.~Fig.~\ref{fig:embed}).
+Other choices of $D$ and $\tau$ would commonly yields non-zero values for the period-3 regime.
+For the three other cases of band merging, laminar state and outer crisis, the values of all SCMs significantly differ from zero.
+The pattern frequency based SCM $\mathcal{C}_O$ shows smaller values than the other two measures for all regimes, which implies a positioning in the lower right parts of the complexity--entropy planes in Fig.~\ref{fig:fBmNoise}(a).
+However, for the transition matrix based SCMs, both the estimated entropies $\mathcal{H}_W$, $\mathcal{H}_E$ and associated complexity measures $\mathcal{C}_W$, $\mathcal{C}_E$ show results that are more consistent with the expectations when considering the respective level of chaoticity of the system in those three regimes (i.e., a rising Lyapunov exponent with increasing $r$ in the respective chaotic regimes).
+In particular, the estimated SCMs take their largest values for $r = 4$ in the outer crisis regime, and the SCMs are the smallest when $r = 3.679$ in the band merging case, while the laminar regime displays intermediate SCM values.
 
-When only one unique ordinal pattern is identified following the time series, we set the SCMs values $0$, which is reasonable since only self-loop is observed.
-This happens only if $D$ is an integer factor of the period for periodic windows of the Logistic map (e.g., Fig.
-\ref{fig:embed}(g)).
-Another notion is the data requirements for computing entropies and complexity measures.
-For reliable estimation of SCMs, a condition of the length of time series $L$ larger than $D!$ has been satisfied while $3 \leq D \leq 7$\cite{rossoPRL2007,kowalskiPhyD2007}.
-In the particular example of Logistic map, one may use longer data lengths (i.e., we have tested it up to $5 \times 10^6$), but this does not improve the estimate very much.
+\subsection{Complexity--entropy planes} \label{sec:plane}
 
-Traditionally, SCM have been applied to distinguish chaos from noise by means of the complexity-entropy plane.
-In particular, this plane shows the statistical complexity measures ($\mathcal{C}_O$, $\mathcal{C}_W$ and $\mathcal{C}_E$) as a function of the corresponding normalized entropy values ($\mathcal{H}_O$, $\mathcal{H}_W$ and $\mathcal{H}_E$).
-The above SCMs quantify both randomness and correlation structures of time series, which in consequence result in a non-trivial function of entropy  values.
-Chaotic systems present maximal complexity while stochastic systems have lower values of complexity, hence presenting in different regions in the complexity-entropy plane\cite{rossoPRL2007}.
-Furthermore, SCM values lie in a range of a minimum $\mathcal{C}_{min}$ and a maximum $\mathcal{C}_{max}$.
-A general algorithm for computing the bounds $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$ is provided in \cite{martinPhyA2006}.
-Figure \ref{fig:fBmNoise} shows the complexity-entropy planes for embedding parameters $D = \tau = 6$.
-We emphasize that the results in Fig.
-\ref{fig:fBmNoise} are qualitatively the same when $D$ and $\tau$ are varied in the interval $[3, 7]$, and we choose $D = 4, 5, 6$ as illustrative examples as shown in Fig.
-\ref{fig:fBmNoise}.
-Furthermore, the computation of $\mathcal{C}_{max}$ depends on the embedding $D$ since $D$ determines the number of patterns and pattern transitions which are included in the definition of SCMs.
-However, $\mathcal{C}_{min}$ shows less dependence on $D$.
+In order to put our previous results into a broader context, we next show the complexity--entropy planes obtained when varying the control parameter $r$ in the logistic map.
+It is notable that the logistic map has already been widely discussed in the framework of SCMs \cite{RossoPRE2007,MartinPLA2003}, but only for pattern frequency based complexity measures.
+For each value of $r$, we construct an OPTN from a time series of length $L = 31000$.
+In order to exclude transient behavior, we remove the first $1000$ values from the data series.
+We have repeated the following calculations for embedding parameters $D = \tau \in [2, 7]$.
+
+When only one unique ordinal pattern is identified in a given time series, all SCMs take a value of $0$, which is reasonable since only self-loops are observed.
+This however happens only if $D$ is an integer multiple of the period for the periodic windows of the logistic map (e.g., Fig.~\ref{fig:embed}(g)).
+Another practical concern is the data requirements for computing entropies and complexity measures.
+For a reliable estimation of SCMs, a pragmatic condition on the length $L$ of the time series $L$ as being sufficiently larger than $D!$ \cite{rossoPRL2007,kowalskiPhyD2007} has been satisfied in our setting for $D \leq 7$.
+In the particular example of logistic map, one may easily generate realizations of longer lengths, which however do not improve the obtained estimates significantly (as we have tested for $L\leq 5 \times 10^6$, not shown).
+
+In previous works, SCMs have been applied to distinguish chaos from noise by employing the concept of complexity--entropy plane~\cite{rossoPRL2007}.
+Generalizing this idea to our OPTN based SCMs, this plane shows the statistical complexity measures ($\mathcal{C}_O$, $\mathcal{C}_W$ and $\mathcal{C}_E$) as a function of the corresponding normalized entropy values ($\mathcal{H}_O$, $\mathcal{H}_W$ and $\mathcal{H}_E$).
+The three SCMs characterize both randomness and correlation structures in a time series, which in consequence result in a non-trivial dependence on the associated entropy values.
+Specifically, chaotic systems present maximal complexity while stochastic systems have lower values of complexity, hence appearing in different regions of the complexity--entropy plane~\cite{rossoPRL2007}.
+Furthermore, for a given entropy value, the range of possible SCM values is bound by a minimum $\mathcal{C}_{min}$ and a maximum $\mathcal{C}_{max}$.
+A general algorithm for computing the bounds $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$ has been provided in Ref.~\cite{martinPhyA2006}.
+
+Figure~\ref{fig:fBmNoise} shows the complexity--entropy planes for embedding parameters $D = \tau$ varied between $4$ and $6$.
+We emphasize that the results are qualitatively similar when $D$ and $\tau$ are varied in the interval $[3, 7]$.
+Furthermore, it is notable that the values of $\mathcal{C}_{max}$ depend on the embedding dimension $D$ since $D$ determines the number of patterns and pattern transitions which are considered in the definition of SCMs.
+By contrast, $\mathcal{C}_{min}$ shows less dependence on $D$.
+
 \begin{figure*}
@@
 	\includegraphics[width=2\columnwidth]{CompEntropy_LogisticHC.pdf}
-\caption{\small{Complexity-entropy planes (Jensen-Shannon divergence) to show statistical complexity measures for example series of the logistic map for varying the control parameter $r$.
-(a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
+\caption{\small{Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
 The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy.
-(a-c) for embedding $D = 4, 5$ and (d-f) for $D = 6$.
-In (a, d), the bottom-right rectangles of dotted lines highlighted the ambiguous regions for the traditional pattern frequency based SCM analysis.
-}  \label{fig:fBmNoise}}
+Panels (a-c) show the results for embedding dimensions $D = 4,5$ and (d-f) for $D = 6$.
+In panels (a,d), the bottom-right rectangles of dotted red lines highlight the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
+\textcolor{red}{I don't quite get where this information comes from.
+Is it taken from Rosso's works or own results of this study?
+I do not see stochastic processes like fBm/fGn being discussed here, which is probably also not necessary.
+But we should clarify this.}}  \label{fig:fBmNoise}}
 \end{figure*}
 
-In this complexity-entropy plane, pattern frequency based SCM values lie around the maximum curve $\mathcal{C}_{max}$.
-It is clear to see that the displacements of the dots depend on the embedding $D$ as shown in Fig.
-\ref{fig:fBmNoise}(a, d).
-In addition, there are many dots in the bottom-right position of the plane, which have been highlighted by rectangles of dotted lines.
-Indeed, this is a challenging region for SCMs to distinguish chaotic time series from random noise in Fig.
-\ref{fig:fBmNoise}(a, d), which has been reported in \cite{BorgesAMC2019}, too.
-It is not an easy task to correctly tell the difference between chaos (i.e., chaotic regimes of large $r$ in the Logistic map) and noise.
-This is because the SCMs of $\mathcal{H}_O$ and $\mathcal{C}_O$ are static in the sense that only pattern frequencies have been considered in the definitions.
-The results that  $\mathcal{H}_O$ and $\mathcal{C}_O$ are displaced close to $\mathcal{C}_{max}$ curves have been observed when the embedding $D$ is increased (Fig.
+First of all, we recover the known result that pattern frequency based SCM values lie around the maximum curve $\mathcal{C}_{max}$, while the displacements of the estimated values from their theoretical upper bounds depend on the embedding dimension $D$ as shown in Fig.~\ref{fig:fBmNoise}(a,d).
+Most notable, we find many combinations of complexity--entropy values in the lower right part of the plane (high entropy, low complexity), which have been highlighted by rectangles of dotted lines.
+This region of the plane is challenging for SCMs to distinguish between chaotic time series and random noise~\cite{BorgesAMC2019} and is in our case occupied by values corresponding to chaotic regimes of the logistic map at large $r$ values.
+The corresponding ambiguity results from the fact that both $\mathcal{H}_O$ and $\mathcal{C}_O$ only take static information in the sense that only pattern frequencies have been considered in their definitions.
+%The results that  $\mathcal{H}_O$ and $\mathcal{C}_O$ are displaced close to $\mathcal{C}_{max}$ curves have been observed when the embedding $D$ is increased (Fig.
 \ref{fig:fBmNoise}(a, d)).
 
-For other pattern transition based SCMs of $\mathcal{C}_W$ vs $\mathcal{H}_W$ and $\mathcal{C}_E$ vs $\mathcal{H}_E$, we also find that the dots are distributed close to the respective maximum lines of $\mathcal{C}_{max}$.
-Increasing the embedding dimension further to $D=6, 7$, the dots are covered by the left uprising branch the parabola curve $\mathcal{C}_{max}$ (Fig.
-\ref{fig:fBmNoise}(b, d)).
-The definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes into account the information of both pattern frequencies and pattern transitions, which results in that the dots are closer to $\mathcal{C}_{max}$ in a relative faster manner (Fig.
-\ref{fig:fBmNoise}(c, f)).
-The most interesting result is that no points lie in the bottom-right challenging regions of ambiguity.
-It is the deterministic transition behavior among patterns that makes the distinction from random noise.
+For our new pattern transition based SCMs, we find that in the complexity--entropy planes of $\mathcal{C}_W$ versus $\mathcal{H}_W$ and $\mathcal{C}_E$ versus $\mathcal{H}_E$, the results for the logistic map again closely align with the theoretical upper bounds $\mathcal{C}_{max}$.
+When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely and essentially cover the left uprising branch of the curve for the SCM based exclusively on transition frequencies (Fig.~\ref{fig:fBmNoise}(b,d)).
+By contrast, the definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes information on both, pattern frequencies and transition frequencies into account, which results in the obtained SCM values aligning even more closely with $\mathcal{C}_{max}$ already at lower embedding dimensions (Fig.~\ref{fig:fBmNoise}(c,f)).
+The most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of ambiguity.
+We therefore suggest that it is the deterministic transition behavior among patterns that makes the distinction from random noise.
 
-Based on the results of Fig. \ref{fig:fBmNoise}, we conclude that pattern transition behavior provides novel insights for defining SCMs, which complement the traditional SCMs.
+Based on the results of Fig.~\ref{fig:fBmNoise}, we conclude that the pattern transition behavior encoded in the OPTNs provides novel insights that can be exploited in terms of SCMs that complement their traditional counterparts.
 
-\section{Characterizing dynamical transitions} \label{sec:transi}
-The Logistic map experiences various bifurcation scenarios when the control parameter $r$ is systematically increased.
-However, these dynamical regimes and regime transitions are hidden in the complexity-entropy planes.
-In this subsection, we aim to expand the figures by showing complexity measures depending on the control parameter $r$.
-Our motivation is to verify whether SCMs computed from ordinal pattern transition networks are able to detect the dynamical transitions when the control parameter $r$ is varied with a step size $0.001$.
-Technically, it is a rather simple idea to check whether the normalized entropy $\mathcal{H}$ and SCMs $\mathcal{C}$ are able to track the changes of dynamics.
-Furthermore, we compare our results with the spectrum of the Lyapunov exponents.
+\subsection{Characterizing dynamical transitions} \label{sec:transi}
+The logistic map experiences a sequences of different types of bifurcations when the control parameter $r$ is systematically increased.
+However, the corresponding dynamical regimes and regime transitions are not easy to identify from the complexity--entropy planes discussed above.
+In the following, we therefore aim to expand the previous results by showing the dependence of our SCMs on the control parameter $r$.
+Here, our motivation is to verify whether the normalized entropies and associated SCMs computed from OPTNs are indeed able to detect the dynamical transitions along the complex bifurcation scenarios of the logistic map and hence track qualitative changes in the dynamics, including both period--chaos and chaos--chaos transitions.
+For the sake of comparison, we take the associated Lyapunov exponent of the logistic map as an established measure for characterizing the type of dynamics (regular versus chaotic) along with the degree of chaoticity.
 
@@
 	\includegraphics[width=2\columnwidth]{logisticEntropy.pdf}
-\caption{\small{Behavior of different statistical complexity measures for the Logistic map in dependence on the control parameter $r$.
-(a) bifurcation diagram, (b) the spectrum of the maximal Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{H}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
-(g, h) are entropy $\mathcal{H}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$.
-Several major periodic windows have been highlighted by grey background.
-Vertical dotted lines correspond to the cases as summarized in Tab.
-\ref{tableLog}.
+\caption{\small{Behavior of the different SCMs and associated entropy characteristics for the logistic map in dependence on the control parameter $r$.
+(a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{S}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ based on the globally normalized weight matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g,h) entropy $\mathcal{S}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized weight matrix $\mathbf{W}$.
+Several major periodic windows have been highlighted by grey background shading.
+Vertical dotted lines indicate the cases summarized in Tab.~\ref{tableLog}.
 } \label{fig:bifurcation}}
@@
 
-Overall, all SCMs curves follow the changes in the bifurcation diagram, for instance, showing distinctions in periodic windows  (highlighted by grey color in Fig.
-\ref{fig:bifurcation}).
-Except $\mathcal{C}_O$, all other SCMs behave in a much similar manner as the Lyapunov exponents.
-Namely, as the control parameter $r$ is increased, the chaoticity level grows reaching a maximum at $r = 4$, which, therefore, leads to an increase in the statistical complexity.
-This similar increasing trend is absent in the case of pattern frequency based complexity measure $\mathcal{C}_O$ (Fig.
-\ref{fig:bifurcation}(c)).
-Note that it is a known fact that pattern frequency dependent SCMs often do not follow the growth in the chaoticity in the Logistic map \cite{MartinPLA2003}, which can be improved by Wootter's distance function, instead of the Shannon-Jensen divergence that we used in this work.
-From this point of view, pattern transition frequency based measures of $\mathcal{C}_W$ and $\mathcal{C}_E$ are more reasonable in the sense that they easily track the growth level of chaoticity.
+As a general observation, we find that all three SCMs clearly follow the changes in the bifurcation diagram, for instance, showing distinct values in periodic windows (highlighted by grey color in Fig.~\ref{fig:bifurcation}).
+Other than $\mathcal{C}_O$, the two transition frequency based SCMs behave in a similar way as the Lyapunov exponent.
+Specifically, as the control parameter $r$ is increased, the chaoticity level grows gradually, reaching a maximum at $r = 4$, which is reflected in an increase in the statistical complexity.
+A similar increase is absent in the pattern frequency based complexity measure $\mathcal{C}_O$ (Fig.~\ref{fig:bifurcation}(c)).
+In this context, we note that it is an established fact that pattern frequency based SCMs often do not trace the growth in the chaoticity of the logistic map~\cite{MartinPLA2003}, which can however be improved by using Wooter's distance function instead of the Shannon-Jensen divergence used in this work.
+From this point of view, the pattern transition frequency based SCMs $\mathcal{C}_W$ and $\mathcal{C}_E$ exhibit more informative behavior in the sense that they track the growth of the level of chaoticity.
 
-There is another improvement for $\mathcal{H}_E$ and $\mathcal{C}_E$ when the Logistic map presents a period doubling bifurcation at $r = 3.544$ (Fig.
-\ref{fig:bifurcation}(g, h)).
-The SCMs of $\mathcal{H}_O$,  $\mathcal{C}_O$, $\mathcal{H}_W$ and $\mathcal{C}_W$ show non-zero values in this period doubling window.
-Furthermore, there are jumps at the bifurcation point $r = 3.544$, which have been also reported in \cite{BandtPRL2002}.
-To a large extent, the jumps should not be expected because complexity does not change when $r$ passes this point.
-We show in Fig.
-\ref{fig:transient} that the iterative transient points are responsible to affect ordinal pattern definitions, which results in the peculiar jumps of SCMs at $r = 3.544$.
-For estimations from time series of finite length of this work, $\mathcal{H}_E$ and $\mathcal{C}_E$ are not affected seemly.
-Certainly, we should not over interpret their capabilities since the numerical inaccuracy would be accumulated in a longer iterative process such that different ordinal patterns are identified.
+Another conceptual improvement is found for $\mathcal{H}_E$ and $\mathcal{C}_E$ in the parameter range where the logistic map presents period doubling bifurcations, for example, at $r = 3.544$ (Fig.~\ref{fig:bifurcation}(g,h)).
+It may be noticed that $\mathcal{H}_O$, $\mathcal{C}_O$, $\mathcal{H}_W$ and $\mathcal{C}_W$ all exhibit non-zero values in this parameter range despite a purely periodic dynamics.
+Furthermore, there are jumps of all four measures at the points of period doubling bifurcations, which have been also reported earlier in Ref.~\cite{BandtPRL2002}.
+These jumps however are not desirable since the complexity of the dynamics does not change when $r$ passes nay of those points.
+For the case of the period doubling bifurcation at $r=3.544$ (replacing a period-4 by a period-8 solution), we show in Fig.~\ref{fig:transient} that the observed jump in the entropy and SCM values may be explained by recurrent iterations leading to long transients.
+\textcolor{red}{I do not fully get this reasoning...} For estimations based on time series of finite length as used in this work, $\mathcal{H}_E$ and $\mathcal{C}_E$ are however not affected in the same way.
+Certainly, we should not over-interpret their capabilities since numerical inaccuracy would be accumulated in a longer iterative process such that different ordinal patterns are identified.
 In addition, there are many periodic windows of different periods along the axis of control parameter $r$, which prevents us from using a unique predefined number of iterations as initial transients for time series of different period length.
+Notably, similar jumps have also been observed when the system bifurcates from a period-3 to a period-6 solution.
+\textcolor{red}{Yong: give corresponding $r$ value here.}
+
 \begin{figure}
@@
 	\includegraphics[width=\columnwidth]{period4_exampleTransients.pdf}
-\caption{\small{Example to show the transient effects in generating proper bifurcation diagrams of the Logistic map, focusing on the window of period-4 to period-8.
-The time points get blurred when $r$ is close to $3.544$ when a short transient of $50$ iterates are removed from the analysis (a).
-In (b) $2000$ initial iterates are removed.
+\caption{\small{Transient effects in the generation of proper bifurcation diagrams for the logistic map, exemplified by the period doubling bifurcation from a period-4 to a period-8 solution.
+(a) The actual bifurcation point gets blurred when $r$ is close to $3.544$ if only a short transient of $50$ iterations is removed from the time series.
+(b) A large number of initial iterations ($2000$) are removed, the correct bifurcation point becomes visible.
 }\label{fig:transient}}
 \end{figure}
-Meanwhile, these jumps have also been observed when the system bifurcates from period-3 to period-6.
 
-\section{Transitions in continuous system}
-The previous results have been further verified in the chaotic R\"ossler system \cite{Roessler1976} while varying the control parameter, namely, $(\dot{x}, \dot{y}, \dot{z}) = (-y-z, x+0.2y, b+z(x-5.7))$, where $b$ is the control parameter.
-There are some challenging problems for constructing OPTNs for time series from continuous systems, for instance, proper choice of sampling frequency and embedding parameters, which must depend on the particular time scales of the system.
-To minimize the discussions on proper choices of further algorithmic parameters, we introduce Poincar\'e sections at $y=0$, $\dot{y}<0$ (for each value of $b$) and construct OPTNs from $N = 10000$ intersection points.
-The results have been shown in Fig.
+\subsection{Dynamical transitions in continuous system}\label{sec:cont}
+
+While we have focused exclusively on the case of the time-discrete logistic map so far, it may be interesting to identify whether a similar behavior can also be obtained for time-continuous deterministic dynamical systems.
+To this end, we will illustrate a corresponding analysis for the example of the chaotic R\"ossler system \cite{Roessler1976}
+\begin{eqnarray}
+\dot{x} &=& -y-z, \nonumber \\
+\dot{y} &=& x+0.2y, \\
+\dot{z} &=& b+z(x-5.7) \nonumber
+\end{eqnarray}
+while varying the control parameter $b$.
+From a conceptual perspective, constructing OPTNs for time series from continuous systems faces certain additional practical challenges, including the proper choice of sampling frequency and embedding parameters, which depend on the particular time scales of the system.
+To avoid any corresponding discussion on proper choices of further algorithmic parameters, we employ here a Poincar\'e section to each sample trajectory of the system at $y=0$, $\dot{y}<0$ (for different values of $b$) and construct OPTNs from $N = 10,000$ intersection points.
+The results are shown in Fig.
 \ref{fig:bifurRossler}.
+To this end, we conclude that the general behavior of the different SCMs and associated entropy measures resembles that previously observed for the logistic map in Fig.~\ref{fig:bifurcation}.
+Specifically, all measures trace the succession of bifurcations in the considered range of $b$.
+We outline more detailed follow-up investigations on the behavior of our OPTN based SCMs as relevant topics for future work.
+
 \begin{figure*}
@@
 	\includegraphics[width=2\columnwidth]{rosslerEntropy.pdf}
-\caption{\small{Similar with the caption of Fig.
-\ref{fig:bifurcation}.
-Behavior of different SCMs for the R\"ossler system varying the control parameter $b$.
-(a) bifurcation diagram ($x$-component on the Poinar\'e sections), (b) the maximal Lyapunov exponent $\lambda$, (c) pattern frequency based permutation entropy $\mathcal{H}_O$ and (d) complexity $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ by the globally normalized weighting matrix $\mathbf{W}$ and (f) the corresponding complexity $\mathcal{C}_W$.
-(g, h) are entropy $\mathcal{H}_E$ and complexity $\mathcal{C}_E$ which are similar with (e, f) but based on node-wise out-link normalized weighting matrix $\mathbf{W}$.
-Several major periodic windows have been highlighted by grey background.
-} \label{fig:bifurRossler}}
+\caption{\small{Same as Fig.
+\ref{fig:bifurcation}, but for Poincar\'e sections of the R\"ossler system while varying the control parameter $b$ (see text for details).
+Panel (a) presents a bifurcation diagram based on the $x$ components of all points in the Poincar\'e sections.
+Panel (b) shows the largest Lyapunov exponent of the system in dependence of $b$.} \label{fig:bifurRossler}}
 \end{figure*}
-Rather similar results as Fig. \ref{fig:bifurcation} have been obtained in the chaotic R\"ossler system, identifying different dynamic transitions, where some periodic windows have been highlighted.
 
-\section{Real time series of fluid flows and ECG} \label{sec:time}
-In this section, we use two different experimental time series to demonstrate that SCMs successfully characterize the complexity properties of the underlying systems, showing distinct values for different dynamics.
+\section{Real-world examples} \label{sec:time}
 
-The first example is from a laboratory fluid experiment of Baroclinic instability which is used to study patterns of cyclones and anticyclones in climate systems \cite{Read_jfm_1992,ZouEPJST2008}.
-Depending on the parameters of rotation rate, temperature difference, viscosity and fluid density, this experimental system exhibits a rich variety of flow regimes.
-In particular, we focus on: (i) {Stable fluid:} the temperature signal of the stable flow exhibits periodic oscillation as the wave drift patterns arrive at the measured fixed point.
-(ii) {Quasi-periodic 2- frequency Amplitude vacillation (AV-2):} this case is identified as a 2-frequency quasi-periodic amplitude flux.
+While we have previously demonstrated the suitability of OPTN based SCMs for tracing changes in dynamical complexity for time series from deterministic dynamical systems, in the following we will use two different experimental time series to demonstrate that SCMs also successfully characterize the complexity properties of real-world systems, showing distinct values for different types of dynamics.
+
+The first example originates from a laboratory fluid experiment of baroclinic instability which is used to study patterns of cyclones and anticyclones in the Earth's atmosphere~\cite{Read_jfm_1992,ZouEPJST2008}.
+Depending on the parameters rotation rate, temperature difference, viscosity and fluid density, this experimental system exhibits a rich variety of flow regimes.
+In particular, we focus on the following regimes:
+\begin{enumerate}[(i)]
+\item {Stable fluid:} The temperature signal of the stable flow exhibits periodic oscillations as the wave drift patterns arrive at the measured fixed point. 
+\item {Quasi-periodic 2-frequency amplitude vacillation (AV-2):} This case is identified as a 2-frequency quasi-periodic amplitude flux.
 The wave drift is composed of slow and regular oscillations in the temperature signal, but a fast modulation in the amplitude is also visible.
-AV-2 is characterized by periodic growth and a drop in wave amplitude with little change in waveform.
-(iii) {Quasi-periodic 3-frequency amplitude vacillation (AV-3): } We also analyze a quasi-periodic 3-frequency amplitude vacillation (AV-3) time series which is obtained from a three-dimensional direct numerical simulation in an air-filled rotating Baroclinic instability experimental \cite{Read_jfm_1992}.
-(iv) {Modulated amplitude vacillation (MAV):} this case is identified as a low-dimension stream, chaotically modulated and with waves of varying amplitudes.
+AV-2 is characterized by a periodic growth and drop in wave amplitude with little change in waveform.
+\item {Quasi-periodic 3-frequency amplitude vacillation (AV-3):} We also analyze a quasi-periodic 3-frequency amplitude vacillation time series which is obtained from a three-dimensional direct numerical simulation in an air-filled rotating baroclinic instability experiment~\cite{Read_jfm_1992}.
+\item {Modulated amplitude vacillation (MAV):} This case is identified as a low-dimensional stream, chaotically modulated and with waves of varying amplitudes.
 Amplitude modulation results in a complex temperature signal.
+\end{enumerate}
 
-The second example time series comes from the physiological signals of human electrocardiogram (ECG) recordings that are collected from patients in a Coronary Care Unit and the Creighton University Cardiac Center.
-Ventricular arrhythmia are a very significant but poorly understood dynamical system, showing rich nonlinear properties.
+The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients in a Coronary Care Unit and the Creighton University Cardiac Center.
+Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant but still relatively poorly understood dynamical system, showing rich nonlinear properties.
 In this work, we focus on three different states: normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}.
-From the data bank of recordings of ECG during various rhythms, we selected 14 time series showing sinus rhythm (SR) prior to onset of arrhythmia, 12 time series of VT and 17 VF.
+From the data bank of recordings of ECG during various rhythms, we selected 14 time series showing sinus rhythm (SR) prior to onset of arrhythmia, 12 time series of VT and 17 of VF.
 The ECG recordings have been annotated with SR, VT and VF by skilled cardiologists on the basis of waveform morphology.
-Each recording consists of at least $10,000$ data points (20s) in length and 250 Hz in sampling frequency.
+Each recording consists of at least $10,000$ data points (20~s) in length at a 250~Hz sampling frequency.
 
-For choosing embedding parameters for both example time series, we follow the suggestions that are proposed in the respective literature \cite{ZouEPJST2008,smallCSF2002}.
-Although we are facing with the inevitable presence of non-stationary and noise in the experimental data, we have verified that the following results do not change significantly while embedding parameters are varied.
-In particular, we have treated embedding dimension as a free parameter which is varied in the interval as $D = 2, 3, \dots, 7$, while the embedding delay $\tau$ is the following.
-In the case of temperature time series from the fluid flows, embedding delay $\tau_{i} = 218s$ for stable solution, $\tau_{ii} = 136.5s$  for AV-2, $\tau_{iii} = 185.625$s for AV-3, and $\tau_{iv} = 118s$ for MAV.
-In the case of recordings of ECG time series, embedding delay $\tau = 5$.
+For choosing the embedding parameters for both sets of example time series, we follow the suggestions from earlier studies~\cite{ZouEPJST2008,smallCSF2002}.
+Although we are facing the inevitable presence of nonstationarity and noise in the experimental data, we have verified that the following results do not change significantly while our embedding parameters are varied.
+In particular, we have treated the embedding dimension as a free parameter which has been varied in the range of $D = 2, 3, \ldots, 7$, while the embedding delay $\tau$ has been kept fixed as follows.
+In the case of temperature time series from the fluid experiments, $\tau_{i} = 218$~s for the stable solution, $\tau_{ii} = 136.5$~s  for AV-2, $\tau_{iii} = 185.625$~s for AV-3, and $\tau_{iv} = 118$~s for MAV.
+For the ECG time series, we consistently use an embedding delay of $\tau = 5$.
 The results have been averaged over all subjects.
 
-Generally speaking, the OPTN based statistical complexity measures of both the flow data and the ECG recordings enhance our understanding of the underlying nonlinear system, which are shown in Fig.
-\ref{fig:fluid}.
-More specifically, in the case of fluid data (Fig.
-\ref{fig:fluid}(a)), complexity measures of $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ show consistent complexity values for stable wave solution (SW), quasi-periodic 2-frequency Amplitude vacillation (AV-2), quasi-periodic 3-frequency Amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV).
-Namely, the complexity of these four cases has the following order: SW has the smallest complexity since it shows periodic oscillation of trivial recurrences, while MAV has the highest complexity since it is chaotically modulated amplitude vacillation.
-Quasiperiodic states show inter-mediate complexity values since these two cases have non-trivial recurrences but with low complexity.
+In both sets of time series, the OPTN based SCMs enhance our understanding of the underlying nonlinear system, which can be seen from the results summarized in Fig.~\ref{fig:fluid}.
+In the case of the fluid data (Fig.~\ref{fig:fluid}(a)), $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ all show consistent changes of the complexity and entropy values between stable wave solution (SW), quasi-periodic 2-frequency amplitude vacillation (AV-2), quasi-periodic 3-frequency amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV).
+Specifically, the complexity of these four cases has the following order: SW has the smallest complexity since it shows periodic oscillations with trivial recurrence, while MAV has the highest complexity since it presents chaotically modulated amplitude vacillation.
+Quasiperiodic states show intermediate complexity values since these two cases have non-trivial recurrences but with low complexity.
 Interestingly, AV-3 presents higher complexity than AV-2.
-The relationship of complexity between periodic, quasiperiodic and chaotic solutions still holds for $\mathcal{C}_O$ although the corresponding values become comparable.
-In particular, it becomes hard for $\mathcal{C}_O$ to tell the difference between AV-2 and AV-3.
+The order of complexity values between periodic, quasiperiodic and chaotic solutions still holds for $\mathcal{C}_O$ although the corresponding values become comparable.
+In particular, it becomes hard for $\mathcal{C}_O$ to identify the difference between AV-2 and AV-3.
 
-Qualitative similar results have been obtained for ECG recording series (Fig.
-\ref{fig:fluid}(b)).
+Qualitatively similar results are obtained for the ECG recordings (Fig.~\ref{fig:fluid}(b)).
 Clearly, different ECG recordings exhibit remarkably distinct complexity values.
-Specifically, SR recordings show pronounced highest complexity values of $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ since there are rather broad frequency bands between 50 and 100 beats/min in the signals \cite{smallCSF2002}.
-On the other hand, the complexity of VT has the smallest values comparable to the case of VF.
-Similarly to the case of flow data, SR shows the smallest value of $\mathcal{C}_O$ while VT and VF have comparable values that are relative higher than SR.
+Specifically, SR recordings show very pronounced the highest values of $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$, reflecting the existence of significant spectral power in a rather broad frequency band between 50 and 100 beats/min in the signals \cite{smallCSF2002}.
+On the other hand, the complexity of VT has the smallest value, even lower than for the case of VF.
+Unlike the other five measures, we note, however, that SR shows the smallest value of $\mathcal{C}_O$ while VT and VF have comparable values that are higher than SR.
+This difference may be explained by the fact that unlike the two other OPTN based SCMs, this measure only takes static information of pattern frequencies into account and may therefore not be able to properly distinguish subtle differences reflected in the temporal order of patterns.
+
 \begin{figure*}
@@
 	\includegraphics[width=2\columnwidth]{fluidExample.pdf}
-\caption{\small{Application of SCMs to experimental fluid data (a) and human recordings of ECG (b).
-The results have been averaged over embedding delays $D = 2, 3, \dots, 7$.
-In (a), stable wave (SW) $\blacklozenge$, quasiperiodic 2-frequency amplitude vacillation (AV-2) $\bullet$, quasiperiodic 3-frequency amplitude vacillation (AV-3) $\blacktriangle$, and chaotic modulated amplitude vacillation (MAV) $\blacksquare$.
-In (b), normal sinus rhythm (SR) $\blacktriangle$, ventricular tachycardia (VT) $\blacklozenge$ and ventricular fibrillation (VF) $\bullet$.
+\caption{\small{OPTN based entropy and SCM characteristics for experimental fluid data (a) and human ECG recordings (b).
+The results have been averaged over embedding dimensions in the range of $D = 2, 3, \ldots, 7$.
+In (a), stable wave (SW, $\blacklozenge$), quasiperiodic 2-frequency amplitude vacillation (AV-2, $\bullet$), quasiperiodic 3-frequency amplitude vacillation (AV-3, $\blacktriangle$) and chaotic modulated amplitude vacillation (MAV, $\blacksquare$) are distinguished.
+In (b), results are shown for normal sinus rhythm (SR, $\blacktriangle$), ventricular tachycardia (VT, $\blacklozenge$) and ventricular fibrillation (VF, $\bullet$).
 } \label{fig:fluid}}
@@
 \section{Conclusions} \label{sec:con}
-In summary, we have proposed to include ordinal pattern transitions into the traditional ordinal pattern frequency analysis, generalizing statistical complexity measures for nonlinear time series analysis.
-The pattern transition properties of the corresponding time series networks provide novel insights for permutation entropy which relies on the static pattern frequency only.
-Therefore, this is another case showing the usefulness of ordinal pattern transition network for time series analysis \cite{ZouPR2018}.
-In addition, we suggest two slightly different ways to consider the effects of pattern transitions which are based on (i) a globally normalized weight transition matrix $\mathbf{W}$ and (ii) node-wise normalized transition matrix.
-Note that the node-wise based SCMs consider both individual pattern frequencies and the transition frequencies among them.
-We show that these two new statistical complexity measures characterize the complexity properties of deterministic chaos successfully, far away from the lower right bottom of the complexity-entropy plane (Fig.
-\ref{fig:fBmNoise}).
-This helps to obtain deeper understandings of complexity analysis for time series of both deterministic chaos and stochastic systems, for instance, further numerical studies are required to show distinctions of various dynamical systems of chaotic maps, long range correlated fractional Brownian motion and fractional Gaussian noise etc \cite{BorgesAMC2019,RossoPRE2007}.
-We have demonstrated the usefulness of SCMs by further results of time series for experimental fluid flows and ECG recordings, showing distinctive characteristics for different dynamics.
+In summary, we have proposed to expand the traditional ordinal pattern frequency analysis by also including ordinal pattern transitions, thereby generalizing existing statistical complexity measures for nonlinear time series analysis.
+The pattern transition properties encoded in the underlying time series networks provide novel insights beyond that provided by the celebrated permutation entropy which relies on the pattern frequencies only.
+Therefore, this study presents another case showing the usefulness of applying ordinal pattern transition networks for time series analysis \cite{ZouPR2018}.
 
-Some open problems of traditional nonlinear time series analysis should be further considered, for example, length dependence for reliable estimations of SCMs.
-For iterative processes of maps, one may use longer data lengths, which however does not improve the estimate very much due to the accumulation effects of computer iterations (Fig.
-\ref{fig:transient}).
-For continuous systems, together with embedding parameters of $D$ and $\tau$, the sampling frequency is another important one that needs to be further explored.
-Generalization from univariate to bivariate time series analysis will be further discussed.
-We noticed that this remains to be a largely untouched area for statistical complexity analysis.
-It will be an interesting task to show the dependence of complexity on interaction.
+We have suggested two slightly different ways to consider the effects of pattern transitions in statistical complexity measures, which are based on (i) a globally normalized transition matrix and (ii) a node-wise normalized transition matrix.
+Here, the complexity measure based on the node-wise normalized matrix considers both, individual pattern frequencies and the frequencies of transitions among patterns.
+We have shown that the associated two new statistical complexity measures characterize the complexity properties of deterministic chaos successfully and clearly highlight deterministic-chaotic dynamics in the corresponding complexity--entropy planes (Fig.~\ref{fig:fBmNoise}).
+This finding helps to obtain an improved complexity analysis for time series of both deterministic chaos and stochastic systems, fostering their better understanding.
+However, further more systematic numerical performance tests employing different types of dynamical systems are required to provide compelling evidence for distinctions between different dynamics beyond the cases presented in the current work, including various models of chaotic maps, long-range correlated stochastic processes like fractional Brownian motion and fractional Gaussian noise, etc.~\cite{BorgesAMC2019,RossoPRE2007}.
+While this opens several opportunities for follow-up research, we have demonstrated here the practical usefulness of our improved SCMs by employing this framework to time series for experimental fluid flows and ECG recordings, showing distinctive characteristics for different dynamics.
+
+Some open problems common to other nonlinear time series analysis methods remain to be further addressed in future work, for example, the dependence of reliable estimates of SCMs on the available time series lengths.
+For time-discrete dynamical systems like chaotic maps, one may easily use longer data series, which however does not improve the estimate very much due to the accumulation of effects due to a limited numerical precision (Fig.~\ref{fig:transient}).
+For continuous systems, together with the embedding parameters, the sampling frequency is another important characteristic which needs to be further explored.
+Moreover, the generalization of our approach from univariate to bivariate time series analysis shall be further discussed.
+We noticed that this aspect has remained a largely untouched subject for statistical complexity analysis so far, while it presents an interesting task to show the dependence of complexity on interaction.
+%Finally, this work has focused on complexity measures exploiting the classical notion of Shannon entropy in various ways.
+In turn, a plethora of dynamical systems exhibiting particularly complex behavior may present multifractal characteristics and would therefore benefit from considering the full spectrum of generalized (R\'enyi) entropies and associated statistical complexity measures to completely capture the information on dynamical complexity encoded in ordinal pattern sequences.
+We outline further work on the aforementioned topics as subjects of future research.
 
@@
 11872182 and 11835003).
-YZ further acknowledges the support by Shanghai Municipal Science and Technology Major Project (grant no.
+YZ further acknowledges support by the Shanghai Municipal Science and Technology Major Project (grant no.
 2018SHZDZX01) and ZJ Lab.
@@
 \bibliographystyle{unsrt}
-\bibliography{ref_Zou}
+\bibliography{ref_Zou,ref_Donner}
 

### 2020-11-24 manuscript_Huang.tex
new figures included.

@@
 On the other hand, for evaluating the statistical properties of either pattern frequencies or transition frequencies, this comes on the cost of requiring increasingly longer time series.
-In the following, we will use realizations of the logistic map comprising \textcolor{red}{$L=xxx$} iterations.
-\textcolor{red}{Yong: please add the used time series length!
-Is it the same as in the next subsection...?}
+In the following, we will use realizations of the logistic map comprising $L = 1 \times 10^6$ iterations.
 
@@
 For the three other cases of band merging, laminar state and outer crisis, the values of all SCMs significantly differ from zero.
-The pattern frequency based SCM $\mathcal{C}_O$ shows smaller values than the other two measures for all regimes, which implies a positioning in the lower right parts of the complexity--entropy planes in Fig.~\ref{fig:fBmNoise}(a).
+The pattern frequency based SCM $\mathcal{C}_O$ shows smaller values than the other two measures for all regimes, which implies a positioning in the lower right parts of the complexity--entropy planes in Fig.~\ref{fig:CElogistic}(a).
 However, for the transition matrix based SCMs, both the estimated entropies $\mathcal{H}_W$, $\mathcal{H}_E$ and associated complexity measures $\mathcal{C}_W$, $\mathcal{C}_E$ show results that are more consistent with the expectations when considering the respective level of chaoticity of the system in those three regimes (i.e., a rising Lyapunov exponent with increasing $r$ in the respective chaotic regimes).
@@
 It is notable that the logistic map has already been widely discussed in the framework of SCMs \cite{RossoPRE2007,MartinPLA2003}, but only for pattern frequency based complexity measures.
-For each value of $r$, we construct an OPTN from a time series of length $L = 31000$.
-In order to exclude transient behavior, we remove the first $1000$ values from the data series.
+For each value of $r$, we construct an OPTN from a time series of length $L = 1 \times 10 ^ 6$, after removing the first $1000$ values of transients from the data series.
 We have repeated the following calculations for embedding parameters $D = \tau \in [2, 7]$.
@@
 For a reliable estimation of SCMs, a pragmatic condition on the length $L$ of the time series $L$ as being sufficiently larger than $D!$ \cite{rossoPRL2007,kowalskiPhyD2007} has been satisfied in our setting for $D \leq 7$.
-In the particular example of logistic map, one may easily generate realizations of longer lengths, which however do not improve the obtained estimates significantly (as we have tested for $L\leq 5 \times 10^6$, not shown).
+In the particular example of logistic map, one may easily generate realizations of longer lengths, which however do not improve the obtained estimates significantly (as we have tested for $L \leq 5 \times 10^6$, not shown).
 
@@
 
-Figure~\ref{fig:fBmNoise} shows the complexity--entropy planes for embedding parameters $D = \tau$ varied between $4$ and $6$.
+Figure~\ref{fig:CElogistic} shows the complexity--entropy planes for embedding parameters $D = \tau$ varied between $4$ and $6$.
 We emphasize that the results are qualitatively similar when $D$ and $\tau$ are varied in the interval $[3, 7]$.
@@
 	\includegraphics[width=2\columnwidth]{CompEntropy_LogisticHC.pdf}
-\caption{\small{Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
+\caption{\small{(Color online)Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
 The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy.
@@
 In panels (a,d), the bottom-right rectangles of dotted red lines highlight the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
-\textcolor{red}{I don't quite get where this information comes from.
-Is it taken from Rosso's works or own results of this study?
-I do not see stochastic processes like fBm/fGn being discussed here, which is probably also not necessary.
-But we should clarify this.}}  \label{fig:fBmNoise}}
+}  \label{fig:CElogistic}}
 \end{figure*}
 
-First of all, we recover the known result that pattern frequency based SCM values lie around the maximum curve $\mathcal{C}_{max}$, while the displacements of the estimated values from their theoretical upper bounds depend on the embedding dimension $D$ as shown in Fig.~\ref{fig:fBmNoise}(a,d).
+First of all, we recover the known result that pattern frequency based SCM values lie around the maximum curve $\mathcal{C}_{max}$, while the displacements of the estimated values from their theoretical upper bounds depend on the embedding dimension $D$ as shown in Fig.~\ref{fig:CElogistic}(a,d).
 Most notable, we find many combinations of complexity--entropy values in the lower right part of the plane (high entropy, low complexity), which have been highlighted by rectangles of dotted lines.
@@
 The corresponding ambiguity results from the fact that both $\mathcal{H}_O$ and $\mathcal{C}_O$ only take static information in the sense that only pattern frequencies have been considered in their definitions.
-%The results that  $\mathcal{H}_O$ and $\mathcal{C}_O$ are displaced close to $\mathcal{C}_{max}$ curves have been observed when the embedding $D$ is increased (Fig.
-\ref{fig:fBmNoise}(a, d)).
 
 For our new pattern transition based SCMs, we find that in the complexity--entropy planes of $\mathcal{C}_W$ versus $\mathcal{H}_W$ and $\mathcal{C}_E$ versus $\mathcal{H}_E$, the results for the logistic map again closely align with the theoretical upper bounds $\mathcal{C}_{max}$.
-When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely and essentially cover the left uprising branch of the curve for the SCM based exclusively on transition frequencies (Fig.~\ref{fig:fBmNoise}(b,d)).
-By contrast, the definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes information on both, pattern frequencies and transition frequencies into account, which results in the obtained SCM values aligning even more closely with $\mathcal{C}_{max}$ already at lower embedding dimensions (Fig.~\ref{fig:fBmNoise}(c,f)).
+When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely and essentially cover the left uprising branch of the curve for the SCM based exclusively on transition frequencies (Fig.~\ref{fig:CElogistic}(b,d)).
+By contrast, the definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes information on both, pattern frequencies and transition frequencies into account, which results in the obtained SCM values aligning even more closely with $\mathcal{C}_{max}$ already at lower embedding dimensions (Fig.~\ref{fig:CElogistic}(c,f)).
 The most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of ambiguity.
@@
 
-Based on the results of Fig.~\ref{fig:fBmNoise}, we conclude that the pattern transition behavior encoded in the OPTNs provides novel insights that can be exploited in terms of SCMs that complement their traditional counterparts.
+Based on the results of Fig.~\ref{fig:CElogistic}, we conclude that the pattern transition behavior encoded in the OPTNs provides novel insights that can be exploited in terms of SCMs that complement their traditional counterparts.
+
+%% update but not properly integrated 
+We compare the results with fractional Brownian motion (fBm), which has long-range temporal correlations \cite{Mandelbrot1968}.
+Specifically, for an fBm process the long range of the process is characterized by the Hurst exponent $\alpha$ when positively correlated (persistence) for $1/2 < \alpha < 1$, while suppressed (anti-persistence) for $0< \alpha < 1/2$.
+$\alpha=1/2$ corresponds to the classical Brownian motion.
+In an analogy to the previous case of Logistic map of varying control parameter $r$, we generate time series for varying $\alpha \in (0, 1)$ with step size 0.05 and the results are shown in Fig.
+\ref{fig:CEfBm}.
+
+By the same time, we transform the fBm series into a stationary one by a first-order difference filter, i.e., by considering its increments $x_{i+1}- x_i$.
+The transformed series is commonly referred to as fractional Gaussian noise (fGn).
+Notably, fGn retains the long-range correlations and Gaussian probability density function (PDF) from the underlying
+fBm process. Note that the time delay embedding issues of both fBm and fGn have been discussed in detail in \cite{Zou2015}. The complexity--entropy planes of fGn are shown in Fig. \ref{fig:CEfGn}.
+
+First, the traditional pattern frequency based SCMs values are reproduced as shown in the Figures \ref{fig:CEfBm}(a, d) and \ref{fig:CEfGn}(a, d) fGn, which agree with that have been reported in \cite{RossoPRE2007,rossoPRL2007}.
+
+\begin{figure*}
+	\centering 
+	\includegraphics[width=2\columnwidth]{CompEntropy_fBm.pdf}
+\caption{\small{(Color online) Similar with Fig.
+\ref{fig:CElogistic} but for fractional Brownian motion of Hurst exponent $\alpha \in (0, 1)$ with step size 0.05.
+(a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
+The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy.
+Panels (a-c) show the results for embedding dimensions $D = 4,5$ and (d-f) for $D = 6$.
+}  \label{fig:CEfBm}}
+\end{figure*}
+
+\begin{figure*}
+	\centering 
+	\includegraphics[width=2\columnwidth]{CompEntropy_fGn.pdf}
+\caption{\small{(Color online) Similar with Fig.
+\ref{fig:CElogistic} but for fractional Gaussian noise Hurst exponent $\alpha \in (0, 1)$ with step size 0.05.
+(a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
+The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy.
+Panels (a-c) show the results for embedding dimensions $D = 4,5$ and (d-f) for $D = 6$.
+}  \label{fig:CEfGn}}
+\end{figure*}
 
@@
 Furthermore, there are jumps of all four measures at the points of period doubling bifurcations, which have been also reported earlier in Ref.~\cite{BandtPRL2002}.
-These jumps however are not desirable since the complexity of the dynamics does not change when $r$ passes nay of those points.
-For the case of the period doubling bifurcation at $r=3.544$ (replacing a period-4 by a period-8 solution), we show in Fig.~\ref{fig:transient} that the observed jump in the entropy and SCM values may be explained by recurrent iterations leading to long transients.
+These jumps however are not desirable since the complexity of the dynamics does not change when $r$ passes any of those points.
+For the case of the period doubling bifurcation at $r=3.544$ (replacing a period-4 by a period-8 solution), we show in Fig.~\ref{fig:transient} that the observed jump in the entropy and SCM values may be explained by accumulated numerical errors during the iterations attracted to the periodic-4 points, which yields long transients.
 \textcolor{red}{I do not fully get this reasoning...} For estimations based on time series of finite length as used in this work, $\mathcal{H}_E$ and $\mathcal{C}_E$ are however not affected in the same way.
@@
 In addition, there are many periodic windows of different periods along the axis of control parameter $r$, which prevents us from using a unique predefined number of iterations as initial transients for time series of different period length.
-Notably, similar jumps have also been observed when the system bifurcates from a period-3 to a period-6 solution.
-\textcolor{red}{Yong: give corresponding $r$ value here.}
+Notably, similar jumps have also been observed when the system bifurcates from a period-3 to a period-6 solution at $r = 3.842$.
 
@@
 Here, the complexity measure based on the node-wise normalized matrix considers both, individual pattern frequencies and the frequencies of transitions among patterns.
-We have shown that the associated two new statistical complexity measures characterize the complexity properties of deterministic chaos successfully and clearly highlight deterministic-chaotic dynamics in the corresponding complexity--entropy planes (Fig.~\ref{fig:fBmNoise}).
+We have shown that the associated two new statistical complexity measures characterize the complexity properties of deterministic chaos successfully and clearly highlight deterministic-chaotic dynamics in the corresponding complexity--entropy planes (Fig.~\ref{fig:CElogistic}).
 This finding helps to obtain an improved complexity analysis for time series of both deterministic chaos and stochastic systems, fostering their better understanding.

### 2020-11-26 manuscript_Huang.tex
Changes made by Donner

@@
 In this work, we focus on one particular approach: the application of ordinal pattern transition networks for characterizing time series data.
-More specifically, we generalize some traditional estimators of statistical complexity measures (SCMs), which explicitly disclose heterogeneous frequencies of ordinal pattern transitions in the time series networks.
-To demonstrate the usefulness of the estimators, we employ SCMs to characterize different dynamical transitions in the logistic map as a paradigmatic model system, as well as real-world time series of fluid experiments and ECG data.
-The obtained results for both, artificial and experimental data demonstrate that the consideration of transition frequencies between different ordinal patterns leads to dynamically meaningful estimates of SCMs, which provide prospective tools for the analysis of time series with finite length.
+More specifically, we generalize a traditional statistical complexity measure (SCM) based on permutation entropy by explicitly disclosing heterogeneous frequencies of ordinal pattern transitions.
+To demonstrate the usefulness of these generalized SCMs, we employ them to characterize different dynamical transitions in the logistic map as a paradigmatic model system, as well as real-world time series of fluid experiments and electrocardiogram recordings.
+The obtained results for both, artificial and experimental data demonstrate that the consideration of transition frequencies between different ordinal patterns leads to dynamically meaningful estimates of SCMs, which provide prospective tools for the analysis of observational time series.
 \end{abstract}
@@
 \begin{quotation}
-In the recent decade, traditional tools of nonlinear time series analysis have been undergoing fast developments benefiting from concepts from complex network theory.
-Along this line of research topics, ordinal pattern transition networks have been expanding the established concept of permutation entropy and provide complementary insights into the dynamical organization underlying time series data.
-Traditionally, permutations based on order patterns are simple and easy to implement concepts providing estimators of statistical complexity measures (SCMs), which however rely on static pattern frequencies only.
-Yet, much additional information can be exploited by including order pattern transitions into the definitions of SCMs -- an idea that however has not been widely developed and applied so far.
-In this work, we generalize existing estimators of SCMs by means of ordinal pattern transition networks which take into account the pattern transition properties explicitly.
-The usefulness of our generalizations of existing SCMs is demonstrated by using time series of both, model and experimental data.
+In the recent decade, the field of nonlinear time series analysis has been undergoing fast developments benefiting from concepts from complex network theory.
+Along this line of research, ordinal pattern transition networks have been expanding the established concept of ordinal time series analysis and provide new insights into the dynamical organization underlying time series data that complement existing methods like permutation entropy.
+Permutations based on ordinal patterns are a simple and easy to implement concept that naturally provides statistical complexity measures (SCMs), which in the case of permutation entropy rely on pattern frequencies only.
+Yet, much additional information can be exploited by including order pattern transition frequencies into the definitions of SCMs -- an idea that however has not been widely developed and applied so far.
+In this work, we generalize existing permutation based SCMs by means of ordinal pattern transition networks which take into account the pattern transition properties explicitly.
+The usefulness of our generalizations is demonstrated by using time series of both, model and experimental data.
 \end{quotation}
@@
 \section{Introduction}
-In the recent decade, complex network based approaches have emerged as prominent tools for nonlinear time series analysis \cite{Kantz97,Sprott2003}, which have found many interesting applications to experimental data from various origins \cite{ZouPR2018}.
-In this line of research, there exist several major methods that are of particular importance, including recurrence networks \cite{MarwanPLA2009,Donner2010a}, visibility graphs \cite{Lacasa2008,Nunez2012}, transition networks \cite{Nicolis2005,MichaelChaos2015} and cycle networks \cite{ZhangPRL2006}.
+Complex network approaches \cite{ZouPR2018} have recently emerged as prominent tools for nonlinear time series analysis \cite{Kantz97,Sprott2003} and already found many interesting applications to observational and experimental data from various fields.
+In this line of research, there exist several major methods that are of particular practical importance, including cycle networks \cite{ZhangPRL2006}, recurrence networks \cite{MarwanPLA2009,Donner2010a}, visibility graphs \cite{Lacasa2008,Nunez2012}, and transition networks \cite{Nicolis2005,MichaelChaos2015}.
 Depending on the particular questions at hand, each of these complex network approaches exhibit various variants that can be employed for different application purposes.
@@
 For a recent systematic review, see Ref.~\cite{ZouPR2018}.
-In this work, we focus on constructing ordinal pattern transition networks (OPTNs) for time series data, which present the advantages of easy implementation and a wide range of existing application to data from different origins, including correlated stochastic processes, neurophysiological (EEG) and human cardiac activity (ECG) series \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019,Subramaniyam2020}.
+In this work, we focus on constructing ordinal pattern transition networks (OPTNs) derived from time series data, which present the advantage of an easy implementation and a wide range of existing applications to data from different origins, including correlated stochastic processes, neurophysiological (electroencephalogram, EEG) and human cardiac activity (electrocardiogram, ECG) series \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019,Subramaniyam2020}.
 
@@
 
-More specifically, given a one-dimensional time series $\{ x_i\}_{i=1, \cdots, L}$ comprising $L$ observations from a dynamical system, we first reconstruct the corresponding phase space by time delay embedding $\vec{x}_i = [x_i, x_{i+\tau}, \ldots, x_{i+(D-1)\tau}]$ with an embedding dimension $D$ \cite{Takens1981,Kantz97}.
-Next, we represent each embedding vector $\vec{x}(i)$ by the corresponding rank order of its components, which is encoded into a symbol $\pi_{i}$.
-Hence, when sliding windows from $i=1$ to $N = L - (D - 1)\tau$ in the embedding space, a symbolic representation of the trajectory $\{\pi_i\}$ is produced.
+Given a one-dimensional time series $\{ x_i\}_{i=1, \ldots, L}$ comprising $L$ observations from a dynamical system, we first reconstruct the corresponding phase space by time delay embedding $\vec{x}_i = [x_i, x_{i+\tau}, \ldots, x_{i+(D-1)\tau}]$ with an embedding dimension $D$ \cite{Takens1981,Kantz97}.
+Next, we represent each embedding vector $\vec{x}_i$ by the corresponding rank order of its components, which is encoded into a symbol $\pi_{i}$ indicating the corresponding ordinal pattern.
+Hence, when sliding windows from $i=1$ to $N = L - (D - 1)\tau$ in the embedding space, a symbolic representation $\{\pi_i\}_{i=1, \ldots, N}$ of the trajectory is produced.
 It is possible to derive information about the dynamics of the underlying system by assessing the probabilities of occurrence of different ordinal patterns.
-For example, for time series resulting from deterministic dynamical systems, certain patterns may not occur at all.
-More generally, we obtain an empirical probability distribution $P$ whose elements $p_{\pi_{i}}$ are the frequencies associated to the different patterns $\pi_{i}$, $i = 1, \cdots, D!$.
-Clearly, $P$ provides a significant and feasible way to estimate a characteristic probability distribution function for a given time series, which plays a crucial role in extracting statistical complexity information on the underlying dynamical systems \cite{BandtPRL2002,AmigoPTRSA2014}.
+For example, for time series resulting from deterministic dynamical systems, certain patterns may not occur at all \cite{AmigoEPL2007,AmigoEPL2008,ZaninChaos2008,RossoEPJB2012}.
+More generally, we obtain an empirical probability distribution $P$ whose elements $p_{\pi_{i}}$ are the frequencies associated with the different patterns $\pi_{i}$, $i = 1, \ldots, D!$.
+Clearly, $P$ provides a significant and feasible way to estimate a characteristic probability distribution function for a given time series, which plays a crucial role in extracting statistical complexity information on the underlying dynamical system \cite{BandtPRL2002,AmigoPTRSA2014}.
 
 The conceptually relatively simple computation of $P$ based on ordinal patterns for the underlying system prompts to another successful application of this framework in terms of statistical complexity measures (SCMs) \cite{kowalskiEntropy2011}.
-One common way of defining SCMs is taking the product of a normalized Shannon entropy and an associated measure of disequilibrium \cite{LopezPLA1995,kowalskiEntropy2011}, which captures specific organizational properties of structure and patterns in a process.
-Following the vast amount of previous work on this topic  \cite{rossoPRL2007,LopezPLA1995,kowalskiEntropy2011,Lange2013}, in this work we consider a variant of SCMs based on the concept of Jensen-Shannon divergence, which is defined as $\mathcal{C}_{JS}[P] = \mathcal{H}[P] \cdot  Q_{JS}[P, P_e]$, where $\mathcal{H}[P]$ is the normalized Shannon entropy and $Q_{JS}[P, P_e]$ is the so-called disequilibrium function measuring the distance of the given distribution $P$ to the uniform distribution $P_e$.
+One common way of defining SCMs is taking the product of a normalized Shannon entropy and an associated measure of disequilibrium \cite{LopezPLA1995,kowalskiEntropy2011}, which captures specific organizational properties of structure and patterns in the observed process.
+Following the vast amount of previous work on this topic  \cite{rossoPRL2007,LopezPLA1995,kowalskiEntropy2011,Lange2013}, we consider here a variant of SCMs based on the concept of Jensen-Shannon divergence, which is defined as $\mathcal{C}_{JS}[P] = \mathcal{H}[P] \cdot  Q_{JS}[P, P_e]$, where $\mathcal{H}[P]$ is the normalized Shannon entropy and $Q_{JS}[P, P_e]$ is the so-called disequilibrium function measuring the distance of the given distribution $P$ to the uniform distribution $P_e$.
 Accordingly, we define $Q_{JS}[P, P_e] = Q_{0} \cdot JS[P, P_e]$ where $JS[P, P_e]$ is the Jensen-Shannon divergence between $P$ and $P_e$ and $Q_{0}$ is a normalization factor.
@@
 The analysis of complexity--entropy planes has already taken advantage from the ordinal pattern based symbolic encoding of time series and has found various applications \cite{rossoPRL2007,kowalskiEntropy2011}.
-However, in the complexity--entropy analysis, there are some chaotic maps that could be easily confused with random noise because no clearly separable domains of values are available to differentiate those different types of dynamics as reported in \cite{BorgesAMC2019}.
+However, in the classical complexity--entropy analysis, there are some chaotic maps that could be easily confused with random noise because no clearly separable domains of values are available to differentiate those different types of dynamics as reported in Ref.~\cite{BorgesAMC2019}.
 So far, it remains a challenging task to disclose a possible non-monotonic relationship between complexity and entropy \cite{MartinPLA2003}.
@@
 
-In this work, we draw upon the recent success of both, complexity--entropy plain analysis and OPTNs, to generalize the traditional SCMs by incorporating pattern transition probabilities.
-The advantages of this approach are as follows: (i) All SCMs successfully show clear distinction between different types of dynamics, which is demonstrated for time series from numerical models and experimental data.
+In this work, we draw upon the recent success of both, complexity--entropy plane analysis and OPTNs, to generalize an existing SCM by incorporating pattern transition probabilities.
+The advantages of this approach, which will be demonstrated in the course of this work, are as follows: (i) The modified SCMs allow for a clear distinction between different types of dynamics, which is demonstrated for time series from numerical models and experimental data.
 (ii) The generalized SCMs help to capture a consistent relationship between SCMs and chaoticity.
@@
 Based on this matrix, we propose in Section~\ref{sec:SCM} to compute SCMs from three different perspectives, including static pattern frequency properties, dynamic pattern transition properties, and a combination of both.
-In Section~\ref{sec:results1}, we then proceed with describing some selected results for the logistic map as as paradigmatic model system exhibiting various types of periodic and chaotic solutions.
+In Section~\ref{sec:results1}, we then proceed with describing some selected results for the logistic map as a paradigmatic model system exhibiting various types of periodic and chaotic solutions.
 We first discuss in some detail one example of a periodic orbit to emphasize the important effects of embedding parameters in Section~\ref{sec:embeddings}.
 The practical usefulness of SCMs will be further illustrated by studying four exemplary time series from different typical dynamical regimes (Section~\ref{sec:four}).
-In Section~\ref{sec:plane}, the complexity--entropy planes associated with each SCM will be discussed while the control parameter of the logistic map is systematically increased.
+In Section~\ref{sec:plane}, the complexity--entropy planes associated with each SCM will be discussed as the control parameter of the logistic map is systematically varied.
 In order to characterize the dynamical transitions in this model, we further study the behavior of the SCM values in dependence on the control parameter in Section~\ref{sec:transi} and show that similar results can also be obtained for time-continuous dynamical systems like the R\"ossler oscillator in Section~\ref{sec:cont}.
@@
 Most recent studies employing the concept of permutation entropy $\mathcal{S}_O$ have defined this measure by focusing exclusively on the frequencies of ordinal patterns, which however disregards the transition behavior between subsequent patterns.
-To emphasize the transition frequencies between any pair of patterns, measures of transitional complexity have been further proposed in Refs.~\cite{zhangSciRep2017,McCullough2017b} to quantify the properties of the associated OPTNs.
-To this end, we associate each directed link in an OPTN with its transition frequency $w_{ij} = p({\pi_i \to \pi_j})$.
-Calculating the transition frequencies for each pair of patterns, we obtain a weighted directed network characterized by a weighted adjacency matrix $\mathbf{W} = \{ w_{ij} \}, i, j \in [1, D!]$.
+To emphasize on the transition frequencies between any pair of patterns, measures of transitional complexity have been further proposed in Refs.~\cite{zhangSciRep2017,McCullough2017b} to quantify the properties of the associated OPTNs.
+To this end, we associate each directed link in an OPTN with a transition frequency $w_{ij} = p_{\pi_i \to \pi_j}$.
+Calculating the transition frequencies for each pair of patterns, we hence obtain a weighted and directed network characterized by a weighted adjacency matrix $\mathbf{W} = \{ w_{ij} \}$ with $i, j \in \{1,\ldots, D!\}$.
 
 Based on this transition matrix $\mathbf{W}$ of an OPTN, we note that there two slightly different ways to introduce normalization factors.
-The first option is to use a global normalization to ensure $\sum_{i,j}^{D!} w_{ij} = 1$ \cite{zhangSciRep2017}, which results in a globally normalized transition entropy as discussed below.
-The second option has been proposed by McCullough {\textit{et al}}~\cite{McCullough2017b} who first considered the local out-link transition frequency from pattern $\pi_i$ to $\pi_j$
+The first option \cite{zhangSciRep2017} is to use a global normalization to ensure that $\sum_{i,j}^{D!} w_{ij} = 1$, which results in a globally normalized transition entropy as discussed below.
+The second option has been proposed by McCullough {\textit{et al.}}~\cite{McCullough2017b} who first considered the local out-link transition frequency from pattern $\pi_i$ to $\pi_j$
 \begin{equation} \label{eq:localTp}
@@
 \noindent
-Note that the transition frequency of Eq.~\eqref{eq:localTp} is pattern (row-) wise normalized.
+Note that the transition frequency of Eq.~\eqref{eq:localTp} is pattern (row) wise normalized.
 This case will be referred to as the node-wise out-link normalized transition matrix $\mathbf{W}$ in the following.
-In this way, it easily captures the heterogeneous behavior of both (static) occurrence frequencies of different patterns and (dynamic) transition frequencies between patterns.
+In this way, it easily captures the heterogeneous behavior of both, (static) occurrence frequencies of different patterns and (dynamic) transition frequencies between patterns.
 
 In addition to the proper normalization, we need to consider self-loops that may affect the numerical estimation of $\mathbf{W}$ especially for time series from continuous systems with a given sampling rate.
-For example, it has been demonstrated in Ref.~\cite{zhangSciRep2017} that there are about $99\%$ self-loops in the time series of R\"ossler system when integrated with a step size $h = 0.01$, while the about $1\%$ non-self-loops are hidden in $\mathbf{W}$.
+For example, it has been demonstrated in Ref.~\cite{zhangSciRep2017} that there are about $99\%$ self-loops in time series of the R\"ossler system when integrated with a step size $h = 0.01$, while the about $1\%$ non-self-loops are hidden in $\mathbf{W}$.
 From this example, one may easily see that self-loops are related to the serial correlation of the time series under study \cite{BorgesAMC2019}.
 On the other hand, neglecting self-loops in $\mathbf{W}$ emphasizes the transition behavior between different patterns, which has therefore been adopted in most existing studies of complex networks for computational simplicity and theoretical concerns \cite{CostaADPhy2007}.
-Accordingly, we also remove self-loops in the present study before computing the weighted matrix $\mathbf{W}$ while acknowledging that in other cases of stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results \cite{BorgesAMC2019}.
+Accordingly, we also remove self-loops in the present study before computing $\mathbf{W}$ while acknowledging that in other cases of stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results \cite{BorgesAMC2019}.
 Arguably, whether to remove or to consider self-loops depends on the particular process under study.
-In any of the applications discussed in the rest of this paper, we have ensured that our results do not change qualitatively when considering the effects of self-loops.
+In any of the applications discussed in the rest of this paper, we have ensured that our results do not change qualitatively when including self-loops.
 
@@
 
-In the following, we will briefly review the existing ordinal pattern frequency based SCM before proposing two novel ways to define SCMs, which explicitly consider transition frequencies of the resulting OPTN representations for time series.
+In the following, we briefly review the existing ordinal pattern frequency based SCM before proposing two novel ways to define SCMs, which explicitly consider transition frequencies of the resulting OPTN representations for time series.
 
@@
 
-The permutation entropy of a time series is the Shannon entropy of the distribution of patterns $P = \{p_{\pi_1}, p_{\pi_2},\ldots,p_{\pi_i} \}$, which is defined as
+The permutation entropy of a time series is the Shannon entropy of the distribution of patterns $P = \{p_{\pi_1}, p_{\pi_2},\ldots,p_{\pi_{D!}} \}$, which is defined as
 \begin{equation}
@@
 
-In the next step, we use some appropriate statistical concept to measure the distance of the given empirical distribution $P$ to the uniform distribution $P_e$, which is the essential idea behind the class of SCMs considered in this work.
-Several different distance functions have been considered in the literature, including the Euclidean norm, Wooters' distance, Kullback-Leibler relative entropy and Jensen-Shannon divergence \cite{kowalskiEntropy2011}.
+In the next step, we use some appropriate statistical concept to measure the distance of the empirical distribution $P$ to a uniform distribution $P_e$, which is the essential idea behind the class of SCMs considered in this work.
+Several different distance functions have been employed in the literature, including the Euclidean norm, Wooters' distance, Kullback-Leibler relative entropy and Jensen-Shannon divergence \cite{kowalskiEntropy2011}.
 In this work, the distance between the distributions $P$ and $P_e$ is characterized by the Jensen-Shannon divergence $Q_{JS}[P] = Q_0 \cdot JS[P, P_e]$ with
@@
 The function $Q_{JS}[P]$ is different from zero if the system has a kind of preference of certain patterns among the accessible ones, reflecting determinism of the system.
-Finally, the SCM based on permutation entropy of node (pattern) frequencies is defined as
+Finally, the SCM based on permutation entropy estimated from the empirical node (pattern) frequencies is defined as
 \begin{equation}
@@
 \end{equation}
-This quantity is static in the sense that it quantifies the amount of information stored by ordinal frequencies in the system and its disequilibrium of the observed parts of its accessible states in comparison to a uniform distribution \cite{LopezPLA1995}.
+This quantity is static in the sense that it quantifies the amount of information stored by ordinal pattern frequencies in the system and its disequilibrium of the observed parts of its accessible states in comparison to a uniform distribution \cite{LopezPLA1995}.
 
@@
 (D!
-- 1)$ possible pattern transitions in the uniformly distribution function $P_e$.
-Thus, in a full analogy with the node-based SCM, the Jensen-Shannon distance is computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$ with
+- 1)$ possible pattern transitions in the uniform distribution function $P_e$.
+Thus, in a full analogy with the permutation entropy based SCM, the Jensen-Shannon divergence is computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$ with
 \begin{equation}
@@
 \end{equation}
-We note that we have the same expression for the normalization factor $Q_0$ as above while replacing $N = D! (D! - 1)$ in Eq.~\eqref{eq:Q0}. This further yields the SCM 
+We note that we have the same expression for the normalization factor $Q_0$ as above when setting $N = D! (D! - 1)$ in Eq.~\eqref{eq:Q0}. This further yields the SCM 
 \begin{equation}
@@
 
-We finally consider the case where the ordinal transition matrix is node-wised normalized, i.e., the row sum of $\mathbf{W}$ is 1.
-Accordingly, we again first introduce the node-wise out-edge transition entropy
+We finally consider the case where the ordinal pattern transition matrix is node-wised normalized, i.e., the row sum of $\mathbf{W}$ is 1.
+Accordingly, we again first introduce the node-wise out-link transition entropy
 \begin{equation}
@@
 Therefore, the superscript index can be suppressed in the above definition.
-In addition, the node-wise Jensen-Shannon distance is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$ with the disequilibrium measure
+In addition, the node-wise Jensen-Shannon divergence is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$ with the disequilibrium measure
 \begin{eqnarray}
@@
 \end{eqnarray}
-In this case, we obtain the normalization factor $Q_0$ by replacing $N = D! - 1$ in Eq.~\eqref{eq:Q0}. Taking into account node frequencies, the global Jensen-Shannon divergence is computed as 
+In this case, we obtain the normalization factor $Q_0$ by setting $N = D! - 1$ in Eq.~\eqref{eq:Q0}. Taking into account node frequencies, the global Jensen-Shannon divergence is computed as 
 \begin{equation}
@@
 \end{equation}
-which therefore yields an alternative definition of SCM of the OPTN as 
+which therefore yields an alternative definition of an SCM of the OPTN as 
 \begin{equation}
@@
 where $\mathcal{H}_{E}[W] = \sum_{i} p_{\pi_i} \mathcal{H}_{E}^{\pi_i}[W_{\pi_i}]$.
-This quantity is dynamic in the sense that it takes both the static ordinal state frequencies and the dynamic state transition information stored in the system into account.
+This quantity is dynamic in the sense that it takes both the static ordinal pattern frequencies and the dynamic state transition information into account.
 
@@
 There are two important algorithmic parameters in constructing OPTNs -- the embedding parameters $D$ and $\tau$.
-The choice of $D$ and $\tau$ may markedly affect the results for the associated SCMs since $D$ and $\tau$ should compromise the essential time scales of the dynamics in the underlying time series.
-In common applications of ordinal pattern analysis (e.g., estimation of the permutation entropy $\mathcal{S}_O$), the choice of optimal (or just pragmatic) values of $D$ and $\tau$ depends on the field of application, which requires experience of experts as systematically summarized in \cite{Riedl2013}.
+The choice of $D$ and $\tau$ can markedly affect the results for the associated SCMs since $D$ and $\tau$ should compromise the essential time scales of the dynamics in the underlying time series.
+In common applications of ordinal pattern analysis (e.g., the estimation of the permutation entropy $\mathcal{S}_O$), the choice of optimal (or just pragmatic) values of $D$ and $\tau$ depends on the field of application as systematically summarized in Ref.~\cite{Riedl2013}.
 
@@
 Other values of $D$ and $\tau$ however yield non-zero entropy values.
-The results of Fig.~\ref{fig:embed} may raise concern since different choices if $D$ and $\tau$ will change the placement of the system in the complexity--entropy plane.
+The results of Fig.~\ref{fig:embed} may raise concern since different choices of $D$ and $\tau$ will change the placement of the system in the complexity--entropy plane.
 In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for periodic windows when the control parameter $r$ is varied.
-In such cases, we should focus more on the transitions between different dynamic regimes.
-In the following, we will show that the results are robust when the control parameter is systematically changed while keeping the same parameter settings of $D$ and $\tau$.
+In such cases, we should focus more on the transitions between different dynamical regimes.
+In the following, we will show that the results are robust when the control parameter is systematically changed while keeping the same embedding parameters.
 
-\subsection{Results for four typical dynamic regimes} \label{sec:four}
+\subsection{Results for four typical dynamical regimes} \label{sec:four}
 
-We further illustrate the potential of the proposed SCMs to characterize different dynamical regimes and transitions for the logistic map with the control parameter $r$ varying in the range $[3.5, 4]$ with a step size of $\Delta r = 0.001$.
-In this range of $r$, various dynamical regimes and transitions between them can be found, for instance, period doublings, band merging points, inner and outer crises, and intermittency \cite{Kantz97}, which has made this system serving as a paradigmatic model for assessing nonlinear time series analysis methods \cite{MarwanPLA2009}.
+We further illustrate the potential of the proposed SCMs to characterize the logistic map with the control parameter $r$ varying in the range $[3.5, 4]$ with a step size of $\Delta r = 0.001$.
+In this range of $r$, various dynamical regimes and transitions between them can be found, for instance, period doubling cascades, band merging points, inner and outer crises, and intermittency \cite{Kantz97}, which has made this system serving as a paradigmatic model for assessing nonlinear time series analysis methods \cite{MarwanPLA2009}.
 
@@
 On the other hand, for evaluating the statistical properties of either pattern frequencies or transition frequencies, this comes on the cost of requiring increasingly longer time series.
-In the following, we will use realizations of the logistic map comprising $L = 1 \times 10^6$ iterations.
+In the following, we will use realizations of the logistic map comprising $L = 10^6$ iterations (after removing the first $1000$ iterations of each realization which might reflect transient behavior of the system before reaching its respective attractor).
 
 For all four situations, we investigate the SCMs along with their dual entropy characteristics in detail and summarize the obtained results in Tab.~\ref{tableLog}.
-We note that several types of chaos--chaos transitions will be discussed in the following: the band merging crisis corresponds to intermittency, the inner crisis to certain chaos-chaos transition and the outer crisis to fully chaotic dynamics.
+We note that several types of chaos--chaos transitions will be discussed in the following: the band merging crisis corresponds to intermittency, the inner crisis to some more subtle chaos-chaos transition and the outer crisis to fully chaotic dynamics.
 
@@
     \end{tabular}}
-   \caption{Statistical complexity measures for four different values of the control parameter $r$ of the logistic map (embedding parameters: $D = \tau = 6$).   \label{tableLog}}    
+   \caption{Values of the three considered SCMs along with their adjoint entropy measures for four different values of the control parameter $r$ of the logistic map (embedding parameters: $D = \tau = 6$).
+\label{tableLog}}
 \end{table}
@@
 Due to our choice of $\tau$ and $D$, all SCMs take values of zero for the period-3 series because the temporal distance between each pair of components of the embedding vector covers a multiple of two full periods (cf.~Fig.~\ref{fig:embed}).
-Other choices of $D$ and $\tau$ would commonly yields non-zero values for the period-3 regime.
+Other choices of $D$ and $\tau$ would commonly yield non-zero values for the period-3 regime.
 For the three other cases of band merging, laminar state and outer crisis, the values of all SCMs significantly differ from zero.
-The pattern frequency based SCM $\mathcal{C}_O$ shows smaller values than the other two measures for all regimes, which implies a positioning in the lower right parts of the complexity--entropy planes in Fig.~\ref{fig:CElogistic}(a).
+The pattern frequency based SCM $\mathcal{C}_O$ shows smaller values than the other two measures for all regimes, which implies a positioning in the lower right part of the complexity--entropy plane in Fig.~\ref{fig:CElogistic}(a).
 However, for the transition matrix based SCMs, both the estimated entropies $\mathcal{H}_W$, $\mathcal{H}_E$ and associated complexity measures $\mathcal{C}_W$, $\mathcal{C}_E$ show results that are more consistent with the expectations when considering the respective level of chaoticity of the system in those three regimes (i.e., a rising Lyapunov exponent with increasing $r$ in the respective chaotic regimes).
@@
 In order to put our previous results into a broader context, we next show the complexity--entropy planes obtained when varying the control parameter $r$ in the logistic map.
-It is notable that the logistic map has already been widely discussed in the framework of SCMs \cite{RossoPRE2007,MartinPLA2003}, but only for pattern frequency based complexity measures.
-For each value of $r$, we construct an OPTN from a time series of length $L = 1 \times 10 ^ 6$, after removing the first $1000$ values of transients from the data series.
-We have repeated the following calculations for embedding parameters $D = \tau \in [2, 7]$.
+This simple nonlinear system has already been widely discussed in the framework of SCMs \cite{RossoPRE2007,MartinPLA2003}, but only for pattern frequency based complexity measures.
+For each value of $r$, we construct an OPTN from a time series of length $L = 10 ^ 6$ with the embedding parameters $D = \tau \in \{2,\ldots,7\}$.
 
-When only one unique ordinal pattern is identified in a given time series, all SCMs take a value of $0$, which is reasonable since only self-loops are observed.
-This however happens only if $D$ is an integer multiple of the period for the periodic windows of the logistic map (e.g., Fig.~\ref{fig:embed}(g)).
-Another practical concern is the data requirements for computing entropies and complexity measures.
-For a reliable estimation of SCMs, a pragmatic condition on the length $L$ of the time series $L$ as being sufficiently larger than $D!$ \cite{rossoPRL2007,kowalskiPhyD2007} has been satisfied in our setting for $D \leq 7$.
-In the particular example of logistic map, one may easily generate realizations of longer lengths, which however do not improve the obtained estimates significantly (as we have tested for $L \leq 5 \times 10^6$, not shown).
+As already stated above, when only one unique ordinal pattern is identified in a given time series, all SCMs take a value of $0$, which is reasonable since only self-loops are observed.
+This however happens only if the embedding parameters are chosen such that they exactly coincide with the periodicity of a certain periodic window (see Fig.~\ref{fig:embed}(g)).
+Another practical aspect concerns the data requirements for computing entropies and complexity measures.
+For a reliable estimation of SCMs, a pragmatic condition on the length $L$ of the time series to be sufficiently larger than $D!$ \cite{rossoPRL2007,kowalskiPhyD2007} has been satisfied in our setting for $D \leq 7$.
+In the particular example of the logistic map, one may easily generate much longer realizations, which however do not improve the obtained entropy and SCM estimates significantly (as we have tested for $L \leq 5 \times 10^6$, not shown).
 
 In previous works, SCMs have been applied to distinguish chaos from noise by employing the concept of complexity--entropy plane~\cite{rossoPRL2007}.
-Generalizing this idea to our OPTN based SCMs, this plane shows the statistical complexity measures ($\mathcal{C}_O$, $\mathcal{C}_W$ and $\mathcal{C}_E$) as a function of the corresponding normalized entropy values ($\mathcal{H}_O$, $\mathcal{H}_W$ and $\mathcal{H}_E$).
-The three SCMs characterize both randomness and correlation structures in a time series, which in consequence result in a non-trivial dependence on the associated entropy values.
-Specifically, chaotic systems present maximal complexity while stochastic systems have lower values of complexity, hence appearing in different regions of the complexity--entropy plane~\cite{rossoPRL2007}.
-Furthermore, for a given entropy value, the range of possible SCM values is bound by a minimum $\mathcal{C}_{min}$ and a maximum $\mathcal{C}_{max}$.
-A general algorithm for computing the bounds $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$ has been provided in Ref.~\cite{martinPhyA2006}.
+Generalizing this idea to our OPTN based SCMs, we can define three such planes showing the statistical complexity measures ($\mathcal{C}_O$, $\mathcal{C}_W$ and $\mathcal{C}_E$) as a function of the corresponding normalized entropy values ($\mathcal{H}_O$, $\mathcal{H}_W$ and $\mathcal{H}_E$).
+The three SCMs characterize both randomness and correlation structure in a time series, which as a consequence result in a non-trivial dependence on the associated entropy values.
+Specifically, chaotic systems present high complexity while stochastic systems have lower values of complexity, hence ideally appearing in distinct regions of the complexity--entropy plane~\cite{rossoPRL2007}.
+Furthermore, for a given entropy value, it has been shown previously that at a given entropy value, the range of possible SCM values is bound by a minimum $\mathcal{C}_{min}$ and a maximum $\mathcal{C}_{max}$.
+A general algorithm for computing these bounds has been provided in Ref.~\cite{martinPhyA2006}.
 
 Figure~\ref{fig:CElogistic} shows the complexity--entropy planes for embedding parameters $D = \tau$ varied between $4$ and $6$.
-We emphasize that the results are qualitatively similar when $D$ and $\tau$ are varied in the interval $[3, 7]$.
+We emphasize that the results are qualitatively similar when $D$ and $\tau$ are varied in the range $\{3,\ldots, 7\}$.
 Furthermore, it is notable that the values of $\mathcal{C}_{max}$ depend on the embedding dimension $D$ since $D$ determines the number of patterns and pattern transitions which are considered in the definition of SCMs.
@@
 	\includegraphics[width=2\columnwidth]{CompEntropy_LogisticHC.pdf}
-\caption{\small{(Color online)Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
-The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy.
+\caption{\small{(Color online) Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
+The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values in dependence on the respective entropy.
 Panels (a-c) show the results for embedding dimensions $D = 4,5$ and (d-f) for $D = 6$.
-In panels (a,d), the bottom-right rectangles of dotted red lines highlight the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
+In panels (a,d), the bottom-right rectangles of dotted red lines indicate the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
 }  \label{fig:CElogistic}}
@@
 
-First of all, we recover the known result that pattern frequency based SCM values lie around the maximum curve $\mathcal{C}_{max}$, while the displacements of the estimated values from their theoretical upper bounds depend on the embedding dimension $D$ as shown in Fig.~\ref{fig:CElogistic}(a,d).
-Most notable, we find many combinations of complexity--entropy values in the lower right part of the plane (high entropy, low complexity), which have been highlighted by rectangles of dotted lines.
+First of all, we recover the known result that pattern frequency based SCM values lie close to the maximum $\mathcal{C}_{max}$, while the displacements of the estimated values from their theoretical upper bounds depend on the embedding dimension $D$ as shown in Fig.~\ref{fig:CElogistic}(a,d).
+Most notably, we find many combinations of complexity--entropy values in the lower right part of the plane (high entropy, low complexity), which have been highlighted by rectangles of dotted lines.
 This region of the plane is challenging for SCMs to distinguish between chaotic time series and random noise~\cite{BorgesAMC2019} and is in our case occupied by values corresponding to chaotic regimes of the logistic map at large $r$ values.
@@
 
-For our new pattern transition based SCMs, we find that in the complexity--entropy planes of $\mathcal{C}_W$ versus $\mathcal{H}_W$ and $\mathcal{C}_E$ versus $\mathcal{H}_E$, the results for the logistic map again closely align with the theoretical upper bounds $\mathcal{C}_{max}$.
-When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely and essentially cover the left uprising branch of the curve for the SCM based exclusively on transition frequencies (Fig.~\ref{fig:CElogistic}(b,d)).
-By contrast, the definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes information on both, pattern frequencies and transition frequencies into account, which results in the obtained SCM values aligning even more closely with $\mathcal{C}_{max}$ already at lower embedding dimensions (Fig.~\ref{fig:CElogistic}(c,f)).
-The most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of ambiguity.
-We therefore suggest that it is the deterministic transition behavior among patterns that makes the distinction from random noise.
+For our new pattern transition based SCMs, we find that in the complexity--entropy planes $\mathcal{H}_W \times \mathcal{C}_W$ and $\mathcal{H}_E \times \mathcal{C}_E$, the results for the logistic map again closely align with the theoretical upper bounds $\mathcal{C}_{max}$.
+When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely and essentially cover exclusively the left uprising branch of the curve for the SCM based solely on transition frequencies (Figs.~\ref{fig:CElogistic}(b,e) and \ref{fig:CElogistic}(c,f)), which distinctively differs from the original permutation entropy based SCM.
+Specifically, the definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes information on both, pattern frequencies and transition frequencies into account, which results in the obtained SCM values aligning even more closely with $\mathcal{C}_{max}$ already at lower embedding dimensions (Fig.~\ref{fig:CElogistic}(c,f)).
+The in our opinion most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of possible ambiguity between deterministic-chaotic and stochastic dynamics.
+We therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
 
-Based on the results of Fig.~\ref{fig:CElogistic}, we conclude that the pattern transition behavior encoded in the OPTNs provides novel insights that can be exploited in terms of SCMs that complement their traditional counterparts.
-
-%% update but not properly integrated 
-We compare the results with fractional Brownian motion (fBm), which has long-range temporal correlations \cite{Mandelbrot1968}.
-Specifically, for an fBm process the long range of the process is characterized by the Hurst exponent $\alpha$ when positively correlated (persistence) for $1/2 < \alpha < 1$, while suppressed (anti-persistence) for $0< \alpha < 1/2$.
-$\alpha=1/2$ corresponds to the classical Brownian motion.
-In an analogy to the previous case of Logistic map of varying control parameter $r$, we generate time series for varying $\alpha \in (0, 1)$ with step size 0.05 and the results are shown in Fig.
-\ref{fig:CEfBm}.
-
-By the same time, we transform the fBm series into a stationary one by a first-order difference filter, i.e., by considering its increments $x_{i+1}- x_i$.
-The transformed series is commonly referred to as fractional Gaussian noise (fGn).
-Notably, fGn retains the long-range correlations and Gaussian probability density function (PDF) from the underlying
-fBm process. Note that the time delay embedding issues of both fBm and fGn have been discussed in detail in \cite{Zou2015}. The complexity--entropy planes of fGn are shown in Fig. \ref{fig:CEfGn}.
-
-First, the traditional pattern frequency based SCMs values are reproduced as shown in the Figures \ref{fig:CEfBm}(a, d) and \ref{fig:CEfGn}(a, d) fGn, which agree with that have been reported in \cite{RossoPRE2007,rossoPRL2007}.
+In order to further analyze the capability of our modified complexity--entropy planes to distinguish chaotic from stochastic dynamics, we compare the results for the logistic map with those obtained for fractional Brownian motion (fBm), i.e., stochastic processes with long-range temporal correlations.
+Specifically, for an fBm process the long-range correlations of the process are uniquely characterized by the Hurst exponent $H$ when positively correlated (persistence) for $1/2 < H < 1$, while being suppressed (anti-persistence) for $0< H < 1/2$.
+$H=1/2$ corresponds to the classical Brownian motion.
+Similar to the previous case of the logistic map with varying control parameter $r$, we generate time series for varying $H \in (0, 1)$ with a step size of $\Delta H=0.05$ and present the resulting complexity--entropy planes in Fig.~\ref{fig:CEfBm}.
+Moreover, since fBm series in the persistent regime are nonstationary, we further transform all fBm realizations into stationary time series by employing first-order difference filtering, i.e., by considering the increments $x_{i+1}- x_i$.
+The transformed series are commonly referred to as fractional Gaussian noise (fGn), and the associated complexity--entropy planes are shown in Fig.~\ref{fig:CEfGn}.
+Notably, fGn retains the long-range correlations and Gaussian probability density function (PDF) of the underlying
+fBm process. For a discussion on practical concerns regarding time delay embedding of fBm and fGn processes, we refer to Ref.~\cite{Zou2015}.
 
@@
 	\includegraphics[width=2\columnwidth]{CompEntropy_fBm.pdf}
-\caption{\small{(Color online) Similar with Fig.
-\ref{fig:CElogistic} but for fractional Brownian motion of Hurst exponent $\alpha \in (0, 1)$ with step size 0.05.
-(a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
-The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy.
-Panels (a-c) show the results for embedding dimensions $D = 4,5$ and (d-f) for $D = 6$.
-}  \label{fig:CEfBm}}
+\caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Brownian motion with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$.
+Note that the normalized entropy is generally decreasing with increasing $H$.}  \label{fig:CEfBm}}
 \end{figure*}
@@
 	\includegraphics[width=2\columnwidth]{CompEntropy_fGn.pdf}
-\caption{\small{(Color online) Similar with Fig.
-\ref{fig:CElogistic} but for fractional Gaussian noise Hurst exponent $\alpha \in (0, 1)$ with step size 0.05.
-(a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
-The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values for a fixed value of the entropy.
-Panels (a-c) show the results for embedding dimensions $D = 4,5$ and (d-f) for $D = 6$.
+\caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Gaussian noise with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$.
+Note that only a small region of the complexity--entropy planes is covered in this case.
 }  \label{fig:CEfGn}}
@@
 
+Comparing the results for both types of stochastic processes with those for the logistic map, a few important observations are made.
+First of all, the traditional pattern frequency based SCM values as originally reported by Rosso and co-workers \cite{RossoPRE2007,rossoPRL2007} are well reproduced as expected (Figs.~\ref{fig:CEfBm}(a,d) and \ref{fig:CEfGn}(a,d)).
+Second, all complexity--entropy pairs of fBm and fGn clearly differ from the theoretical maximum complexity values, which does not change with increasing embedding dimension $D$.
+Especially for the new OPTN based SCMs, we find a partial overlap between the entropy values of fBm and the logistic map, while the associated SCM values are however clearly distinct except for the cases with the lowest entropy, which correspond to the most nonstationary situations ($H$ close to 1).
+Accordingly, we suggest that combining information on the position in the complexity--entropy plane and on the stationarity could allow a clear distinction between chaotic and stochastic dynamics even in those extreme situations.
+Notably, for the stationary fGn, all entropy measures are confined to very high entropy and low complexity values -- a range where no combinations have been found for the new transition frequency based SCMs for the logistic map.
+
+Taken all reported results on the complexity--entropy planes for the logistic map, fBm and fGn together, we conclude that the modified SCMs indeed allow distinguishing chaotic from stochastic dynamics more precisely than the original approach.
+We therefore suggest that the pattern transition behavior encoded in the OPTNs provides novel insights that can be exploited in terms of SCMs that complement their traditional counterparts.
+
 \subsection{Characterizing dynamical transitions} \label{sec:transi}
-The logistic map experiences a sequences of different types of bifurcations when the control parameter $r$ is systematically increased.
+The logistic map experiences a sequence of different types of bifurcations when the control parameter $r$ is systematically increased.
 However, the corresponding dynamical regimes and regime transitions are not easy to identify from the complexity--entropy planes discussed above.
-In the following, we therefore aim to expand the previous results by showing the dependence of our SCMs on the control parameter $r$.
+In the following, we therefore explicitly study the dependence of our SCMs on the control parameter $r$.
 Here, our motivation is to verify whether the normalized entropies and associated SCMs computed from OPTNs are indeed able to detect the dynamical transitions along the complex bifurcation scenarios of the logistic map and hence track qualitative changes in the dynamics, including both period--chaos and chaos--chaos transitions.
-For the sake of comparison, we take the associated Lyapunov exponent of the logistic map as an established measure for characterizing the type of dynamics (regular versus chaotic) along with the degree of chaoticity.
+As a benchmark, we take the associated Lyapunov exponent as an established measure for characterizing the type of dynamics (regular versus chaotic) along with the degree of chaoticity.
 
@@
 \caption{\small{Behavior of the different SCMs and associated entropy characteristics for the logistic map in dependence on the control parameter $r$.
-(a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{S}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ based on the globally normalized weight matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g,h) entropy $\mathcal{S}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized weight matrix $\mathbf{W}$.
+(a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{S}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ based on the globally normalized transition matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g,h) entropy $\mathcal{S}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized transition matrix $\mathbf{W}$.
 Several major periodic windows have been highlighted by grey background shading.
@@
 Other than $\mathcal{C}_O$, the two transition frequency based SCMs behave in a similar way as the Lyapunov exponent.
-Specifically, as the control parameter $r$ is increased, the chaoticity level grows gradually, reaching a maximum at $r = 4$, which is reflected in an increase in the statistical complexity.
+Specifically, as the control parameter $r$ is increased, the chaoticity level grows gradually, reaching a maximum at $r = 4$, which is reflected by an increase in statistical complexity.
 A similar increase is absent in the pattern frequency based complexity measure $\mathcal{C}_O$ (Fig.~\ref{fig:bifurcation}(c)).
-In this context, we note that it is an established fact that pattern frequency based SCMs often do not trace the growth in the chaoticity of the logistic map~\cite{MartinPLA2003}, which can however be improved by using Wooter's distance function instead of the Shannon-Jensen divergence used in this work.
-From this point of view, the pattern transition frequency based SCMs $\mathcal{C}_W$ and $\mathcal{C}_E$ exhibit more informative behavior in the sense that they track the growth of the level of chaoticity.
+In this context, we note that it is an established fact that pattern frequency based SCMs often do not trace the growth in the chaoticity of the logistic map~\cite{MartinPLA2003}, which can however be improved by using Wooter's distance function instead of the Shannon-Jensen divergence employed in this work.
+From this point of view, the pattern transition frequency based SCMs $\mathcal{C}_W$ and $\mathcal{C}_E$ exhibit some more informative behavior in the sense that they track the growth of the level of chaoticity.
 
@@
 It may be noticed that $\mathcal{H}_O$, $\mathcal{C}_O$, $\mathcal{H}_W$ and $\mathcal{C}_W$ all exhibit non-zero values in this parameter range despite a purely periodic dynamics.
-Furthermore, there are jumps of all four measures at the points of period doubling bifurcations, which have been also reported earlier in Ref.~\cite{BandtPRL2002}.
+Furthermore, there are jumps of all four measures at the points of period doubling bifurcations, which have been already reported in Ref.~\cite{BandtPRL2002}.
 These jumps however are not desirable since the complexity of the dynamics does not change when $r$ passes any of those points.
 For the case of the period doubling bifurcation at $r=3.544$ (replacing a period-4 by a period-8 solution), we show in Fig.~\ref{fig:transient} that the observed jump in the entropy and SCM values may be explained by accumulated numerical errors during the iterations attracted to the periodic-4 points, which yields long transients.
-\textcolor{red}{I do not fully get this reasoning...} For estimations based on time series of finite length as used in this work, $\mathcal{H}_E$ and $\mathcal{C}_E$ are however not affected in the same way.
+For estimations based on time series of finite length as used in this work, $\mathcal{H}_E$ and $\mathcal{C}_E$ are however not affected in the same way.
 Certainly, we should not over-interpret their capabilities since numerical inaccuracy would be accumulated in a longer iterative process such that different ordinal patterns are identified.
-In addition, there are many periodic windows of different periods along the axis of control parameter $r$, which prevents us from using a unique predefined number of iterations as initial transients for time series of different period length.
+In addition, we pass through many periodic windows of different periods as the control parameter $r$ is varied, which prevents us from using a unique predefined number of iterations as initial transients for time series of different period length.
 Notably, similar jumps have also been observed when the system bifurcates from a period-3 to a period-6 solution at $r = 3.842$.
@@
 	\includegraphics[width=\columnwidth]{period4_exampleTransients.pdf}
-\caption{\small{Transient effects in the generation of proper bifurcation diagrams for the logistic map, exemplified by the period doubling bifurcation from a period-4 to a period-8 solution.
+\caption{\small{Transient behavior affecting the generation of proper bifurcation diagrams for the logistic map, exemplified here for the period doubling bifurcation from a period-4 to a period-8 solution.
 (a) The actual bifurcation point gets blurred when $r$ is close to $3.544$ if only a short transient of $50$ iterations is removed from the time series.
-(b) A large number of initial iterations ($2000$) are removed, the correct bifurcation point becomes visible.
+(b) When a sufficiently large number of initial iterations ($2000$) are removed, the correct bifurcation point becomes visible.
 }\label{fig:transient}}
@@
 
-While we have focused exclusively on the case of the time-discrete logistic map so far, it may be interesting to identify whether a similar behavior can also be obtained for time-continuous deterministic dynamical systems.
+While we have focused so far exclusively on the case of the time-discrete logistic map, it may be interesting to study whether a similar behavior can also be obtained for time-continuous deterministic dynamical systems.
 To this end, we will illustrate a corresponding analysis for the example of the chaotic R\"ossler system \cite{Roessler1976}
@@
 To avoid any corresponding discussion on proper choices of further algorithmic parameters, we employ here a Poincar\'e section to each sample trajectory of the system at $y=0$, $\dot{y}<0$ (for different values of $b$) and construct OPTNs from $N = 10,000$ intersection points.
-The results are shown in Fig.
-\ref{fig:bifurRossler}.
-To this end, we conclude that the general behavior of the different SCMs and associated entropy measures resembles that previously observed for the logistic map in Fig.~\ref{fig:bifurcation}.
+The results are shown in Fig.~\ref{fig:bifurRossler}.
+To this end, we conclude that the general behavior of the different SCMs and associated entropy measures closely resembles that reported for the logistic map in Fig.~\ref{fig:bifurcation}.
 Specifically, all measures trace the succession of bifurcations in the considered range of $b$.
-We outline more detailed follow-up investigations on the behavior of our OPTN based SCMs as relevant topics for future work.
+We outline more detailed follow-up investigations on the behavior of our OPTN based SCMs to such time-continuous systems as relevant topics for future work.
 
@@
 \begin{enumerate}[(i)]
-\item {Stable fluid:} The temperature signal of the stable flow exhibits periodic oscillations as the wave drift patterns arrive at the measured fixed point. 
-\item {Quasi-periodic 2-frequency amplitude vacillation (AV-2):} This case is identified as a 2-frequency quasi-periodic amplitude flux.
+\item {Stable fluid:} The temperature signal of the stable flow exhibits periodic oscillations as the wave drift patterns arrive at the fixed point of measurements. 
+\item {Quasi-periodic 2-frequency amplitude vacillation:} This case is identified as a 2-frequency quasi-periodic amplitude flux.
 The wave drift is composed of slow and regular oscillations in the temperature signal, but a fast modulation in the amplitude is also visible.
 AV-2 is characterized by a periodic growth and drop in wave amplitude with little change in waveform.
-\item {Quasi-periodic 3-frequency amplitude vacillation (AV-3):} We also analyze a quasi-periodic 3-frequency amplitude vacillation time series which is obtained from a three-dimensional direct numerical simulation in an air-filled rotating baroclinic instability experiment~\cite{Read_jfm_1992}.
-\item {Modulated amplitude vacillation (MAV):} This case is identified as a low-dimensional stream, chaotically modulated and with waves of varying amplitudes.
-Amplitude modulation results in a complex temperature signal.
+\item {Quasi-periodic 3-frequency amplitude vacillation:} We also analyze a quasi-periodic 3-frequency amplitude vacillation time series which is obtained from a three-dimensional direct numerical simulation in an air-filled rotating baroclinic instability experiment~\cite{Read_jfm_1992}.
+\textcolor{red}{Yong: This reads like only this series is from a numerical model while the others are experimental data.
+Please clarify.}
+\item {Modulated amplitude vacillation:} This case is identified as a low-dimensional stream, chaotically modulated and with waves of varying amplitudes.
+Amplitude modulation results in a complex temperature dynamics.
 \end{enumerate}
@@
 The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients in a Coronary Care Unit and the Creighton University Cardiac Center.
-Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant but still relatively poorly understood dynamical system, showing rich nonlinear properties.
+\textcolor{red}{Yong: please provide here the reference to the original data (prior to Michael's paper), if there is any...} Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant anomaly that is still relatively poorly understood from a dynamical system perspective, showing rich nonlinear properties.
 In this work, we focus on three different states: normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}.
-From the data bank of recordings of ECG during various rhythms, we selected 14 time series showing sinus rhythm (SR) prior to onset of arrhythmia, 12 time series of VT and 17 of VF.
+From the data base of ECG recordings during various rhythms, we selected 14 time series showing sinus rhythm (SR) prior to onset of arrhythmia, 12 time series of VT and 17 of VF.
 The ECG recordings have been annotated with SR, VT and VF by skilled cardiologists on the basis of waveform morphology.
@@
 In the case of temperature time series from the fluid experiments, $\tau_{i} = 218$~s for the stable solution, $\tau_{ii} = 136.5$~s  for AV-2, $\tau_{iii} = 185.625$~s for AV-3, and $\tau_{iv} = 118$~s for MAV.
-For the ECG time series, we consistently use an embedding delay of $\tau = 5$.
+For the ECG time series, we consistently use an embedding delay of $\tau = 5$ time steps.
 The results have been averaged over all subjects.
@@
 In both sets of time series, the OPTN based SCMs enhance our understanding of the underlying nonlinear system, which can be seen from the results summarized in Fig.~\ref{fig:fluid}.
-In the case of the fluid data (Fig.~\ref{fig:fluid}(a)), $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ all show consistent changes of the complexity and entropy values between stable wave solution (SW), quasi-periodic 2-frequency amplitude vacillation (AV-2), quasi-periodic 3-frequency amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV).
-Specifically, the complexity of these four cases has the following order: SW has the smallest complexity since it shows periodic oscillations with trivial recurrence, while MAV has the highest complexity since it presents chaotically modulated amplitude vacillation.
-Quasiperiodic states show intermediate complexity values since these two cases have non-trivial recurrences but with low complexity.
+For the fluid data (Fig.~\ref{fig:fluid}(a)), $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ all show consistent changes of the complexity and entropy values between stable wave solution (SW), quasi-periodic 2-frequency amplitude vacillation (AV-2), quasi-periodic 3-frequency amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV).
+Specifically, the complexity of these four cases has the following order: SW has the lowest complexity since it shows periodic oscillations with trivial recurrence, while MAV has the highest complexity since it presents chaotically modulated amplitude vacillation.
+Quasiperiodic states exhibit intermediate complexity values since these two cases have non-trivial recurrences while still featuring a certain degree of regularity.
 Interestingly, AV-3 presents higher complexity than AV-2.
-The order of complexity values between periodic, quasiperiodic and chaotic solutions still holds for $\mathcal{C}_O$ although the corresponding values become comparable.
+The order of complexity values between periodic, quasiperiodic and chaotic solutions still holds for $\mathcal{C}_O$, but the corresponding values become very similar.
 In particular, it becomes hard for $\mathcal{C}_O$ to identify the difference between AV-2 and AV-3.
@@
 On the other hand, the complexity of VT has the smallest value, even lower than for the case of VF.
-Unlike the other five measures, we note, however, that SR shows the smallest value of $\mathcal{C}_O$ while VT and VF have comparable values that are higher than SR.
+Unlike the other five measures, we note, however, that SR shows the smallest value of $\mathcal{C}_O$ while VT and VF have comparable values that are higher than for SR.
 This difference may be explained by the fact that unlike the two other OPTN based SCMs, this measure only takes static information of pattern frequencies into account and may therefore not be able to properly distinguish subtle differences reflected in the temporal order of patterns.
@@
 In summary, we have proposed to expand the traditional ordinal pattern frequency analysis by also including ordinal pattern transitions, thereby generalizing existing statistical complexity measures for nonlinear time series analysis.
-The pattern transition properties encoded in the underlying time series networks provide novel insights beyond that provided by the celebrated permutation entropy which relies on the pattern frequencies only.
+The pattern transition properties encoded in the underlying time series networks allow to gain insights beyond those provided by the celebrated permutation entropy method which relies on the pattern frequencies only.
 Therefore, this study presents another case showing the usefulness of applying ordinal pattern transition networks for time series analysis \cite{ZouPR2018}.
 
-We have suggested two slightly different ways to consider the effects of pattern transitions in statistical complexity measures, which are based on (i) a globally normalized transition matrix and (ii) a node-wise normalized transition matrix.
-Here, the complexity measure based on the node-wise normalized matrix considers both, individual pattern frequencies and the frequencies of transitions among patterns.
+We have suggested two slightly different ways to incorporate the effects of pattern transitions in statistical complexity measures, which are based on (i) a globally normalized transition matrix and (ii) a node-wise normalized transition matrix.
+Here, the complexity measure based on the node-wise normalized matrix combines both, individual pattern frequencies and the frequencies of transitions among patterns.
 We have shown that the associated two new statistical complexity measures characterize the complexity properties of deterministic chaos successfully and clearly highlight deterministic-chaotic dynamics in the corresponding complexity--entropy planes (Fig.~\ref{fig:CElogistic}).
-This finding helps to obtain an improved complexity analysis for time series of both deterministic chaos and stochastic systems, fostering their better understanding.
-However, further more systematic numerical performance tests employing different types of dynamical systems are required to provide compelling evidence for distinctions between different dynamics beyond the cases presented in the current work, including various models of chaotic maps, long-range correlated stochastic processes like fractional Brownian motion and fractional Gaussian noise, etc.~\cite{BorgesAMC2019,RossoPRE2007}.
-While this opens several opportunities for follow-up research, we have demonstrated here the practical usefulness of our improved SCMs by employing this framework to time series for experimental fluid flows and ECG recordings, showing distinctive characteristics for different dynamics.
+This finding helps to obtain an improved complexity analysis for time series of both, deterministic chaos and stochastic systems, fostering their better discrimination.
+However, further more systematic numerical performance tests employing different types of dynamical systems are required to provide compelling evidence for distinctions between different types of dynamics beyond the cases presented in the current work~\cite{BorgesAMC2019,RossoPRE2007}.
+While this opens several opportunities for follow-up research, we have already demonstrated here the practical usefulness of our improved SCMs by employing this framework to time series of experimental fluid flows and ECG recordings, showing distinctive characteristics for different cases.
 
-Some open problems common to other nonlinear time series analysis methods remain to be further addressed in future work, for example, the dependence of reliable estimates of SCMs on the available time series lengths.
-For time-discrete dynamical systems like chaotic maps, one may easily use longer data series, which however does not improve the estimate very much due to the accumulation of effects due to a limited numerical precision (Fig.~\ref{fig:transient}).
+Some open problems also common to other nonlinear time series analysis methods remain to be further addressed, for example, the dependence of reliable estimates of SCMs on the available time series length.
+For time-discrete dynamical systems like chaotic maps, one may easily use longer data series, which however does not seem to improve the estimates very much due to the accumulation of effects due to a limited numerical precision (Fig.~\ref{fig:transient}).
 For continuous systems, together with the embedding parameters, the sampling frequency is another important characteristic which needs to be further explored.

### 2020-11-26 manuscript_Huang.tex
almost done!

@@
 
-\author{Jie Zhang}
-	\affiliation{Institute of Science and Technology for Brain-Inspired Intellegence, Fudan University, Shanghai 200433,  China}
-
 \author{Reik V. Donner}
@@
         \affiliation{Potsdam Institute for Climate Impact Research (PIK) -- Member of the Leibniz Association, Telegrafenberg A31, 14473 Potsdam, Germany}
+
+\author{Jie Zhang}
+	\affiliation{Institute of Science and Technology for Brain-Inspired Intellegence, Fudan University, Shanghai 200433,  China}
 
@@
 
-In this work, we draw upon the recent success of both, complexity--entropy plane analysis and OPTNs, to generalize an existing SCM by incorporating pattern transition probabilities.
+In this work, we draw upon the recent success of both, complexity--entropy plane analysis and OPTNs, to generalize  existing SCMs by incorporating pattern transition probabilities.
 The advantages of this approach, which will be demonstrated in the course of this work, are as follows: (i) The modified SCMs allow for a clear distinction between different types of dynamics, which is demonstrated for time series from numerical models and experimental data.
@@
 	\includegraphics[width=2\columnwidth]{CompEntropy_LogisticHC.pdf}
-\caption{\small{(Color online) Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a,d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b,e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c,f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
+\caption{\small{(Color online) Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
 The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values in dependence on the respective entropy.
 Panels (a-c) show the results for embedding dimensions $D = 4,5$ and (d-f) for $D = 6$.
-In panels (a,d), the bottom-right rectangles of dotted red lines indicate the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
+In panels (a, d), the bottom-right rectangles of dotted red lines indicate the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
 }  \label{fig:CElogistic}}
@@
 For our new pattern transition based SCMs, we find that in the complexity--entropy planes $\mathcal{H}_W \times \mathcal{C}_W$ and $\mathcal{H}_E \times \mathcal{C}_E$, the results for the logistic map again closely align with the theoretical upper bounds $\mathcal{C}_{max}$.
-When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely and essentially cover exclusively the left uprising branch of the curve for the SCM based solely on transition frequencies (Figs.~\ref{fig:CElogistic}(b,e) and \ref{fig:CElogistic}(c,f)), which distinctively differs from the original permutation entropy based SCM.
-Specifically, the definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes information on both, pattern frequencies and transition frequencies into account, which results in the obtained SCM values aligning even more closely with $\mathcal{C}_{max}$ already at lower embedding dimensions (Fig.~\ref{fig:CElogistic}(c,f)).
-The in our opinion most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of possible ambiguity between deterministic-chaotic and stochastic dynamics.
+When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely and essentially cover exclusively the left uprising branch of the curve for the SCM based solely on transition frequencies (Figs.~\ref{fig:CElogistic}(b, e) and \ref{fig:CElogistic}(c, f)), which distinctively differs from the original permutation entropy based SCM.
+Specifically, the definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes information on both, pattern frequencies and transition frequencies into account, which results in the obtained SCM values aligning even more closely with $\mathcal{C}_{max}$ already at lower embedding dimensions (Fig.~\ref{fig:CElogistic}(c, f)).
+In our opinion, the most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of possible ambiguity between deterministic-chaotic and stochastic dynamics.
 We therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
@@
 The transformed series are commonly referred to as fractional Gaussian noise (fGn), and the associated complexity--entropy planes are shown in Fig.~\ref{fig:CEfGn}.
-Notably, fGn retains the long-range correlations and Gaussian probability density function (PDF) of the underlying
-fBm process. For a discussion on practical concerns regarding time delay embedding of fBm and fGn processes, we refer to Ref.~\cite{Zou2015}.
+Notably, fGn retains the long-range correlations and Gaussian probability density function (PDF) of the underlying fBm process.
+For a discussion on practical concerns regarding time delay embedding of fBm and fGn processes, we refer to Ref.~\cite{Zou2015}.
 
@@
 \caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Brownian motion with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$.
-Note that the normalized entropy is generally decreasing with increasing $H$.}  \label{fig:CEfBm}}
+Note that the normalized entropy is generally decreasing with increasing $H$.
+{\color{red} Should the text arrows are confusing while indicating the increasing direction of $H$?
+}}  \label{fig:CEfBm}}
 \end{figure*}
@@
 Comparing the results for both types of stochastic processes with those for the logistic map, a few important observations are made.
-First of all, the traditional pattern frequency based SCM values as originally reported by Rosso and co-workers \cite{RossoPRE2007,rossoPRL2007} are well reproduced as expected (Figs.~\ref{fig:CEfBm}(a,d) and \ref{fig:CEfGn}(a,d)).
+First of all, the traditional pattern frequency based SCM values as originally reported by Rosso and co-workers \cite{RossoPRE2007,rossoPRL2007} are well reproduced as expected (Figs.~\ref{fig:CEfBm}(a, d) and \ref{fig:CEfGn}(a, d)).
 Second, all complexity--entropy pairs of fBm and fGn clearly differ from the theoretical maximum complexity values, which does not change with increasing embedding dimension $D$.
@@
 \caption{\small{Behavior of the different SCMs and associated entropy characteristics for the logistic map in dependence on the control parameter $r$.
-(a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{S}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{S}_W$ based on the globally normalized transition matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g,h) entropy $\mathcal{S}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized transition matrix $\mathbf{W}$.
+(a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{H}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ based on the globally normalized transition matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g, h) entropy $\mathcal{H}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized transition matrix $\mathbf{W}$.
 Several major periodic windows have been highlighted by grey background shading.
@@
 
-Another conceptual improvement is found for $\mathcal{H}_E$ and $\mathcal{C}_E$ in the parameter range where the logistic map presents period doubling bifurcations, for example, at $r = 3.544$ (Fig.~\ref{fig:bifurcation}(g,h)).
+Another conceptual improvement is found for $\mathcal{H}_E$ and $\mathcal{C}_E$ in the parameter range where the logistic map presents period doubling bifurcations, for example, at $r = 3.544$ (Fig.~\ref{fig:bifurcation}(g, h)).
 It may be noticed that $\mathcal{H}_O$, $\mathcal{C}_O$, $\mathcal{H}_W$ and $\mathcal{C}_W$ all exhibit non-zero values in this parameter range despite a purely periodic dynamics.
@@
 The results are shown in Fig.~\ref{fig:bifurRossler}.
-To this end, we conclude that the general behavior of the different SCMs and associated entropy measures closely resembles that reported for the logistic map in Fig.~\ref{fig:bifurcation}.
+Based on these results, we conclude that the general behavior of the different SCMs and associated entropy measures closely resembles that reported for the logistic map in Fig.~\ref{fig:bifurcation}.
 Specifically, all measures trace the succession of bifurcations in the considered range of $b$.
-We outline more detailed follow-up investigations on the behavior of our OPTN based SCMs to such time-continuous systems as relevant topics for future work.
+We outline more detailed follow-up investigations on the behavior of our OPTN based SCMs to such time-continuous systems directly as relevant topics for future work.
 
@@
 The first example originates from a laboratory fluid experiment of baroclinic instability which is used to study patterns of cyclones and anticyclones in the Earth's atmosphere~\cite{Read_jfm_1992,ZouEPJST2008}.
-Depending on the parameters rotation rate, temperature difference, viscosity and fluid density, this experimental system exhibits a rich variety of flow regimes.
+Depending on the parameters of rotation rate, temperature difference, viscosity and fluid density, this experimental system exhibits a rich variety of flow regimes.
 In particular, we focus on the following regimes:
@@
 \item {Quasi-periodic 3-frequency amplitude vacillation:} We also analyze a quasi-periodic 3-frequency amplitude vacillation time series which is obtained from a three-dimensional direct numerical simulation in an air-filled rotating baroclinic instability experiment~\cite{Read_jfm_1992}.
-\textcolor{red}{Yong: This reads like only this series is from a numerical model while the others are experimental data.
-Please clarify.}
 \item {Modulated amplitude vacillation:} This case is identified as a low-dimensional stream, chaotically modulated and with waves of varying amplitudes.
@@
 
-The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients in a Coronary Care Unit and the Creighton University Cardiac Center.
-\textcolor{red}{Yong: please provide here the reference to the original data (prior to Michael's paper), if there is any...} Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant anomaly that is still relatively poorly understood from a dynamical system perspective, showing rich nonlinear properties.
+The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients from MIT-BIH Database and the Creighton University Cardiac Center \cite{GoldbergerMITBIH2000}.
+Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant anomaly that is still relatively poorly understood from a dynamical system perspective, showing rich nonlinear properties.
 In this work, we focus on three different states: normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}.
@@
 We noticed that this aspect has remained a largely untouched subject for statistical complexity analysis so far, while it presents an interesting task to show the dependence of complexity on interaction.
-%Finally, this work has focused on complexity measures exploiting the classical notion of Shannon entropy in various ways.
-In turn, a plethora of dynamical systems exhibiting particularly complex behavior may present multifractal characteristics and would therefore benefit from considering the full spectrum of generalized (R\'enyi) entropies and associated statistical complexity measures to completely capture the information on dynamical complexity encoded in ordinal pattern sequences.
 We outline further work on the aforementioned topics as subjects of future research.
@@
 2018SHZDZX01) and ZJ Lab.
+The authors thank J.
+B.
+Borges and O.
+A.
+Rosso for sharing algorithms for computing $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$.
 

### 2020-11-26 manuscript_Huang.tex
Updated by Reik.

@@
 \caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Brownian motion with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$.
-Note that the normalized entropy is generally decreasing with increasing $H$.
-{\color{red} Should the text arrows are confusing while indicating the increasing direction of $H$?
-}}  \label{fig:CEfBm}}
+Note that the normalized entropy is generally decreasing with increasing $H$ (red arrow).}  \label{fig:CEfBm}}
 \end{figure*}
@@
 
-While we have previously demonstrated the suitability of OPTN based SCMs for tracing changes in dynamical complexity for time series from deterministic dynamical systems, in the following we will use two different experimental time series to demonstrate that SCMs also successfully characterize the complexity properties of real-world systems, showing distinct values for different types of dynamics.
+While we have previously demonstrated the suitability of OPTN based SCMs for tracing changes in dynamical complexity for time series from deterministic dynamical systems, in the following we will use two different sets of experimental time series to demonstrate that SCMs also successfully characterize the complexity properties of real-world systems, showing distinct values for different types of dynamics.
 
@@
 AV-2 is characterized by a periodic growth and drop in wave amplitude with little change in waveform.
-\item {Quasi-periodic 3-frequency amplitude vacillation:} We also analyze a quasi-periodic 3-frequency amplitude vacillation time series which is obtained from a three-dimensional direct numerical simulation in an air-filled rotating baroclinic instability experiment~\cite{Read_jfm_1992}.
+\item {Quasi-periodic 3-frequency amplitude vacillation:} For the sake of completeness, we also analyze a quasi-periodic 3-frequency amplitude vacillation time series.
+Unlike for the other studied cases, this time series has not been obtained from experimental results, but from a three-dimensional direct numerical simulation of the air-filled rotating baroclinic instability experiment~\cite{Read_jfm_1992}.
 \item {Modulated amplitude vacillation:} This case is identified as a low-dimensional stream, chaotically modulated and with waves of varying amplitudes.
@@
 Quasiperiodic states exhibit intermediate complexity values since these two cases have non-trivial recurrences while still featuring a certain degree of regularity.
-Interestingly, AV-3 presents higher complexity than AV-2.
+In our case, AV-3 presents a higher complexity than AV-2, which should however be interpreted with care since the AV-2 series originates from an experiment, while the AV-3 series has been taken from a numerical simulation of the same system (hence, there may be a certain level of observational noise present in AV-2 but not in AV-3).
 The order of complexity values between periodic, quasiperiodic and chaotic solutions still holds for $\mathcal{C}_O$, but the corresponding values become very similar.

### 2020-11-27 manuscript_Huang.tex
updates after the submission.

@@
-\documentclass[12pt,aip,cha,reprint,nofootinbib]{revtex4-1}
+\documentclass[aip,cha,reprint,nofootinbib]{revtex4-1}
 \usepackage{amsfonts,amssymb,amsmath,times}%
@@
 
-\bibliographystyle{unsrt}
-\bibliography{ref_Zou,ref_Donner}
+% \bibliography{ref_Zou,ref_Donner}
+
+%merlin.mbs aipnum4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
+%Control: key (0)
+%Control: author (8) initials jnrlst
+%Control: editor formatted (1) identically to author
+%Control: production of article title (0) allowed
+%Control: page (1) range
+%Control: year (1) truncated
+%Control: production of eprint (0) enabled
+\begin{thebibliography}{39}%
+\makeatletter
+\providecommand \@ifxundefined [1]{%
+ \@ifx{#1\undefined}
+}%
+\providecommand \@ifnum [1]{%
+ \ifnum #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \@ifx [1]{%
+ \ifx #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \natexlab [1]{#1}%
+\providecommand \enquote  [1]{``#1''}%
+\providecommand \bibnamefont  [1]{#1}%
+\providecommand \bibfnamefont [1]{#1}%
+\providecommand \citenamefont [1]{#1}%
+\providecommand \href@noop [0]{\@secondoftwo}%
+\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
+\providecommand \@href[1]{\@@startlink{#1}\@@href}%
+\providecommand \@@href[1]{\endgroup#1\@@endlink}%
+\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
+  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
+\providecommand \@@startlink[1]{}%
+\providecommand \@@endlink[0]{}%
+\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
+\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
+\providecommand \urlprefix  [0]{URL }%
+\providecommand \Eprint [0]{\href }%
+\providecommand \doibase [0]{http://dx.doi.org/}%
+\providecommand \selectlanguage [0]{\@gobble}%
+\providecommand \bibinfo  [0]{\@secondoftwo}%
+\providecommand \bibfield  [0]{\@secondoftwo}%
+\providecommand \translation [1]{[#1]}%
+\providecommand \BibitemOpen [0]{}%
+\providecommand \bibitemStop [0]{}%
+\providecommand \bibitemNoStop [0]{.\EOS\space}%
+\providecommand \EOS [0]{\spacefactor3000\relax}%
+\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
+\let\auto@bib@innerbib\@empty
+%</preamble>
+\bibitem [{\citenamefont {Zou}\ \emph {et~al.}(2019)\citenamefont {Zou},
+  \citenamefont {Donner}, \citenamefont {Marwan}, \citenamefont {Donges},\ and\
+  \citenamefont {Kurths}}]{ZouPR2018}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Zou}}, \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont {Donner}},
+  \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Marwan}}, \bibinfo
+  {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Complex network approaches to nonlinear time
+  series analysis},}\ }\href {\doibase 10.1016/j.physrep.2018.10.005}
+  {\bibfield  {journal} {\bibinfo  {journal} {Physics Reports}\ }\textbf
+  {\bibinfo {volume} {787}},\ \bibinfo {pages} {1--97} (\bibinfo {year}
+  {2019})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kantz}\ and\ \citenamefont
+  {Schreiber}(2004)}]{Kantz97}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
+  {Kantz}}\ and\ \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Schreiber}},\ }\href@noop {} {\emph {\bibinfo {title} {Nonlinear Time Series
+  Analysis}}},\ \bibinfo {edition} {2nd}\ ed.\ (\bibinfo  {publisher}
+  {Cambridge University Press, Cambridge},\ \bibinfo {year} {2004})\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Sprott}(2003)}]{Sprott2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~C.}\ \bibnamefont
+  {Sprott}},\ }\href@noop {} {\emph {\bibinfo {title} {Chaos and Time-Series
+  Analysis}}}\ (\bibinfo  {publisher} {Oxford University Press, Oxford},\
+  \bibinfo {year} {2003})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zhang}\ and\ \citenamefont
+  {Small}(2006)}]{ZhangPRL2006}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Zhang}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {{Complex network from
+  pseudoperiodic time series: Topology versus dynamics}},}\ }\href {\doibase
+  10.1103/PhysRevLett.96.238701} {\bibfield  {journal} {\bibinfo  {journal}
+  {{Phys. Rev. Lett.}}\ }\textbf {\bibinfo {volume} {{96}}},\ \bibinfo {pages}
+  {238701} (\bibinfo {year} {{2006}})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Marwan}\ \emph {et~al.}(2009)\citenamefont {Marwan},
+  \citenamefont {Donges}, \citenamefont {Zou}, \citenamefont {Donner},\ and\
+  \citenamefont {Kurths}}]{MarwanPLA2009}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
+  {Marwan}}, \bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}},
+  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Zou}}, \bibinfo {author}
+  {\bibfnamefont {R.~V.}\ \bibnamefont {Donner}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Complex network approach for recurrence analysis of time
+  series},}\ }\href {\doibase https://doi.org/10.1016/j.physleta.2009.09.042}
+  {\bibfield  {journal} {\bibinfo  {journal} {Phys. Lett. A}\ }\textbf
+  {\bibinfo {volume} {373}},\ \bibinfo {pages} {4246 -- 4254} (\bibinfo {year}
+  {2009})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Donner}\ \emph {et~al.}(2010)\citenamefont {Donner},
+  \citenamefont {Zou}, \citenamefont {Donges}, \citenamefont {Marwan},\ and\
+  \citenamefont {Kurths}}]{Donner2010a}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont
+  {Donner}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Zou}},
+  \bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}}, \bibinfo
+  {author} {\bibfnamefont {N.}~\bibnamefont {Marwan}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {{Recurrence networks --- a novel paradigm for nonlinear
+  time series analysis}},}\ }\href {\doibase 10.1088/1367-2630/12/3/033025}
+  {\bibfield  {journal} {\bibinfo  {journal} {New J. Phys.}\ }\textbf {\bibinfo
+  {volume} {12}},\ \bibinfo {pages} {033025} (\bibinfo {year}
+  {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Lacasa}\ \emph {et~al.}(2008)\citenamefont {Lacasa},
+  \citenamefont {Luque}, \citenamefont {Ballesteros}, \citenamefont {Luque},\
+  and\ \citenamefont {Nuno}}]{Lacasa2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Lacasa}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Luque}},
+  \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Ballesteros}}, \bibinfo
+  {author} {\bibfnamefont {J.}~\bibnamefont {Luque}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.~C.}\ \bibnamefont {Nuno}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {{From time series to complex networks: The visibility
+  graph}},}\ }\href {\doibase 10.1073/pnas.0709247105} {\bibfield  {journal}
+  {\bibinfo  {journal} {Proc. Natl. Acad. Sci.}\ }\textbf {\bibinfo {volume}
+  {105}},\ \bibinfo {pages} {4972--4975} (\bibinfo {year} {2008})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Nu{\~{n}}ez}\ \emph {et~al.}(2012)\citenamefont
+  {Nu{\~{n}}ez}, \citenamefont {Lacasa}, \citenamefont {Patricio},\ and\
+  \citenamefont {Luque}}]{Nunez2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Nu{\~{n}}ez}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Lacasa}},
+  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Patricio}}, \ and\
+  \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Luque}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {{Visibility Algorithms: A Short
+  Review}},}\ }in\ \href {\doibase 10.5772/34810} {\emph {\bibinfo {booktitle}
+  {New Front. Graph Theory}}}\ (\bibinfo  {publisher} {InTech},\ \bibinfo
+  {year} {2012})\ Chap.~\bibinfo {chapter} {6}, pp.\ \bibinfo {pages}
+  {119--152}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Nicolis}, \citenamefont {Cantu},\ and\ \citenamefont
+  {Nicolis}(2005)}]{Nicolis2005}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Nicolis}}, \bibinfo {author} {\bibfnamefont {A.~G.}\ \bibnamefont {Cantu}},
+  \ and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Nicolis}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {{Dynamical aspects of
+  interaction networks}},}\ }\href {\doibase 10.1142/S0218127405014167}
+  {\bibfield  {journal} {\bibinfo  {journal} {Int. J. Bifurc. Chaos}\ }\textbf
+  {\bibinfo {volume} {15}},\ \bibinfo {pages} {3467--3480} (\bibinfo {year}
+  {2005})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {McCullough}\ \emph {et~al.}(2015)\citenamefont
+  {McCullough}, \citenamefont {Small}, \citenamefont {Stemler},\ and\
+  \citenamefont {Iu}}]{MichaelChaos2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {McCullough}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},
+  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Stemler}}, \ and\
+  \bibinfo {author} {\bibfnamefont {H.~H.-C.}\ \bibnamefont {Iu}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Time lagged ordinal partition networks
+  for capturing dynamics of continuous dynamical systems},}\ }\href {\doibase
+  10.1063/1.4919075} {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\
+  }\textbf {\bibinfo {volume} {25}},\ \bibinfo {pages} {053101} (\bibinfo
+  {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kulp}\ \emph {et~al.}(2016)\citenamefont {Kulp},
+  \citenamefont {Chobot}, \citenamefont {Freitas},\ and\ \citenamefont
+  {Sprechini}}]{KulpChaos2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.~W.}\ \bibnamefont
+  {Kulp}}, \bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont {Chobot}},
+  \bibinfo {author} {\bibfnamefont {H.~R.}\ \bibnamefont {Freitas}}, \ and\
+  \bibinfo {author} {\bibfnamefont {G.~D.}\ \bibnamefont {Sprechini}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Using ordinal partition
+  transition networks to analyze {ECG} data},}\ }\href {\doibase
+  10.1063/1.4959537} {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\
+  }\textbf {\bibinfo {volume} {26}},\ \bibinfo {pages} {073114} (\bibinfo
+  {year} {2016})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zhang}\ \emph {et~al.}(2017)\citenamefont {Zhang},
+  \citenamefont {Zhou}, \citenamefont {Tang}, \citenamefont {Guo},
+  \citenamefont {Small},\ and\ \citenamefont {Zou}}]{zhangSciRep2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~Y.}\ \bibnamefont
+  {Zhang}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Zhou}},
+  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Tang}}, \bibinfo {author}
+  {\bibfnamefont {H.}~\bibnamefont {Guo}}, \bibinfo {author} {\bibfnamefont
+  {M.}~\bibnamefont {Small}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {Y.}~\bibnamefont {Zou}},\ }\bibfield  {title} {\enquote {\bibinfo {title}
+  {Constructing ordinal partition transition networks from multivariate time
+  series},}\ }\href {\doibase 10.1038/s41598-017-08245-x} {\bibfield  {journal}
+  {\bibinfo  {journal} {Scientific Reports}\ }\textbf {\bibinfo {volume} {7}},\
+  \bibinfo {pages} {7795} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {McCullough}\ \emph {et~al.}(2017)\citenamefont
+  {McCullough}, \citenamefont {Small}, \citenamefont {Iu},\ and\ \citenamefont
+  {Stemler}}]{McCullough2017b}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {McCullough}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},
+  \bibinfo {author} {\bibfnamefont {H.~H.~C.}\ \bibnamefont {Iu}}, \ and\
+  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Stemler}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {{Multiscale ordinal network analysis of
+  human cardiac dynamics}},}\ }\href {\doibase 10.1098/rsta.2016.0292}
+  {\bibfield  {journal} {\bibinfo  {journal} {Philos. Trans. R. Soc. A Math.
+  Phys. Eng. Sci.}\ }\textbf {\bibinfo {volume} {375}},\ \bibinfo {pages}
+  {20160292} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Borges}\ \emph {et~al.}(2019)\citenamefont {Borges},
+  \citenamefont {Ramos}, \citenamefont {Mini}, \citenamefont {Rosso},
+  \citenamefont {Frery},\ and\ \citenamefont {Loureiro}}]{BorgesAMC2019}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~B.}\ \bibnamefont
+  {Borges}}, \bibinfo {author} {\bibfnamefont {H.~S.}\ \bibnamefont {Ramos}},
+  \bibinfo {author} {\bibfnamefont {R.~A.}\ \bibnamefont {Mini}}, \bibinfo
+  {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}}, \bibinfo {author}
+  {\bibfnamefont {A.~C.}\ \bibnamefont {Frery}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {A.~A.~F.}\ \bibnamefont {Loureiro}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Learning and distinguishing time series dynamics
+  via ordinal patterns transition graphs},}\ }\href {\doibase
+  https://doi.org/10.1016/j.amc.2019.06.068} {\bibfield  {journal} {\bibinfo
+  {journal} {Applied Mathematics and Computation}\ }\textbf {\bibinfo {volume}
+  {362}},\ \bibinfo {pages} {124554} (\bibinfo {year} {2019})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Subramaniyam}\ \emph {et~al.}(view)\citenamefont
+  {Subramaniyam}, \citenamefont {Donner}, \citenamefont {Caron}, \citenamefont
+  {Panuccio},\ and\ \citenamefont {Hyttinen}}]{Subramaniyam2020}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~P.}\ \bibnamefont
+  {Subramaniyam}}, \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont
+  {Donner}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Caron}},
+  \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Panuccio}}, \ and\
+  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Hyttinen}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Causal coupling inference from
+  multivariate time series based on ordinal partition transition networks},}\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Nonlinear
+  Dynamics}\ } (\bibinfo {year} {in review})},\ \bibinfo {note} {arXiv preprint
+  2010.00948}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {{Bandt}}\ and\ \citenamefont
+  {{Pompe}}(2002)}]{BandtPRL2002}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {{Bandt}}}\ and\ \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont
+  {{Pompe}}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Permutation
+  entropy: A natural complexity measure for time series},}\ }\href {\doibase
+  10.1103/PhysRevLett.88.174102} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {88}},\ \bibinfo {pages}
+  {174102} (\bibinfo {year} {2002})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Amig{\'o}}, \citenamefont {Keller},\ and\
+  \citenamefont {Unakafova}(2014)}]{AmigoPTRSA2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
+  {Amig{\'o}}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Keller}}, \
+  and\ \bibinfo {author} {\bibfnamefont {V.~A.}\ \bibnamefont {Unakafova}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Ordinal symbolic analysis
+  and its application to biomedical recordings},}\ }\href {\doibase
+  10.1098/rsta.2014.0091} {\bibfield  {journal} {\bibinfo  {journal} {Phil.
+  Trans. R. Soc. A}\ }\textbf {\bibinfo {volume} {373}},\ \bibinfo {pages}
+  {20140091} (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zanin}(2008)}]{ZaninChaos2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Zanin}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Forbidden
+  patterns in financial time series},}\ }\href {\doibase 10.1063/1.2841197}
+  {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\ }\textbf {\bibinfo
+  {volume} {18}},\ \bibinfo {pages} {013119} (\bibinfo {year}
+  {2008})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Barreiro}, \citenamefont {Marti},\ and\ \citenamefont
+  {Masoller}(2011)}]{BarreiroChaos2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Barreiro}}, \bibinfo {author} {\bibfnamefont {A.~C.}\ \bibnamefont {Marti}},
+  \ and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Masoller}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Inferring long memory
+  processes in the climate network via ordinal pattern analysis},}\ }\href
+  {\doibase 10.1063/1.3545273} {\bibfield  {journal} {\bibinfo  {journal}
+  {Chaos}\ }\textbf {\bibinfo {volume} {21}},\ \bibinfo {pages} {013101}
+  (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Takens}(1981)}]{Takens1981}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
+  {Takens}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {{Detecting
+  strange attractors in turbulence}},}\ }in\ \href {\doibase
+  10.1007/BFb0091924} {\emph {\bibinfo {booktitle} {Dynamical Systems and
+  Turbulence, Warwick 1980}}},\ \bibinfo {series} {Lecture Notes in
+  Mathematics}, Vol.\ \bibinfo {volume} {898},\ \bibinfo {editor} {edited by\
+  \bibinfo {editor} {\bibfnamefont {D.}~\bibnamefont {Rand}}\ and\ \bibinfo
+  {editor} {\bibfnamefont {L.-S.}\ \bibnamefont {Young}}}\ (\bibinfo
+  {publisher} {Springer, New York},\ \bibinfo {year} {1981})\ pp.\ \bibinfo
+  {pages} {366--381}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Amig\'o}, \citenamefont {Zambrano},\ and\
+  \citenamefont {Sanju\'an}(2007)}]{AmigoEPL2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
+  {Amig\'o}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Zambrano}}, \
+  and\ \bibinfo {author} {\bibfnamefont {M.~A.~F.}\ \bibnamefont {Sanju\'an}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {True and false forbidden
+  patterns in deterministic and random dynamics},}\ }\href {\doibase
+  https://doi.org/10.1209/0295-5075/79/50001} {\bibfield  {journal} {\bibinfo
+  {journal} {Europhys. Lett.}\ }\textbf {\bibinfo {volume} {79}},\ \bibinfo
+  {pages} {50001} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Amig\'o}, \citenamefont {Zambrano},\ and\
+  \citenamefont {Sanju\'an}(2008)}]{AmigoEPL2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
+  {Amig\'o}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Zambrano}}, \
+  and\ \bibinfo {author} {\bibfnamefont {M.~A.~F.}\ \bibnamefont {Sanju\'an}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Combinatorial detection of
+  determinism in noisy time series},}\ }\href {\doibase
+  https://doi.org/10.1209/0295-5075/83/60005} {\bibfield  {journal} {\bibinfo
+  {journal} {Europhys. Lett.}\ }\textbf {\bibinfo {volume} {83}},\ \bibinfo
+  {pages} {60005} (\bibinfo {year} {2008})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Rosso}\ \emph {et~al.}(2012)\citenamefont {Rosso},
+  \citenamefont {Carpi}, \citenamefont {Saco}, \citenamefont {Ravetti},
+  \citenamefont {Larrondo},\ and\ \citenamefont {Plastino}}]{RossoEPJB2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont
+  {Rosso}}, \bibinfo {author} {\bibfnamefont {L.~C.}\ \bibnamefont {Carpi}},
+  \bibinfo {author} {\bibfnamefont {P.~M.}\ \bibnamefont {Saco}}, \bibinfo
+  {author} {\bibfnamefont {M.~G.}\ \bibnamefont {Ravetti}}, \bibinfo {author}
+  {\bibfnamefont {H.~A.}\ \bibnamefont {Larrondo}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Plastino}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {The {A}mig{\'o} paradigm of forbidden/missing patterns: a
+  detailed analysis},}\ }\href {\doibase 10.1140/epjb/e2012-30307-8} {\bibfield
+   {journal} {\bibinfo  {journal} {Eur. Phys. J. B}\ }\textbf {\bibinfo
+  {volume} {85}},\ \bibinfo {pages} {419} (\bibinfo {year} {2012})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Kowalski}\ \emph {et~al.}(2011)\citenamefont
+  {Kowalski}, \citenamefont {Martin}, \citenamefont {Plastino}, \citenamefont
+  {Rosso},\ and\ \citenamefont {Casas}}]{kowalskiEntropy2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~M.}\ \bibnamefont
+  {Kowalski}}, \bibinfo {author} {\bibfnamefont {M.~T.}\ \bibnamefont
+  {Martin}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}},
+  \bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}}, \ and\
+  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Casas}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Distances in probability space and the
+  statistical complexity setup},}\ }\href {\doibase 10.3390/e13061055}
+  {\bibfield  {journal} {\bibinfo  {journal} {Entropy}\ }\textbf {\bibinfo
+  {volume} {13}},\ \bibinfo {pages} {1055--1075} (\bibinfo {year}
+  {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {L\'opez-Ruiz}, \citenamefont {Mancini},\ and\
+  \citenamefont {Calbet}(1995)}]{LopezPLA1995}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {L\'opez-Ruiz}}, \bibinfo {author} {\bibfnamefont {H.~L.}\ \bibnamefont
+  {Mancini}}, \ and\ \bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
+  {Calbet}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {A statistical
+  measure of complexity},}\ }\href {\doibase
+  https://doi.org/10.1016/0375-9601(95)00867-5} {\bibfield  {journal} {\bibinfo
+   {journal} {Phys. Lett. A}\ }\textbf {\bibinfo {volume} {209}},\ \bibinfo
+  {pages} {321 -- 326} (\bibinfo {year} {1995})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Rosso}\ \emph
+  {et~al.}(2007{\natexlab{a}})\citenamefont {Rosso}, \citenamefont {Larrondo},
+  \citenamefont {Martin}, \citenamefont {Plastino},\ and\ \citenamefont
+  {Fuentes}}]{rossoPRL2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont
+  {Rosso}}, \bibinfo {author} {\bibfnamefont {H.~A.}\ \bibnamefont {Larrondo}},
+  \bibinfo {author} {\bibfnamefont {M.~T.}\ \bibnamefont {Martin}}, \bibinfo
+  {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {M.~A.}\ \bibnamefont {Fuentes}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Distinguishing noise from chaos},}\
+  }\href {\doibase 10.1103/PhysRevLett.99.154102} {\bibfield  {journal}
+  {\bibinfo  {journal} {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {99}},\
+  \bibinfo {pages} {154102} (\bibinfo {year} {2007}{\natexlab{a}})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Lange}, \citenamefont {Rosso},\ and\ \citenamefont
+  {Hauhs}(2013)}]{Lange2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
+  {Lange}}, \bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}}, \
+  and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Hauhs}},\ }\bibfield
+   {title} {\enquote {\bibinfo {title} {Ordinal pattern and statistical
+  complexity analysis of daily stream flow time series},}\ }\href {\doibase
+  10.1140/epjst/e2013-01858-3} {\bibfield  {journal} {\bibinfo  {journal} {Eur.
+  Phys. J. S.T.}\ }\textbf {\bibinfo {volume} {222}},\ \bibinfo {pages}
+  {535--552} (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Martin}, \citenamefont {Plastino},\ and\
+  \citenamefont {Rosso}(2003)}]{MartinPLA2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Martin}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}}, \
+  and\ \bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Statistical complexity and
+  disequilibrium},}\ }\href {\doibase
+  https://doi.org/10.1016/S0375-9601(03)00491-2} {\bibfield  {journal}
+  {\bibinfo  {journal} {Phys. Lett. A}\ }\textbf {\bibinfo {volume} {311}},\
+  \bibinfo {pages} {126 -- 132} (\bibinfo {year} {2003})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {da~F.~Costa}\ \emph {et~al.}(2007)\citenamefont
+  {da~F.~Costa}, \citenamefont {Rodrigues}, \citenamefont {Travieso},\ and\
+  \citenamefont {Boas}}]{CostaADPhy2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {da~F.~Costa}}, \bibinfo {author} {\bibfnamefont {F.~A.}\ \bibnamefont
+  {Rodrigues}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Travieso}},
+  \ and\ \bibinfo {author} {\bibfnamefont {P.~R.~V.}\ \bibnamefont {Boas}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Characterization of complex
+  networks: A survey of measurements},}\ }\href {\doibase
+  10.1080/00018730601170527} {\bibfield  {journal} {\bibinfo  {journal}
+  {Advances in Physics}\ }\textbf {\bibinfo {volume} {56}},\ \bibinfo {pages}
+  {167--242} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Riedl}, \citenamefont {M\"uller},\ and\ \citenamefont
+  {Wessel}(2013)}]{Riedl2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Riedl}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {M\"uller}}, \
+  and\ \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Wessel}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Practical considerations of
+  permutation entropy},}\ }\href {\doibase 10.1140/epjst/e2013-01862-7}
+  {\bibfield  {journal} {\bibinfo  {journal} {Eur. Phys. J. Spec. Top.}\
+  }\textbf {\bibinfo {volume} {222}},\ \bibinfo {pages} {249 -- 262} (\bibinfo
+  {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Rosso}\ \emph
+  {et~al.}(2007{\natexlab{b}})\citenamefont {Rosso}, \citenamefont {Zunino},
+  \citenamefont {P\'erez}, \citenamefont {Figliola}, \citenamefont {Larrondo},
+  \citenamefont {Garavaglia}, \citenamefont {Mart\'{\i}n},\ and\ \citenamefont
+  {Plastino}}]{RossoPRE2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont
+  {Rosso}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Zunino}},
+  \bibinfo {author} {\bibfnamefont {D.~G.}\ \bibnamefont {P\'erez}}, \bibinfo
+  {author} {\bibfnamefont {A.}~\bibnamefont {Figliola}}, \bibinfo {author}
+  {\bibfnamefont {H.~A.}\ \bibnamefont {Larrondo}}, \bibinfo {author}
+  {\bibfnamefont {M.}~\bibnamefont {Garavaglia}}, \bibinfo {author}
+  {\bibfnamefont {M.~T.}\ \bibnamefont {Mart\'{\i}n}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Plastino}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Extracting features of gaussian self-similar stochastic
+  processes via the bandt-pompe approach},}\ }\href {\doibase
+  10.1103/PhysRevE.76.061114} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. E}\ }\textbf {\bibinfo {volume} {76}},\ \bibinfo {pages} {061114}
+  (\bibinfo {year} {2007}{\natexlab{b}})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kowalski}\ \emph {et~al.}(2007)\citenamefont
+  {Kowalski}, \citenamefont {Martín}, \citenamefont {Plastino},\ and\
+  \citenamefont {Rosso}}]{kowalskiPhyD2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Kowalski}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Martín}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}}, \ and\
+  \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Rosso}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Bandt-pompe approach to the
+  classical-quantum transition},}\ }\href {\doibase
+  https://doi.org/10.1016/j.physd.2007.06.015} {\bibfield  {journal} {\bibinfo
+  {journal} {Physica D}\ }\textbf {\bibinfo {volume} {233}},\ \bibinfo {pages}
+  {21 -- 31} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Martin}, \citenamefont {Plastino},\ and\
+  \citenamefont {Rosso}(2006)}]{martinPhyA2006}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Martin}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}}, \
+  and\ \bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Generalized statistical
+  complexity measures: Geometrical and analytical properties},}\ }\href
+  {\doibase https://doi.org/10.1016/j.physa.2005.11.053} {\bibfield  {journal}
+  {\bibinfo  {journal} {Physica A}\ }\textbf {\bibinfo {volume} {369}},\
+  \bibinfo {pages} {439 -- 462} (\bibinfo {year} {2006})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zou}, \citenamefont {Donner},\ and\ \citenamefont
+  {Kurths}(2015)}]{Zou2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Zou}}, \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont {Donner}}, \
+  and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Kurths}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {{Analyzing long-term
+  correlated stochastic processes by means of recurrence networks: Potentials
+  and pitfalls}},}\ }\href {\doibase 10.1103/PhysRevE.91.022926} {\bibfield
+  {journal} {\bibinfo  {journal} {Phys. Rev. E}\ }\textbf {\bibinfo {volume}
+  {91}},\ \bibinfo {pages} {022926} (\bibinfo {year} {2015})},\ \Eprint
+  {http://arxiv.org/abs/1409.3613} {1409.3613} \BibitemShut {NoStop}%
+\bibitem [{\citenamefont {R{\"o}ssler}(1976)}]{Roessler1976}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~E.}\ \bibnamefont
+  {R{\"o}ssler}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {An equation
+  for continuous chaos},}\ }\href {\doibase 10.1016/0375-9601(76)90101-8}
+  {\bibfield  {journal} {\bibinfo  {journal} {Phys. Lett. A}\ }\textbf
+  {\bibinfo {volume} {57}},\ \bibinfo {pages} {397--398} (\bibinfo {year}
+  {1976})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Read}\ \emph {et~al.}(1992)\citenamefont {Read},
+  \citenamefont {Bel}, \citenamefont {Johnson},\ and\ \citenamefont
+  {Small}}]{Read_jfm_1992}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.~L.}\ \bibnamefont
+  {Read}}, \bibinfo {author} {\bibfnamefont {M.~J.}\ \bibnamefont {Bel}},
+  \bibinfo {author} {\bibfnamefont {D.~W.}\ \bibnamefont {Johnson}}, \ and\
+  \bibinfo {author} {\bibfnamefont {R.~M.}\ \bibnamefont {Small}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Quasi-periodic and chaotic flow regimes
+  in a thermally driven, rotating fluid a nnulus},}\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Journal of Fluid Mechanics}\ }\textbf
+  {\bibinfo {volume} {238}},\ \bibinfo {pages} {599--632} (\bibinfo {year}
+  {1992})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zou}\ \emph {et~al.}(2008)\citenamefont {Zou},
+  \citenamefont {Thiel}, \citenamefont {Romano}, \citenamefont {Read},\ and\
+  \citenamefont {Kurths}}]{ZouEPJST2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Zou}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Thiel}}, \bibinfo
+  {author} {\bibfnamefont {M.}~\bibnamefont {Romano}}, \bibinfo {author}
+  {\bibfnamefont {P.}~\bibnamefont {Read}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Recurrence analysis of quasiperiodicity in experimental
+  fluid data},}\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Eur. Phys. J. Spec. Top.}\ }\textbf {\bibinfo {volume} {164}},\ \bibinfo
+  {pages} {23--33} (\bibinfo {year} {2008})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Goldberger}\ \emph {et~al.}(2000)\citenamefont
+  {Goldberger}, \citenamefont {Amaral}, \citenamefont {Glass}, \citenamefont
+  {Hausdorff}, \citenamefont {Ivanov}, \citenamefont {Mark}, \citenamefont
+  {Mietus}, \citenamefont {Moody}, \citenamefont {Peng},\ and\ \citenamefont
+  {Stanley}}]{GoldbergerMITBIH2000}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~L.}\ \bibnamefont
+  {Goldberger}}, \bibinfo {author} {\bibfnamefont {L.~A.~N.}\ \bibnamefont
+  {Amaral}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Glass}},
+  \bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont {Hausdorff}}, \bibinfo
+  {author} {\bibfnamefont {P.~C.}\ \bibnamefont {Ivanov}}, \bibinfo {author}
+  {\bibfnamefont {R.~G.}\ \bibnamefont {Mark}}, \bibinfo {author}
+  {\bibfnamefont {J.~E.}\ \bibnamefont {Mietus}}, \bibinfo {author}
+  {\bibfnamefont {G.~B.}\ \bibnamefont {Moody}}, \bibinfo {author}
+  {\bibfnamefont {C.-K.}\ \bibnamefont {Peng}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {H.~E.}\ \bibnamefont {Stanley}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Physiobank, physiotoolkit, and physionet},}\
+  }\href {\doibase 10.1161/01.CIR.101.23.e215} {\bibfield  {journal} {\bibinfo
+  {journal} {Circulation}\ }\textbf {\bibinfo {volume} {101}},\ \bibinfo
+  {pages} {e215--e220} (\bibinfo {year} {2000})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Small}\ \emph {et~al.}(2002)\citenamefont {Small},
+  \citenamefont {Yu}, \citenamefont {Simonotto}, \citenamefont {Harrison},
+  \citenamefont {Grubb},\ and\ \citenamefont {Fox}}]{smallCSF2002}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Small}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Yu}}, \bibinfo
+  {author} {\bibfnamefont {J.}~\bibnamefont {Simonotto}}, \bibinfo {author}
+  {\bibfnamefont {R.~G.}\ \bibnamefont {Harrison}}, \bibinfo {author}
+  {\bibfnamefont {N.}~\bibnamefont {Grubb}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {K.}~\bibnamefont {Fox}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Uncovering non-linear structure in human ecg
+  recordings},}\ }\href {\doibase
+  https://doi.org/10.1016/S0960-0779(01)00168-0} {\bibfield  {journal}
+  {\bibinfo  {journal} {Chaos, Solitons \& Fractals}\ }\textbf {\bibinfo
+  {volume} {13}},\ \bibinfo {pages} {1755 -- 1762} (\bibinfo {year}
+  {2002})}\BibitemShut {NoStop}%
+\end{thebibliography}%
 

### 2021-01-17 manuscript_Huang.tex
Revision starts...

--- 2020-11-26 manuscript_Huang.tex
@@
 
-It may be noted that in the above discussion on ordinal pattern based time series analysis methods, the transition behavior between patterns has largely remained unconsidered.
-Recently, OPTN representations have been proposed to precisely capture this transition behavior of the ordinal patterns \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}, which opens a broader perspective beyond the standard ordinal symbolic analysis of time series.
+It may be noted that in the above discussion on ordinal pattern based time series analysis methods, the transition behavior between patterns {\color{red} has been discussed as early as in \cite{batesPLA1993}, but since then has largely remained unconsidered systematically from the viewpoint of time series network perspectives.} Recently, OPTN representations have been proposed to precisely capture this transition behavior of the ordinal patterns \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}, which opens a broader perspective beyond the standard ordinal symbolic analysis of time series.
 In this framework, each ordinal pattern is considered as a vertex in a graph, and a directed and weighted edge connecting two patterns in the graph is established according to the corresponding transition frequency (i.e., the probability with which a given permutation is followed by a certain other one).
@@
 From this example, one may easily see that self-loops are related to the serial correlation of the time series under study \cite{BorgesAMC2019}.
-On the other hand, neglecting self-loops in $\mathbf{W}$ emphasizes the transition behavior between different patterns, which has therefore been adopted in most existing studies of complex networks for computational simplicity and theoretical concerns \cite{CostaADPhy2007}.
-Accordingly, we also remove self-loops in the present study before computing $\mathbf{W}$ while acknowledging that in other cases of stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results \cite{BorgesAMC2019}.
+{\color{red} From the viewpoint of statistical measures of complex network theory}, neglecting self-loops in $\mathbf{W}$ emphasizes the transition behavior between different patterns, which has therefore been adopted in most existing studies of complex networks for computational simplicity and theoretical concerns \cite{CostaADPhy2007}.
+Accordingly, we also remove self-loops in the present study before computing $\mathbf{W}$ while acknowledging that in other cases of {\color{red} highly correlated} stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results \cite{BorgesAMC2019}.
 Arguably, whether to remove or to consider self-loops depends on the particular process under study.
-In any of the applications discussed in the rest of this paper, we have ensured that our results do not change qualitatively when including self-loops.
+In any of the applications discussed in the rest of this paper, we have ensured that our results do not change qualitatively when including self-loops{\color{red}, and further numerical results will be provided in Supplementary Materials when necessary.
+}
 
@@
 - 1)$ possible pattern transitions in the uniform distribution function $P_e$.
-Thus, in a full analogy with the permutation entropy based SCM, the Jensen-Shannon divergence is computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$ with
+{\color{red}Note that when self loops are considered in the transition matrix, the normalization factor $\mathcal{S}_{W, max} = \log D!
+\times D!$ is used and there are $D!
+\times D!$ maximal accessible patterns in $P_e$.
+} Thus, in a full analogy with the permutation entropy based SCM, the Jensen-Shannon divergence is computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$ with
 \begin{equation}
@@
 Therefore, the superscript index can be suppressed in the above definition.
-In addition, the node-wise Jensen-Shannon divergence is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$ with the disequilibrium measure
+{\color{red}When self loops are considered in the transition matrix, the normalization factor $\mathcal{S}_{E, max} = \log D!
+$ is used.} In addition, the node-wise Jensen-Shannon divergence is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$ with the disequilibrium measure
 \begin{eqnarray}
@@
 In common applications of ordinal pattern analysis (e.g., the estimation of the permutation entropy $\mathcal{S}_O$), the choice of optimal (or just pragmatic) values of $D$ and $\tau$ depends on the field of application as systematically summarized in Ref.~\cite{Riedl2013}.
+{\color{red}Throughout this work, we have considered $D$ as a kind of free parameter which is varied as $D \in [2, 6]$, although we illustrate the results of $D = 6$ mainly due to the page limit of the paper.
+The choice of $\tau$ originates from the fact that lower entropy values are expected for periodic windows of the logistic map as it has been illustrated in Fig.
+\ref{fig:embed}.
+}
 
@@
 (a) Time series comprising 30 iterations of the period-4 orbit.
-The corresponding ordinal pattern series are shown for different combinations of $D$ and $\tau$: (b) $D = \tau = 3$, (c) $D = \tau = 4$, (d) $D = \tau = 5$, and (e) $D = \tau = 6$.
+The corresponding ordinal pattern series are shown for different combinations of $D$ and $\tau$: (b) $\tau = D = 3$, (c) $\tau = D = 4$, (d) $ \tau = D = 5$, and (e) $\tau = D = 6$.
 The respective permutation entropy values normalized by $\log D!$ as in Eq.~\eqref{eq:Ho} are indicated in the insets of each panel.
@@
 The weights of directed edges are not shown (recall that two slightly different normalizations are considered in this work).
-Note that in panel (g), there exist exclusively self-loops since $D = \tau = 4$ results in a unique, constant pattern.
+Note that in panel (g), there exist exclusively self-loops since $\tau = D = 4$ results in a unique, constant pattern.
 Missing patterns are suppressed.}
@@
 with $r= 3.5$.
-For the sake of simplicity, we choose $D = \tau $.
-In this case, Fig.~\ref{fig:embed} demonstrates that the resulting ordinal pattern sequences are significantly affected by the different parameter settings, which further result in different permutation entropy values $\mathcal{H}_O$.
-For instance, only one unique ordinal pattern is observed when $D = \tau = 4$ (Fig.~\ref{fig:embed}(c,g)) resulting in $\mathcal{H}_O = 0$, which is due to the coverage of exactly one full period by the particular embedding parameters.
+{\color{red}In this example, we propose to choose $\tau = D$, which may largely capture the period-4 and period-3 windows of the logistic map.
+More specifically,} Fig.~\ref{fig:embed} demonstrates that the resulting ordinal pattern sequences are significantly affected by the different parameter settings, which further result in different permutation entropy values $\mathcal{H}_O$.
+For instance, only one unique ordinal pattern is observed when $\tau = D = 4$ (Fig.~\ref{fig:embed}(c, g)) resulting in $\mathcal{H}_O = 0$, which is due to the coverage of exactly one full period by the particular embedding parameters.
 Other values of $D$ and $\tau$ however yield non-zero entropy values.
-The results of Fig.~\ref{fig:embed} may raise concern since different choices of $D$ and $\tau$ will change the placement of the system in the complexity--entropy plane.
-In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for periodic windows when the control parameter $r$ is varied.
-In such cases, we should focus more on the transitions between different dynamical regimes.
-In the following, we will show that the results are robust when the control parameter is systematically changed while keeping the same embedding parameters.
+The results of Fig.~\ref{fig:embed} may raise concern since different choices of $D$ and $\tau$ will change the placement of the system in the complexity--entropy plane{\color{red}, which is an open problem for ordinal pattern analysis \cite{AmigoPTRSA2014,Riedl2013}.} In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for {\color{red}all different} periodic windows when the control parameter $r$ is varied.
+When the control parameter $r$ is changed, we should focus more on the transitions between different dynamical regimes.
+% In the following, we will show that the results are robust when the control parameter is systematically changed while keeping the same embedding parameters.
 
 \subsection{Results for four typical dynamical regimes} \label{sec:four}
-
 We further illustrate the potential of the proposed SCMs to characterize the logistic map with the control parameter $r$ varying in the range $[3.5, 4]$ with a step size of $\Delta r = 0.001$.
@@
 Specifically, chaotic systems present high complexity while stochastic systems have lower values of complexity, hence ideally appearing in distinct regions of the complexity--entropy plane~\cite{rossoPRL2007}.
-Furthermore, for a given entropy value, it has been shown previously that at a given entropy value, the range of possible SCM values is bound by a minimum $\mathcal{C}_{min}$ and a maximum $\mathcal{C}_{max}$.
+Furthermore, {\color{red}it has been shown previously that at a given entropy value,} the range of possible SCM values is bound by a minimum $\mathcal{C}_{min}$ and a maximum $\mathcal{C}_{max}$.
 A general algorithm for computing these bounds has been provided in Ref.~\cite{martinPhyA2006}.
@@
 In our opinion, the most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of possible ambiguity between deterministic-chaotic and stochastic dynamics.
-We therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
+{\color{red}On the other hand, there are many points in the lower left part of the plane showing small values of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$.
+} We therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
 
@@
 For a discussion on practical concerns regarding time delay embedding of fBm and fGn processes, we refer to Ref.~\cite{Zou2015}.
-
 \begin{figure*}
@@
 Especially for the new OPTN based SCMs, we find a partial overlap between the entropy values of fBm and the logistic map, while the associated SCM values are however clearly distinct except for the cases with the lowest entropy, which correspond to the most nonstationary situations ($H$ close to 1).
-Accordingly, we suggest that combining information on the position in the complexity--entropy plane and on the stationarity could allow a clear distinction between chaotic and stochastic dynamics even in those extreme situations.
+{\color{red}Taking into account the effects of self loops in the transition matrix, the results of Fig.
+\ref{fig:CEfBm} do not change significantly although the SCM values extend slightly in the left half of the complexity--entropy plane as $H$ is increased, which are shown in Fig.
+S1 of Supplementary Material (SM).
+Note that the discriminatory power of } Accordingly, we suggest that combining information on the position in the complexity--entropy plane and on the stationarity could allow a clear distinction between chaotic and stochastic dynamics even in those extreme situations.
 Notably, for the stationary fGn, all entropy measures are confined to very high entropy and low complexity values -- a range where no combinations have been found for the new transition frequency based SCMs for the logistic map.
+{\color{red}In the case of fGn, almost the same complexity--entropy plane has been obtained when self loops are included, which has been shown in Fig.
+S3 of SM.
+}
 
@@
 \section{Real-world examples} \label{sec:time}
-
 While we have previously demonstrated the suitability of OPTN based SCMs for tracing changes in dynamical complexity for time series from deterministic dynamical systems, in the following we will use two different sets of experimental time series to demonstrate that SCMs also successfully characterize the complexity properties of real-world systems, showing distinct values for different types of dynamics.
@@
 \begin{enumerate}[(i)]
-\item {Stable fluid:} The temperature signal of the stable flow exhibits periodic oscillations as the wave drift patterns arrive at the fixed point of measurements. 
-\item {Quasi-periodic 2-frequency amplitude vacillation:} This case is identified as a 2-frequency quasi-periodic amplitude flux.
+\item {Stable fluid (SW):} The temperature signal of the stable flow exhibits periodic oscillations as the wave drift patterns arrive at the fixed point of measurements. 
+\item {Quasi-periodic 2-frequency amplitude vacillation (AV-2):} This case is identified as a 2-frequency quasi-periodic amplitude flux.
 The wave drift is composed of slow and regular oscillations in the temperature signal, but a fast modulation in the amplitude is also visible.
 AV-2 is characterized by a periodic growth and drop in wave amplitude with little change in waveform.
-\item {Quasi-periodic 3-frequency amplitude vacillation:} For the sake of completeness, we also analyze a quasi-periodic 3-frequency amplitude vacillation time series.
+\item {Quasi-periodic 3-frequency amplitude vacillation (AV-3):} For the sake of completeness, we also analyze a quasi-periodic 3-frequency amplitude vacillation time series.
 Unlike for the other studied cases, this time series has not been obtained from experimental results, but from a three-dimensional direct numerical simulation of the air-filled rotating baroclinic instability experiment~\cite{Read_jfm_1992}.
-\item {Modulated amplitude vacillation:} This case is identified as a low-dimensional stream, chaotically modulated and with waves of varying amplitudes.
+\item {Modulated amplitude vacillation (MAV):} This case is identified as a low-dimensional stream, chaotically modulated and with waves of varying amplitudes.
 Amplitude modulation results in a complex temperature dynamics.
 \end{enumerate}
+{\color{red}One temperature time series of each flow regime has been measured at $2$s for periods of up to $9 \times 10^4$s.
+Therefore, each recording consists of $N = 45000$ data points.
+The examples of segmented time series have been shown in Fig.
+\ref{tseriesFluid} for the illustration purpose.
+Further details of the lower dimensional chaotic properties have been reported in \cite{Read_jfm_1992,thiel2004a}.
+}
+\begin{figure}
+	\centering 
+	\includegraphics[width=\columnwidth]{TS_fluid.pdf}
+\caption{\small{Sample segments of the temperature recordings from the flow experiment: (a) SW, (b) AV-2, (c) AV-3 and (d) MAV. The vertical axis is temperature $T$. } \label{tseriesFluid}}
+\end{figure}
 
 The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients from MIT-BIH Database and the Creighton University Cardiac Center \cite{GoldbergerMITBIH2000}.
-Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant anomaly that is still relatively poorly understood from a dynamical system perspective, showing rich nonlinear properties.
-In this work, we focus on three different states: normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}.
-From the data base of ECG recordings during various rhythms, we selected 14 time series showing sinus rhythm (SR) prior to onset of arrhythmia, 12 time series of VT and 17 of VF.
-The ECG recordings have been annotated with SR, VT and VF by skilled cardiologists on the basis of waveform morphology.
-Each recording consists of at least $10,000$ data points (20~s) in length at a 250~Hz sampling frequency.
+Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant anomaly that is still relatively poorly understood from a dynamical system perspective, showing rich nonlinear properties \cite{smallCSF2002}.
+{\color{red}The ECG recordings have been measured at a 250~Hz sampling frequency for about 8.5 minutes \cite{GoldbergerMITBIH2000}.
+From these long term recordings, we focus on three different rhythmic states of normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}, which have been annotated by skilled cardiologists on the basis of waveform morphology.
+To reduce the possible finite size effects in calculating SCMs, we choose each time series of at least $10000$ time points.
+In consequence, out of 18 different patients we obtain 14 time series showing sinus rhythm (SR) prior to onset of arrhythmia, 12 records of VT and 17 records of VF.
+After the annotations, each time series may have different length, ranging from $14600$ to $159600$ data points in SR, $16000\sim 132000$ in VT, and $14500 \sim 124000$ in VF.
+We do not apply any preprocessing to the data, except checking the length requirements.
+We show some sections of time series of each case in Fig.
+\ref{tseriesECG}.
+}
+\begin{figure}
+	\centering 
+	\includegraphics[width=\columnwidth]{TS_ECG.pdf}
+\caption{\small{Three representative ECG recording sections (randomly chosen from different rhythmic states): (a) sinus rhythm (SR), (b) VF and VT.
+The horizontal axis of each panel is the datum number (at 250 Hz) and the vertical axis is ECG surface voltage (in milli-volts).
+} \label{tseriesECG}}
+\end{figure}
 
 For choosing the embedding parameters for both sets of example time series, we follow the suggestions from earlier studies~\cite{ZouEPJST2008,smallCSF2002}.
-Although we are facing the inevitable presence of nonstationarity and noise in the experimental data, we have verified that the following results do not change significantly while our embedding parameters are varied.
-In particular, we have treated the embedding dimension as a free parameter which has been varied in the range of $D = 2, 3, \ldots, 7$, while the embedding delay $\tau$ has been kept fixed as follows.
-In the case of temperature time series from the fluid experiments, $\tau_{i} = 218$~s for the stable solution, $\tau_{ii} = 136.5$~s  for AV-2, $\tau_{iii} = 185.625$~s for AV-3, and $\tau_{iv} = 118$~s for MAV.
-For the ECG time series, we consistently use an embedding delay of $\tau = 5$ time steps.
+Although we are facing the inevitable presence of nonstationarity and noise in the experimental data, we have verified that the following results do not change {\color{red}qualitatively} while the embedding parameters are varied.
+In particular, we have treated the embedding dimension as a free parameter which has been varied in the range of $D = 2, 3, \ldots, 6$, while the embedding delay $\tau$ has been {\color{red} determined by the first zero of the autocorrelation function \cite{Kantz97}.
+In the case of temperature time series from the fluid experiments, $\tau_{1} = 110$ time steps (data points) for SW, AV-2 and AV-3, while $\tau_{2} = 60$ for MAV,  which have been consistent with the literature \cite{Read_jfm_1992,thiel2004a}.} For the ECG time series, we consistently use an embedding delay of $\tau = 5$ time steps as suggested in \cite{smallCSF2002}.
 The results have been averaged over all subjects.
 
-In both sets of time series, the OPTN based SCMs enhance our understanding of the underlying nonlinear system, which can be seen from the results summarized in Fig.~\ref{fig:fluid}.
-For the fluid data (Fig.~\ref{fig:fluid}(a)), $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ all show consistent changes of the complexity and entropy values between stable wave solution (SW), quasi-periodic 2-frequency amplitude vacillation (AV-2), quasi-periodic 3-frequency amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV).
+In both sets of time series, the OPTN based SCMs enhance our understanding of the underlying nonlinear system, which can be seen from the results summarized in Figs.
+(\ref{fig:fluid}, \ref{fig:ecg}).
+For the fluid data (Fig.~\ref{fig:fluid}), $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$ all show consistent changes of the complexity and entropy values between stable wave solution (SW), quasi-periodic 2-frequency amplitude vacillation (AV-2), quasi-periodic 3-frequency amplitude vacillation (AV-3) and chaotic modulated amplitude vacillation (MAV).
 Specifically, the complexity of these four cases has the following order: SW has the lowest complexity since it shows periodic oscillations with trivial recurrence, while MAV has the highest complexity since it presents chaotically modulated amplitude vacillation.
@@
 In particular, it becomes hard for $\mathcal{C}_O$ to identify the difference between AV-2 and AV-3.
-
-Qualitatively similar results are obtained for the ECG recordings (Fig.~\ref{fig:fluid}(b)).
-Clearly, different ECG recordings exhibit remarkably distinct complexity values.
-Specifically, SR recordings show very pronounced the highest values of $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$, reflecting the existence of significant spectral power in a rather broad frequency band between 50 and 100 beats/min in the signals \cite{smallCSF2002}.
-On the other hand, the complexity of VT has the smallest value, even lower than for the case of VF.
-Unlike the other five measures, we note, however, that SR shows the smallest value of $\mathcal{C}_O$ while VT and VF have comparable values that are higher than for SR.
-This difference may be explained by the fact that unlike the two other OPTN based SCMs, this measure only takes static information of pattern frequencies into account and may therefore not be able to properly distinguish subtle differences reflected in the temporal order of patterns.
-
 \begin{figure*}
@@
 	\includegraphics[width=2\columnwidth]{fluidExample.pdf}
-\caption{\small{OPTN based entropy and SCM characteristics for experimental fluid data (a) and human ECG recordings (b).
-The results have been averaged over embedding dimensions in the range of $D = 2, 3, \ldots, 7$.
-In (a), stable wave (SW, $\blacklozenge$), quasiperiodic 2-frequency amplitude vacillation (AV-2, $\bullet$), quasiperiodic 3-frequency amplitude vacillation (AV-3, $\blacktriangle$) and chaotic modulated amplitude vacillation (MAV, $\blacksquare$) are distinguished.
-In (b), results are shown for normal sinus rhythm (SR, $\blacktriangle$), ventricular tachycardia (VT, $\blacklozenge$) and ventricular fibrillation (VF, $\bullet$).
+\caption{\small{OPTN based entropy and SCM characteristics for experimental fluid data, which are shown for embedding dimensions of $D = 4, 5, 6$.
+In all subfigures, stable wave (SW, $\blacklozenge$), quasiperiodic 2-frequency amplitude vacillation (AV-2, $\bullet$), quasiperiodic 3-frequency amplitude vacillation (AV-3, $\blacktriangle$) and chaotic modulated amplitude vacillation (MAV, $\blacksquare$) are distinguished.
 } \label{fig:fluid}}
+\end{figure*}
+
+Qualitatively similar results are obtained for the ECG recordings, as shown in Fig.~\ref{fig:ecg} for three different embedding dimensions.
+Generally speaking, different ECG recordings exhibit remarkably distinct complexity values since the ensemble medians are well separated between different time series groups of SR, VF and VT as shown by the box plots in Fig.
+\ref{fig:ecg}.
+More specifically, SR recordings show very pronounced the highest median of $\mathcal{H}_O$, $\mathcal{H}_W$, $\mathcal{C}_W$, $\mathcal{H}_E$ and $\mathcal{C}_E$, reflecting the existence of significant spectral power in a rather broad frequency band between 50 and 100 beats/min in the signals \cite{smallCSF2002}.
+On the other hand, the complexity of VT has the smallest value, even lower than for the case of VF.
+Unlike the other five measures, we note, however, that SR shows the smallest median values of $\mathcal{C}_O$ while VT has the largest values (Fig.
+\ref{fig:ecg}(b)).
+
+ {\color{red}Meanwhile, we find mild dependence of the discriminatory power of $\mathcal{C}_O$ and $\mathcal{C}_W$ on the embedding dimension $D$.
+In particular, in the case of $D=6$ the median values of $\mathcal{C}_O$ of VF are comparable to that of VT (Fig.~\ref{fig:ecg}(b)).
+On the other hand for $D=4$, the median values of $\mathcal{C}_W$ are all comparable for SR, VF and VT (Fig.~\ref{fig:ecg}(d)).
+This insufficiency of $\mathcal{C}_O$ and $\mathcal{C}_W$ due to embedding may be explained by the fact that $\mathcal{C}_O$ takes only static pattern frequencies into account, while $\mathcal{C}_W$ considers dynamic pattern transitions, which may therefore not be able to properly distinguish subtle differences reflected in the temporal order of patterns.
+In contrast, $\mathcal{C}_E$ considers both individual pattern frequencies and dynamic pattern transition frequencies.
+}
+\begin{figure*}
+	\centering 
+	\includegraphics[width=2\columnwidth]{ecgExample.pdf}
+\caption{\small{Box plots for OPTN based entropy and SCM characteristics for human ECG recordings, showing well separated medians for each subject group SR, VF and VT.
+Entropy and SCM scores are shown for embedding dimensions of $D = 4, 5, 6$.
+} \label{fig:ecg}}
 \end{figure*}
@@
 However, further more systematic numerical performance tests employing different types of dynamical systems are required to provide compelling evidence for distinctions between different types of dynamics beyond the cases presented in the current work~\cite{BorgesAMC2019,RossoPRE2007}.
-While this opens several opportunities for follow-up research, we have already demonstrated here the practical usefulness of our improved SCMs by employing this framework to time series of experimental fluid flows and ECG recordings, showing distinctive characteristics for different cases.
+While this opens several opportunities for follow-up research, we have already demonstrated here the practical usefulness of our improved SCMs by employing this framework to time series of experimental fluid flows and ECG recordings, showing {\color{red} statistically significant discrimination} for different cases.
 
 Some open problems also common to other nonlinear time series analysis methods remain to be further addressed, for example, the dependence of reliable estimates of SCMs on the available time series length.
-For time-discrete dynamical systems like chaotic maps, one may easily use longer data series, which however does not seem to improve the estimates very much due to the accumulation of effects due to a limited numerical precision (Fig.~\ref{fig:transient}).
-For continuous systems, together with the embedding parameters, the sampling frequency is another important characteristic which needs to be further explored.
+{\color{red}For reliable estimations of traditional entropy $\mathcal{H}_O$ and SCM $\mathcal{C}_O$, the length $N > 5 D!$ has been largely suggested \cite{BandtPRL2002,AmigoPTRSA2014}.
+However, even longer length is required for the newly proposed SCMs based on transition frequencies because the transition matrix has the dimension $D!\times (D!
+- 1)$ if self loops are excluded.
+This should not be a big problem when working with time series from numerical models provided with the rapid increased computing power nowadays.
+} For time-discrete dynamical systems like chaotic maps, one may easily use longer data series, which however does not seem to improve the estimates very much due to the accumulation of effects due to a limited numerical precision (Fig.~\ref{fig:transient}).
+{\color{red} The dependence of the new SCMs on the length $N$ becomes crucial for experimental time series studies.} For continuous systems, together with the embedding parameters, the sampling frequency is another important characteristic which needs to be further explored.
 Moreover, the generalization of our approach from univariate to bivariate time series analysis shall be further discussed.
 We noticed that this aspect has remained a largely untouched subject for statistical complexity analysis so far, while it presents an interesting task to show the dependence of complexity on interaction.
-We outline further work on the aforementioned topics as subjects of future research.
+{\color{red} We may integrate some of the traditional network topological statistical measures of higher order into the SCMs, for instance, clustering coefficients or path lengths, which have been widely discussed to characterize the dynamical properties of the system generating the time series at hand \cite{ZouPR2018}.} We outline further work on the aforementioned topics as subjects of future research.
 
@@
 Parts of this work have been financially supported by the National Natural Science Foundation of China (grant no.
-11872182 and 11835003).
+11872182,11835003, and 11875132), the Natural Science Foundation of Shanghai (grant no.
+18ZR1411800).
 YZ further acknowledges support by the Shanghai Municipal Science and Technology Major Project (grant no.
@@
 
-\bibliographystyle{unsrt}
+\section*{Data Availability Statement}
+The fluid data that support the findings of this study are available from the corresponding author upon reasonable request.
+The ECG data that support the findings of this study are openly available in {\tt https://www.physionet.org} from MIT-BIH Database and the Creighton University Cardiac Center \cite{GoldbergerMITBIH2000}.
+
 \bibliography{ref_Zou,ref_Donner}

### 2021-01-26 manuscript_Huang.tex
revision almost done!

@@
 We note that several types of chaos--chaos transitions will be discussed in the following: the band merging crisis corresponds to intermittency, the inner crisis to some more subtle chaos-chaos transition and the outer crisis to fully chaotic dynamics.
-
 \begin{table}[htb]
@@
     \hline
-    band merging    & 3.679 & 0.978 & 0.043 & 0.595 & 0.577 & 0.211 & 0.205  \\
+    band merging    & 3.679 & 0.978 & 0.043 & 0.639 & 0.582 & 0.302 & 0.277  \\
     \hline
-    laminar state    & 3.791 & 0.986 & 0.029 & 0.633 & 0.609 & 0.280 & 0.269 \\
+    laminar state    & 3.791 & 0.989 & 0.026 & 0.671 & 0.618 & 0.354 & 0.328 \\
     \hline
-    outer crisis  & 4.0 & 0.994 & 0.014 & 0.677 & 0.638 & 0.359 & 0.339 \\
+    outer crisis  & 4.0 & 1.0 & 0 & 0.761 & 0.662 & 0.522 & 0.458 \\
     \hline
@@
 For the three other cases of band merging, laminar state and outer crisis, the values of all SCMs significantly differ from zero.
-The pattern frequency based SCM $\mathcal{C}_O$ shows smaller values than the other two measures for all regimes, which implies a positioning in the lower right part of the complexity--entropy plane in Fig.~\ref{fig:CElogistic}(a).
+The pattern frequency based SCM $\mathcal{C}_O$ shows zero value and $\mathcal{H}_{O} = 1$, which implies a positioning in the lower right corner of the complexity--entropy plane in Fig.~\ref{fig:CElogistic}(a).
 However, for the transition matrix based SCMs, both the estimated entropies $\mathcal{H}_W$, $\mathcal{H}_E$ and associated complexity measures $\mathcal{C}_W$, $\mathcal{C}_E$ show results that are more consistent with the expectations when considering the respective level of chaoticity of the system in those three regimes (i.e., a rising Lyapunov exponent with increasing $r$ in the respective chaotic regimes).
@@
 \subsection{Complexity--entropy planes} \label{sec:plane}
-
+\subsubsection{Logistic map} 
 In order to put our previous results into a broader context, we next show the complexity--entropy planes obtained when varying the control parameter $r$ in the logistic map.
@@
 For a reliable estimation of SCMs, a pragmatic condition on the length $L$ of the time series to be sufficiently larger than $D!$ \cite{rossoPRL2007,kowalskiPhyD2007} has been satisfied in our setting for $D \leq 7$.
-In the particular example of the logistic map, one may easily generate much longer realizations, which however do not improve the obtained entropy and SCM estimates significantly (as we have tested for $L \leq 5 \times 10^6$, not shown).
+In the particular example of the logistic map, one may easily generate much longer realizations, which however {\color{red} is subjected to finite size effects in experimental time series analysis}.
 
@@
 Furthermore, it is notable that the values of $\mathcal{C}_{max}$ depend on the embedding dimension $D$ since $D$ determines the number of patterns and pattern transitions which are considered in the definition of SCMs.
-By contrast, $\mathcal{C}_{min}$ shows less dependence on $D$.
-
+{\color{red}The curves of $\mathcal{C}_{max}$ for different embeddings have been further annotated in Fig.
+\ref{fig:CElogistic}.
+In contrast, the $\mathcal{C}_{min}$ values are close to each other, showing less dependence on $D$, and therefore they are not annotated.
+}
 \begin{figure*}
@@
 \caption{\small{(Color online) Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
-The dashed lines correspond to the maximum $\mathcal{C}_{max}$ and minimum complexity $\mathcal{C}_{min}$ values in dependence on the respective entropy.
-Panels (a-c) show the results for embedding dimensions $D = 4,5$ and (d-f) for $D = 6$.
+The {\color{red} dotted lines correspond to $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ in dependence on the respective entropy for $D=4$, and the dashed lines are $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ for $D=5$.
+Panels (a-c) show the results for embedding dimensions $D = 4,5$ with annotations for the respective $\mathcal{C}_{max}$ only} and (d-f) for $D = 6$.
 In panels (a, d), the bottom-right rectangles of dotted red lines indicate the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
+However, the lower left rectangles of (b, c, e, f) correspond to regions of lower complexities.
 }  \label{fig:CElogistic}}
@@
 In our opinion, the most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of possible ambiguity between deterministic-chaotic and stochastic dynamics.
-{\color{red}On the other hand, there are many points in the lower left part of the plane showing small values of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$.
-} We therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
 
+{\color{red}On the other hand, there are many values in the lower left corner of the plane showing small values of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ (Fig.
+\ref{fig:CElogistic}(b, c, e, f)).
+This region corresponds to regular orbits of low complexities.
+Along with the choice of embedding parameters, the length of time series and long term accumulations of the computation precisions play important roles in the lower left region of the complexity planes, which will be further illustrated by Figs.
+\ref{fig:bifurcation} and \ref{fig:transient}.
+Out of the left and right corners, complexities of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ show consistent trends along with the degree of chaoticity as will be explained later in Fig.
+\ref{fig:bifurcation}.
+} In general, we therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
+
+{\color{red}In addition, we observe a linear correlation between $\mathcal{C}_W$ and $\mathcal{H}_{W}$ (Fig.
+\ref{fig:CElogistic}(e)), respectively, $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ (Fig.
+\ref{fig:CElogistic}(f)).
+However, we note that this linear correlation appears only for large embedding $D = 6$.
+This linear correlation coincides with the up-rising brach of the maximal complexity curve $\mathcal{C}_{max}$, which is expected by the theory \cite{SMCbook2010}.
+}
+
+{\color{red} @ Reik: I'm not sure whether to include the following discussion in the main text as suggested by the referee.
+The discriminatory power of different SCMs have been further compared in Fig.
+\ref{fig:CHlogCmax}, which characterize the distances to the maximal complexity $\mathcal{C}_{max}$.
+In particular, for a given $\mathcal{H}$ we compute a normalized complexity $\mathcal{C}^{N} = (\mathcal{C} - \mathcal{C}_{min}) / (\mathcal{C}_{max} - \mathcal{C}_{min})$.
+In turn, $\mathcal{C}^{N}$ is bounded in the interval of $[0, 1]$ and the alignment along $\mathcal{C}_{max}$ is reflected by large values of $\mathcal{C}^{N}$.
+The results are shown in Fig.
+\ref{fig:CHlogCmax}.
+The dots larger than 1 are numerical artefacts Fig.
+\ref{fig:CHlogCmax}(c) which are due to  improper choice of embeddings when the Logistic map shows high periodic iterations.
+\begin{figure*}
+        \centering
+        \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_Logistic.pdf}
+\caption{(Color online) Three SCMs which are normalized by $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$  at a given $\mathcal{H}$, for the logistic map.  \label{fig:CHlogCmax}}
+\end{figure*}
+}
+
+\subsubsection{Fractional Brownian motion and fractional Gaussian noise}
 In order to further analyze the capability of our modified complexity--entropy planes to distinguish chaotic from stochastic dynamics, we compare the results for the logistic map with those obtained for fractional Brownian motion (fBm), i.e., stochastic processes with long-range temporal correlations.
@@
 
+{\color{red} For fBM, we can include another figure showing normalized complexity measures by the corresponding $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ as Fig.
+\ref{fig:CHlogCmax}.
+As the embedding dimension $D$ is increased, the dots are more aligned to $\mathcal{C}_{max}$ and extend further to the left half of the complexity--entropy plane.
+\begin{figure*}
+        \centering
+        \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_fBm.pdf}
+\caption{(Color online) Three SCMs which are normalized by $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$  at a given $\mathcal{H}$, for the fractional Brownian motion.  \label{fig:CHlogCmax}}
+\end{figure*}
+}
+
 Comparing the results for both types of stochastic processes with those for the logistic map, a few important observations are made.
 First of all, the traditional pattern frequency based SCM values as originally reported by Rosso and co-workers \cite{RossoPRE2007,rossoPRL2007} are well reproduced as expected (Figs.~\ref{fig:CEfBm}(a, d) and \ref{fig:CEfGn}(a, d)).
-Second, all complexity--entropy pairs of fBm and fGn clearly differ from the theoretical maximum complexity values, which does not change with increasing embedding dimension $D$.
+{\color{red}All three pairs of complexity measures lead to parabola-shaped curves as a function of the respective entropy.
+In addition, the newly proposed complexities of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ have higher values, extending to the left plane of lower values for the entropy, which increases chances to discriminate chaos from stochastic processes in comparison to $\mathcal{C}_O$.
+}
+
+Second, all complexity--entropy pairs of fBm clearly differ from the theoretical maximum complexity values, which does not change with increasing embedding dimension $D$.
 Especially for the new OPTN based SCMs, we find a partial overlap between the entropy values of fBm and the logistic map, while the associated SCM values are however clearly distinct except for the cases with the lowest entropy, which correspond to the most nonstationary situations ($H$ close to 1).
 {\color{red}Taking into account the effects of self loops in the transition matrix, the results of Fig.
-\ref{fig:CEfBm} do not change significantly although the SCM values extend slightly in the left half of the complexity--entropy plane as $H$ is increased, which are shown in Fig.
-S1 of Supplementary Material (SM).
-Note that the discriminatory power of } Accordingly, we suggest that combining information on the position in the complexity--entropy plane and on the stationarity could allow a clear distinction between chaotic and stochastic dynamics even in those extreme situations.
+\ref{fig:CEfBm} do not change significantly although the SCM values extend slightly in the left half of the complexity--entropy plane as $H$ is increased, which are further shown in Fig.
+S1 of Supplementary Material (SM).} Accordingly, we suggest that combining information on the position in the complexity--entropy plane and on the stationarity could allow a clear distinction between chaotic and stochastic dynamics even in those extreme situations.
+
 Notably, for the stationary fGn, all entropy measures are confined to very high entropy and low complexity values -- a range where no combinations have been found for the new transition frequency based SCMs for the logistic map.
-{\color{red}In the case of fGn, almost the same complexity--entropy plane has been obtained when self loops are included, which has been shown in Fig.
+{\color{red}Furthermore, almost the same complexity--entropy plane has been obtained when self loops are included, which has been shown in Fig.
 S3 of SM.
@@
 	\includegraphics[width=2\columnwidth]{logisticEntropy.pdf}
-\caption{\small{Behavior of the different SCMs and associated entropy characteristics for the logistic map in dependence on the control parameter $r$.
-(a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{H}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ based on the globally normalized transition matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g, h) entropy $\mathcal{H}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized transition matrix $\mathbf{W}$.
+\caption{\small{Behavior of the different SCMs and associated entropy characteristics for the logistic map in dependence on the control parameter $r$ {\color{red}(embedding dimension $D = 6$).} (a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{H}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ based on the globally normalized transition matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g, h) entropy $\mathcal{H}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized transition matrix $\mathbf{W}$.
 Several major periodic windows have been highlighted by grey background shading.
@@
 \caption{\small{Same as Fig.
-\ref{fig:bifurcation}, but for Poincar\'e sections of the R\"ossler system while varying the control parameter $b$ (see text for details).
-Panel (a) presents a bifurcation diagram based on the $x$ components of all points in the Poincar\'e sections.
+\ref{fig:bifurcation}, but for Poincar\'e sections of the R\"ossler system while varying the control parameter $b$ (see text for details {\color{red}and embedding dimension $D = 6$).} Panel (a) presents a bifurcation diagram based on the $x$ components of all points in the Poincar\'e sections.
 Panel (b) shows the largest Lyapunov exponent of the system in dependence of $b$.} \label{fig:bifurRossler}}
@@
 } For time-discrete dynamical systems like chaotic maps, one may easily use longer data series, which however does not seem to improve the estimates very much due to the accumulation of effects due to a limited numerical precision (Fig.~\ref{fig:transient}).
-{\color{red} The dependence of the new SCMs on the length $N$ becomes crucial for experimental time series studies.} For continuous systems, together with the embedding parameters, the sampling frequency is another important characteristic which needs to be further explored.
+{\color{red} The dependence of the new SCMs on the length $N$ becomes crucial for experimental time series studies.
+One of more interesting problems is to understand the extra information of transition probability based complexity measures since linear correlations have been found for $\mathcal{C}_{E}$ and small entropy $H$, which are expected for a large number of patterns, i.e., $N = D!$ for high embedding $D$ or equivalently $N = D!
+\times (D!-1)$ transitions.
+} For continuous systems, together with the embedding parameters, the sampling frequency is another important characteristic which needs to be further explored.
 Moreover, the generalization of our approach from univariate to bivariate time series analysis shall be further discussed.
@@
 Parts of this work have been financially supported by the National Natural Science Foundation of China (grant no.
-11872182,11835003, and 11875132), the Natural Science Foundation of Shanghai (grant no.
+11872182, 11835003, 11875132, 61973086), the Natural Science Foundation of Shanghai (grant no.
 18ZR1411800).
-YZ further acknowledges support by the Shanghai Municipal Science and Technology Major Project (grant no.
+JZ and YZ further acknowledge support by the Shanghai Municipal Science and Technology Major Project (grant no.
 2018SHZDZX01) and ZJ Lab.

### 2021-01-26 manuscript_Huang.tex
revision almost done.

@@
         \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_Logistic.pdf}
-\caption{(Color online) Three SCMs which are normalized by $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$  at a given $\mathcal{H}$, for the logistic map.  \label{fig:CHlogCmax}}
+\caption{(Color online) Three SCMs which are normalized by $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$  at a given $\mathcal{H}$, for the logistic map.
+Embedding dimensions are indicated by legends.
+\label{fig:CHlogCmax}}
 \end{figure*}
@@
 
-{\color{red} For fBM, we can include another figure showing normalized complexity measures by the corresponding $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ as Fig.
+{\color{red} For fBM, we show another figure of normalized complexity measures by the corresponding $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ as Fig.
 \ref{fig:CHlogCmax}.

### 2021-01-27 manuscript_Huang.tex
revision is done!

@@
 This region corresponds to regular orbits of low complexities.
-Along with the choice of embedding parameters, the length of time series and long term accumulations of the computation precisions play important roles in the lower left region of the complexity planes, which will be further illustrated by Figs.
-\ref{fig:bifurcation} and \ref{fig:transient}.
-Out of the left and right corners, complexities of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ show consistent trends along with the degree of chaoticity as will be explained later in Fig.
-\ref{fig:bifurcation}.
-} In general, we therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
+Along with the choice of embedding parameters, the length of time series and long term accumulations of the computation precisions play important roles in the lower left region of the complexity planes, which will be further further discussed in Sec.
+\ref{sec:transi}.
+Out of the left and right corners, complexities of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ show consistent trends along with the degree of chaoticity as will be explained later in Sec.
+\ref{sec:transi}.} In general, we therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
 
@@
 \ref{fig:CElogistic}(f)).
-However, we note that this linear correlation appears only for large embedding $D = 6$.
-This linear correlation coincides with the up-rising brach of the maximal complexity curve $\mathcal{C}_{max}$, which is expected by the theory \cite{SMCbook2010}.
-}
+However, we note that this linear correlation appears only for large embedding $D = 6$ and coincides with the up-rising brach of the maximal complexity curve $\mathcal{C}_{max}$, which is expected by the theory \cite{SMCbook2010}.}
 
-{\color{red} @ Reik: I'm not sure whether to include the following discussion in the main text as suggested by the referee.
-The discriminatory power of different SCMs have been further compared in Fig.
-\ref{fig:CHlogCmax}, which characterize the distances to the maximal complexity $\mathcal{C}_{max}$.
+{\color{red} The discriminatory power of different SCMs have been further compared in Fig.
+\ref{fig:CHlogCmax}, which characterize how well the alignments are to the maximal complexity $\mathcal{C}_{max}$.
 In particular, for a given $\mathcal{H}$ we compute a normalized complexity $\mathcal{C}^{N} = (\mathcal{C} - \mathcal{C}_{min}) / (\mathcal{C}_{max} - \mathcal{C}_{min})$.
-In turn, $\mathcal{C}^{N}$ is bounded in the interval of $[0, 1]$ and the alignment along $\mathcal{C}_{max}$ is reflected by large values of $\mathcal{C}^{N}$.
+In consequence, $\mathcal{C}^{N}$ is bounded in the interval of $[0, 1]$ and the alignments along $\mathcal{C}_{max}$ are reflected by large values of $\mathcal{C}^{N}$.
 The results are shown in Fig.
 \ref{fig:CHlogCmax}.
-The dots larger than 1 are numerical artefacts Fig.
-\ref{fig:CHlogCmax}(c) which are due to  improper choice of embeddings when the Logistic map shows high periodic iterations.
+For the traditional $\mathcal{C}_{O}$, again we find there are many points in the right half plane (Fig.
+\ref{fig:CHlogCmax}(a) for all embedding dimensions.
+In contrast,  there are no points in the region of $\mathcal{H} >0.8$ for $\mathcal{C}_W$ and $\mathcal{C}_E$ (Fig.
+\ref{fig:CHlogCmax}(b,c)).
+Note that in Fig.
+\ref{fig:CHlogCmax}(c), the dots larger than 1 are numerical artefacts when the control parameters $r$ are close bifurcation points resulting in long transients (as will be explained by Fig.
+\ref{fig:transient} in Sec.
+\ref{sec:transi}).
 \begin{figure*}
@@
         \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_Logistic.pdf}
-\caption{(Color online) Three SCMs which are normalized by $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$  at a given $\mathcal{H}$, for the logistic map.
+\caption{(Color online) Normalized complexity -- entropy planes of three SCMs for the logistic map, which are normalized by $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$  at a given $\mathcal{H}$.
+(a) $\mathcal{C}_O$, (b) $\mathcal{C}_W$, and (c) $\mathcal{C}_E$.
+In (c), the dots larger than 1 are numerical artefacts when the control parameters $r$ are close bifurcation points resulting in long transients.
 Embedding dimensions are indicated by legends.
@@
 
-{\color{red} For fBM, we show another figure of normalized complexity measures by the corresponding $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ as Fig.
-\ref{fig:CHlogCmax}.
+{\color{red} For fBM, we show the normalized complexity measures which are obtained by the corresponding $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ as Fig.
+\ref{fig:CHfbmCmax}.
 As the embedding dimension $D$ is increased, the dots are more aligned to $\mathcal{C}_{max}$ and extend further to the left half of the complexity--entropy plane.
@@
         \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_fBm.pdf}
-\caption{(Color online) Three SCMs which are normalized by $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$  at a given $\mathcal{H}$, for the fractional Brownian motion.  \label{fig:CHlogCmax}}
+\caption{(Color online) Caption is similar with Fig. \ref{fig:CHlogCmax} for normalized complexity -- entropy planes of three SCMs for the fractional Brownian motion.  \label{fig:CHfbmCmax}}
 \end{figure*}
@@
 For the case of the period doubling bifurcation at $r=3.544$ (replacing a period-4 by a period-8 solution), we show in Fig.~\ref{fig:transient} that the observed jump in the entropy and SCM values may be explained by accumulated numerical errors during the iterations attracted to the periodic-4 points, which yields long transients.
-For estimations based on time series of finite length as used in this work, $\mathcal{H}_E$ and $\mathcal{C}_E$ are however not affected in the same way.
+For estimations based on time series of finite length as used in this work, $\mathcal{H}_E$ and $\mathcal{C}_E$ are however not affected in the same way showing much smaller values (Fig.~\ref{fig:bifurcation}(g, h)).
 Certainly, we should not over-interpret their capabilities since numerical inaccuracy would be accumulated in a longer iterative process such that different ordinal patterns are identified.
@@
 
-Some open problems also common to other nonlinear time series analysis methods remain to be further addressed, for example, the dependence of reliable estimates of SCMs on the available time series length.
+Some open problems also com	mon to other nonlinear time series analysis methods remain to be further addressed, for example, the dependence of reliable estimates of SCMs on the available time series length.
 {\color{red}For reliable estimations of traditional entropy $\mathcal{H}_O$ and SCM $\mathcal{C}_O$, the length $N > 5 D!$ has been largely suggested \cite{BandtPRL2002,AmigoPTRSA2014}.

### 2021-02-10 manuscript_Huang_rev.tex
Reik's changes.

--- 2021-01-27 manuscript_Huang.tex
@@
 Along this line of research, ordinal pattern transition networks have been expanding the established concept of ordinal time series analysis and provide new insights into the dynamical organization underlying time series data that complement existing methods like permutation entropy.
-Permutations based on ordinal patterns are a simple and easy to implement concept that naturally provides statistical complexity measures (SCMs), which in the case of permutation entropy rely on pattern frequencies only.
+Permutation based on ordinal patterns are a simple and easy to implement concept that naturally provides statistical complexity measures (SCMs), which in the case of permutation entropy rely on pattern frequencies only.
 Yet, much additional information can be exploited by including order pattern transition frequencies into the definitions of SCMs -- an idea that however has not been widely developed and applied so far.
@@
 
-The conceptually relatively simple computation of $P$ based on ordinal patterns for the underlying system prompts to another successful application of this framework in terms of statistical complexity measures (SCMs) \cite{kowalskiEntropy2011}.
+The conceptually relatively simple \textcolor{red}{estimation} of $P$ based on ordinal patterns for the underlying system prompts to another successful application of this framework in terms of statistical complexity measures (SCMs) \cite{kowalskiEntropy2011}.
 One common way of defining SCMs is taking the product of a normalized Shannon entropy and an associated measure of disequilibrium \cite{LopezPLA1995,kowalskiEntropy2011}, which captures specific organizational properties of structure and patterns in the observed process.
@@
 Accordingly, we define $Q_{JS}[P, P_e] = Q_{0} \cdot JS[P, P_e]$ where $JS[P, P_e]$ is the Jensen-Shannon divergence between $P$ and $P_e$ and $Q_{0}$ is a normalization factor.
-Furthermore, the complexity--entropy plane $\mathcal{H} \times \mathcal{C}_{JS}$ has been intensively used to distinguish chaotic systems from stochastic ones \cite{rossoPRL2007}.
+Furthermore, the complexity--entropy plane $\mathcal{H} \times \mathcal{C}_{JS}$ has been \textcolor{red}{widely} used to distinguish chaotic systems from stochastic ones \cite{rossoPRL2007}.
 The analysis of complexity--entropy planes has already taken advantage from the ordinal pattern based symbolic encoding of time series and has found various applications \cite{rossoPRL2007,kowalskiEntropy2011}.
@@
 
-It may be noted that in the above discussion on ordinal pattern based time series analysis methods, the transition behavior between patterns {\color{red} has been discussed as early as in \cite{batesPLA1993}, but since then has largely remained unconsidered systematically from the viewpoint of time series network perspectives.} Recently, OPTN representations have been proposed to precisely capture this transition behavior of the ordinal patterns \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}, which opens a broader perspective beyond the standard ordinal symbolic analysis of time series.
+It may be noted that in the \textcolor{red}{context of symbolic dynamics} based time series analysis methods, the transition behavior between patterns {\color{red} has already been discussed as early as in Ref.~\cite{batesPLA1993}, where the authors introduced the concept of fluctuation complexity.
+However, in combination with ordinal patterns, the corresponding idea has long remained not being systematically exploited, especially from a time series network perspective.} Recently, OPTN representations have been proposed to precisely capture this transition behavior of the ordinal patterns \cite{MichaelChaos2015,KulpChaos2016,zhangSciRep2017,McCullough2017b,BorgesAMC2019}, which opens a broader perspective beyond the standard ordinal symbolic analysis of time series.
 In this framework, each ordinal pattern is considered as a vertex in a graph, and a directed and weighted edge connecting two patterns in the graph is established according to the corresponding transition frequency (i.e., the probability with which a given permutation is followed by a certain other one).
@@
 Based on this matrix, we propose in Section~\ref{sec:SCM} to compute SCMs from three different perspectives, including static pattern frequency properties, dynamic pattern transition properties, and a combination of both.
-In Section~\ref{sec:results1}, we then proceed with describing some selected results for the logistic map as a paradigmatic model system exhibiting various types of periodic and chaotic solutions.
-We first discuss in some detail one example of a periodic orbit to emphasize the important effects of embedding parameters in Section~\ref{sec:embeddings}.
+In Section~\ref{sec:results1}, we proceed with describing some selected results for the logistic map as a paradigmatic model system exhibiting various types of periodic and chaotic solutions.
+We first discuss in some detail one example of a periodic orbit to emphasize the important effects of \textcolor{red}{the} embedding \textcolor{red}{dimension} in Section~\ref{sec:embeddings}.
 The practical usefulness of SCMs will be further illustrated by studying four exemplary time series from different typical dynamical regimes (Section~\ref{sec:four}).
@@
 In order to characterize the dynamical transitions in this model, we further study the behavior of the SCM values in dependence on the control parameter in Section~\ref{sec:transi} and show that similar results can also be obtained for time-continuous dynamical systems like the R\"ossler oscillator in Section~\ref{sec:cont}.
-Finally, we show that SCMs successfully distinguish different dynamical regimes in experimental time series from flow data and human ECG measurements in Section~\ref{sec:time}.
+Finally, we \textcolor{red}{demonstrate} that SCMs successfully distinguish different dynamical regimes in experimental time series from flow data and human ECG measurements in Section~\ref{sec:time}.
 Some concluding remarks and discussions will be provided in Section~\ref{sec:con}.
@@
 
-Based on this transition matrix $\mathbf{W}$ of an OPTN, we note that there two slightly different ways to introduce normalization factors.
+Based on this transition matrix $\mathbf{W}$ of an OPTN, we note that there \textcolor{red}{are} two slightly different ways to introduce normalization factors.
 The first option \cite{zhangSciRep2017} is to use a global normalization to ensure that $\sum_{i,j}^{D!} w_{ij} = 1$, which results in a globally normalized transition entropy as discussed below.
-The second option has been proposed by McCullough {\textit{et al.}}~\cite{McCullough2017b} who first considered the local out-link transition frequency from pattern $\pi_i$ to $\pi_j$
+The second option has been proposed by McCullough {\textit{et al.}}~\cite{McCullough2017b} who considered the local out-link transition frequency from pattern $\pi_i$ to $\pi_j$
 \begin{equation} \label{eq:localTp}
@@
 Arguably, whether to remove or to consider self-loops depends on the particular process under study.
-In any of the applications discussed in the rest of this paper, we have ensured that our results do not change qualitatively when including self-loops{\color{red}, and further numerical results will be provided in Supplementary Materials when necessary.
-}
+In any of the applications discussed in the rest of this paper, our results do not change qualitatively when including self-loops{\color{red} (see Supplementary Information Fig.~S1 to S3 for some illustrative examples).}
 
@@
 
-In the following, we briefly review the existing ordinal pattern frequency based SCM before proposing two novel ways to define SCMs, which explicitly consider transition frequencies of the resulting OPTN representations for time series.
+In the following, we briefly review the existing ordinal pattern frequency based SCM before proposing two \textcolor{red}{alternatives} to define SCMs, which explicitly consider transition frequencies of the resulting OPTN representations for time series.
 
@@
 \begin{equation} \label{eq:Q0}
-Q_0 = \frac{-2}{\left(\frac{N+1}{N}\right) \log (N+1) -2 \log (2N) + \log (N)},
+Q_0 = \frac{-2}{\left(\frac{{\color{red}M}+1}{{\color{red}M}}\right) \log ({\color{red}M}+1) -2 \log (2{\color{red}M}) + \log ({\color{red}M})},
 \end{equation}
-where $N = D!$ is the total number of possible patterns. For large $D$, we have $Q_0 \approx 1 /  \log 2$.
+where ${\color{red}M} = D!$ is the total number of possible patterns. For large $D$, we have $Q_0 \approx 1 /  \log 2$.
 
@@
 - 1)$ possible pattern transitions in the uniform distribution function $P_e$.
-{\color{red}Note that when self loops are considered in the transition matrix, the normalization factor $\mathcal{S}_{W, max} = \log D!
-\times D!$ is used and there are $D!
-\times D!$ maximal accessible patterns in $P_e$.
-} Thus, in a full analogy with the permutation entropy based SCM, the Jensen-Shannon divergence is computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$ with
+{\color{red}If self-loops are considered in the transition matrix, the normalization factor $\mathcal{S}_{W, max} = \log D!
+\cdot D!$ is used instead and there are at most $D!^2$ possible pattern transitions in $P_e$.
+}
+
+In full analogy with the permutation entropy based SCM, the Jensen-Shannon divergence is computed as $Q_{JS}[W] = Q_0 \cdot JS[W, P_e]$ with
 \begin{equation}
@@
 \end{equation}
-We note that we have the same expression for the normalization factor $Q_0$ as above when setting $N = D! (D! - 1)$ in Eq.~\eqref{eq:Q0}. This further yields the SCM 
+We note that we have the same expression for the normalization factor $Q_0$ as above when setting ${\color{red}M} = D! (D! - 1)$ in Eq.~\eqref{eq:Q0}. This further yields the SCM 
 \begin{equation}
@@
 Therefore, the superscript index can be suppressed in the above definition.
-{\color{red}When self loops are considered in the transition matrix, the normalization factor $\mathcal{S}_{E, max} = \log D!
-$ is used.} In addition, the node-wise Jensen-Shannon divergence is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$ with the disequilibrium measure
+{\color{red}If self-loops are considered in the transition matrix, the normalization factor $\mathcal{S}_{E, max} = \log D!
+$ is used.}
+
+In addition, the node-wise Jensen-Shannon divergence is defined as $Q_{JS}^{\pi_i}[W_{\pi_i}] = Q_0 \cdot JS^{\pi_i} [W_{\pi_i}, P_e]$ with the disequilibrium measure
 \begin{eqnarray}
@@
 \end{eqnarray}
-In this case, we obtain the normalization factor $Q_0$ by setting $N = D! - 1$ in Eq.~\eqref{eq:Q0}. Taking into account node frequencies, the global Jensen-Shannon divergence is computed as 
+In this case, we obtain the normalization factor $Q_0$ by setting ${\color{red}M} = D!
+- 1$ in Eq.~\eqref{eq:Q0}.
+Taking into account node frequencies, the global Jensen-Shannon divergence is computed as
 \begin{equation}
@@
 
+{\color{red}
+\subsection{Practical considerations}\label{sec:tslength}
+
+One important practical aspect of the proposed methodology concerns the data requirements for computing entropies and complexity measures.
+It is easily understandable that a higher value of the embedding dimension $D$ is generally preferable since it leads to a larger variety of ordinal patterns and, hence, a more robust statistics over the frequencies of such patterns.
+On the other hand, in order to sample all possible patterns sufficiently, this implies the requirement of increasingly longer time series as $D$ rises.
+For a reliable estimation of the classical SCMs from symbolic dynamics that are based on pattern frequencies only, a pragmatic condition suggested in the literature would be considering time series of a length $L$ that is sufficiently larger than $D!$ \cite{rossoPRL2007,kowalskiPhyD2007} (e.g., $N>5D!$).
+
+The situation becomes even worse when our modified SCMs are considered, since they are based on transition frequencies instead of pattern frequencies.
+For $D!$ different ordinal patterns for embedding dimension $D$, there may exist $D!^2$ possible source--destination pairs.
+However, if we consider $\tau=1$ and transitions corresponding to shifting the time index by one time step only, it is clear that from the $D$ components of the source permutation, $D-1$ also contribute to the corresponding destination pattern wíthout changing their mutual order relationships.
+Hence, the new state contributing to the destination pattern can only lead to at most $D$ different order patterns, i.e., we have at most $D\cdot D!$ different pattern transitions.
+However, for other cases like $\tau>1$ and/or larger time shifts between ordinal patterns to be considered, up to $D!^2$ different pairs of patterns and, hence, transitions could occur in the worst case.
+In general, we conclude that the  far higher number of pattern transitions as compared to patterns themselves puts severe restrictions to the choice of $D$ in practical applications of the new SCMs in the context of time series of a given length $L$.
+}
+
 \section{Numerical examples}\label{sec:results1}
 
-\subsection{Choice of embedding parameters} \label{sec:embeddings}
+\subsection{\textcolor{red}{Effect} of embedding parameters} \label{sec:embeddings}
 There are two important algorithmic parameters in constructing OPTNs -- the embedding parameters $D$ and $\tau$.
@@
 In common applications of ordinal pattern analysis (e.g., the estimation of the permutation entropy $\mathcal{S}_O$), the choice of optimal (or just pragmatic) values of $D$ and $\tau$ depends on the field of application as systematically summarized in Ref.~\cite{Riedl2013}.
-{\color{red}Throughout this work, we have considered $D$ as a kind of free parameter which is varied as $D \in [2, 6]$, although we illustrate the results of $D = 6$ mainly due to the page limit of the paper.
-The choice of $\tau$ originates from the fact that lower entropy values are expected for periodic windows of the logistic map as it has been illustrated in Fig.
-\ref{fig:embed}.
-}
+{\color{red}Throughout the remainder of this work, we will consider $D$ as a free parameter, which is varied in the range of $D \in \{2,\ldots,7\}$, although for the sake of brevity, we will mainly show the results for $D = 6$ as illustration.
+The associated choice of the embedding delay $\tau$ is commonly less crucial in nonlinear time series analysis methods.
+In our numerical examples discussed below, the corresponding selection will be mainly guided by the attempt to highlight the fact that lower entropy values are expected for periodic as compared to chaotic behavior.
+In order to cope with a single parameter only, we will particularly use $\tau=D$ for the artificial (deterministic and stochastic) model systems considered in the following, independent of the particular value of $D$.
+We emphasize that this choice is not based on specific considerations of performance optimization, but simply reflects convenience and serves solely illustrative purposes.}
 
@@
 (a) Time series comprising 30 iterations of the period-4 orbit.
-The corresponding ordinal pattern series are shown for different combinations of $D$ and $\tau$: (b) $\tau = D = 3$, (c) $\tau = D = 4$, (d) $ \tau = D = 5$, and (e) $\tau = D = 6$.
+The corresponding ordinal pattern series are shown for different combinations of $D$ and $\tau$ \textcolor{red}{selected for illustrative purposes}: (b) $\tau = D = 3$, (c) $\tau = D = 4$, (d) $ \tau = D = 5$, and (e) $\tau = D = 6$.
 The respective permutation entropy values normalized by $\log D!$ as in Eq.~\eqref{eq:Ho} are indicated in the insets of each panel.
@@
 
-As an illustrative example highlighting the important roles of $D$ and $\tau$ for ordinal pattern based analysis, we consider a period-4 solution of the logistic map 
+As an illustrative example highlighting the important roles of $D$ and $\tau$ for ordinal pattern based \textcolor{red}{time series} analysis, we consider a period-4 solution of the logistic map 
 \begin{equation}
@@
 with $r= 3.5$.
-{\color{red}In this example, we propose to choose $\tau = D$, which may largely capture the period-4 and period-3 windows of the logistic map.
-More specifically,} Fig.~\ref{fig:embed} demonstrates that the resulting ordinal pattern sequences are significantly affected by the different parameter settings, which further result in different permutation entropy values $\mathcal{H}_O$.
+{\color{red}In this example, the choice of $\tau = D$ reflects the particular role of specific choices of the embedding parameters in the context of low-order periodic windows (here, period-4 and period-3) in the bifurcation sequence of the logistic map.
+As expected,} Fig.~\ref{fig:embed} demonstrates that the resulting ordinal pattern sequences are significantly affected by the different parameter settings, which further result in different permutation entropy values $\mathcal{H}_O$.
 For instance, only one unique ordinal pattern is observed when $\tau = D = 4$ (Fig.~\ref{fig:embed}(c, g)) resulting in $\mathcal{H}_O = 0$, which is due to the coverage of exactly one full period by the particular embedding parameters.
 Other values of $D$ and $\tau$ however yield non-zero entropy values.
-The results of Fig.~\ref{fig:embed} may raise concern since different choices of $D$ and $\tau$ will change the placement of the system in the complexity--entropy plane{\color{red}, which is an open problem for ordinal pattern analysis \cite{AmigoPTRSA2014,Riedl2013}.} In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for {\color{red}all different} periodic windows when the control parameter $r$ is varied.
-When the control parameter $r$ is changed, we should focus more on the transitions between different dynamical regimes.
+The results of Fig.~\ref{fig:embed} may raise concern since different choices of $D$ and $\tau$ will change the placement of the system in the complexity--entropy plane.
+%{\color{red}, which is an open problem for ordinal pattern analysis \cite{AmigoPTRSA2014,Riedl2013}.}
+In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for {\color{red}all different} periodic windows when the control parameter $r$ is varied.
+When the control parameter $r$ is changed, we should \textcolor{red}{rather} focus on the transitions between different dynamical regimes.
 % In the following, we will show that the results are robust when the control parameter is systematically changed while keeping the same embedding parameters.
@@
 In all examples, we will consistently use the embedding parameters $D = \tau = 6$.
-In general, we emphasize that a higher value of $D$ leads to a larger variety of ordinal patterns and, hence, a more robust statistics.
+%In general, we emphasize that a higher value of $D$ leads to a larger variety of ordinal patterns and, hence, a more robust statistics.
 On the other hand, for evaluating the statistical properties of either pattern frequencies or transition frequencies, this comes on the cost of requiring increasingly longer time series.
-In the following, we will use realizations of the logistic map comprising $L = 10^6$ iterations (after removing the first $1000$ iterations of each realization which might reflect transient behavior of the system before reaching its respective attractor).
+In the following, we will use realizations of the logistic map comprising $L = 10^6$ iterations (after removing the first $1000$ iterations of each realization which might reflect transient behavior of the system before reaching its respective attractor)\textcolor{red}{, which meets the conditions for the necessary time series length for $D=6$ as discussed in Sect.~\ref{sec:tslength}}.
 
@@
 This however happens only if the embedding parameters are chosen such that they exactly coincide with the periodicity of a certain periodic window (see Fig.~\ref{fig:embed}(g)).
-Another practical aspect concerns the data requirements for computing entropies and complexity measures.
+%Another practical aspect concerns the data requirements for computing entropies and complexity measures.
 For a reliable estimation of SCMs, a pragmatic condition on the length $L$ of the time series to be sufficiently larger than $D!$ \cite{rossoPRL2007,kowalskiPhyD2007} has been satisfied in our setting for $D \leq 7$.

### 2021-02-10 manuscript_Huang_rev.tex
changes.

@@
 
-The situation becomes even worse when our modified SCMs are considered, since they are based on transition frequencies instead of pattern frequencies.
+The situation becomes even more challenging when our modified SCMs are considered, since they are based on transition frequencies instead of pattern frequencies.
 For $D!$ different ordinal patterns for embedding dimension $D$, there may exist $D!^2$ possible source--destination pairs.
-However, if we consider $\tau=1$ and transitions corresponding to shifting the time index by one time step only, it is clear that from the $D$ components of the source permutation, $D-1$ also contribute to the corresponding destination pattern wíthout changing their mutual order relationships.
+However, if we consider $\tau=1$ and transitions corresponding to shifting the time index by one time step only, it is clear that from the $D$ components of the source permutation, $D-1$ also contribute to the corresponding destination pattern without changing their mutual order relationships.
 Hence, the new state contributing to the destination pattern can only lead to at most $D$ different order patterns, i.e., we have at most $D\cdot D!$ different pattern transitions.

### 2021-02-10 manuscript_Huang_rev.tex
Updates from Overleaf

@@
 
-Due to our choice of $\tau$ and $D$, all SCMs take values of zero for the period-3 series because the temporal distance between each pair of components of the embedding vector covers a multiple of two full periods (cf.~Fig.~\ref{fig:embed}).
+Due to our \textcolor{red}{specific} choice of $\tau$ and $D$, all SCMs take values of zero for the period-3 series because the temporal distance between each pair of components of the embedding vector covers a multiple of two full periods (cf.~Fig.~\ref{fig:embed}).
 Other choices of $D$ and $\tau$ would commonly yield non-zero values for the period-3 regime.
@@
 This simple nonlinear system has already been widely discussed in the framework of SCMs \cite{RossoPRE2007,MartinPLA2003}, but only for pattern frequency based complexity measures.
-For each value of $r$, we construct an OPTN from a time series of length $L = 10 ^ 6$ with the embedding parameters $D = \tau \in \{2,\ldots,7\}$.
+For each value of $r$, we \textcolor{red}{again} construct OPTN\textcolor{red}{s} from time series of length $L = 10 ^ 6$ with the embedding parameters $D = \tau \in \{2,\ldots,7\}$.
+{\color{red}As far as the classical ordinal pattern frequency based SCM is concerned, the given time series length is compatible with the chosen range of $D$ values, while the largest embedding dimension of $7$ is at the edge of the range where the ``optimistic'' estimate in Sec.~\ref{sec:tslength} can be considered to apply for transition frequency based SCMs as well.
 
-As already stated above, when only one unique ordinal pattern is identified in a given time series, all SCMs take a value of $0$, which is reasonable since only self-loops are observed.
+In the context of finite sample size effects, we note that for the particular example of the logistic map, one could easily generate realizations of much longer lengths.
+Our numerical results show, however, that longer realizations of a length of up to $L\leq5\cdot 10^6$ do not change the obtained entropy and SCM \emph{estimates} significantly (not shown).
+From a practical perspective, two aspects call for attention in interpreting the obtained entropy and SCM values quantitatively.
+On the one hand, due to the finite numerical precision of the computed trajectories, any chaotic regime may in fact turn out to appear as periodic (with arbitrarily long periods); this can lead to numerical artifacts that will be further discussed in Sec.~\ref{sec:transi}.
+On the other hand, there may exist asymptotic analytical results for the case of $L\to\infty$ and $D\to\infty$, e.g., for the case of $r=4$ where one would expect $\mathcal{H}_O\to 1$ and $\mathcal{C}_O\to 0$, which substantially differ from our numerical results.
+In this regard, all estimated values of the different entropy and complexity measures always need to be considered as finite sample estimates that may well deviate from their expected asymptotic values and thereby depend on the specific values of $L$ and $D$ available for the analysis.
+This is, however, a common feature of other existing symbolic dynamics based entropies and SCMs.
+
+As another introductory side note, we recall that as} already stated above, when only one unique ordinal pattern is identified in a given time series, all SCMs take a value of $0$, which is reasonable since only self-loops are observed.
 This however happens only if the embedding parameters are chosen such that they exactly coincide with the periodicity of a certain periodic window (see Fig.~\ref{fig:embed}(g)).
-%Another practical aspect concerns the data requirements for computing entropies and complexity measures.
-For a reliable estimation of SCMs, a pragmatic condition on the length $L$ of the time series to be sufficiently larger than $D!$ \cite{rossoPRL2007,kowalskiPhyD2007} has been satisfied in our setting for $D \leq 7$.
-In the particular example of the logistic map, one may easily generate much longer realizations, which however {\color{red} is subjected to finite size effects in experimental time series analysis}.
+\textcolor{red}{Hence, it cannot be expected that a given fixed choice of the embedding parameters will be able to highlight all periodic windows of arbitrary order in exactly the same quantitative way.}
 
@@
 Figure~\ref{fig:CElogistic} shows the complexity--entropy planes for embedding parameters $D = \tau$ varied between $4$ and $6$.
-We emphasize that the results are qualitatively similar when $D$ and $\tau$ are varied in the range $\{3,\ldots, 7\}$.
+We emphasize that the results are qualitatively similar \textcolor{red}{for $D=\tau=3,\ 7$ as well (not shown)}.
 Furthermore, it is notable that the values of $\mathcal{C}_{max}$ depend on the embedding dimension $D$ since $D$ determines the number of patterns and pattern transitions which are considered in the definition of SCMs.

### 2021-02-10 manuscript_Huang_rev.tex
Reik's changes.

@@
 Arguably, whether to remove or to consider self-loops depends on the particular process under study.
-In any of the applications discussed in the rest of this paper, our results do not change qualitatively when including self-loops{\color{red} (see Supplementary Information Fig.~S1 to S3 for some illustrative examples).}
+In any of the applications discussed in the rest of this paper, our results do not change qualitatively when including self-loops {\color{red}(see Supplementary Material Sec.~SM1 with Figs.~S1 to S3 for some illustrative examples).}
 
@@
 However, if we consider $\tau=1$ and transitions corresponding to shifting the time index by one time step only, it is clear that from the $D$ components of the source permutation, $D-1$ also contribute to the corresponding destination pattern without changing their mutual order relationships.
-Hence, the new state contributing to the destination pattern can only lead to at most $D$ different order patterns, i.e., we have at most $D\cdot D!$ different pattern transitions.
+Hence, the new state contributing to the destination pattern can only lead to at most $D$ different ordinal patterns, i.e., we have at most $D\cdot D!$ different pattern transitions.
 However, for other cases like $\tau>1$ and/or larger time shifts between ordinal patterns to be considered, up to $D!^2$ different pairs of patterns and, hence, transitions could occur in the worst case.
@@
 In order to cope with a single parameter only, we will particularly use $\tau=D$ for the artificial (deterministic and stochastic) model systems considered in the following, independent of the particular value of $D$.
-We emphasize that this choice is not based on specific considerations of performance optimization, but simply reflects convenience and serves solely illustrative purposes.}
+We emphasize that this choice is not based on specific considerations of performance optimization, but simply reflects convenience and serves solely illustrative purposes.
+Two examples for results obtained with $D=6$ and $\tau=1$ for two well-known model systems can be found in the Supplementary Material Sec.~SM2, Figs.~S4 and S5.}
 
@@
 The results of Fig.~\ref{fig:embed} may raise concern since different choices of $D$ and $\tau$ will change the placement of the system in the complexity--entropy plane.
-%{\color{red}, which is an open problem for ordinal pattern analysis \cite{AmigoPTRSA2014,Riedl2013}.}
 In the special case of the logistic map, we do not have a unique choice of $D$ and $\tau$ for {\color{red}all different} periodic windows when the control parameter $r$ is varied.
 When the control parameter $r$ is changed, we should \textcolor{red}{rather} focus on the transitions between different dynamical regimes.
-% In the following, we will show that the results are robust when the control parameter is systematically changed while keeping the same embedding parameters.
 
@@
 In all examples, we will consistently use the embedding parameters $D = \tau = 6$.
-%In general, we emphasize that a higher value of $D$ leads to a larger variety of ordinal patterns and, hence, a more robust statistics.
-On the other hand, for evaluating the statistical properties of either pattern frequencies or transition frequencies, this comes on the cost of requiring increasingly longer time series.
 In the following, we will use realizations of the logistic map comprising $L = 10^6$ iterations (after removing the first $1000$ iterations of each realization which might reflect transient behavior of the system before reaching its respective attractor)\textcolor{red}{, which meets the conditions for the necessary time series length for $D=6$ as discussed in Sect.~\ref{sec:tslength}}.
@@
 In this regard, all estimated values of the different entropy and complexity measures always need to be considered as finite sample estimates that may well deviate from their expected asymptotic values and thereby depend on the specific values of $L$ and $D$ available for the analysis.
-This is, however, a common feature of other existing symbolic dynamics based entropies and SCMs.
+This is, however, a feature that is also common to other existing symbolic dynamics based entropies and SCMs.
 
@@
 Generalizing this idea to our OPTN based SCMs, we can define three such planes showing the statistical complexity measures ($\mathcal{C}_O$, $\mathcal{C}_W$ and $\mathcal{C}_E$) as a function of the corresponding normalized entropy values ($\mathcal{H}_O$, $\mathcal{H}_W$ and $\mathcal{H}_E$).
-The three SCMs characterize both randomness and correlation structure in a time series, which as a consequence result in a non-trivial dependence on the associated entropy values.
+The three SCMs characterize both randomness and correlation structure in a time series, which as a consequence result\textcolor{red}{s} in a non-trivial dependence on the associated entropy values.
 Specifically, chaotic systems present high complexity while stochastic systems have lower values of complexity, hence ideally appearing in distinct regions of the complexity--entropy plane~\cite{rossoPRL2007}.
@@
 Furthermore, it is notable that the values of $\mathcal{C}_{max}$ depend on the embedding dimension $D$ since $D$ determines the number of patterns and pattern transitions which are considered in the definition of SCMs.
-{\color{red}The curves of $\mathcal{C}_{max}$ for different embeddings have been further annotated in Fig.
-\ref{fig:CElogistic}.
-In contrast, the $\mathcal{C}_{min}$ values are close to each other, showing less dependence on $D$, and therefore they are not annotated.
-}
+{\color{red}Along with the specific results for the logistic map, Fig.~\ref{fig:CElogistic} also shows the theoretical limiting curves of $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$ for three different embedding dimensions ($D=4,\ 5,\ 6$).
+While the lower bounds show only little variation with increasing $D$, the upper limits are more strongly affected and gradually shift upward for larger embedding dimensions.}
+
 \begin{figure*}
@@
 \caption{\small{(Color online) Three different complexity--entropy planes (based on the Jensen-Shannon divergence) illustrating the behavior of the three SCMs for example series of the logistic map with varying control parameter $r$: (a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
-The {\color{red} dotted lines correspond to $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ in dependence on the respective entropy for $D=4$, and the dashed lines are $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ for $D=5$.
-Panels (a-c) show the results for embedding dimensions $D = 4,5$ with annotations for the respective $\mathcal{C}_{max}$ only} and (d-f) for $D = 6$.
-In panels (a, d), the bottom-right rectangles of dotted red lines indicate the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
+The {\color{red} dotted lines correspond to $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ in dependence on the respective entropy for $D=4$ (panels a-c), and the dashed lines indicate $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ for $D=5$ (a-c) and $D=6$ (d-f), respectively.} In panels (a, d), the bottom-right rectangles of dotted red lines indicate the regions for the traditional pattern frequency based SCM analysis where the differentiation between stochastic and deterministic--chaotic dynamics is ambiguous.
 However, the lower left rectangles of (b, c, e, f) correspond to regions of lower complexities.
@@
 
-First of all, we recover the known result that pattern frequency based SCM values lie close to the maximum $\mathcal{C}_{max}$, while the displacements of the estimated values from their theoretical upper bounds depend on the embedding dimension $D$ as shown in Fig.~\ref{fig:CElogistic}(a,d).
-Most notably, we find many combinations of complexity--entropy values in the lower right part of the plane (high entropy, low complexity), which have been highlighted by rectangles of dotted lines.
-This region of the plane is challenging for SCMs to distinguish between chaotic time series and random noise~\cite{BorgesAMC2019} and is in our case occupied by values corresponding to chaotic regimes of the logistic map at large $r$ values.
-The corresponding ambiguity results from the fact that both $\mathcal{H}_O$ and $\mathcal{C}_O$ only take static information in the sense that only pattern frequencies have been considered in their definitions.
+\textcolor{red}{As a first general observation}, we recover the known result that \textcolor{red}{the established} pattern frequency based SCM values \textcolor{red}{for the logistic map} lie close to the \textcolor{red}{theoretical} maximum $\mathcal{C}_{max}$ \textcolor{red}{for a given entropy value} \textcolor{red}{(Fig.~\ref{fig:CElogistic}(a, d))}, while \textcolor{red}{some residual variation of the} displacement of the estimated values from \textcolor{red}{those} upper bounds \textcolor{red}{is retained for all considered values of} the embedding dimension $D$.
+Most notably, we find many combinations of complexity--entropy values in the lower \emph{right} part of the plane (high entropy, low complexity), which \textcolor{red}{is commonly indicative of stochastic processes and has been visually} highlighted by rectangles of dotted lines.
+This region of the plane is challenging for SCMs to distinguish between chaotic time series and random noise~\cite{BorgesAMC2019} and is in our case occupied by values corresponding to chaotic regimes of the logistic map at large $r$ values \textcolor{red}{(see also our previous comment on the expected asymptotic value for $r=4$)}.
+The corresponding ambiguity results from the fact that both $\mathcal{H}_O$ and $\mathcal{C}_O$ only take static information \textcolor{red}{into account} in the sense that only pattern frequencies have been considered in their definitions.
+\textcolor{red}{The observed behavior confirms previous findings that distinguishing (especially high-dimensional) chaotic from stochastic dynamics can be a challenging task in established ordinal pattern based time series analysis methods~\cite{AmigoPTRSA2014,Riedl2013}.}
 
 For our new pattern transition based SCMs, we find that in the complexity--entropy planes $\mathcal{H}_W \times \mathcal{C}_W$ and $\mathcal{H}_E \times \mathcal{C}_E$, the results for the logistic map again closely align with the theoretical upper bounds $\mathcal{C}_{max}$.
-When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely and essentially cover exclusively the left uprising branch of the curve for the SCM based solely on transition frequencies (Figs.~\ref{fig:CElogistic}(b, e) and \ref{fig:CElogistic}(c, f)), which distinctively differs from the original permutation entropy based SCM.
-Specifically, the definition of $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ takes information on both, pattern frequencies and transition frequencies into account, which results in the obtained SCM values aligning even more closely with $\mathcal{C}_{max}$ already at lower embedding dimensions (Fig.~\ref{fig:CElogistic}(c, f)).
-In our opinion, the most interesting result is, however, that we do not find any values in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of possible ambiguity between deterministic-chaotic and stochastic dynamics.
+When increasing the embedding dimension, the estimated values approach the theoretical $\mathcal{C}_{max}$ curve more and more closely\textcolor{red}{, implying that the SCM values are (almost) completely determined by the underlying entropy values as $D$ becomes large.
+For $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ the obtained SCM values align particularly closely with $\mathcal{C}_{max}$ already at low embedding dimensions (Fig.~\ref{fig:CElogistic}(c, f)).
+We hypothesize that this is due to the fact that the definition of these characteristics takes information on both, pattern frequencies and transition frequencies into account, while the other two complexity--entropy pairs focus on only one if those two aspects.}
 
-{\color{red}On the other hand, there are many values in the lower left corner of the plane showing small values of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ (Fig.
-\ref{fig:CElogistic}(b, c, e, f)).
-This region corresponds to regular orbits of low complexities.
-Along with the choice of embedding parameters, the length of time series and long term accumulations of the computation precisions play important roles in the lower left region of the complexity planes, which will be further further discussed in Sec.
-\ref{sec:transi}.
-Out of the left and right corners, complexities of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ show consistent trends along with the degree of chaoticity as will be explained later in Sec.
-\ref{sec:transi}.} In general, we therefore suggest that it is the deterministic transition behavior among patterns that allows an improved distinction from random noise.
+\textcolor{red}{More importantly, for sufficiently large $D$ the resulting complexity--entropy pairs} essentially cover exclusively the left uprising branch of the curve for the SCM \textcolor{red}{accounting for} transition frequencies (Figs.~\ref{fig:CElogistic}(b, e) and \ref{fig:CElogistic}(c, f)), which distinctively differs from the original permutation entropy based SCM.
+\textcolor{red}{In this regard, one may note that this branch implies an almost linear dependency between $\mathcal{C}_W$ and $\mathcal{H}_{W}$ (Fig.~\ref{fig:CElogistic}(e)), respectively, $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ (Fig.~\ref{fig:CElogistic}(f)), which especially applies for small entropy values within an increasingly wide range as $D$ increases.
+This linear increase of the maximum complexity versus entropy curve for sufficiently large embedding dimension can be expected from theoretical considerations~\cite{SMCbook2010}.}
 
-{\color{red}In addition, we observe a linear correlation between $\mathcal{C}_W$ and $\mathcal{H}_{W}$ (Fig.
-\ref{fig:CElogistic}(e)), respectively, $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ (Fig.
-\ref{fig:CElogistic}(f)).
-However, we note that this linear correlation appears only for large embedding $D = 6$ and coincides with the up-rising brach of the maximal complexity curve $\mathcal{C}_{max}$, which is expected by the theory \cite{SMCbook2010}.}
+In our opinion, the most interesting result is, however, that we do not find any \textcolor{red}{complexity--entropy pairs} in the lower right part of the plane, which for the classical SCM based on permutation entropy would present the challenging region of possible ambiguity between deterministic-chaotic and stochastic dynamics.
+{\color{red}By contrast, we find many values in the lower \emph{left} corner of the plane showing small values of both, $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ (Fig.~\ref{fig:CElogistic}(b, c, e, f)).
+This region of the complexity--entropy plane may be indicative of regular (periodic) orbits with low complexity.
+As already emphasized above, along with the choice of embedding parameters, the length of time series and the finite computational precision play important roles in the practical estimation process, possibly mimicking regular behavior in case of actually chaotic dynamics, which could be a reason for the unexpected population of the lower left part of the complexity--entropy planes.
+The mentioned numerical effects will be further discussed in Sec.~\ref{sec:transi}.
+Starting out of both, the lower left and right corners, the transition frequency based SCMs $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ show consistent trends along with the degree of chaoticity, as will also be detailed in Sec.~\ref{sec:transi} below.
+Taken together,} we suggest that the deterministic transition behavior among patterns allows an improved distinction from random noise \textcolor{red}{as compared to pattern frequency based approaches}.
 
-{\color{red} The discriminatory power of different SCMs have been further compared in Fig.
-\ref{fig:CHlogCmax}, which characterize how well the alignments are to the maximal complexity $\mathcal{C}_{max}$.
-In particular, for a given $\mathcal{H}$ we compute a normalized complexity $\mathcal{C}^{N} = (\mathcal{C} - \mathcal{C}_{min}) / (\mathcal{C}_{max} - \mathcal{C}_{min})$.
-In consequence, $\mathcal{C}^{N}$ is bounded in the interval of $[0, 1]$ and the alignments along $\mathcal{C}_{max}$ are reflected by large values of $\mathcal{C}^{N}$.
-The results are shown in Fig.
-\ref{fig:CHlogCmax}.
-For the traditional $\mathcal{C}_{O}$, again we find there are many points in the right half plane (Fig.
-\ref{fig:CHlogCmax}(a) for all embedding dimensions.
-In contrast,  there are no points in the region of $\mathcal{H} >0.8$ for $\mathcal{C}_W$ and $\mathcal{C}_E$ (Fig.
-\ref{fig:CHlogCmax}(b,c)).
-Note that in Fig.
-\ref{fig:CHlogCmax}(c), the dots larger than 1 are numerical artefacts when the control parameters $r$ are close bifurcation points resulting in long transients (as will be explained by Fig.
-\ref{fig:transient} in Sec.
-\ref{sec:transi}).
 \begin{figure*}
@@
         \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_Logistic.pdf}
-\caption{(Color online) Normalized complexity -- entropy planes of three SCMs for the logistic map, which are normalized by $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$  at a given $\mathcal{H}$.
-(a) $\mathcal{C}_O$, (b) $\mathcal{C}_W$, and (c) $\mathcal{C}_E$.
-In (c), the dots larger than 1 are numerical artefacts when the control parameters $r$ are close bifurcation points resulting in long transients.
-Embedding dimensions are indicated by legends.
-\label{fig:CHlogCmax}}
+\caption{\textcolor{red}{(Color online) Normalized complexity -- entropy planes of the three SCMs for the logistic map: (a) $\mathcal{C}_O$, (b) $\mathcal{C}_W$, and (c) $\mathcal{C}_E$.
+In (c), normalized complexity values larger than 1 indicate numerical artifacts when the control parameter $r$ approaches a bifurcation point, resulting in long transients.
+The respective embedding dimensions are indicated by different colors (see legend).
+\label{fig:CHlogCmax}}}
 \end{figure*}
-}
+
+{\color{red}The behavior of the different SCMs relative to their respective lower and upper bounds can be further highlighted by considering normalized quantities.
+Specifically, for a given $\mathcal{H}$ and $D$, we define a normalized complexity $\mathcal{C}^{N} = (\mathcal{C} - \mathcal{C}_{min}) / (\mathcal{C}_{max} - \mathcal{C}_{min})$, which is theoretically bound to the interval $[0,1]$.
+Large values of those normalized complexities close to 1 indicate a good alignment with the maximum complexity curve.
+The corresponding results for the logistic map are shown in Fig.~\ref{fig:CHlogCmax}.
+As expected, for the traditional $\mathcal{C}_{O}$, the behavior is characterized by many points in the right half plane (large entropies, see Fig.~\ref{fig:CHlogCmax}(a)) for all considered embedding dimensions.
+In contrast, there are no points in the range of $\mathcal{H} >0.8$ for $\mathcal{C}_W$ and $\mathcal{C}_E$ (Fig.~\ref{fig:CHlogCmax}(b, c)).
+Note that in Fig.~\ref{fig:CHlogCmax}(c), some normalized complexity values larger than 1 arise at small entropies due to numerical artifacts when the control parameter $r$ is close to a bifurcation point, which results in long transients (as will be explained in Sec.~\ref{sec:transi}).}
 
 \subsubsection{Fractional Brownian motion and fractional Gaussian noise}
-In order to further analyze the capability of our modified complexity--entropy planes to distinguish chaotic from stochastic dynamics, we compare the results for the logistic map with those obtained for fractional Brownian motion (fBm), i.e., stochastic processes with long-range temporal correlations.
+In order to further analyze the capability of our modified complexity--entropy planes to distinguish chaotic from stochastic dynamics, we \textcolor{red}{next} compare the results for the logistic map with those obtained for fractional Brownian motion (fBm), i.e., stochastic processes with long-range temporal correlations.
 Specifically, for an fBm process the long-range correlations of the process are uniquely characterized by the Hurst exponent $H$ when positively correlated (persistence) for $1/2 < H < 1$, while being suppressed (anti-persistence) for $0< H < 1/2$.
@@
 Similar to the previous case of the logistic map with varying control parameter $r$, we generate time series for varying $H \in (0, 1)$ with a step size of $\Delta H=0.05$ and present the resulting complexity--entropy planes in Fig.~\ref{fig:CEfBm}.
+
 Moreover, since fBm series in the persistent regime are nonstationary, we further transform all fBm realizations into stationary time series by employing first-order difference filtering, i.e., by considering the increments $x_{i+1}- x_i$.
@@
 For a discussion on practical concerns regarding time delay embedding of fBm and fGn processes, we refer to Ref.~\cite{Zou2015}.
+
 \begin{figure*}
@@
 
-{\color{red} For fBM, we show the normalized complexity measures which are obtained by the corresponding $\mathcal{C}_{max}$ and $\mathcal{C}_{min}$ as Fig.
-\ref{fig:CHfbmCmax}.
-As the embedding dimension $D$ is increased, the dots are more aligned to $\mathcal{C}_{max}$ and extend further to the left half of the complexity--entropy plane.
+Comparing the results for both types of stochastic processes with those for the logistic map, a few important observations are made.
+First of all, the traditional pattern frequency based SCM values as originally reported by Rosso and co-workers \cite{RossoPRE2007,rossoPRL2007} are well reproduced as expected (Figs.~\ref{fig:CEfBm}(a, d) and \ref{fig:CEfGn}(a, d)).
+Second, \textcolor{red}{unlike for the logistic map} all complexity--entropy pairs of fBm \textcolor{red}{and fGn} clearly differ from the theoretical maximum complexity values, which \textcolor{red}{is a general behavior that} does not change with increasing embedding dimension $D$.
+{\color{red}From a more qualitative perspective, all three complexity--entropy pairs lead to parabola-shaped curves similar to those of the associated lower and upper bounds.
+However, it is notable that the transition based SCMs $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ tend to exhibit higher numerical values than $\mathcal{C}_O$.}
+
+Especially for the new OPTN based SCMs, we find a partial overlap between the entropy values of fBm and the logistic map, while the associated SCM values are however clearly distinct except for the \textcolor{red}{fBm realizations} with the lowest entropy \textcolor{red}{values}, which correspond to the most nonstationary situations ($H$ close to 1).
+\textcolor{red}{In general, we observe for all three complexity--entropy planes of fBm that with increasing $D$, the associated range of entropy values extends further towards the left (lower entropies).
+Taken both, upward and leftward shift of the complexity--entropy curves individually may therefore enhance the risk of confusing deterministic-chaotic and stochastic processes in the complexity--entropy plane even with our alternative SCMs.
+However, our results indicate that} combining information on the position in the complexity--entropy plane and on the stationarity could allow a clear distinction between chaotic and stochastic dynamics even in extreme situations \textcolor{red}{(e.g., a large value of $H$)}.
+Notably, for the stationary fGn, all \textcolor{red}{three complexity--entropy planes} are confined to very high entropy and low complexity values -- a range where no combinations have been found for the new transition frequency based SCMs for the logistic map.
+
+\textcolor{red}{While we have excluded self-loops in the definitions of the considered SCMs so far, including the latter does not modify the results discussed above markedly.
+In the Supplementary Material Sec.~SM1, Figs.~S1-S3, we provide additional illustrations showing the complexity--entropy planes for fBm and fGn for this methodological variant, revealing no qualitative and only minor quantitative changes with respect to the results discussed above.
+This may however differ for strongly correlated deterministic systems, a type of dynamics that we have not yet investigated in the course of the present study.}
+
+{\color{red}Finally, to highlight the difference in complexity values between logistic map and fBm, we show in Fig.~\ref{fig:CHfbmCmax} the behavior of the normalized SCMs with entropy for different embedding dimensions $D$.
+It can be seen that increasing $D$ gradually shifts the normalized complexity values closer to $1$ while at the same time allowing for gradually lower entropy values, which is consistent with the findings in the standard (non-normalized) complexity--entropy planes.
+However, the obtained normalized curves still deviate significantly from those previously found for the logistic map, indicating a good separation between both types of dynamics using each of the three considered SCMs.}
+
 \begin{figure*}
@@
         \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_fBm.pdf}
-\caption{(Color online) Caption is similar with Fig. \ref{fig:CHlogCmax} for normalized complexity -- entropy planes of three SCMs for the fractional Brownian motion.  \label{fig:CHfbmCmax}}
+\caption{\color{red}(Color online) Same as in Fig.~\ref{fig:CHlogCmax}, but for different realizations of fractional Brownian motion with varying Hurst exponent $H$.  \label{fig:CHfbmCmax}}
 \end{figure*}
-}
 
-Comparing the results for both types of stochastic processes with those for the logistic map, a few important observations are made.
-First of all, the traditional pattern frequency based SCM values as originally reported by Rosso and co-workers \cite{RossoPRE2007,rossoPRL2007} are well reproduced as expected (Figs.~\ref{fig:CEfBm}(a, d) and \ref{fig:CEfGn}(a, d)).
-{\color{red}All three pairs of complexity measures lead to parabola-shaped curves as a function of the respective entropy.
-In addition, the newly proposed complexities of $\mathcal{C}_{W}$ and $\mathcal{C}_{E}$ have higher values, extending to the left plane of lower values for the entropy, which increases chances to discriminate chaos from stochastic processes in comparison to $\mathcal{C}_O$.
-}
-
-Second, all complexity--entropy pairs of fBm clearly differ from the theoretical maximum complexity values, which does not change with increasing embedding dimension $D$.
-Especially for the new OPTN based SCMs, we find a partial overlap between the entropy values of fBm and the logistic map, while the associated SCM values are however clearly distinct except for the cases with the lowest entropy, which correspond to the most nonstationary situations ($H$ close to 1).
-{\color{red}Taking into account the effects of self loops in the transition matrix, the results of Fig.
-\ref{fig:CEfBm} do not change significantly although the SCM values extend slightly in the left half of the complexity--entropy plane as $H$ is increased, which are further shown in Fig.
-S1 of Supplementary Material (SM).} Accordingly, we suggest that combining information on the position in the complexity--entropy plane and on the stationarity could allow a clear distinction between chaotic and stochastic dynamics even in those extreme situations.
-
-Notably, for the stationary fGn, all entropy measures are confined to very high entropy and low complexity values -- a range where no combinations have been found for the new transition frequency based SCMs for the logistic map.
-{\color{red}Furthermore, almost the same complexity--entropy plane has been obtained when self loops are included, which has been shown in Fig.
-S3 of SM.
-}
-
-Taken all reported results on the complexity--entropy planes for the logistic map, fBm and fGn together, we conclude that the modified SCMs indeed allow distinguishing chaotic from stochastic dynamics more precisely than the original approach.
+Taken all reported results on the complexity--entropy planes for the logistic map, fBm and fGn together, we conclude that the modified SCMs indeed \textcolor{red}{tend to distinguish} chaotic from stochastic dynamics more precisely than the original approach.
 We therefore suggest that the pattern transition behavior encoded in the OPTNs provides novel insights that can be exploited in terms of SCMs that complement their traditional counterparts.
@@
 	\includegraphics[width=2\columnwidth]{logisticEntropy.pdf}
-\caption{\small{Behavior of the different SCMs and associated entropy characteristics for the logistic map in dependence on the control parameter $r$ {\color{red}(embedding dimension $D = 6$).} (a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{H}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ based on the globally normalized transition matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g, h) entropy $\mathcal{H}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized transition matrix $\mathbf{W}$.
+\caption{\small{Behavior of the different SCMs and associated entropy characteristics for the logistic map in dependence on the control parameter $r$ {\color{red}(embedding dimension $D = 6$, embedding delay $\tau=6$).} (a) Bifurcation diagram, (b) Lyapunov exponent, (c) pattern frequency based (permutation) entropy $\mathcal{H}_O$ and (d) associated complexity measure $\mathcal{C}_O$; (e) pattern transition frequency based entropy $\mathcal{H}_W$ based on the globally normalized transition matrix $\mathbf{W}$ and (f) the corresponding complexity measure $\mathcal{C}_W$, and (g, h) entropy $\mathcal{H}_E$ and complexity measure $\mathcal{C}_E$ based on the node-wise out-link normalized transition matrix $\mathbf{W}$.
 Several major periodic windows have been highlighted by grey background shading.
@@
 
-As a general observation, we find that all three SCMs clearly follow the changes in the bifurcation diagram, for instance, showing distinct values in periodic windows (highlighted by grey color in Fig.~\ref{fig:bifurcation}).
+As a general observation, we find that all three SCMs clearly follow the changes in the bifurcation diagram, for instance, showing distinct values in periodic windows (highlighted by grey color in Fig.~\ref{fig:bifurcation} \textcolor{red}{and Supplementary Material Fig.~S4, where the obtained results are shown for the same embedding dimension $D=6$, but two different values of the embedding delay $\tau$}).
 Other than $\mathcal{C}_O$, the two transition frequency based SCMs behave in a similar way as the Lyapunov exponent.
@@
 It may be noticed that $\mathcal{H}_O$, $\mathcal{C}_O$, $\mathcal{H}_W$ and $\mathcal{C}_W$ all exhibit non-zero values in this parameter range despite a purely periodic dynamics.
-Furthermore, there are jumps of all four measures at the points of period doubling bifurcations, which have been already reported in Ref.~\cite{BandtPRL2002}.
+Furthermore, there are jumps of all four measures at the points of period doubling bifurcations, which have already \textcolor{red}{been} reported in Ref.~\cite{BandtPRL2002}.
 These jumps however are not desirable since the complexity of the dynamics does not change when $r$ passes any of those points.
@@
 Certainly, we should not over-interpret their capabilities since numerical inaccuracy would be accumulated in a longer iterative process such that different ordinal patterns are identified.
-In addition, we pass through many periodic windows of different periods as the control parameter $r$ is varied, which prevents us from using a unique predefined number of iterations as initial transients for time series of different period length.
+In addition, we pass through many periodic windows of different periods as the control parameter $r$ is varied, which prevents us from using a unique predefined number of iterations as initial transients for time series \textcolor{red}{with} different period length.
 Notably, similar jumps have also been observed when the system bifurcates from a period-3 to a period-6 solution at $r = 3.842$.
@@
 To avoid any corresponding discussion on proper choices of further algorithmic parameters, we employ here a Poincar\'e section to each sample trajectory of the system at $y=0$, $\dot{y}<0$ (for different values of $b$) and construct OPTNs from $N = 10,000$ intersection points.
-The results are shown in Fig.~\ref{fig:bifurRossler}.
+The results are shown in Fig.~\ref{fig:bifurRossler} \textcolor{red}{and Supplementary Material Fig.~S5 for the same embedding dimension $D=6$, but two different values of the embedding delay $\tau$}.
 Based on these results, we conclude that the general behavior of the different SCMs and associated entropy measures closely resembles that reported for the logistic map in Fig.~\ref{fig:bifurcation}.
@@
 \caption{\small{Same as Fig.
-\ref{fig:bifurcation}, but for Poincar\'e sections of the R\"ossler system while varying the control parameter $b$ (see text for details {\color{red}and embedding dimension $D = 6$).} Panel (a) presents a bifurcation diagram based on the $x$ components of all points in the Poincar\'e sections.
+\ref{fig:bifurcation}, but for Poincar\'e sections of the R\"ossler system while varying the control parameter $b$ (see text for details{\color{red}, embedding dimension $D = 6$).} Panel (a) presents a bifurcation diagram based on the $x$ components of all points in the Poincar\'e sections.
 Panel (b) shows the largest Lyapunov exponent of the system in dependence of $b$.} \label{fig:bifurRossler}}
@@
 \section{Real-world examples} \label{sec:time}
-While we have previously demonstrated the suitability of OPTN based SCMs for tracing changes in dynamical complexity for time series from deterministic dynamical systems, in the following we will use two different sets of experimental time series to demonstrate that SCMs also successfully characterize the complexity properties of real-world systems, showing distinct values for different types of dynamics.
+While we have previously demonstrated the suitability of OPTN based SCMs for tracing changes in dynamical complexity for time series from deterministic dynamical systems, in the following we will use two different sets of experimental time series to demonstrate that \textcolor{red}{our} SCMs also successfully characterize the complexity properties of real-world systems, showing distinct values for different types of dynamics.
 
 The first example originates from a laboratory fluid experiment of baroclinic instability which is used to study patterns of cyclones and anticyclones in the Earth's atmosphere~\cite{Read_jfm_1992,ZouEPJST2008}.
-Depending on the parameters of rotation rate, temperature difference, viscosity and fluid density, this experimental system exhibits a rich variety of flow regimes.
+Depending on the parameters rotation rate, temperature difference, viscosity and fluid density, this experimental system exhibits a rich variety of flow regimes.
 In particular, we focus on the following regimes:
@@
 \end{enumerate}
-{\color{red}One temperature time series of each flow regime has been measured at $2$s for periods of up to $9 \times 10^4$s.
-Therefore, each recording consists of $N = 45000$ data points.
-The examples of segmented time series have been shown in Fig.
-\ref{tseriesFluid} for the illustration purpose.
-Further details of the lower dimensional chaotic properties have been reported in \cite{Read_jfm_1992,thiel2004a}.
+{\color{red}One temperature time series of each flow regime has been measured at $2$~s resolution for periods of up to $9 \cdot 10^4$s.
+Therefore, each recording consists of $N = 45,000$ data points.
+Example segments of the resulting time series are shown in Fig.~\ref{tseriesFluid} for illustrative purposes.
+Further details on the lower-dimensional chaotic properties have been reported in Refs.~\cite{Read_jfm_1992,thiel2004a}.
 }
@@
 	\includegraphics[width=\columnwidth]{TS_fluid.pdf}
-\caption{\small{Sample segments of the temperature recordings from the flow experiment: (a) SW, (b) AV-2, (c) AV-3 and (d) MAV. The vertical axis is temperature $T$. } \label{tseriesFluid}}
+\caption{\small{Sample segments of the temperature recordings from the flow experiment: (a) SW, (b) AV-2, (c) AV-3 and (d) MAV.
+The vertical axis i\textcolor{red}{describes the} temperature $T$.
+} \label{tseriesFluid}}
 \end{figure}
 
-The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients from MIT-BIH Database and the Creighton University Cardiac Center \cite{GoldbergerMITBIH2000}.
+The second set of example time series comprises physiological signals of human electrocardiogram (ECG) recordings collected from patients from \textcolor{red}{the} MIT-BIH Database and the Creighton University Cardiac Center \cite{GoldbergerMITBIH2000}.
 Heart rate variability suffering from ventricular arrhythmia presents a physiologically highly significant anomaly that is still relatively poorly understood from a dynamical system perspective, showing rich nonlinear properties \cite{smallCSF2002}.
-{\color{red}The ECG recordings have been measured at a 250~Hz sampling frequency for about 8.5 minutes \cite{GoldbergerMITBIH2000}.
-From these long term recordings, we focus on three different rhythmic states of normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}, which have been annotated by skilled cardiologists on the basis of waveform morphology.
-To reduce the possible finite size effects in calculating SCMs, we choose each time series of at least $10000$ time points.
-In consequence, out of 18 different patients we obtain 14 time series showing sinus rhythm (SR) prior to onset of arrhythmia, 12 records of VT and 17 records of VF.
-After the annotations, each time series may have different length, ranging from $14600$ to $159600$ data points in SR, $16000\sim 132000$ in VT, and $14500 \sim 124000$ in VF.
-We do not apply any preprocessing to the data, except checking the length requirements.
-We show some sections of time series of each case in Fig.
+{\color{red}The ECG recordings have been obtained at a 250~Hz sampling frequency for about 8.5~minutes \cite{GoldbergerMITBIH2000}.
+From these long-term recordings, we focus on the three different rhythmic states of normal sinus rhythm (SR), ventricular tachycardia (VT) and ventricular fibrillation (VF) \cite{smallCSF2002}, which have been annotated by skilled cardiologists based on waveform morphology.
+To reduce possible finite-size effects in estimating SCMs, we study time series of at least $10,000$ time points each.
+In consequence, out of 18 different patients, we obtain 14 time series showing sinus rhythm (SR) prior to the onset of arrhythmia, 12 records of VT and 17 records of VF.
+After the annotations, all time series have different lengths, ranging from $14600$ to $159600$ data points in SR, $16000$ to $132000$ in VT, and $14500$ to $124000$ in VF.
+We do not apply any preprocessing to the data, except for checking the length requirements.
+We show some segments of time series of each case in Fig.
 \ref{tseriesECG}.
@@
 \caption{\small{Three representative ECG recording sections (randomly chosen from different rhythmic states): (a) sinus rhythm (SR), (b) VF and VT.
-The horizontal axis of each panel is the datum number (at 250 Hz) and the vertical axis is ECG surface voltage (in milli-volts).
+The horizontal axis of each panel is the datum number (at 250~Hz \textcolor{red}{sampling rate}) and the vertical axis \textcolor{red}{describes the} ECG surface voltage (in \textcolor{red}{mV}).
 } \label{tseriesECG}}
@@
 For choosing the embedding parameters for both sets of example time series, we follow the suggestions from earlier studies~\cite{ZouEPJST2008,smallCSF2002}.
-Although we are facing the inevitable presence of nonstationarity and noise in the experimental data, we have verified that the following results do not change {\color{red}qualitatively} while the embedding parameters are varied.
-In particular, we have treated the embedding dimension as a free parameter which has been varied in the range of $D = 2, 3, \ldots, 6$, while the embedding delay $\tau$ has been {\color{red} determined by the first zero of the autocorrelation function \cite{Kantz97}.
-In the case of temperature time series from the fluid experiments, $\tau_{1} = 110$ time steps (data points) for SW, AV-2 and AV-3, while $\tau_{2} = 60$ for MAV,  which have been consistent with the literature \cite{Read_jfm_1992,thiel2004a}.} For the ECG time series, we consistently use an embedding delay of $\tau = 5$ time steps as suggested in \cite{smallCSF2002}.
-The results have been averaged over all subjects.
+Although we are facing the inevitable presence of nonstationarity and noise in the experimental data, we have verified that the following results do not change {\color{red}qualitatively as} the embedding parameters are varied.
+In particular, we have \textcolor{red}{again} treated the embedding dimension as a free parameter, which has been varied in the range of $D = 2, 3, \ldots, 6$.
+\textcolor{red}{In turn,} the embedding delay $\tau$ has been {\color{red}fixed as determined by the first root of the autocorrelation function \cite{Kantz97}.
+In the case of the temperature time series from the fluid experiments, $\tau_{1} = 110$ time steps (data points) for SW, AV-2 and AV-3, while $\tau_{2} = 60$ for MAV,  which is consistent with the existing literature~\cite{Read_jfm_1992,thiel2004a}.} For the ECG time series, we consistently use an embedding delay of $\tau = 5$ time steps as suggested in \textcolor{red}{Refs.}~\cite{smallCSF2002}.
+The results have been \textcolor{red}{aggregated} over all subjects.
 
@@
 	\includegraphics[width=2\columnwidth]{fluidExample.pdf}
-\caption{\small{OPTN based entropy and SCM characteristics for experimental fluid data, which are shown for embedding dimensions of $D = 4, 5, 6$.
+\caption{\color{red}\small{OPTN based entropy and SCM characteristics for experimental fluid data, which are shown for embedding dimensions of $D = 4, 5, 6$.
 In all subfigures, stable wave (SW, $\blacklozenge$), quasiperiodic 2-frequency amplitude vacillation (AV-2, $\bullet$), quasiperiodic 3-frequency amplitude vacillation (AV-3, $\blacktriangle$) and chaotic modulated amplitude vacillation (MAV, $\blacksquare$) are distinguished.
@@
 
- {\color{red}Meanwhile, we find mild dependence of the discriminatory power of $\mathcal{C}_O$ and $\mathcal{C}_W$ on the embedding dimension $D$.
+ {\color{red}Meanwhile, we find a mild dependence of the discriminatory power of $\mathcal{C}_O$ and $\mathcal{C}_W$ on the embedding dimension $D$.
 In particular, in the case of $D=6$ the median values of $\mathcal{C}_O$ of VF are comparable to that of VT (Fig.~\ref{fig:ecg}(b)).
 On the other hand for $D=4$, the median values of $\mathcal{C}_W$ are all comparable for SR, VF and VT (Fig.~\ref{fig:ecg}(d)).
-This insufficiency of $\mathcal{C}_O$ and $\mathcal{C}_W$ due to embedding may be explained by the fact that $\mathcal{C}_O$ takes only static pattern frequencies into account, while $\mathcal{C}_W$ considers dynamic pattern transitions, which may therefore not be able to properly distinguish subtle differences reflected in the temporal order of patterns.
+This insufficiency of $\mathcal{C}_O$ and $\mathcal{C}_W$ to detect differences in case of improper embedding dimension may be explained by the fact that $\mathcal{C}_O$ takes only static pattern frequencies into account, while $\mathcal{C}_W$ considers dynamic pattern transitions, which may therefore not be able to properly distinguish subtle differences reflected in the temporal order of patterns.
 In contrast, $\mathcal{C}_E$ considers both individual pattern frequencies and dynamic pattern transition frequencies.
@@
 	\includegraphics[width=2\columnwidth]{ecgExample.pdf}
-\caption{\small{Box plots for OPTN based entropy and SCM characteristics for human ECG recordings, showing well separated medians for each subject group SR, VF and VT.
+\caption{\color{red}\small{Box plots for OPTN based entropy and SCM characteristics for human ECG recordings, showing well separated medians for each subject group SR, VF and VT.
 Entropy and SCM scores are shown for embedding dimensions of $D = 4, 5, 6$.
@@
 However, further more systematic numerical performance tests employing different types of dynamical systems are required to provide compelling evidence for distinctions between different types of dynamics beyond the cases presented in the current work~\cite{BorgesAMC2019,RossoPRE2007}.
-While this opens several opportunities for follow-up research, we have already demonstrated here the practical usefulness of our improved SCMs by employing this framework to time series of experimental fluid flows and ECG recordings, showing {\color{red} statistically significant discrimination} for different cases.
+While this opens several opportunities for follow-up research, we have already demonstrated here the practical usefulness of our improved SCMs by employing this framework to time series of experimental fluid flows and ECG recordings, showing {\color{red} statistically significant discriminatory power} for different cases.
 
-Some open problems also com	mon to other nonlinear time series analysis methods remain to be further addressed, for example, the dependence of reliable estimates of SCMs on the available time series length.
-{\color{red}For reliable estimations of traditional entropy $\mathcal{H}_O$ and SCM $\mathcal{C}_O$, the length $N > 5 D!$ has been largely suggested \cite{BandtPRL2002,AmigoPTRSA2014}.
-However, even longer length is required for the newly proposed SCMs based on transition frequencies because the transition matrix has the dimension $D!\times (D!
-- 1)$ if self loops are excluded.
-This should not be a big problem when working with time series from numerical models provided with the rapid increased computing power nowadays.
-} For time-discrete dynamical systems like chaotic maps, one may easily use longer data series, which however does not seem to improve the estimates very much due to the accumulation of effects due to a limited numerical precision (Fig.~\ref{fig:transient}).
-{\color{red} The dependence of the new SCMs on the length $N$ becomes crucial for experimental time series studies.
-One of more interesting problems is to understand the extra information of transition probability based complexity measures since linear correlations have been found for $\mathcal{C}_{E}$ and small entropy $H$, which are expected for a large number of patterns, i.e., $N = D!$ for high embedding $D$ or equivalently $N = D!
-\times (D!-1)$ transitions.
-} For continuous systems, together with the embedding parameters, the sampling frequency is another important characteristic which needs to be further explored.
+Some open problems also common to other nonlinear time series analysis methods remain to be further addressed, for example, the dependence of reliable estimates of SCMs on the available time series length.
+{\color{red}For reliable estimations of traditional entropy $\mathcal{H}_O$ and SCM $\mathcal{C}_O$, a minimum time series length of $L > 5 D!$ has been largely suggested \cite{BandtPRL2002,AmigoPTRSA2014}.
+However, even much longer lengths are required for the newly proposed SCMs based on transition frequencies, since the transition matrix has the dimensions $D!\cdot (D!
+- 1)$ if self-loops are excluded.
+This should however not present a problematic issue when working with time series from numerical models provided with the rapidly increasing computing power nowadays.} For time-discrete dynamical systems like chaotic maps, one may easily \textcolor{red}{generate} longer data series, which however does not \textcolor{red}{necessarily} improve the estimates very much due to the accumulation of effects \textcolor{red}{resulting from} a limited numerical precision (Fig.~\ref{fig:transient}).
+{\color{red}The dependence of the new SCMs on the time series length $L$ becomes crucial for experimental time series studies.
+One interesting problem in this context is to understand the extra information of transition probability based complexity measures since strong linear correlations have been found between complexity and entropy at small entropy values, which are expected for a large number of patterns (i.e., $M = D!$) for high embedding dimension $D$ (or, in a similar spirit, $M = D!
+\cdot (D!-1)$ transitions).} For continuous systems, together with the embedding parameters, the sampling frequency is another important characteristic which needs to be further explored.
 Moreover, the generalization of our approach from univariate to bivariate time series analysis shall be further discussed.
 We noticed that this aspect has remained a largely untouched subject for statistical complexity analysis so far, while it presents an interesting task to show the dependence of complexity on interaction.
-{\color{red} We may integrate some of the traditional network topological statistical measures of higher order into the SCMs, for instance, clustering coefficients or path lengths, which have been widely discussed to characterize the dynamical properties of the system generating the time series at hand \cite{ZouPR2018}.} We outline further work on the aforementioned topics as subjects of future research.
+{\color{red}Finally, it appears an interesting aspect to integrate some of the established higher-order network topological characteristics into the definition of additional SCMs, for instance, clustering coefficients or average path lengths, which have been widely discussed to characterize the dynamical properties of the system generating the time series at hand \cite{ZouPR2018}.} We outline further work on the aforementioned topics as subjects of future research.
 
@@
 A.
-Rosso for sharing algorithms for computing $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$.
+Rosso for sharing algorithms for computing $\mathcal{C}_{min}$ and $\mathcal{C}_{max}$\textcolor{red}{, and two anonymous reviewers for helpful comments on the original version of this manuscript}.
 

### 2021-02-11 manuscript_Huang_rev.tex
files updated.

@@
 {\color{red} From the viewpoint of statistical measures of complex network theory}, neglecting self-loops in $\mathbf{W}$ emphasizes the transition behavior between different patterns, which has therefore been adopted in most existing studies of complex networks for computational simplicity and theoretical concerns \cite{CostaADPhy2007}.
-Accordingly, we also remove self-loops in the present study before computing $\mathbf{W}$ while acknowledging that in other cases of {\color{red} highly correlated} stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results \cite{BorgesAMC2019}.
+Accordingly, we also remove self-loops in the present study before computing $\mathbf{W}$ while acknowledging that in other cases of {\color{red} highly correlated} stochastic processes and discrete systems, self-loops have been included in the analysis showing some interesting results \cite{BorgesAMC2019,PessaPRE2019}.
 Arguably, whether to remove or to consider self-loops depends on the particular process under study.
@@
 For $D!$ different ordinal patterns for embedding dimension $D$, there may exist $D!^2$ possible source--destination pairs.
-However, if we consider $\tau=1$ and transitions corresponding to shifting the time index by one time step only, it is clear that from the $D$ components of the source permutation, $D-1$ also contribute to the corresponding destination pattern without changing their mutual order relationships.
+However, if we consider $\tau=1$ and transitions corresponding to shifting the time index by one time step only, it is clear that from the $D$ components of the source permutation, $D-1$ also contribute to the corresponding destination pattern without changing their mutual order relationships \cite{BorgesAMC2019,PessaPRE2019}.
 Hence, the new state contributing to the destination pattern can only lead to at most $D$ different ordinal patterns, i.e., we have at most $D\cdot D!$ different pattern transitions.
@@
 \textcolor{red}{In this regard, one may note that this branch implies an almost linear dependency between $\mathcal{C}_W$ and $\mathcal{H}_{W}$ (Fig.~\ref{fig:CElogistic}(e)), respectively, $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ (Fig.~\ref{fig:CElogistic}(f)), which especially applies for small entropy values within an increasingly wide range as $D$ increases.
-This linear increase of the maximum complexity versus entropy curve for sufficiently large embedding dimension can be expected from theoretical considerations~\cite{SMCbook2010}.}
+This linear increase of the maximum complexity versus entropy curve for sufficiently large embedding dimension is expected from theoretical considerations~\cite{SMCbook2010}.}
 
@@
 \caption{\small{Sample segments of the temperature recordings from the flow experiment: (a) SW, (b) AV-2, (c) AV-3 and (d) MAV.
-The vertical axis i\textcolor{red}{describes the} temperature $T$.
+The vertical axis \textcolor{red}{describes the} temperature $T$.
 } \label{tseriesFluid}}

### 2021-02-12 manuscript_Huang_rev.tex
Reik's changes.

@@
 On the one hand, due to the finite numerical precision of the computed trajectories, any chaotic regime may in fact turn out to appear as periodic (with arbitrarily long periods); this can lead to numerical artifacts that will be further discussed in Sec.~\ref{sec:transi}.
-On the other hand, there may exist asymptotic analytical results for the case of $L\to\infty$ and $D\to\infty$, e.g., for the case of $r=4$ where one would expect $\mathcal{H}_O\to 1$ and $\mathcal{C}_O\to 0$, which substantially differ from our numerical results.
-In this regard, all estimated values of the different entropy and complexity measures always need to be considered as finite sample estimates that may well deviate from their expected asymptotic values and thereby depend on the specific values of $L$ and $D$ available for the analysis.
+On the other hand,
+%there may exist asymptotic analytical results for the case of $L\to\infty$ and $D\to\infty$, e.g., for the case of $r=4$ where one would expect $\mathcal{H}_O\to 1$ and $\mathcal{C}_O\to 0$, which substantially differ from our numerical results.
+In this regard,
+all estimated values of the different entropy and complexity measures always need to be considered as finite sample estimates that may well deviate from their expected asymptotic values and thereby depend on the specific values of $L$ and $D$ available for the analysis.
 This is, however, a feature that is also common to other existing symbolic dynamics based entropies and SCMs.
@@
 Most notably, we find many combinations of complexity--entropy values in the lower \emph{right} part of the plane (high entropy, low complexity), which \textcolor{red}{is commonly indicative of stochastic processes and has been visually} highlighted by rectangles of dotted lines.
-This region of the plane is challenging for SCMs to distinguish between chaotic time series and random noise~\cite{BorgesAMC2019} and is in our case occupied by values corresponding to chaotic regimes of the logistic map at large $r$ values \textcolor{red}{(see also our previous comment on the expected asymptotic value for $r=4$)}.
+This region of the plane is challenging for SCMs to distinguish between chaotic time series and random noise~\cite{BorgesAMC2019} and is in our case occupied by values corresponding to chaotic regimes of the logistic map at large $r$ values.
 The corresponding ambiguity results from the fact that both $\mathcal{H}_O$ and $\mathcal{C}_O$ only take static information \textcolor{red}{into account} in the sense that only pattern frequencies have been considered in their definitions.

### 2021-02-12 manuscript_Huang_rev.tex
Updated.

@@
 On the one hand, due to the finite numerical precision of the computed trajectories, any chaotic regime may in fact turn out to appear as periodic (with arbitrarily long periods); this can lead to numerical artifacts that will be further discussed in Sec.~\ref{sec:transi}.
-On the other hand,
-%there may exist asymptotic analytical results for the case of $L\to\infty$ and $D\to\infty$, e.g., for the case of $r=4$ where one would expect $\mathcal{H}_O\to 1$ and $\mathcal{C}_O\to 0$, which substantially differ from our numerical results.
-In this regard,
-all estimated values of the different entropy and complexity measures always need to be considered as finite sample estimates that may well deviate from their expected asymptotic values and thereby depend on the specific values of $L$ and $D$ available for the analysis.
+On the other hand, all estimated values of the different entropy and complexity measures always need to be considered as finite sample estimates that may well deviate from their expected asymptotic values and thereby depend on the specific values of $L$ and $D$ available for the analysis.
 This is, however, a feature that is also common to other existing symbolic dynamics based entropies and SCMs.
@@
 \textcolor{red}{In this regard, one may note that this branch implies an almost linear dependency between $\mathcal{C}_W$ and $\mathcal{H}_{W}$ (Fig.~\ref{fig:CElogistic}(e)), respectively, $\mathcal{C}_{E}$ and $\mathcal{H}_{E}$ (Fig.~\ref{fig:CElogistic}(f)), which especially applies for small entropy values within an increasingly wide range as $D$ increases.
-This linear increase of the maximum complexity versus entropy curve for sufficiently large embedding dimension is expected from theoretical considerations~\cite{SMCbook2010}.}
+This linear increase of the maximum complexity versus entropy curve for sufficiently large embedding dimension is expected from theoretical considerations~\cite{SMCbook2010}.
+Some more discussions have been provided in the Supplementary Material (Sec.
+SM3, Fig.
+S6-S8).
+}
 

### 2021-02-12 manuscript_Huang_rev.tex
Reik's final changes.

@@
 For $D!$ different ordinal patterns for embedding dimension $D$, there may exist $D!^2$ possible source--destination pairs.
-However, if we consider $\tau=1$ and transitions corresponding to shifting the time index by one time step only, it is clear that from the $D$ components of the source permutation, $D-1$ also contribute to the corresponding destination pattern without changing their mutual order relationships \cite{BorgesAMC2019,PessaPRE2019}.
-Hence, the new state contributing to the destination pattern can only lead to at most $D$ different ordinal patterns, i.e., we have at most $D\cdot D!$ different pattern transitions.
-However, for other cases like $\tau>1$ and/or larger time shifts between ordinal patterns to be considered, up to $D!^2$ different pairs of patterns and, hence, transitions could occur in the worst case.
+However, if we consider $\tau=1$ and transitions corresponding to shifting the time index by $\Delta t=1$ time step only, from the $D$ components of the source permutation $D-1$ also contribute to the corresponding destination pattern without changing their mutual order relationships \cite{BorgesAMC2019,PessaPRE2019}.
+Hence, in this special case the new state contributing to the destination pattern can only lead to at most $D$ different ordinal patterns, i.e., we have at most $D\cdot D!$ different pattern transitions.
+However, in other situations with $\tau>1$ and/or larger time shifts $\Delta t$ between successive ordinal patterns to be considered, up to $D!^2$ different pairs of patterns and, hence, transitions could occur.
 In general, we conclude that the far higher number of pattern transitions as compared to patterns themselves puts severe restrictions to the choice of $D$ in practical applications of the new SCMs in the context of time series of a given length $L$.
@@
 {\color{red}Throughout the remainder of this work, we will consider $D$ as a free parameter, which is varied in the range of $D \in \{2,\ldots,7\}$, although for the sake of brevity, we will mainly show the results for $D = 6$ as illustration.
+
 The associated choice of the embedding delay $\tau$ is commonly less crucial in nonlinear time series analysis methods.
@@
 We emphasize that this choice is not based on specific considerations of performance optimization, but simply reflects convenience and serves solely illustrative purposes.
-Two examples for results obtained with $D=6$ and $\tau=1$ for two well-known model systems can be found in the Supplementary Material Sec.~SM2, Figs.~S4 and S5.}
+Two examples for results obtained with $D=6$ and $\tau=1$ for two well-known model systems can be found in the Supplementary Material Sec.~SM2, Figs.~S4 and S5.
+In comparison with the corresponding results presented in the following, different values of $\tau$ seem to commonly lead to similar qualitative behaviors of our entropy and SCM characteristics, yet possibly different "resolution" (e.g., when inter-comparing the obtained numerical values for periodic and chaotic time series).}
 
@@
 This linear increase of the maximum complexity versus entropy curve for sufficiently large embedding dimension is expected from theoretical considerations~\cite{SMCbook2010}.
-Some more discussions have been provided in the Supplementary Material (Sec.
-SM3, Fig.
-S6-S8).
-}
+Some additional discussion on this observation can be found in the Supplementary Material Sec.~ SM3, supported by the corresponding Fig.~S6.}
 
@@
 {\color{red}The behavior of the different SCMs relative to their respective lower and upper bounds can be further highlighted by considering normalized quantities.
-Specifically, for a given $\mathcal{H}$ and $D$, we define a normalized complexity $\mathcal{C}^{N} = (\mathcal{C} - \mathcal{C}_{min}) / (\mathcal{C}_{max} - \mathcal{C}_{min})$, which is theoretically bound to the interval $[0,1]$.
+Specifically, for a given $\mathcal{H}$ and $D$, we define a normalized complexity
+\begin{equation}
+\mathcal{C}^{N} = \frac{\mathcal{C} - \mathcal{C}_{min}}{\mathcal{C}_{max} - \mathcal{C}_{min}},
+\end{equation}
+\noindent
+which is theoretically bound to the interval $[0,1]$.
 Large values of those normalized complexities close to 1 indicate a good alignment with the maximum complexity curve.
 The corresponding results for the logistic map are shown in Fig.~\ref{fig:CHlogCmax}.
-As expected, for the traditional $\mathcal{C}_{O}$, the behavior is characterized by many points in the right half plane (large entropies, see Fig.~\ref{fig:CHlogCmax}(a)) for all considered embedding dimensions.
+As expected, for the traditional SCM $\mathcal{C}_{O}$, the behavior is characterized by many points in the right half plane (large entropies, see Fig.~\ref{fig:CHlogCmax}(a)) for all considered embedding dimensions.
 In contrast, there are no points in the range of $\mathcal{H} >0.8$ for $\mathcal{C}_W$ and $\mathcal{C}_E$ (Fig.~\ref{fig:CHlogCmax}(b, c)).
@@
 
-Especially for the new OPTN based SCMs, we find a partial overlap between the entropy values of fBm and the logistic map, while the associated SCM values are however clearly distinct except for the \textcolor{red}{fBm realizations} with the lowest entropy \textcolor{red}{values}, which correspond to the most nonstationary situations ($H$ close to 1).
-\textcolor{red}{In general, we observe for all three complexity--entropy planes of fBm that with increasing $D$, the associated range of entropy values extends further towards the left (lower entropies).
+Especially for the new OPTN based SCMs, we find a partial overlap between the entropy values of fBm and the logistic map, while the associated SCM values are however clearly distinct except for the \textcolor{red}{fBm realizations} with the lowest entropy \textcolor{red}{values}, which correspond to the most nonstationary situations ($H$ close to 1), \textcolor{red}{which can be further seen in the Supplementary Material Fig.~S7.
+In general, we observe for all three complexity--entropy planes of fBm that with increasing $D$, the associated range of entropy values extends further towards the left (lower entropies).
 Taken both, upward and leftward shift of the complexity--entropy curves individually may therefore enhance the risk of confusing deterministic-chaotic and stochastic processes in the complexity--entropy plane even with our alternative SCMs.
 However, our results indicate that} combining information on the position in the complexity--entropy plane and on the stationarity could allow a clear distinction between chaotic and stochastic dynamics even in extreme situations \textcolor{red}{(e.g., a large value of $H$)}.
-Notably, for the stationary fGn, all \textcolor{red}{three complexity--entropy planes} are confined to very high entropy and low complexity values -- a range where no combinations have been found for the new transition frequency based SCMs for the logistic map.
+Notably, for the stationary fGn, all \textcolor{red}{three complexity--entropy planes} are confined to very high entropy and low complexity values \textcolor{red}{(see also Supplementary Material Fig.~S8)} -- a range where no combinations have been found for the new transition frequency based SCMs for the logistic map.
 
@@
 
-As a general observation, we find that all three SCMs clearly follow the changes in the bifurcation diagram, for instance, showing distinct values in periodic windows (highlighted by grey color in Fig.~\ref{fig:bifurcation} \textcolor{red}{and Supplementary Material Fig.~S4, where the obtained results are shown for the same embedding dimension $D=6$, but two different values of the embedding delay $\tau$}).
-Other than $\mathcal{C}_O$, the two transition frequency based SCMs behave in a similar way as the Lyapunov exponent.
+As a general observation, we find that all three SCMs clearly follow the changes in the bifurcation diagram, for instance, showing distinct values in periodic windows (highlighted by grey color in Fig.~\ref{fig:bifurcation} \textcolor{red}{and Supplementary Material Fig.~S4, where the obtained results are shown for the same embedding dimension $D=6$, but two different values of the embedding delay $\tau$).
+Notably, the distinction between periodic and chaotic windows in terms of all entropy and SCM values is more pronounced for $\tau=6$ as compared to $\tau=1$, which supports our choice of a higher embedding delay for obtaining the results discussed in the main text.
+However,} other than $\mathcal{C}_O$ the two transition frequency based SCMs behave in a similar way as the Lyapunov exponent.
 Specifically, as the control parameter $r$ is increased, the chaoticity level grows gradually, reaching a maximum at $r = 4$, which is reflected by an increase in statistical complexity.
@@
 To avoid any corresponding discussion on proper choices of further algorithmic parameters, we employ here a Poincar\'e section to each sample trajectory of the system at $y=0$, $\dot{y}<0$ (for different values of $b$) and construct OPTNs from $N = 10,000$ intersection points.
-The results are shown in Fig.~\ref{fig:bifurRossler} \textcolor{red}{and Supplementary Material Fig.~S5 for the same embedding dimension $D=6$, but two different values of the embedding delay $\tau$}.
-Based on these results, we conclude that the general behavior of the different SCMs and associated entropy measures closely resembles that reported for the logistic map in Fig.~\ref{fig:bifurcation}.
+The results are shown in Fig.~\ref{fig:bifurRossler} \textcolor{red}{and Supplementary Material Fig.~S5 for the same embedding dimension $D=6$, but two different values of the embedding delay $\tau$.
+Again, we observe that the differences between periodic and chaotic windows become more pronounced for the larger value of the embedding delay $\tau$.
+In general, we find} that the \textcolor{red}{overall} behavior of the different SCMs and associated entropy measures closely resembles that reported for the logistic map in Fig.~\ref{fig:bifurcation}.
 Specifically, all measures trace the succession of bifurcations in the considered range of $b$.

### 2021-02-16 manuscript_Huang_rev.tex
After submission.

@@
         \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_Logistic.pdf}
-\caption{\textcolor{red}{(Color online) Normalized complexity -- entropy planes of the three SCMs for the logistic map: (a) $\mathcal{C}_O$, (b) $\mathcal{C}_W$, and (c) $\mathcal{C}_E$.
+\caption{\textcolor{red}{(Color online) Normalized complexity -- entropy planes of the three SCMs for the logistic map: (a) $\mathcal{C}_O^N$, (b) $\mathcal{C}_W^N$, and (c) $\mathcal{C}_E^N$.
 In (c), normalized complexity values larger than 1 indicate numerical artifacts when the control parameter $r$ approaches a bifurcation point, resulting in long transients.
@@
 \caption{\small{(Color online) Same as in Fig.~\ref{fig:CElogistic}, but for fractional Brownian motion with a Hurst exponent $H \in (0, 1)$ varying with a step size of $\Delta H=0.05$.
-Note that the normalized entropy is generally decreasing with increasing $H$ (red arrow).}  \label{fig:CEfBm}}
+Note that the normalized entropy is generally decreasing with increasing $H$ (red arrow).
+(a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
+(a-c) for $D = 4, 5$, and (d-f) for $D = 6$.
+}  \label{fig:CEfBm}}
 \end{figure*}
@@
 Note that only a small region of the complexity--entropy planes is covered in this case.
+(a, d) $\mathcal{C}_O$ versus $\mathcal{H}_O$, (b, e) $\mathcal{C}_{W}$ versus $\mathcal{H}_{W}$, and (c, f) $\mathcal{C}_{E}$ versus $\mathcal{H}_{E}$.
+(a-c) for $D = 4, 5$, and (d-f) for $D = 6$.
 }  \label{fig:CEfGn}}
@@
         \includegraphics[width=2\columnwidth]{CompEntropyCNormalized_fBm.pdf}
-\caption{\color{red}(Color online) Same as in Fig.~\ref{fig:CHlogCmax}, but for different realizations of fractional Brownian motion with varying Hurst exponent $H$.  \label{fig:CHfbmCmax}}
+\caption{\color{red}(Color online) Same as in Fig.~\ref{fig:CHlogCmax}, but for different realizations of fractional Brownian motion with varying Hurst exponent $H$.
+(a) $\mathcal{C}_O^N$, (b) $\mathcal{C}_W^N$, and (c) $\mathcal{C}_E^N$.
+\label{fig:CHfbmCmax}}
 \end{figure*}
@@
 \ref{fig:bifurcation}, but for Poincar\'e sections of the R\"ossler system while varying the control parameter $b$ (see text for details{\color{red}, embedding dimension $D = 6$).} Panel (a) presents a bifurcation diagram based on the $x$ components of all points in the Poincar\'e sections.
-Panel (b) shows the largest Lyapunov exponent of the system in dependence of $b$.} \label{fig:bifurRossler}}
+Panel (b) shows the largest Lyapunov exponent of the system in dependence of $b$.
+(c) $\mathcal{H}_O$ and (d) $\mathcal{C}_O$; (e) $\mathcal{H}_W$ and (f) $\mathcal{C}_W$; (g) $\mathcal{H}_E$ and (h) $\mathcal{C}_E$.
+} \label{fig:bifurRossler}}
 \end{figure*}
@@
 	\includegraphics[width=\columnwidth]{TS_ECG.pdf}
-\caption{\small{Three representative ECG recording sections (randomly chosen from different rhythmic states): (a) sinus rhythm (SR), (b) VF and VT.
+\caption{\small{Three representative ECG recording sections (randomly chosen from different rhythmic states): (a) sinus rhythm (SR), (b) ventricular fibrillation (VF) and (c) ventricular tachycardia (VT).
 The horizontal axis of each panel is the datum number (at 250~Hz \textcolor{red}{sampling rate}) and the vertical axis \textcolor{red}{describes the} ECG surface voltage (in \textcolor{red}{mV}).
@@
 In our case, AV-3 presents a higher complexity than AV-2, which should however be interpreted with care since the AV-2 series originates from an experiment, while the AV-3 series has been taken from a numerical simulation of the same system (hence, there may be a certain level of observational noise present in AV-2 but not in AV-3).
-The order of complexity values between periodic, quasiperiodic and chaotic solutions still holds for $\mathcal{C}_O$, but the corresponding values become very similar.
+The order of complexity values between periodic, quasiperiodic and chaotic solutions generally still holds for $\mathcal{C}_O$, but the corresponding values become very similar.
 In particular, it becomes hard for $\mathcal{C}_O$ to identify the difference between AV-2 and AV-3.
@@
 	\includegraphics[width=2\columnwidth]{fluidExample.pdf}
-\caption{\color{red}\small{OPTN based entropy and SCM characteristics for experimental fluid data, which are shown for embedding dimensions of $D = 4, 5, 6$.
+\caption{\color{red}\small{OPTN based entropy and SCM characteristics for experimental fluid data, which are shown for different embedding dimensions, (a) $D = 4$, (b) $D = 5$, and (c) $D = 6$.
 In all subfigures, stable wave (SW, $\blacklozenge$), quasiperiodic 2-frequency amplitude vacillation (AV-2, $\bullet$), quasiperiodic 3-frequency amplitude vacillation (AV-3, $\blacktriangle$) and chaotic modulated amplitude vacillation (MAV, $\blacksquare$) are distinguished.
@@
 Entropy and SCM scores are shown for embedding dimensions of $D = 4, 5, 6$.
+(a) $\mathcal{H}_O$ and (b) $\mathcal{C}_O$; (c) $\mathcal{H}_W$ and (d) $\mathcal{C}_W$; (e) $\mathcal{H}_E$ and (f) $\mathcal{C}_E$.
 } \label{fig:ecg}}
@@
 
+\section*{Supplementary Material}
+See Supplementary material: Sec.~SM1 with Figs.~S1 to S3 for the complexity--entropy planes when the self-loops are included in the transition matrix; Sec.~SM2 with Figs.~S4 and S5 for transition curves for the logistic map and the chaotic R\"ossler system when embedding parameters are $D=6$ and $\tau=1$; Sec.~ SM3 with Fig.~S6 for the linear increase of the maximum complexity $\mathcal{C}_{max}$ versus entropy curve for sufficiently large embedding dimension.
+However, this linear correlation in general is not expected by fractional Brownian motion and fractional Gaussian noise Sec.~ SM3 with Figs.~S7 and S8.
+
 \section*{Acknowledgements}
 Parts of this work have been financially supported by the National Natural Science Foundation of China (grant no.
-11872182, 11835003, 11875132, 61973086), the Natural Science Foundation of Shanghai (grant no.
+11872182, 11972288, 11835003, 11875132, 61973086), the Natural Science Foundation of Shanghai (grant no.
 18ZR1411800).
-JZ and YZ further acknowledge support by the Shanghai Municipal Science and Technology Major Project (grant no.
-2018SHZDZX01) and ZJ Lab.
+JZ and YZ further acknowledge support by the Shanghai Municipal Science and Technology Major Project and ZJ Lab (grant no.
+2018SHZDZX01).
 The authors thank J.
@@
 
-\bibliography{ref_Zou,ref_Donner}
+% \bibliography{ref_Zou,ref_Donner}
+%merlin.mbs aipnum4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
+%Control: key (0)
+%Control: author (8) initials jnrlst
+%Control: editor formatted (1) identically to author
+%Control: production of article title (0) allowed
+%Control: page (1) range
+%Control: year (1) truncated
+%Control: production of eprint (0) enabled
+\begin{thebibliography}{43}%
+\makeatletter
+\providecommand \@ifxundefined [1]{%
+ \@ifx{#1\undefined}
+}%
+\providecommand \@ifnum [1]{%
+ \ifnum #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \@ifx [1]{%
+ \ifx #1\expandafter \@firstoftwo
+ \else \expandafter \@secondoftwo
+ \fi
+}%
+\providecommand \natexlab [1]{#1}%
+\providecommand \enquote  [1]{``#1''}%
+\providecommand \bibnamefont  [1]{#1}%
+\providecommand \bibfnamefont [1]{#1}%
+\providecommand \citenamefont [1]{#1}%
+\providecommand \href@noop [0]{\@secondoftwo}%
+\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
+\providecommand \@href[1]{\@@startlink{#1}\@@href}%
+\providecommand \@@href[1]{\endgroup#1\@@endlink}%
+\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
+  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
+\providecommand \@@startlink[1]{}%
+\providecommand \@@endlink[0]{}%
+\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
+\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
+\providecommand \urlprefix  [0]{URL }%
+\providecommand \Eprint [0]{\href }%
+\providecommand \doibase [0]{http://dx.doi.org/}%
+\providecommand \selectlanguage [0]{\@gobble}%
+\providecommand \bibinfo  [0]{\@secondoftwo}%
+\providecommand \bibfield  [0]{\@secondoftwo}%
+\providecommand \translation [1]{[#1]}%
+\providecommand \BibitemOpen [0]{}%
+\providecommand \bibitemStop [0]{}%
+\providecommand \bibitemNoStop [0]{.\EOS\space}%
+\providecommand \EOS [0]{\spacefactor3000\relax}%
+\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
+\let\auto@bib@innerbib\@empty
+%</preamble>
+\bibitem [{\citenamefont {Zou}\ \emph {et~al.}(2019)\citenamefont {Zou},
+  \citenamefont {Donner}, \citenamefont {Marwan}, \citenamefont {Donges},\ and\
+  \citenamefont {Kurths}}]{ZouPR2018}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Zou}}, \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont {Donner}},
+  \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Marwan}}, \bibinfo
+  {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Complex network approaches to nonlinear time
+  series analysis},}\ }\href {\doibase 10.1016/j.physrep.2018.10.005}
+  {\bibfield  {journal} {\bibinfo  {journal} {Physics Reports}\ }\textbf
+  {\bibinfo {volume} {787}},\ \bibinfo {pages} {1--97} (\bibinfo {year}
+  {2019})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kantz}\ and\ \citenamefont
+  {Schreiber}(2004)}]{Kantz97}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
+  {Kantz}}\ and\ \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont
+  {Schreiber}},\ }\href@noop {} {\emph {\bibinfo {title} {Nonlinear Time Series
+  Analysis}}},\ \bibinfo {edition} {2nd}\ ed.\ (\bibinfo  {publisher}
+  {Cambridge University Press, Cambridge},\ \bibinfo {year} {2004})\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Sprott}(2003)}]{Sprott2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~C.}\ \bibnamefont
+  {Sprott}},\ }\href@noop {} {\emph {\bibinfo {title} {Chaos and Time-Series
+  Analysis}}}\ (\bibinfo  {publisher} {Oxford University Press, Oxford},\
+  \bibinfo {year} {2003})\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zhang}\ and\ \citenamefont
+  {Small}(2006)}]{ZhangPRL2006}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
+  {Zhang}}\ and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {{Complex network from
+  pseudoperiodic time series: Topology versus dynamics}},}\ }\href {\doibase
+  10.1103/PhysRevLett.96.238701} {\bibfield  {journal} {\bibinfo  {journal}
+  {{Phys. Rev. Lett.}}\ }\textbf {\bibinfo {volume} {{96}}},\ \bibinfo {pages}
+  {238701} (\bibinfo {year} {{2006}})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Marwan}\ \emph {et~al.}(2009)\citenamefont {Marwan},
+  \citenamefont {Donges}, \citenamefont {Zou}, \citenamefont {Donner},\ and\
+  \citenamefont {Kurths}}]{MarwanPLA2009}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.}~\bibnamefont
+  {Marwan}}, \bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}},
+  \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Zou}}, \bibinfo {author}
+  {\bibfnamefont {R.~V.}\ \bibnamefont {Donner}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Complex network approach for recurrence analysis of time
+  series},}\ }\href {\doibase https://doi.org/10.1016/j.physleta.2009.09.042}
+  {\bibfield  {journal} {\bibinfo  {journal} {Phys. Lett. A}\ }\textbf
+  {\bibinfo {volume} {373}},\ \bibinfo {pages} {4246 -- 4254} (\bibinfo {year}
+  {2009})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Donner}\ \emph {et~al.}(2010)\citenamefont {Donner},
+  \citenamefont {Zou}, \citenamefont {Donges}, \citenamefont {Marwan},\ and\
+  \citenamefont {Kurths}}]{Donner2010a}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont
+  {Donner}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {Zou}},
+  \bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont {Donges}}, \bibinfo
+  {author} {\bibfnamefont {N.}~\bibnamefont {Marwan}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {{Recurrence networks --- a novel paradigm for nonlinear
+  time series analysis}},}\ }\href {\doibase 10.1088/1367-2630/12/3/033025}
+  {\bibfield  {journal} {\bibinfo  {journal} {New J. Phys.}\ }\textbf {\bibinfo
+  {volume} {12}},\ \bibinfo {pages} {033025} (\bibinfo {year}
+  {2010})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Lacasa}\ \emph {et~al.}(2008)\citenamefont {Lacasa},
+  \citenamefont {Luque}, \citenamefont {Ballesteros}, \citenamefont {Luque},\
+  and\ \citenamefont {Nuno}}]{Lacasa2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {Lacasa}}, \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Luque}},
+  \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {Ballesteros}}, \bibinfo
+  {author} {\bibfnamefont {J.}~\bibnamefont {Luque}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.~C.}\ \bibnamefont {Nuno}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {{From time series to complex networks: The visibility
+  graph}},}\ }\href {\doibase 10.1073/pnas.0709247105} {\bibfield  {journal}
+  {\bibinfo  {journal} {Proc. Natl. Acad. Sci.}\ }\textbf {\bibinfo {volume}
+  {105}},\ \bibinfo {pages} {4972--4975} (\bibinfo {year} {2008})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Nu{\~{n}}ez}\ \emph {et~al.}(2012)\citenamefont
+  {Nu{\~{n}}ez}, \citenamefont {Lacasa}, \citenamefont {Patricio},\ and\
+  \citenamefont {Luque}}]{Nunez2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Nu{\~{n}}ez}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Lacasa}},
+  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Patricio}}, \ and\
+  \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont {Luque}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {{Visibility Algorithms: A Short
+  Review}},}\ }in\ \href {\doibase 10.5772/34810} {\emph {\bibinfo {booktitle}
+  {New Front. Graph Theory}}}\ (\bibinfo  {publisher} {InTech},\ \bibinfo
+  {year} {2012})\ Chap.~\bibinfo {chapter} {6}, pp.\ \bibinfo {pages}
+  {119--152}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Nicolis}, \citenamefont {Cantu},\ and\ \citenamefont
+  {Nicolis}(2005)}]{Nicolis2005}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
+  {Nicolis}}, \bibinfo {author} {\bibfnamefont {A.~G.}\ \bibnamefont {Cantu}},
+  \ and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Nicolis}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {{Dynamical aspects of
+  interaction networks}},}\ }\href {\doibase 10.1142/S0218127405014167}
+  {\bibfield  {journal} {\bibinfo  {journal} {Int. J. Bifurc. Chaos}\ }\textbf
+  {\bibinfo {volume} {15}},\ \bibinfo {pages} {3467--3480} (\bibinfo {year}
+  {2005})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {McCullough}\ \emph {et~al.}(2015)\citenamefont
+  {McCullough}, \citenamefont {Small}, \citenamefont {Stemler},\ and\
+  \citenamefont {Iu}}]{MichaelChaos2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {McCullough}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},
+  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Stemler}}, \ and\
+  \bibinfo {author} {\bibfnamefont {H.~H.-C.}\ \bibnamefont {Iu}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Time lagged ordinal partition networks
+  for capturing dynamics of continuous dynamical systems},}\ }\href {\doibase
+  10.1063/1.4919075} {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\
+  }\textbf {\bibinfo {volume} {25}},\ \bibinfo {pages} {053101} (\bibinfo
+  {year} {2015})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kulp}\ \emph {et~al.}(2016)\citenamefont {Kulp},
+  \citenamefont {Chobot}, \citenamefont {Freitas},\ and\ \citenamefont
+  {Sprechini}}]{KulpChaos2016}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.~W.}\ \bibnamefont
+  {Kulp}}, \bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont {Chobot}},
+  \bibinfo {author} {\bibfnamefont {H.~R.}\ \bibnamefont {Freitas}}, \ and\
+  \bibinfo {author} {\bibfnamefont {G.~D.}\ \bibnamefont {Sprechini}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Using ordinal partition
+  transition networks to analyze {ECG} data},}\ }\href {\doibase
+  10.1063/1.4959537} {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\
+  }\textbf {\bibinfo {volume} {26}},\ \bibinfo {pages} {073114} (\bibinfo
+  {year} {2016})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zhang}\ \emph {et~al.}(2017)\citenamefont {Zhang},
+  \citenamefont {Zhou}, \citenamefont {Tang}, \citenamefont {Guo},
+  \citenamefont {Small},\ and\ \citenamefont {Zou}}]{zhangSciRep2017}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~Y.}\ \bibnamefont
+  {Zhang}}, \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Zhou}},
+  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Tang}}, \bibinfo {author}
+  {\bibfnamefont {H.}~\bibnamefont {Guo}}, \bibinfo {author} {\bibfnamefont
+  {M.}~\bibnamefont {Small}}, \ and\ \bibinfo {author} {\bibfnamefont
+  {Y.}~\bibnamefont {Zou}},\ }\bibfield  {title} {\enquote {\bibinfo {title}
+  {Constructing ordinal partition transition networks from multivariate time
+  series},}\ }\href {\doibase 10.1038/s41598-017-08245-x} {\bibfield  {journal}
+  {\bibinfo  {journal} {Scientific Reports}\ }\textbf {\bibinfo {volume} {7}},\
+  \bibinfo {pages} {7795} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {McCullough}\ \emph {et~al.}(2017)\citenamefont
+  {McCullough}, \citenamefont {Small}, \citenamefont {Iu},\ and\ \citenamefont
+  {Stemler}}]{McCullough2017b}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {McCullough}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Small}},
+  \bibinfo {author} {\bibfnamefont {H.~H.~C.}\ \bibnamefont {Iu}}, \ and\
+  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {Stemler}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {{Multiscale ordinal network analysis of
+  human cardiac dynamics}},}\ }\href {\doibase 10.1098/rsta.2016.0292}
+  {\bibfield  {journal} {\bibinfo  {journal} {Philos. Trans. R. Soc. A Math.
+  Phys. Eng. Sci.}\ }\textbf {\bibinfo {volume} {375}},\ \bibinfo {pages}
+  {20160292} (\bibinfo {year} {2017})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Borges}\ \emph {et~al.}(2019)\citenamefont {Borges},
+  \citenamefont {Ramos}, \citenamefont {Mini}, \citenamefont {Rosso},
+  \citenamefont {Frery},\ and\ \citenamefont {Loureiro}}]{BorgesAMC2019}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~B.}\ \bibnamefont
+  {Borges}}, \bibinfo {author} {\bibfnamefont {H.~S.}\ \bibnamefont {Ramos}},
+  \bibinfo {author} {\bibfnamefont {R.~A.}\ \bibnamefont {Mini}}, \bibinfo
+  {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}}, \bibinfo {author}
+  {\bibfnamefont {A.~C.}\ \bibnamefont {Frery}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {A.~A.~F.}\ \bibnamefont {Loureiro}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Learning and distinguishing time series dynamics
+  via ordinal patterns transition graphs},}\ }\href {\doibase
+  https://doi.org/10.1016/j.amc.2019.06.068} {\bibfield  {journal} {\bibinfo
+  {journal} {Applied Mathematics and Computation}\ }\textbf {\bibinfo {volume}
+  {362}},\ \bibinfo {pages} {124554} (\bibinfo {year} {2019})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Subramaniyam}\ \emph {et~al.}(view)\citenamefont
+  {Subramaniyam}, \citenamefont {Donner}, \citenamefont {Caron}, \citenamefont
+  {Panuccio},\ and\ \citenamefont {Hyttinen}}]{Subramaniyam2020}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {N.~P.}\ \bibnamefont
+  {Subramaniyam}}, \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont
+  {Donner}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Caron}},
+  \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Panuccio}}, \ and\
+  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Hyttinen}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Causal coupling inference from
+  multivariate time series based on ordinal partition transition networks},}\
+  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Nonlinear
+  Dynamics}\ } (\bibinfo {year} {in review})},\ \bibinfo {note} {arXiv preprint
+  2010.00948}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {{Bandt}}\ and\ \citenamefont
+  {{Pompe}}(2002)}]{BandtPRL2002}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {C.}~\bibnamefont
+  {{Bandt}}}\ and\ \bibinfo {author} {\bibfnamefont {B.}~\bibnamefont
+  {{Pompe}}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Permutation
+  entropy: A natural complexity measure for time series},}\ }\href {\doibase
+  10.1103/PhysRevLett.88.174102} {\bibfield  {journal} {\bibinfo  {journal}
+  {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {88}},\ \bibinfo {pages}
+  {174102} (\bibinfo {year} {2002})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Amig{\'o}}, \citenamefont {Keller},\ and\
+  \citenamefont {Unakafova}(2014)}]{AmigoPTRSA2014}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
+  {Amig{\'o}}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {Keller}}, \
+  and\ \bibinfo {author} {\bibfnamefont {V.~A.}\ \bibnamefont {Unakafova}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Ordinal symbolic analysis
+  and its application to biomedical recordings},}\ }\href {\doibase
+  10.1098/rsta.2014.0091} {\bibfield  {journal} {\bibinfo  {journal} {Phil.
+  Trans. R. Soc. A}\ }\textbf {\bibinfo {volume} {373}},\ \bibinfo {pages}
+  {20140091} (\bibinfo {year} {2014})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zanin}(2008)}]{ZaninChaos2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Zanin}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Forbidden
+  patterns in financial time series},}\ }\href {\doibase 10.1063/1.2841197}
+  {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\ }\textbf {\bibinfo
+  {volume} {18}},\ \bibinfo {pages} {013119} (\bibinfo {year}
+  {2008})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Barreiro}, \citenamefont {Marti},\ and\ \citenamefont
+  {Masoller}(2011)}]{BarreiroChaos2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Barreiro}}, \bibinfo {author} {\bibfnamefont {A.~C.}\ \bibnamefont {Marti}},
+  \ and\ \bibinfo {author} {\bibfnamefont {C.}~\bibnamefont {Masoller}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Inferring long memory
+  processes in the climate network via ordinal pattern analysis},}\ }\href
+  {\doibase 10.1063/1.3545273} {\bibfield  {journal} {\bibinfo  {journal}
+  {Chaos}\ }\textbf {\bibinfo {volume} {21}},\ \bibinfo {pages} {013101}
+  (\bibinfo {year} {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Takens}(1981)}]{Takens1981}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
+  {Takens}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {{Detecting
+  strange attractors in turbulence}},}\ }in\ \href {\doibase
+  10.1007/BFb0091924} {\emph {\bibinfo {booktitle} {Dynamical Systems and
+  Turbulence, Warwick 1980}}},\ \bibinfo {series} {Lecture Notes in
+  Mathematics}, Vol.\ \bibinfo {volume} {898},\ \bibinfo {editor} {edited by\
+  \bibinfo {editor} {\bibfnamefont {D.}~\bibnamefont {Rand}}\ and\ \bibinfo
+  {editor} {\bibfnamefont {L.-S.}\ \bibnamefont {Young}}}\ (\bibinfo
+  {publisher} {Springer, New York},\ \bibinfo {year} {1981})\ pp.\ \bibinfo
+  {pages} {366--381}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Amig\'o}, \citenamefont {Zambrano},\ and\
+  \citenamefont {Sanju\'an}(2007)}]{AmigoEPL2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
+  {Amig\'o}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Zambrano}}, \
+  and\ \bibinfo {author} {\bibfnamefont {M.~A.~F.}\ \bibnamefont {Sanju\'an}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {True and false forbidden
+  patterns in deterministic and random dynamics},}\ }\href {\doibase
+  https://doi.org/10.1209/0295-5075/79/50001} {\bibfield  {journal} {\bibinfo
+  {journal} {Europhys. Lett.}\ }\textbf {\bibinfo {volume} {79}},\ \bibinfo
+  {pages} {50001} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Amig\'o}, \citenamefont {Zambrano},\ and\
+  \citenamefont {Sanju\'an}(2008)}]{AmigoEPL2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont
+  {Amig\'o}}, \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {Zambrano}}, \
+  and\ \bibinfo {author} {\bibfnamefont {M.~A.~F.}\ \bibnamefont {Sanju\'an}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Combinatorial detection of
+  determinism in noisy time series},}\ }\href {\doibase
+  https://doi.org/10.1209/0295-5075/83/60005} {\bibfield  {journal} {\bibinfo
+  {journal} {Europhys. Lett.}\ }\textbf {\bibinfo {volume} {83}},\ \bibinfo
+  {pages} {60005} (\bibinfo {year} {2008})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Rosso}\ \emph {et~al.}(2012)\citenamefont {Rosso},
+  \citenamefont {Carpi}, \citenamefont {Saco}, \citenamefont {Ravetti},
+  \citenamefont {Larrondo},\ and\ \citenamefont {Plastino}}]{RossoEPJB2012}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont
+  {Rosso}}, \bibinfo {author} {\bibfnamefont {L.~C.}\ \bibnamefont {Carpi}},
+  \bibinfo {author} {\bibfnamefont {P.~M.}\ \bibnamefont {Saco}}, \bibinfo
+  {author} {\bibfnamefont {M.~G.}\ \bibnamefont {Ravetti}}, \bibinfo {author}
+  {\bibfnamefont {H.~A.}\ \bibnamefont {Larrondo}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Plastino}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {The {A}mig{\'o} paradigm of forbidden/missing patterns: a
+  detailed analysis},}\ }\href {\doibase 10.1140/epjb/e2012-30307-8} {\bibfield
+   {journal} {\bibinfo  {journal} {Eur. Phys. J. B}\ }\textbf {\bibinfo
+  {volume} {85}},\ \bibinfo {pages} {419} (\bibinfo {year} {2012})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Kowalski}\ \emph {et~al.}(2011)\citenamefont
+  {Kowalski}, \citenamefont {Martin}, \citenamefont {Plastino}, \citenamefont
+  {Rosso},\ and\ \citenamefont {Casas}}]{kowalskiEntropy2011}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~M.}\ \bibnamefont
+  {Kowalski}}, \bibinfo {author} {\bibfnamefont {M.~T.}\ \bibnamefont
+  {Martin}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}},
+  \bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}}, \ and\
+  \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Casas}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Distances in probability space and the
+  statistical complexity setup},}\ }\href {\doibase 10.3390/e13061055}
+  {\bibfield  {journal} {\bibinfo  {journal} {Entropy}\ }\textbf {\bibinfo
+  {volume} {13}},\ \bibinfo {pages} {1055--1075} (\bibinfo {year}
+  {2011})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {L\'opez-Ruiz}, \citenamefont {Mancini},\ and\
+  \citenamefont {Calbet}(1995)}]{LopezPLA1995}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {L\'opez-Ruiz}}, \bibinfo {author} {\bibfnamefont {H.~L.}\ \bibnamefont
+  {Mancini}}, \ and\ \bibinfo {author} {\bibfnamefont {X.}~\bibnamefont
+  {Calbet}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {A statistical
+  measure of complexity},}\ }\href {\doibase
+  https://doi.org/10.1016/0375-9601(95)00867-5} {\bibfield  {journal} {\bibinfo
+   {journal} {Phys. Lett. A}\ }\textbf {\bibinfo {volume} {209}},\ \bibinfo
+  {pages} {321 -- 326} (\bibinfo {year} {1995})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Rosso}\ \emph
+  {et~al.}(2007{\natexlab{a}})\citenamefont {Rosso}, \citenamefont {Larrondo},
+  \citenamefont {Martin}, \citenamefont {Plastino},\ and\ \citenamefont
+  {Fuentes}}]{rossoPRL2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont
+  {Rosso}}, \bibinfo {author} {\bibfnamefont {H.~A.}\ \bibnamefont {Larrondo}},
+  \bibinfo {author} {\bibfnamefont {M.~T.}\ \bibnamefont {Martin}}, \bibinfo
+  {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}}, \ and\ \bibinfo
+  {author} {\bibfnamefont {M.~A.}\ \bibnamefont {Fuentes}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Distinguishing noise from chaos},}\
+  }\href {\doibase 10.1103/PhysRevLett.99.154102} {\bibfield  {journal}
+  {\bibinfo  {journal} {Phys. Rev. Lett.}\ }\textbf {\bibinfo {volume} {99}},\
+  \bibinfo {pages} {154102} (\bibinfo {year} {2007}{\natexlab{a}})}\BibitemShut
+  {NoStop}%
+\bibitem [{\citenamefont {Lange}, \citenamefont {Rosso},\ and\ \citenamefont
+  {Hauhs}(2013)}]{Lange2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
+  {Lange}}, \bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}}, \
+  and\ \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Hauhs}},\ }\bibfield
+   {title} {\enquote {\bibinfo {title} {Ordinal pattern and statistical
+  complexity analysis of daily stream flow time series},}\ }\href {\doibase
+  10.1140/epjst/e2013-01858-3} {\bibfield  {journal} {\bibinfo  {journal} {Eur.
+  Phys. J. S.T.}\ }\textbf {\bibinfo {volume} {222}},\ \bibinfo {pages}
+  {535--552} (\bibinfo {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Martin}, \citenamefont {Plastino},\ and\
+  \citenamefont {Rosso}(2003)}]{MartinPLA2003}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Martin}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}}, \
+  and\ \bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Statistical complexity and
+  disequilibrium},}\ }\href {\doibase
+  https://doi.org/10.1016/S0375-9601(03)00491-2} {\bibfield  {journal}
+  {\bibinfo  {journal} {Phys. Lett. A}\ }\textbf {\bibinfo {volume} {311}},\
+  \bibinfo {pages} {126 -- 132} (\bibinfo {year} {2003})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Bates}\ and\ \citenamefont
+  {Shepard}(1993)}]{batesPLA1993}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~E.}\ \bibnamefont
+  {Bates}}\ and\ \bibinfo {author} {\bibfnamefont {H.~K.}\ \bibnamefont
+  {Shepard}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {Measuring
+  complexity using information fluctuation},}\ }\href {\doibase
+  https://doi.org/10.1016/0375-9601(93)90232-O} {\bibfield  {journal} {\bibinfo
+   {journal} {Phys. Lett. A}\ }\textbf {\bibinfo {volume} {172}},\ \bibinfo
+  {pages} {416 -- 425} (\bibinfo {year} {1993})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {da~F.~Costa}\ \emph {et~al.}(2007)\citenamefont
+  {da~F.~Costa}, \citenamefont {Rodrigues}, \citenamefont {Travieso},\ and\
+  \citenamefont {Boas}}]{CostaADPhy2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {L.}~\bibnamefont
+  {da~F.~Costa}}, \bibinfo {author} {\bibfnamefont {F.~A.}\ \bibnamefont
+  {Rodrigues}}, \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont {Travieso}},
+  \ and\ \bibinfo {author} {\bibfnamefont {P.~R.~V.}\ \bibnamefont {Boas}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Characterization of complex
+  networks: A survey of measurements},}\ }\href {\doibase
+  10.1080/00018730601170527} {\bibfield  {journal} {\bibinfo  {journal}
+  {Advances in Physics}\ }\textbf {\bibinfo {volume} {56}},\ \bibinfo {pages}
+  {167--242} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Pessa}\ and\ \citenamefont
+  {Ribeiro}(2019)}]{PessaPRE2019}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~A.~B.}\
+  \bibnamefont {Pessa}}\ and\ \bibinfo {author} {\bibfnamefont {H.~V.}\
+  \bibnamefont {Ribeiro}},\ }\bibfield  {title} {\enquote {\bibinfo {title}
+  {Characterizing stochastic time series with ordinal networks},}\ }\href
+  {\doibase 10.1103/PhysRevE.100.042304} {\bibfield  {journal} {\bibinfo
+  {journal} {Phys. Rev. E}\ }\textbf {\bibinfo {volume} {100}},\ \bibinfo
+  {pages} {042304} (\bibinfo {year} {2019})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Kowalski}\ \emph {et~al.}(2007)\citenamefont
+  {Kowalski}, \citenamefont {Martín}, \citenamefont {Plastino},\ and\
+  \citenamefont {Rosso}}]{kowalskiPhyD2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
+  {Kowalski}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Martín}},
+  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}}, \ and\
+  \bibinfo {author} {\bibfnamefont {O.}~\bibnamefont {Rosso}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Bandt-pompe approach to the
+  classical-quantum transition},}\ }\href {\doibase
+  https://doi.org/10.1016/j.physd.2007.06.015} {\bibfield  {journal} {\bibinfo
+  {journal} {Physica D}\ }\textbf {\bibinfo {volume} {233}},\ \bibinfo {pages}
+  {21 -- 31} (\bibinfo {year} {2007})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Riedl}, \citenamefont {M\"uller},\ and\ \citenamefont
+  {Wessel}(2013)}]{Riedl2013}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Riedl}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {M\"uller}}, \
+  and\ \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {Wessel}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Practical considerations of
+  permutation entropy},}\ }\href {\doibase 10.1140/epjst/e2013-01862-7}
+  {\bibfield  {journal} {\bibinfo  {journal} {Eur. Phys. J. Spec. Top.}\
+  }\textbf {\bibinfo {volume} {222}},\ \bibinfo {pages} {249 -- 262} (\bibinfo
+  {year} {2013})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Rosso}\ \emph
+  {et~al.}(2007{\natexlab{b}})\citenamefont {Rosso}, \citenamefont {Zunino},
+  \citenamefont {P\'erez}, \citenamefont {Figliola}, \citenamefont {Larrondo},
+  \citenamefont {Garavaglia}, \citenamefont {Mart\'{\i}n},\ and\ \citenamefont
+  {Plastino}}]{RossoPRE2007}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont
+  {Rosso}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Zunino}},
+  \bibinfo {author} {\bibfnamefont {D.~G.}\ \bibnamefont {P\'erez}}, \bibinfo
+  {author} {\bibfnamefont {A.}~\bibnamefont {Figliola}}, \bibinfo {author}
+  {\bibfnamefont {H.~A.}\ \bibnamefont {Larrondo}}, \bibinfo {author}
+  {\bibfnamefont {M.}~\bibnamefont {Garavaglia}}, \bibinfo {author}
+  {\bibfnamefont {M.~T.}\ \bibnamefont {Mart\'{\i}n}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {A.}~\bibnamefont {Plastino}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Extracting features of gaussian self-similar stochastic
+  processes via the bandt-pompe approach},}\ }\href {\doibase
+  10.1103/PhysRevE.76.061114} {\bibfield  {journal} {\bibinfo  {journal} {Phys.
+  Rev. E}\ }\textbf {\bibinfo {volume} {76}},\ \bibinfo {pages} {061114}
+  (\bibinfo {year} {2007}{\natexlab{b}})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Martin}, \citenamefont {Plastino},\ and\
+  \citenamefont {Rosso}(2006)}]{martinPhyA2006}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Martin}}, \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {Plastino}}, \
+  and\ \bibinfo {author} {\bibfnamefont {O.~A.}\ \bibnamefont {Rosso}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {Generalized statistical
+  complexity measures: Geometrical and analytical properties},}\ }\href
+  {\doibase https://doi.org/10.1016/j.physa.2005.11.053} {\bibfield  {journal}
+  {\bibinfo  {journal} {Physica A}\ }\textbf {\bibinfo {volume} {369}},\
+  \bibinfo {pages} {439 -- 462} (\bibinfo {year} {2006})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {L.-Ruiz}, \citenamefont {Mancini},\ and\
+  \citenamefont {Calbet}(2010)}]{SMCbook2010}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {R.}~\bibnamefont
+  {L.-Ruiz}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {Mancini}}, \
+  and\ \bibinfo {author} {\bibfnamefont {X.}~\bibnamefont {Calbet}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {A statistical measure of
+  complexity},}\ }in\ \href {\doibase 10.2174/9781608057603113010012} {\emph
+  {\bibinfo {booktitle} {Concepts and Recent Advances in Generalized
+  Information Measures and Statistics}}},\ \bibinfo {editor} {edited by\
+  \bibinfo {editor} {\bibfnamefont {A.~M.}\ \bibnamefont {Kowalski}}, \bibinfo
+  {editor} {\bibfnamefont {R.~D.}\ \bibnamefont {Rossignoli}}, \ and\ \bibinfo
+  {editor} {\bibfnamefont {E.~M.~F.}\ \bibnamefont {Curado}}}\ (\bibinfo
+  {publisher} {Bentham Science Publishers},\ \bibinfo {year} {2010})\ pp.\
+  \bibinfo {pages} {147--168}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zou}, \citenamefont {Donner},\ and\ \citenamefont
+  {Kurths}(2015)}]{Zou2015}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Zou}}, \bibinfo {author} {\bibfnamefont {R.~V.}\ \bibnamefont {Donner}}, \
+  and\ \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Kurths}},\
+  }\bibfield  {title} {\enquote {\bibinfo {title} {{Analyzing long-term
+  correlated stochastic processes by means of recurrence networks: Potentials
+  and pitfalls}},}\ }\href {\doibase 10.1103/PhysRevE.91.022926} {\bibfield
+  {journal} {\bibinfo  {journal} {Phys. Rev. E}\ }\textbf {\bibinfo {volume}
+  {91}},\ \bibinfo {pages} {022926} (\bibinfo {year} {2015})},\ \Eprint
+  {http://arxiv.org/abs/1409.3613} {1409.3613} \BibitemShut {NoStop}%
+\bibitem [{\citenamefont {R{\"o}ssler}(1976)}]{Roessler1976}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {O.~E.}\ \bibnamefont
+  {R{\"o}ssler}},\ }\bibfield  {title} {\enquote {\bibinfo {title} {An equation
+  for continuous chaos},}\ }\href {\doibase 10.1016/0375-9601(76)90101-8}
+  {\bibfield  {journal} {\bibinfo  {journal} {Phys. Lett. A}\ }\textbf
+  {\bibinfo {volume} {57}},\ \bibinfo {pages} {397--398} (\bibinfo {year}
+  {1976})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Read}\ \emph {et~al.}(1992)\citenamefont {Read},
+  \citenamefont {Bel}, \citenamefont {Johnson},\ and\ \citenamefont
+  {Small}}]{Read_jfm_1992}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.~L.}\ \bibnamefont
+  {Read}}, \bibinfo {author} {\bibfnamefont {M.~J.}\ \bibnamefont {Bel}},
+  \bibinfo {author} {\bibfnamefont {D.~W.}\ \bibnamefont {Johnson}}, \ and\
+  \bibinfo {author} {\bibfnamefont {R.~M.}\ \bibnamefont {Small}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {Quasi-periodic and chaotic flow regimes
+  in a thermally driven, rotating fluid a nnulus},}\ }\href@noop {} {\bibfield
+  {journal} {\bibinfo  {journal} {Journal of Fluid Mechanics}\ }\textbf
+  {\bibinfo {volume} {238}},\ \bibinfo {pages} {599--632} (\bibinfo {year}
+  {1992})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Zou}\ \emph {et~al.}(2008)\citenamefont {Zou},
+  \citenamefont {Thiel}, \citenamefont {Romano}, \citenamefont {Read},\ and\
+  \citenamefont {Kurths}}]{ZouEPJST2008}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
+  {Zou}}, \bibinfo {author} {\bibfnamefont {M.}~\bibnamefont {Thiel}}, \bibinfo
+  {author} {\bibfnamefont {M.}~\bibnamefont {Romano}}, \bibinfo {author}
+  {\bibfnamefont {P.}~\bibnamefont {Read}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Recurrence analysis of quasiperiodicity in experimental
+  fluid data},}\ }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal}
+  {Eur. Phys. J. Spec. Top.}\ }\textbf {\bibinfo {volume} {164}},\ \bibinfo
+  {pages} {23--33} (\bibinfo {year} {2008})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Thiel}\ \emph {et~al.}(2004)\citenamefont {Thiel},
+  \citenamefont {Romano}, \citenamefont {Read},\ and\ \citenamefont
+  {Kurths}}]{thiel2004a}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Thiel}}, \bibinfo {author} {\bibfnamefont {M.~C.}\ \bibnamefont {Romano}},
+  \bibinfo {author} {\bibfnamefont {P.~L.}\ \bibnamefont {Read}}, \ and\
+  \bibinfo {author} {\bibfnamefont {J.}~\bibnamefont {Kurths}},\ }\bibfield
+  {title} {\enquote {\bibinfo {title} {{Estimation of dynamical invariants
+  without embedding by recurrence plots}},}\ }\href {\doibase
+  10.1063/1.1667633} {\bibfield  {journal} {\bibinfo  {journal} {Chaos}\
+  }\textbf {\bibinfo {volume} {14}},\ \bibinfo {pages} {234--243} (\bibinfo
+  {year} {2004})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Goldberger}\ \emph {et~al.}(2000)\citenamefont
+  {Goldberger}, \citenamefont {Amaral}, \citenamefont {Glass}, \citenamefont
+  {Hausdorff}, \citenamefont {Ivanov}, \citenamefont {Mark}, \citenamefont
+  {Mietus}, \citenamefont {Moody}, \citenamefont {Peng},\ and\ \citenamefont
+  {Stanley}}]{GoldbergerMITBIH2000}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.~L.}\ \bibnamefont
+  {Goldberger}}, \bibinfo {author} {\bibfnamefont {L.~A.~N.}\ \bibnamefont
+  {Amaral}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {Glass}},
+  \bibinfo {author} {\bibfnamefont {J.~M.}\ \bibnamefont {Hausdorff}}, \bibinfo
+  {author} {\bibfnamefont {P.~C.}\ \bibnamefont {Ivanov}}, \bibinfo {author}
+  {\bibfnamefont {R.~G.}\ \bibnamefont {Mark}}, \bibinfo {author}
+  {\bibfnamefont {J.~E.}\ \bibnamefont {Mietus}}, \bibinfo {author}
+  {\bibfnamefont {G.~B.}\ \bibnamefont {Moody}}, \bibinfo {author}
+  {\bibfnamefont {C.-K.}\ \bibnamefont {Peng}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {H.~E.}\ \bibnamefont {Stanley}},\ }\bibfield  {title}
+  {\enquote {\bibinfo {title} {Physiobank, physiotoolkit, and physionet},}\
+  }\href {\doibase 10.1161/01.CIR.101.23.e215} {\bibfield  {journal} {\bibinfo
+  {journal} {Circulation}\ }\textbf {\bibinfo {volume} {101}},\ \bibinfo
+  {pages} {e215--e220} (\bibinfo {year} {2000})}\BibitemShut {NoStop}%
+\bibitem [{\citenamefont {Small}\ \emph {et~al.}(2002)\citenamefont {Small},
+  \citenamefont {Yu}, \citenamefont {Simonotto}, \citenamefont {Harrison},
+  \citenamefont {Grubb},\ and\ \citenamefont {Fox}}]{smallCSF2002}%
+  \BibitemOpen
+  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {M.}~\bibnamefont
+  {Small}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {Yu}}, \bibinfo
+  {author} {\bibfnamefont {J.}~\bibnamefont {Simonotto}}, \bibinfo {author}
+  {\bibfnamefont {R.~G.}\ \bibnamefont {Harrison}}, \bibinfo {author}
+  {\bibfnamefont {N.}~\bibnamefont {Grubb}}, \ and\ \bibinfo {author}
+  {\bibfnamefont {K.}~\bibnamefont {Fox}},\ }\bibfield  {title} {\enquote
+  {\bibinfo {title} {Uncovering non-linear structure in human ecg
+  recordings},}\ }\href {\doibase
+  https://doi.org/10.1016/S0960-0779(01)00168-0} {\bibfield  {journal}
+  {\bibinfo  {journal} {Chaos, Solitons \& Fractals}\ }\textbf {\bibinfo
+  {volume} {13}},\ \bibinfo {pages} {1755 -- 1762} (\bibinfo {year}
+  {2002})}\BibitemShut {NoStop}%
+\end{thebibliography}%
 
